---
title: "00_ADHDCYC_Setup_Pkgs"
author: "Tory Eisenlohr-Moul"
date: "`r Sys.Date()`"
output: html_document
---

# ---- Load Libraries ----


## Project Management, Version Control, and Reproducibility
```{r project-management}
# Libraries for project and version control
#require(usethis)     # Setting up GitHub repositories and managing projects
#require(gitcreds)    # Managing Git credentials
#require(targets)     # Workflow management for reproducible research
#require(roxygen2)    # Documentation generation for R packages
#require(testthat)    # Unit testing for R packages to ensure code quality
require(conflicted)  # Manage conflicts between functions from different packages
#require(rprojroot)   # Finding the root of an R project
#require(here)        # Finding files within an R project
#require(renv)        # Dependency management for reproducibility
#renv::config() # Display the renv configuration
#renv::activate() # Activate the renv environment
#renv::deactivate(clean = TRUE)


```



## Data Handling, Plot Arrangement, and Visualization
```{r data-visualization}
# Libraries for data manipulation, cleaning, arranging, and visualization

require(patchwork)        # Easily combine multiple ggplot2 plots
require(ggplot2)          # Core data visualization library
require(ggdist)           # Visualizing distributions and uncertainty
require(ggforce)          # Extending ggplot2 with advanced features
require(ggrepel)          # Non-overlapping text labels for ggplot2
require(sjPlot)           # Plotting for statistical models (e.g., mixed models)
require(corrplot)         # Correlation matrix visualizations
require(see)              # Visualizing model checks and diagnostics
require(ggthemes)         # Additional themes for ggplot2
require(ggpubr)           # Publication-ready plots with ggplot2
#require(gganimate)        # Animated visualizations with ggplot2
#require(scales)           # Customizing axis scales and labels in ggplot2
#require(ggridges)         # Ridge plots for visualizing distributions
#require(tidybayes)        # Bayesian model visualizations
require(ggVennDiagram)    # Alternative Venn diagrams with ggplot2
#require(lemon)            # Advanced facetting and axis manipulation
#require(paletteer)        # Comprehensive color palettes for ggplot2
require(plot3D)           # 3D scatter plots and surfaces
require(plotly)           # Interactive plots with ggplot2
#require(ggalluvial)       # Alluvial plots for visualizing categorical data
require(ggtext)           # Rich text rendering in ggplot2
#require(ggupset)          # Upset plots for visualizing set intersections
#require(ggbeeswarm)       # Beeswarm plots for visualizing categorical data
#require(ggpointdensity)   # Point density plots for visualizing data density
require(ggdist)           # Visualizing distributions and uncertainty
require(ggside)           # Side plots for ggplot2
#require(ggpattern)        # Geoms for patterned fills in ggplot2
#require(ggimage)          # Geoms for displaying images in ggplot2
#require(ggpmisc)          # Miscellaneous extensions for ggplot2
#require(gganimate)        # Animated visualizations with ggplot2
require(gridExtra)  # For combining the plots



```

## File Handling
```{r file-handling}
# Libraries for reading and writing various file formats
require(haven)       # Reading SPSS, Stata, and SAS files
require(readxl)      # Reading Excel files
require(readr)       # Reading CSV and other text files
require(writexl)     # Writing Excel files
```

## Time Series and Rolling Averages
```{r time-series}

require(zoo)         # Rolling means/averages and time series
require(lubridate)   # Date and time manipulation
#require(timetk)      # Time series analysis tools
#require(tsibble)     # Tidy temporal data frames
#require(forecast)    # Time series forecasting
#require(feasts)      # Feature extraction and statistics for time series
#require(tsfeatures)  # Time series feature extraction
#require(slider)      # Flexible computation of rolling statistics
```
## Menstrual Cycle Packages

```{r menstrual-cycle}
# Libraries for menstrual cycle analysis
#require(cpass)
```


## Machine Learning and Predictive Modeling
```{r ml-predictive}

#require(caret)       # Unified interface for predictive modeling
#require(e1071)       # Support vector machines and other ML methods
#require(randomForest) # Random forests for classification and regression
#require(xgboost)     # Extreme gradient boosting
#require(lightgbm)    # Light gradient boosting
#require(caretEnsemble) # Ensemble models with caret
#require(tune)        # Hyperparameter tuning
#require(ranger)      # Fast implementation of random forests
#require(keras)       # Deep learning via Keras
```


## Generalized Additive Models and Bayesian Analysis
```{r gam}
# Libraries for GAMs and Bayesian analysis
require(mgcv)        # Generalized additive models
require(marginaleffects) # Marginal effects from regression models
require(gratia)      # Diagnostic tools for GAMs
require(gamair)      # Data and tools for learning about GAMs
require(gam.hp)      # GAMs with hierarchical partitioning

```


## Bayesian Analysis
```{r bayesian}
# Libraries for Bayesian analysis
#require(rstan)       # R interface to Stan
#require(bayesplot)   # Plotting functions for Bayesian analysis
#require(loo)         # Model comparison using PSIS-LOO
#require(bayestestR)  # Tools for Bayesian hypothesis testing
#require(brms)        # Bayesian regression models using Stan

```

## Statistical Tools
```{r statistical-tools}
# Libraries for specific statistical methods
require(psych)       # Descriptive statistics and psychometrics
#require(GPArotation) # Factor analysis rotations
require(rmcorr)      # Repeated measures correlations
require(careless)    # Detecting careless responses in surveys
require(responsePatterns) # Analyzing response patterns
```

## Power Analysis
```{r power-analysis}

#require(longpower)   # Power calculations for longitudinal data
#require(pwr)         # Power analysis for t-tests, ANOVAs, and correlations
#require(simr)        # Power analysis for mixed models
#require(poweRlaw)    # Power laws and exponential distributions
#require(pwr2ppl)     # Power analysis for two-phase pilot studies
#require(pwrFDR)      # Power analysis for false discovery rate control

```

# Clinical Trials Tools
```{r RCT}

#require(BayesCTDesign) # Bayesian clinical trial design
#require(clinfun)     # Clinical trial design and analysis

```


## Mixed-Effects and Generalized Models
```{r mixed-and-generalized-models}
# Libraries for mixed-effects models and generalized models

require(lme4)        # Fitting mixed-effects models
#require(nlme)        # Fitting linear and nonlinear mixed-effects models
#require(merTools)    # Tools for analyzing mixed-effects models
#require(MuMIn)       # Multi-model inference
#require(robustlmm)   # Robust linear mixed models
#require(ordinal)     # Ordinal regression models
#require(bayestestR)  # Tools for Bayesian hypothesis testing
require(lmtest)      # Testing linear models
#require(sjstats)     # Summary statistics for statistical models
require(sandwich)    # Robust covariance matrix estimators
require(lmerTest)    # Adds p-values for lme4 models
require(emmeans)     # Estimated marginal means
require(broom.mixed) # Tidying mixed model results
require(performance) # Model performance metrics
#require(glmmTMB)     # Generalized linear mixed models with flexible distributions
#require(MCMCglmm)    # Generalized linear mixed models using MCMC
#require(blme)        # Bayesian linear mixed models extending lme4
#require(brms)        # Bayesian generalized (non-)linear multivariate multilevel models using Stan
```
## General Libraries and Exploratory Data Analysis
```{r general-and-eda}
# Libraries for general data manipulation and exploratory data analysis

require(devtools) # Required for installing packages from GitHub
require(beepr) # For audio notifications
#require(chattr) # For adding comments to code chunks
require(progress)     # Progress bars for R scripts)
#require(remedy) # For error handling

require(Rmisc)  # For descriptive statistics
require(tidyr)    
require(dplyr)   
require(ggplot2)    
require(stringr) 
require(purrr)
require(forcats)
require(skimr)        # Summarize and inspect data quickly
require(janitor)      # For cleaning data, e.g., renaming columns, removing empty rows
#require(DataExplorer) # Automated EDA report generation
#require(GGally)       # Extension of ggplot2 for pairs plots
#require(Hmisc)        # Descriptive statistics and missing data management
#require(gghighlight)  # Highlight specific groups in ggplot2 visualizations
#require(naniar)       # Visualizing and analyzing missing data
#require(DescTools)    # Tools for descriptive statistics
#require(forcats)      # Tools for working with categorical data
#require(waldo)       # Visualizing differences between data frames
#require(visdat)       # Visualizing missing data and structure
#require(tweenr)       # Interpolating data for animations


```

## Check for Conflicts
```{r}
conflict_scout()


```


# ---- End of Setup ----
