
#NOTE: need to add in variance decomp for the GAMMs. 



# Input dataset

```{r}
# Load your CSV file (adjust the path to your actual file location)
df <- read_csv("~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/adhd_daily_df_20241017_L2N85.csv")
```















# FITTING GAMMs 


# GAMM and initial plot - E2 - Ov to Ov
```{r}

hist(df$E2)
df$E2log <- log(df$E2+1)
hist(df$E2log)

gamm_E2 <- gam(E2log ~ s(id, bs = "re") + 
                 s(scaled_cycleday, id, bs = "re") + 
                 s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_e2)


# Save the model summary as a text file
summary_output <- capture.output(summary(gamm_e2))
summary_file_path <- "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/E2model.txt"
writeLines(summary_output, summary_file_path)
cat("Model summary saved to:", summary_file_path, "\n")

plot.gam(gamm_e2, select = 3)



```



# Model-Implied Plot - E2 - Ov to Ov
```{r}

e2dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_e2, newdata = e2dat, type = "response", transform = function(x) exp(x) - 1)

#don't need to include anything for 'transform' if your outcome is not log-transformed. This undoes the logtransform so model-implied values are not on the log scale 

e2dat$estimate = pred$estimate
e2dat$conf.low = pred$conf.low
e2dat$conf.high = pred$conf.high

# Plotting
e2plot <- ggplot(e2dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Salivary Estradiol") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

e2plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/E2.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)

```



# GAMM and initial plot - LH
```{r}
hist(df$LH)
df$LHlog <- log(df$LH+1)
hist(df$LHlog)
gamm_LH <- gam(LHlog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_LH)


# Save the model summary as a text file
summary_output <- capture.output(summary(gamm_LH))
summary_file_path <- "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/LHmodel.txt"
writeLines(summary_output, summary_file_path)
cat("Model summary saved to:", summary_file_path, "\n")

plot.gam(gamm_LH, select = 3)

hist(df$LH)
```


# Model-Implied Plot - LH
```{r}

LHdat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_LH, newdata = LHdat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

LHdat$estimate = pred$estimate
LHdat$conf.low = pred$conf.low
LHdat$conf.high = pred$conf.high

# Plotting
LHplot <- ggplot(LHdat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "LH") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

LHplot


ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/LH.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)

```


# GAMM and initial plot - P4
```{r}
hist(df$P4)
df$P4log <- log(df$P4+1)
hist(df$P4log)

gamm_P4 <- gam(P4log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 10), data = df, method = "REML")

summary(gamm_P4)


# Save the model summary as a text file
summary_output <- capture.output(summary(gamm_P4))
summary_file_path <- "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/gamm_P4.txt"
writeLines(summary_output, summary_file_path)
cat("Model summary saved to:", summary_file_path, "\n")

plot.gam(gamm_P4, select = 3)

```

# Model-Implied Plot - P4
```{r}

P4dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_P4, newdata = P4dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

P4dat$estimate = pred$estimate
P4dat$conf.low = pred$conf.low
P4dat$conf.high = pred$conf.high


# Plotting
P4plot <- ggplot(P4dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Salivary P4") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

P4plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/P4.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)

```









```{r}

# Function to run the GAMM and generate plots for a given outcome variable
gamm_model_loop <- function(outcome, df, output_path, plot_path) {
  
  # Step 1: Create a histogram for the outcome variable to check its distribution
  hist(df[[outcome]], main = paste("Histogram of", outcome), xlab = outcome)
  
  # Step 2: Log-transform the outcome variable to deal with skewness
  # Add 1 to avoid log(0) issues, which is undefined
  log_outcome <- paste0(outcome, "log")  # Create a name for the log-transformed variable
  df[[log_outcome]] <- log(df[[outcome]] + 1)  # Apply log transformation and store in a new column
  hist(df[[log_outcome]], main = paste("Histogram of Log-Transformed", outcome), xlab = log_outcome)  # Create a histogram for the transformed data
  
  # Step 3: Fit a Generalized Additive Mixed Model (GAMM)
  # The formula includes smooth terms for the id (random effect) and scaled_cycleday
  gamm_model <- gam(as.formula(paste0(log_outcome, " ~ s(id, bs = 're') + s(scaled_cycleday, id, bs = 're') + s(scaled_cycleday, k = 30)")),
                    data = df, method = "REML")  # REML method is recommended for smoother estimation
  
  # Step 4: Display the summary of the model
 summary(gamm_model)
  
  
  # Step 5: Save the model summary to a text file
  summary_output <- capture.output(summary(gamm_model))  # Capture the summary output as a character vector
  summary_file_path <- file.path(output_path, paste0("gamm_", outcome, ".txt"))  # Define the path to save the summary
  writeLines(summary_output, summary_file_path)  # Write the summary to the file
  cat("Model summary saved to:", summary_file_path, "\n")  # Print confirmation
  
  # Step 6: Create a dataset for making predictions from the model
  # We create a grid of values for scaled_cycleday ranging from -1 to 1
  pred_data <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),  # Values of scaled_cycleday for prediction
                           id = 0)  # We use id = 0 (population-level prediction)
  
  # Step 7: Make predictions using the `marginaleffects` package and undo the log transformation
  pred <- predictions(gamm_model, newdata = pred_data, type = "response", transform = function(x) exp(x) - 1)  # Use predictions function
  pred_data$estimate <- pred$estimate  # Store the predicted estimates
  pred_data$conf.low <- pred$conf.low  # Store the lower bound of the 95% confidence interval
  pred_data$conf.high <- pred$conf.high  # Store the upper bound of the 95% confidence interval
  
  # Step 8: Plot the predicted values and confidence intervals
  outcome_plot <- ggplot(pred_data, aes(x = scaled_cycleday, y = estimate)) +
    scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), 
                       labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +  # Custom labels for x-axis
    labs(x = "Cycle Day (Scaled)", y = outcome) +  # Labels for axes
    geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
              fill = "grey70", alpha = 0.2, color = "white") +  # Highlight the menses onset period
    geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
              fill = "grey87", alpha = 0.2, color = "white") +  # Highlight the ovulation period
    geom_line(size = 1, show.legend = TRUE) +  # Line plot for the estimated values
    theme_minimal()  # Use a minimal theme for cleaner visualization
  
  # Display the plot
  print(outcome_plot)
  
  # Step 9: Save the plot to a file
  plot_file_path <- file.path(plot_path, paste0(outcome, "_k30.png"))  # Define the path for saving the plot
  ggsave(filename = plot_file_path, plot = outcome_plot, width = 8, height = 6)  # Save the plot as a PNG file
  cat("Plot saved to:", plot_file_path, "\n")  # Print confirmation
}

# Define paths for saving outputs (modify as per your folder structure)
output_path <- "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/"
plot_path <- file.path(output_path, "GAMMplots_k30")

# List of outcomes to loop through (replace with your actual outcomes)
# already exists as "outcomelist" above

# Loop through each outcome and apply the GAMM function
for (outcome in outcomelist) {
  gamm_model_loop(outcome, df, output_path, plot_path)  # Apply the function to each outcome
}

```




# Trying a version of the loop that merges in E2: 

```{r}


gamm_model_loop <- function(outcome, df, output_path, plot_path) {
  
  # Step 1: Create a histogram for the outcome variable to check its distribution
  hist(df[[outcome]], main = paste("Histogram of", outcome), xlab = outcome)
  
  # Step 2: Log-transform the outcome variable to deal with skewness
  log_outcome <- paste0(outcome, "log")  
  df[[log_outcome]] <- log(df[[outcome]] + 1)  
  hist(df[[log_outcome]], main = paste("Histogram of Log-Transformed", outcome), xlab = log_outcome)
  
  # Step 3: Fit a GAMM
  gamm_model <- gam(as.formula(paste0(log_outcome, " ~ s(id, bs = 're') + s(scaled_cycleday, id, bs = 're') + s(scaled_cycleday, k = 30)")),
                    data = df, method = "REML")
  
  # Step 4: Save the model summary
  summary_output <- summary(gamm_model)
  summary_file_path <- file.path(output_path, paste0("gamm_", outcome, ".txt"))
  writeLines(capture.output(summary_output), summary_file_path)
  
  # Step 4a: Extract p-value and edf for scaled_cycleday
  scaled_cycleday_term <- summary_output$s.table["s(scaled_cycleday)", ]
  p_value <- scaled_cycleday_term["p-value"]
  edf <- scaled_cycleday_term["edf"]

  # Determine significance symbol
  significance <- ""
  if (p_value < 0.001) {
    significance <- "***"
  } else if (p_value < 0.01) {
    significance <- "**"
  } else if (p_value < 0.05) {
    significance <- "*"
  } else if (p_value < 0.15) {
    significance <- "†"
  }

 # Step 5: Create a dataset for predictions from the model
pred_data <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1), id = 0)
pred <- predict(gamm_model, newdata = pred_data, type = "response", se.fit = TRUE) # Predict on the link scale

# Calculate confidence intervals on the link scale
pred_data$conf.low <- pred$fit - 1.96 * pred$se.fit 
pred_data$conf.high <- pred$fit + 1.96 * pred$se.fit

# Transform predictions and confidence intervals to the response scale
pred_data$estimate <- exp(pred$fit) - 1
pred_data$conf.low <- exp(pred_data$conf.low) - 1
pred_data$conf.high <- exp(pred_data$conf.high) - 1

# Scale the transformed values  <-- This is the moved line
pred_data$estimate <- scale(pred_data$estimate)
pred_data$conf.low <- scale(pred_data$conf.low)
pred_data$conf.high <- scale(pred_data$conf.high) 

  # Step 6: Plot predicted values and confidence intervals for the outcome and add E2 curve
  outcome_plot <- ggplot(pred_data, aes(x = scaled_cycleday)) +
    scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), 
                       labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
    labs(x = "Cycle Day (Scaled)", y = "Scaled Predicted Values") +
    geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
              fill = "grey70", alpha = 0.2, color = "white") +
    geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
              fill = "grey87", alpha = 0.2, color = "white") +
    
    # Confidence ribbons for the outcome estimates
    geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "black", alpha = 0.1) +
    
    # Line plot for outcome
    geom_line(aes(y = estimate, color = "outcome"), size = 1, show.legend = TRUE) +  
    
    # Step 6a: Add E2 and P4 curves on top (using e2dat and P4dat)
    geom_line(data = e2dat, aes(x = scaled_cycleday, y = scale(estimate), color = "E2"), size = 1, linetype = "dashed") +  # E2 curve
    geom_line(data = P4dat, aes(x = scaled_cycleday, y = scale(estimate), color = "P4"), size = 1, linetype = "dashed") +  # P4 curve
    
    # Step 6b: Customize legend with desired order
    scale_color_manual(name = "Key", 
                       values = c("outcome" = "black", "E2" = "forestgreen", "P4" = "red"),
                       labels = c("outcome" = outcome, "E2" = "E2", "P4" = "P4"),
                       breaks = c("outcome", "E2", "P4")) +  # Ensure correct legend order
    
    theme_minimal() +
    theme(legend.position = "right")

  # Create a label for EDF and p-value, and add it below the legend
  annotation_text <- paste("EDF:", round(edf, 2), significance)
  annotated_plot <- ggdraw(outcome_plot) +
    draw_label(annotation_text, x = 0.85, y = 0.2, hjust = 0, size = 10, fontface = "bold")

  # Display the plot with annotation
  print(annotated_plot)

  # Step 7: Save the plot to a file
  plot_file_path <- file.path(plot_path, paste0(outcome, "_with_horm.png"))
  ggsave(filename = plot_file_path, plot = annotated_plot, width = 8, height = 6)
  cat("Plot with Hormone overlays saved to:", plot_file_path, "\n")
}

# Define paths for saving outputs (modify as per your folder structure)
output_path <- "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-10/"
plot_path <- file.path(output_path, "Plots with Hormone Overlays")

# List of outcomes to loop through (replace with your actual outcomes)
outcomelist_test <- c("CSS_Inatt")

outcomelist <- c("CSS_Inatt", "CSS_HypImp", "score_pinball", "score_robot", "UPPS_NU_avg", "UPPS_Persev_avg", "UPPS_Premed_avg", "UPPS_Sens_avg", "UPPS_PU_avg", 
  "CSS_Fx_Total",  "CSS_Inatt_Count", "CSS_HypImp_Count", "BDEFS_WM_avg", "BDEFS_RI_avg", "DEBQ_Total", "DRSP_1", "DRSP_7")

# Loop through each outcome and apply the GAMM function
for (outcome in outcomelist) {
  gamm_model_loop(outcome, df, output_path, plot_path)  # Apply the function to each outcome
}

```


Alt Patchwork Code
```{r}

  library(patchwork)

# Create separate plots for Outcome, E2, and P4
outcome_plot <- ggplot(pred_data, aes(x = scaled_cycleday, y = estimate)) +
  geom_line(size = 1, color = "blue") +
  labs(x = "Cycle Day (Scaled)", y = outcome) +
  theme_minimal()

e2_plot <- ggplot(e2dat, aes(x = scaled_cycleday, y = estimate)) +
  geom_line(size = 1, color = "forestgreen", linetype = "dashed") +
  labs(x = "Cycle Day (Scaled)", y = "E2") +
  theme_minimal()

p4_plot <- ggplot(P4dat, aes(x = scaled_cycleday, y = estimate)) +
  geom_line(size = 1, color = "red", linetype = "dashed") +
  labs(x = "Cycle Day (Scaled)", y = "P4") +
  theme_minimal()

# Arrange all three plots in a single view
combined_plot <- outcome_plot / e2_plot / p4_plot
print(combined_plot)
```




# GAMM and initial plot - DRSP_1 - MS F, NS R
```{r}
hist(df$DRSP_1)
df$DRSP_1log <- log(df$DRSP_1+1)
hist(df$DRSP_1log)
gamm_DRSP_1 <- gam(DRSP_1log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_1)

plot.gam(gamm_DRSP_1, select = 3)

hist(df$DRSP_1)
```


# Model-Implied Plot - DRSP_1
```{r}

DRSP_1dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_1, newdata = DRSP_1dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_1dat$estimate = pred$estimate
DRSP_1dat$conf.low = pred$conf.low
DRSP_1dat$conf.high = pred$conf.high

# Plotting
DRSP_1plot <- ggplot(DRSP_1dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Depressed Mood") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_1plot


ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/depblue.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```


# GAMM and initial plot - DRSP_2 - MS F, NS R
```{r}
hist(df$DRSP_2)
df$DRSP_2log <- log(df$DRSP_2+1)
hist(df$DRSP_2log)
gamm_DRSP_2 <- gam(DRSP_2log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_2)

plot.gam(gamm_DRSP_2, select = 3)

hist(df$DRSP_2)
```


# Model-Implied Plot - DRSP2
```{r}

DRSP_2dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_2, newdata = DRSP_2dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_2dat$estimate = pred$estimate
DRSP_2dat$conf.low = pred$conf.low
DRSP_2dat$conf.high = pred$conf.high

# Plotting
DRSP_2plot <- ggplot(DRSP_2dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Hopelessness") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_2plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/hopeless.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```

# GAMM and initial plot - DRSP_3 - NS F, NS R
```{r}
hist(df$DRSP_3)
df$DRSP_3log <- log(df$DRSP_3+1)
hist(df$DRSP_3log)
gamm_DRSP_3 <- gam(DRSP_3log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_3)

plot.gam(gamm_DRSP_3, select = 3)

hist(df$DRSP_3)
```


# Model-Implied Plot - DRSP_3
```{r}

DRSP_3dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_3, newdata = DRSP_3dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_3dat$estimate = pred$estimate
DRSP_3dat$conf.low = pred$conf.low
DRSP_3dat$conf.high = pred$conf.high

# Plotting
DRSP_3plot <- ggplot(DRSP_3dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Worthlessness & Guilt") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_3plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/worthguilt.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```


# GAMM and initial plot - DRSP_4 - MS F, SIG R
```{r}
hist(df$DRSP_4)
df$DRSP_4log <- log(df$DRSP_4+1)
hist(df$DRSP_4log)
gamm_DRSP_4 <- gam(DRSP_4log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_4)

plot.gam(gamm_DRSP_4, select = 3)

hist(df$DRSP_4)
```


# Model-Implied Plot - DRSP_4
```{r}

DRSP_4dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_4, newdata = DRSP_4dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_4dat$estimate = pred$estimate
DRSP_4dat$conf.low = pred$conf.low
DRSP_4dat$conf.high = pred$conf.high

# Plotting
DRSP_4plot <- ggplot(DRSP_4dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Anxiety & Tension") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_4plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/anxiety.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```


# GAMM and initial plot - DRSP_5 - SIG F, SIG R
```{r}
hist(df$DRSP_5)
df$DRSP_5log <- log(df$DRSP_5+1)
hist(df$DRSP_5log)
gamm_DRSP_5 <- gam(DRSP_5log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_5)

plot.gam(gamm_DRSP_5, select = 3)

hist(df$DRSP_5)
```


# Model-Implied Plot - DRSP_5
```{r}

DRSP_5dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_5, newdata = DRSP_5dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_5dat$estimate = pred$estimate
DRSP_5dat$conf.low = pred$conf.low
DRSP_5dat$conf.high = pred$conf.high

# Plotting
DRSP_5plot <- ggplot(DRSP_5dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Mood Swings") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_5plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/moodswings.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```


# GAMM and initial plot - DRSP_6 - NS F, MS R
```{r}
hist(df$DRSP_6)
df$DRSP_6log <- log(df$DRSP_6+1)
hist(df$DRSP_6log)
gamm_DRSP_6 <- gam(DRSP_6log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_6)

plot.gam(gamm_DRSP_6, select = 3)

hist(df$DRSP_6)
```


# Model-Implied Plot - DRSP_6
```{r}

DRSP_6dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_6, newdata = DRSP_6dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_6dat$estimate = pred$estimate
DRSP_6dat$conf.low = pred$conf.low
DRSP_6dat$conf.high = pred$conf.high

# Plotting
DRSP_6plot <- ggplot(DRSP_6dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Rejection Sensitivity") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_6plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/rejsens.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```



# GAMM and initial plot - DRSP_7 - SIG F, NS R
```{r}
hist(df$DRSP_7)
df$DRSP_7log <- log(df$DRSP_7+1)
hist(df$DRSP_7log)
gamm_DRSP_7 <- gam(DRSP_7log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_7)

plot.gam(gamm_DRSP_7, select = 3)

hist(df$DRSP_7)
```


# Model-Implied Plot - DRSP_7
```{r}

DRSP_7dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_7, newdata = DRSP_7dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_7dat$estimate = pred$estimate
DRSP_7dat$conf.low = pred$conf.low
DRSP_7dat$conf.high = pred$conf.high

# Plotting
DRSP_7plot <- ggplot(DRSP_7dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Anger or Irritability") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_7plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/angirr.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```


# GAMM and initial plot - DRSP_8 - NS F, NS R
```{r}
hist(df$DRSP_8)
df$DRSP_8log <- log(df$DRSP_8+1)
hist(df$DRSP_8log)
gamm_DRSP_8 <- gam(DRSP_8log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_8)

plot.gam(gamm_DRSP_8, select = 3)

hist(df$DRSP_8)
```


# Model-Implied Plot - DRSP_8
```{r}

DRSP_8dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_8, newdata = DRSP_8dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_8dat$estimate = pred$estimate
DRSP_8dat$conf.low = pred$conf.low
DRSP_8dat$conf.high = pred$conf.high

# Plotting
DRSP_8plot <- ggplot(DRSP_8dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Interpersonal Conflict") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_8plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/intconf.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```

# GAMM and initial plot - DRSP_9 - Sig F, NS R
```{r}
hist(df$DRSP_9)
df$DRSP_9log <- log(df$DRSP_9+1)
hist(df$DRSP_9log)
gamm_DRSP_9 <- gam(DRSP_9log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_9)

plot.gam(gamm_DRSP_9, select = 3)

hist(df$DRSP_9)
```


# Model-Implied Plot - DRSP_9
```{r}

DRSP_9dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_9, newdata = DRSP_9dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_9dat$estimate = pred$estimate
DRSP_9dat$conf.low = pred$conf.low
DRSP_9dat$conf.high = pred$conf.high

# Plotting
DRSP_9plot <- ggplot(DRSP_9dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Lack of Interest or Motivation") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_9plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/lackintmot.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```

# GAMM and initial plot - DRSP_10 - NS F, NS R
```{r}
hist(df$DRSP_10)
df$DRSP_10log <- log(df$DRSP_10+1)
hist(df$DRSP_10log)
gamm_DRSP_10 <- gam(DRSP_10log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_10)

plot.gam(gamm_DRSP_10, select = 3)

hist(df$DRSP_10)
```


# Model-Implied Plot - DRSP_10
```{r}

DRSP_10dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_10, newdata = DRSP_10dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_10dat$estimate = pred$estimate
DRSP_10dat$conf.low = pred$conf.low
DRSP_10dat$conf.high = pred$conf.high

# Plotting
DRSP_10plot <- ggplot(DRSP_10dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Difficulty Concentrating") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_10plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/diffconc.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```

# GAMM and initial plot - DRSP_11 - MS F, NS R
```{r}
hist(df$DRSP_11)
df$DRSP_11log <- log(df$DRSP_11+1)
hist(df$DRSP_11log)
gamm_DRSP_11 <- gam(DRSP_11log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_11)

plot.gam(gamm_DRSP_11, select = 3)

hist(df$DRSP_11)
```


# Model-Implied Plot - DRSP_11
```{r}

DRSP_11dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_11, newdata = DRSP_11dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_11dat$estimate = pred$estimate
DRSP_11dat$conf.low = pred$conf.low
DRSP_11dat$conf.high = pred$conf.high

# Plotting
DRSP_11plot <- ggplot(DRSP_11dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Lethargy, Fatigue") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_11plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/fatigue.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```


# GAMM and initial plot - DRSP_12 - SIG F, NS R
```{r}
hist(df$DRSP_12)
df$DRSP_12log <- log(df$DRSP_12+1)
hist(df$DRSP_12log)
gamm_DRSP_12 <- gam(DRSP_12log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_12)

plot.gam(gamm_DRSP_12, select = 3)

hist(df$DRSP_12)
```


# Model-Implied Plot - DRSP_12
```{r}

DRSP_12dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_12, newdata = DRSP_12dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_12dat$estimate = pred$estimate
DRSP_12dat$conf.low = pred$conf.low
DRSP_12dat$conf.high = pred$conf.high

# Plotting
DRSP_12plot <- ggplot(DRSP_12dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Increased Appetite, Overeating") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_12plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/Appoverate.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```

# GAMM and initial plot - DRSP_13 - SIG F, NS R
```{r}
hist(df$DRSP_13)
df$DRSP_13log <- log(df$DRSP_13+1)
hist(df$DRSP_13log)
gamm_DRSP_13 <- gam(DRSP_13log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_13)

plot.gam(gamm_DRSP_13, select = 3)

hist(df$DRSP_13)
```


# Model-Implied Plot - DRSP_13
```{r}

DRSP_13dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_13, newdata = DRSP_13dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_13dat$estimate = pred$estimate
DRSP_13dat$conf.low = pred$conf.low
DRSP_13dat$conf.high = pred$conf.high

# Plotting
DRSP_13plot <- ggplot(DRSP_13dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Food Cravings") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_13plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/foodcrave.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```

# GAMM and initial plot - DRSP_14 - NS F, NS R
```{r}
hist(df$DRSP_14)
df$DRSP_14log <- log(df$DRSP_14+1)
hist(df$DRSP_14log)
gamm_DRSP_14 <- gam(DRSP_14log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_14)

plot.gam(gamm_DRSP_14, select = 3)

hist(df$DRSP_14)
```


# Model-Implied Plot - DRSP_14
```{r}

DRSP_14dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_14, newdata = DRSP_14dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_14dat$estimate = pred$estimate
DRSP_14dat$conf.low = pred$conf.low
DRSP_14dat$conf.high = pred$conf.high

# Plotting
DRSP_14plot <- ggplot(DRSP_14dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Hypersomnia") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_14plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/Hypersomnia.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```


# GAMM and initial plot - DRSP_15 - NS F, NS R
```{r}
hist(df$DRSP_15)
df$DRSP_15log <- log(df$DRSP_15+1)
hist(df$DRSP_15log)
gamm_DRSP_15 <- gam(DRSP_15log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_15)

plot.gam(gamm_DRSP_15, select = 3)

hist(df$DRSP_15)
```


# Model-Implied Plot - TEMPLATE
```{r}

DRSP_15dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_15, newdata = DRSP_15dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_15dat$estimate = pred$estimate
DRSP_15dat$conf.low = pred$conf.low
DRSP_15dat$conf.high = pred$conf.high

# Plotting
DRSP_15plot <- ggplot(DRSP_15dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Insomnia") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_15plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/Insomnia.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```


# GAMM and initial plot - DRSP_16 - NS F, NS R
```{r}
hist(df$DRSP_16)
df$DRSP_16log <- log(df$DRSP_16+1)
hist(df$DRSP_16log)
gamm_DRSP_16 <- gam(DRSP_16log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_16)

plot.gam(gamm_DRSP_16, select = 3)

hist(df$DRSP_16)
```


# Model-Implied Plot - DRSP_16
```{r}

DRSP_16dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_16, newdata = DRSP_16dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_16dat$estimate = pred$estimate
DRSP_16dat$conf.low = pred$conf.low
DRSP_16dat$conf.high = pred$conf.high

# Plotting
DRSP_16plot <- ggplot(DRSP_16dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Overwhelmed") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_16plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/overwhelm.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```


# GAMM and initial plot - DRSP_17 - Sig F, NS R
```{r}
hist(df$DRSP_17)
df$DRSP_17log <- log(df$DRSP_17+1)
hist(df$DRSP_17log)
gamm_DRSP_17 <- gam(DRSP_17log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_17)

plot.gam(gamm_DRSP_17, select = 3)

hist(df$DRSP_17)
```


# Model-Implied Plot - TEMPLATE
```{r}

DRSP_17dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_17, newdata = DRSP_17dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_17dat$estimate = pred$estimate
DRSP_17dat$conf.low = pred$conf.low
DRSP_17dat$conf.high = pred$conf.high

# Plotting
DRSP_17plot <- ggplot(DRSP_17dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Out of Control") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_17plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/OOC.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```



# GAMM and initial plot - DRSP_18 - Sig F, Sig R
```{r}
hist(df$DRSP_18)
df$DRSP_18log <- log(df$DRSP_18+1)
hist(df$DRSP_18log)
gamm_DRSP_18 <- gam(DRSP_18log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_18)

plot.gam(gamm_DRSP_18, select = 3)

hist(df$DRSP_18)
```


# Model-Implied Plot - DRSP_18
```{r}

DRSP_18dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_18, newdata = DRSP_18dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_18dat$estimate = pred$estimate
DRSP_18dat$conf.low = pred$conf.low
DRSP_18dat$conf.high = pred$conf.high

# Plotting
DRSP_18plot <- ggplot(DRSP_18dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Breast Tenderness") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_18plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/Breast Tender.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```


# GAMM and initial plot - DRSP_19 - SIG F, NS R
```{r}
hist(df$DRSP_19)
df$DRSP_19log <- log(df$DRSP_19+1)
hist(df$DRSP_19log)
gamm_DRSP_19 <- gam(DRSP_19log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_19)

plot.gam(gamm_DRSP_19, select = 3)

hist(df$DRSP_19)
```


# Model-Implied Plot - DRSP_19
```{r}

DRSP_19dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_19, newdata = DRSP_19dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_19dat$estimate = pred$estimate
DRSP_19dat$conf.low = pred$conf.low
DRSP_19dat$conf.high = pred$conf.high

# Plotting
DRSP_19plot <- ggplot(DRSP_19dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Swelling, Bloating, or Weight Gain") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_19plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/swellbloatgain.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```


# GAMM and initial plot - DSRP_20 - MS F, NS R
```{r}
hist(df$DRSP_20)
df$DRSP_20log <- log(df$DRSP_20+1)
hist(df$DRSP_20log)
gamm_DRSP_20 <- gam(DRSP_20log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_20)

plot.gam(gamm_DRSP_20, select = 3)

hist(df$DRSP_20)
```


# Model-Implied Plot - DSRP_20
```{r}

DRSP_20dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_20, newdata = DRSP_20dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_20dat$estimate = pred$estimate
DRSP_20dat$conf.low = pred$conf.low
DRSP_20dat$conf.high = pred$conf.high

# Plotting
DRSP_20plot <- ggplot(DRSP_20dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Headache") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_20plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/headache.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```




# GAMM and initial plot - DRSP_21 - NS F, NS R
```{r}
hist(df$DRSP_21)
df$DRSP_21log <- log(df$DRSP_21+1)
hist(df$DRSP_21log)
gamm_DRSP_21 <- gam(DRSP_21log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_21)

plot.gam(gamm_DRSP_21, select = 3)

hist(df$DRSP_21)
```


# Model-Implied Plot - DRSP_21
```{r}

DRSP_21dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_21, newdata = DRSP_21dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_21dat$estimate = pred$estimate
DRSP_21dat$conf.low = pred$conf.low
DRSP_21dat$conf.high = pred$conf.high

# Plotting
DRSP_21plot <- ggplot(DRSP_21dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Joint or Muscle Pain") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_21plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/jointmuscpain.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```

# GAMM and initial plot - DRSP_22 - Sig F, NS R
```{r}
hist(df$DRSP_22)
df$DRSP_22log <- log(df$DRSP_22+1)
hist(df$DRSP_22log)
gamm_DRSP_22 <- gam(DRSP_22log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_22)

plot.gam(gamm_DRSP_22, select = 3)

hist(df$DRSP_22)
```


# Model-Implied Plot - DRSP_22
```{r}

DRSP_22dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_22, newdata = DRSP_22dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_22dat$estimate = pred$estimate
DRSP_22dat$conf.low = pred$conf.low
DRSP_22dat$conf.high = pred$conf.high

# Plotting
DRSP_22plot <- ggplot(DRSP_22dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "School or Work Impairment") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_22plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/Workimp.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```



# GAMM and initial plot - DRSP_23 - MS F, NS R
```{r}
hist(df$DRSP_23)
df$DRSP_23log <- log(df$DRSP_23+1)
hist(df$DRSP_23log)
gamm_DRSP_23 <- gam(DRSP_23log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_DRSP_23)

plot.gam(gamm_DRSP_23, select = 3)

hist(df$DRSP_23)
```


# Model-Implied Plot - DRSP_23
```{r}

DRSP_23dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_23, newdata = DRSP_23dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_23dat$estimate = pred$estimate
DRSP_23dat$conf.low = pred$conf.low
DRSP_23dat$conf.high = pred$conf.high

# Plotting
DRSP_23plot <- ggplot(DRSP_23dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Relationship Interference") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_23plot

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/GAMMplots/relimp.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)
```

# Output all Model Files to Box

```{r}

outcomelist <- c(
  "DRSP_1", "DRSP_2", "DRSP_3", "DRSP_4", "DRSP_5", "DRSP_6", "DRSP_7", "DRSP_8", "DRSP_9", 
  "DRSP_10", "DRSP_11", "DRSP_12", "DRSP_13", "DRSP_14", "DRSP_15", "DRSP_16", "DRSP_17", 
  "DRSP_18", "DRSP_19", "DRSP_20", "DRSP_21", "DRSP_22", "DRSP_23", 
  "UPPS_NU_avg", "UPPS_Persev_avg", "UPPS_Premed_avg", "UPPS_Sens_avg", "UPPS_PU_avg", 
  "CSS_Fx_Total", "CSS_Inatt", "CSS_HypImp", "CSS_Inatt_Count", "CSS_HypImp_Count", 
  "score_pinball", "score_robot", "BDEFS_Total", "BDEFS_WM_avg", "BDEFS_RI_avg", 
  "UPPS_Total", "DEBQ_Total"
)


# Define your list of outcomes
print(outcomelist)

# Function to save summary of GAMM models to a file
save_gamm_summary <- function(outcome) {
  # Dynamically create model variable name
  model_var <- paste0("gamm_", outcome)
  
  # Dynamically create the file path for each outcome
  file_path <- paste0("/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-09-30/GAMMmodels/", outcome, "_gam_model_summary.txt")
  
  # Capture and write the summary to the file
  writeLines(capture.output(summary(get(model_var))), file_path)
}

# Loop over each outcome and save the summary
lapply(outcomelist, save_gamm_summary)

```




##### Reference and template info below


# OUTCOMES

BDEFS_1, BDEFS_2, BDEFS_3, BDEFS_4, BDEFS_5, BDEFS_6, DRSP_1, DRSP_2, DRSP_3, DRSP_4, DRSP_5, DRSP_6, DRSP_7, DRSP_8, DRSP_9, DRSP_10, DRSP_11, DRSP_12, DRSP_13, DRSP_14, DRSP_15, DRSP_16, DRSP_17, DRSP_18, DRSP_19, DRSP_20, DRSP_21, DRSP_22, DRSP_23, CSS_B_1, CSS_B_2, CSS_B_3, CSS_B_4, CSS_B_5, CSS_B_6, CSS_B_7, CSS_B_8, CSS_B_9, CSS_B_10, CSS_B_11, CSS_B_12, CSS_B_13, CSS_B_14, CSS_B_15, CSS_B_16, CSS_B_17, CSS_B_18, CSS_Function_1, CSS_Function_2, CSS_Function_3, CSS_Function_4, CSS_Function_5, CSS_Function_6, CSS_Function_7, CSS_Function_8, CSS_Function_9, CSS_Function_10, CSS_B2_1, CSS_B2_2, CSS_B2_3, CSS_B2_4, CSS_B2_5, CSS_B2_6, CSS_B2_7, CSS_B2_8, UPPS_1, UPPS_2, UPPS_3, UPPS_4, UPPS_5, UPPS_6, UPPS_9, UPPS_10, UPPS_11, UPPS_12, UPPS_13, UPPS_14, UPPS_15, DEBQ_1, DEBQ_2, DEBQ_3, DEBQ_4, DEBQ_5, DEBQ_6, DEBQ_7, DEBQ_8, DEBQ_9, DEBQ_10, DEBQ_11, DEBQ_12, DEBQ_13

BDEFS_WM_avg, BDEFS_RI_avg, UPPS_NU_avg, UPPS_Persev_avg, UPPS_Premed_avg, UPPS_Sens_avg, UPPS_PU_avg, game_pinball, game_name_pinball, score_pinball, user_level_pinball, session_level_pinball, game_nth_pinball, game_lpi_pinball, RecordedDate_pinball, recorded_date_pinball, hour_recorded_pinball, midnight_to_seven_pinball, game_type_pinball, date_rated_dontuse_pinball, game_robot, game_name_robot, score_robot, user_level_robot, session_level_robot, game_nth_robot, game_lpi_robot, RecordedDate_robot, recorded_date_robot, hour_recorded_robot, midnight_to_seven_robot, game_type_robot, date_rated_dontuse_robot, nth_pinball_squared, nth_robot_squared, RES_robot, RES_pinball,


DRSP_1, DRSP_2, DRSP_3, DRSP_4, DRSP_5, DRSP_6, DRSP_7, DRSP_8, DRSP_9, DRSP_10, DRSP_11, DRSP_12, DRSP_13, DRSP_14, DRSP_15, DRSP_16, DRSP_17, DRSP_18, DRSP_19, DRSP_20, DRSP_21, DRSP_22, DRSP_23, UPPS_NU_avg, UPPS_Persev_avg, UPPS_Premed_avg, UPPS_Sens_avg, UPPS_PU_avg

# GAMM SCALED CYCLE DAY Effects - TEMPLATE MODEL
```{r}
hist(df$P4)
df$P4log <- log(df$P4+1)
hist(df$P4log)
gamm_p4 <- gam(P4log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = df, method = "REML")

summary(gamm_p4)

plot.gam(gamm_p4, select = 3)

hist(df$P4)
```


# GAMM SCALED CYCLE DAY Effects - TEMPLATE PLOT
```{r}

P4dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_p4, newdata = P4dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

P4dat$estimate = pred$estimate
P4dat$conf.low = pred$conf.low
P4dat$conf.high = pred$conf.high

# Plotting
P4plot <- ggplot(P4dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "P4") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

P4plot


```








# HORMONE MODELS

## Prepare Hormones

```{r}

#df$logE2 = log(df$E2 + 1)
#df$logP4 = log(df$P4 + 1)

df <- df %>%
  group_by(id) %>%
  mutate(
      logE2= log(E2+1), 
    logP4= log(P4+1),
    E2zd = scale(E2),
    P4zd = scale(P4),
    
    # Calculate 5th and 95th percentiles for E2 and P4 within each id
    E2_lower = quantile(E2zd, 0.05, na.rm = TRUE),
    E2_upper = quantile(E2zd, 0.95, na.rm = TRUE),
    P4_lower = quantile(P4zd, 0.05, na.rm = TRUE),
    P4_upper = quantile(P4zd, 0.95, na.rm = TRUE),
    
    # Manually Winsorize by clamping values to the 5th and 95th percentiles
    E2zd = pmin(pmax(E2zd, E2_lower), E2_upper),
    P4zd = pmin(pmax(P4zd, P4_lower), P4_upper),
    
    # Daily derivatives for E2 and P4
    E2zch = c(NA, diff(E2zd)),  # Winsorized E2 derivative
    P4zch = c(NA, diff(P4zd))   # Winsorized P4 derivative
  ) %>%
  filter(
    !is.na(E2zd) & !is.nan(E2zd) & !is.infinite(E2zd) &
      !is.na(P4zd) & !is.nan(P4zd) & !is.infinite(P4zd) &
      !is.na(E2zch) & !is.nan(E2zch) & !is.infinite(E2zch) &
      !is.na(P4zch) & !is.nan(P4zch) & !is.infinite(P4zch)
  ) %>%
  ungroup()


hist(df$E2zd)
hist(df$P4zd)

df_all <- df_all %>%
  group_by(id) %>%
  mutate(
    logE2= log(E2+1), 
    logP4= log(P4+1),
    E2zd = scale(E2),
    P4zd = scale(P4),
    
    # Calculate 5th and 95th percentiles for E2 and P4 within each id
    E2_lower = quantile(E2zd, 0.05, na.rm = TRUE),
    E2_upper = quantile(E2zd, 0.95, na.rm = TRUE),
    P4_lower = quantile(P4zd, 0.05, na.rm = TRUE),
    P4_upper = quantile(P4zd, 0.95, na.rm = TRUE),
    
    # Manually Winsorize by clamping values to the 5th and 95th percentiles
    E2zd = pmin(pmax(E2zd, E2_lower), E2_upper),
    P4zd = pmin(pmax(P4zd, P4_lower), P4_upper),
    
    # Daily derivatives for E2 and P4
    E2zch = c(NA, diff(E2zd)),  # Winsorized E2 derivative
    P4zch = c(NA, diff(P4zd))   # Winsorized P4 derivative
  ) %>%
  filter(
    !is.na(E2zd) & !is.nan(E2zd) & !is.infinite(E2zd) &
      !is.na(P4zd) & !is.nan(P4zd) & !is.infinite(P4zd) &
      !is.na(E2zch) & !is.nan(E2zch) & !is.infinite(E2zch) &
      !is.na(P4zch) & !is.nan(P4zch) & !is.infinite(P4zch)
  ) %>%
  ungroup()


```

# Looping through hormone effects - FIXED: E2zd, P4zd, E2zd*P4zd, E2zch, P4zch

```{r}

# Function to Loop Through Outcomes and Generate GAMM Models
create_gamm_models <- function(outcomelist, df_all, save_dir) {
  for (outcome in outcomelist) {
    # Prepare variables dynamically
    outcome_log <- paste0(outcome, "log")
    df_all[[outcome_log]] <- log(df_all[[outcome]] + 1)

    # Plot histogram for transformed outcome
    hist(df_all[[outcome]], main = paste("Histogram of", outcome))
    hist(df_all[[outcome_log]], main = paste("Histogram of Log-transformed", outcome))

    # GAMM model with hormone z-scores and derivatives
    gamm_formula <- as.formula(paste0(
      outcome_log, " ~ s(id, bs = \"re\") + s(E2zd) +",
      " s(P4zd)  + ti(E2zd, P4zd) +",
      " s(E2zch) + s(P4zch)"
    ))

    gamm_model <- gam(
      gamm_formula,
      data = df_all,
      family = gaussian,
      method = "REML"
    )

    # Summary and plot of the model
    model_summary <- summary(gamm_model)
    print(model_summary)

    # Save the model summary as a text file
    summary_output <- capture.output(model_summary)
    summary_file_path <- file.path(save_dir, paste0(outcome, "_fixed__wzch_85_model_EP_summary.txt"))
    writeLines(summary_output, summary_file_path)
    cat("Model summary saved to:", summary_file_path, "\n")

    # Plot the GAMM effects
    plot.gam(gamm_model, select = 2, main = paste("Estrogen Effect on", outcome))
    plot.gam(gamm_model, select = 3, main = paste("Progesterone Effect on", outcome))
   plot.gam(gamm_model, select = 5, main = paste("Estrogen Change Effect on", outcome))
   plot.gam(gamm_model, select = 6, main = paste("Progesterone Change Effect on", outcome))

    # Simpler model without random effects for visualization
    gamm_simple_formula <- as.formula(paste0(
      outcome_log, " ~ s(E2zd) + s(P4zd) + ti(E2zd, P4zd)"
    ))

    gamm_model_simple <- gam(
      gamm_simple_formula,
      data = df_all,
      family = gaussian,
      method = "REML"
    )

    # Create the grid for contour plot
    grid <- expand.grid(
      E2zd = seq(
        min(df_all$E2zd, na.rm = TRUE),
        max(df_all$E2zd, na.rm = TRUE),
        length.out = 100
      ),
      P4zd = seq(
        min(df_all$P4zd, na.rm = TRUE),
        max(df_all$P4zd, na.rm = TRUE),
        length.out = 100
      )
    )

    # Predict values
    grid$predicted <- predict(gamm_model_simple,
                              newdata = grid,
                              type = "response",
                              transform = function(x) exp(x) - 1)

    # Contour plot
    plot_title <- paste("Predicting", outcome, "from E2 and P4 (Person Standardized)")
    contour_plot <- ggplot() +
      geom_tile(data = grid, aes(x = E2zd, y = P4zd, fill = predicted)) +
      geom_contour(data = grid, aes(x = E2zd, y = P4zd, z = predicted), color = "black") +
      geom_point(data = df_all, aes(x = E2zd, y = P4zd), color = "black", alpha = 0.25) +
      scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = median(grid$predicted)) +
      labs(x = "E2zd", y = "P4zd", fill = "Outcome") +
      theme_minimal() +
      ggtitle(plot_title)

    print(contour_plot)

    # Save the contour plot
    plot_file_path <- file.path(save_dir, paste0(outcome, "_85_wzch_fixed.png"))
    ggsave(filename = plot_file_path, plot = contour_plot, width = 8, height = 6)
    cat("Contour plot saved to:", plot_file_path, "\n")
  }
}

# CALL THE LOOP

#OutcomeList already exists above

create_gamm_models(outcomelist, df, "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/EP_Models_Plots")

```



 
# TEMPLATE GAMM .zd and Derivative effects of Hormones - TEMPLATE

```{r}


hist(df_all$CSS_Inatt)
df_all$CSS_Inattlog <- log(df_all$CSS_Inatt+1)
hist(df_all$CSS_Inattlog)

gamm_hormzd_CSS_Inatt <- gam(
  CSS_Inattlog ~ s(id, bs = "re") + s(E2zd) + s(E2zd, id, bs = "re") +
    s(P4zd) + s(P4zd, id, bs = "re") + ti(E2zd, P4zd) +
    s(E2zch) + s(E2zch, id, bs = "re") + s(P4zch) + s(P4zch, id, bs = "re"),
  data = df_all,
  family = gaussian,
  method = "REML"
)

# Summary and plot of the model
summary(gamm_hormzd_CSS_Inatt)


# Save the model summary as a text file
summary_output <- capture.output(summary(gamm_hormzd_CSS_Inatt))
summary_file_path <- "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/EP_Models_Plots/CSS_Inatt_model_EP_summary.txt"
writeLines(summary_output, summary_file_path)
cat("Model summary saved to:", summary_file_path, "\n")

plot.gam(gamm_hormzd_CSS_Inatt, select=2) #Estrogen
plot.gam(gamm_hormzd_CSS_Inatt, select=4) #Progesterone
plot.gam(gamm_hormzd_CSS_Inatt, select=7) #Estrogen CHANGE
plot.gam(gamm_hormzd_CSS_Inatt, select=8) #Progesterone CHANGE


# Simpler model without random effects for visualization
gamm_horm_CSS_Inatt_simple <- gam(
  CSS_Inattlog ~ s(E2zd) + s(P4zd) + ti(E2zd, P4zd),
  data = df_all,
  family = gaussian,
  method = "REML"
)

# Create the grid for contour plot
grid <- expand.grid(
  E2zd = seq(
    min(df$E2zd, na.rm = TRUE),
    max(df$E2zd, na.rm = TRUE),
    length.out = 100
  ),
  P4zd = seq(
    min(df$P4zd, na.rm = TRUE),
    max(df$P4zd, na.rm = TRUE),
    length.out = 100
  )
)

# Predict values
grid$predicted <- predict(gamm_horm_CSS_Inatt_simple,
                          newdata = grid,
                          type = "response", 
                         transform = function(x) exp(x) - 1)

# Contour plot
ggplot() +
  geom_tile(data = grid, aes(x = E2zd, y = P4zd, fill = predicted)) +
  geom_contour(data = grid,
               aes(x = E2zd, y = P4zd, z = predicted),
               color = "black") +
  geom_point(
    data = df,
    aes(x = E2zd, y = P4zd),
    color = "black",
    alpha = 0.25
  ) +
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = median(grid$predicted)) +
  labs(x = "E2zd", y = "P4zd", fill = "Outcome") +
  theme_minimal() +
  ggtitle("Predicting CSS_Inatt from E2 and P4 (Person Standardized)")

ggsave(filename = "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/EP_Models_Plots/CSS_Inatt.png", 
       plot = last_plot(), 
       width = 8, 
       height = 6)

```



# Looping through hormone effects - FIXED E2zd*P4zd TENSOR PRODUCT SMOOTH

```{r}
library(plotly)
library(htmlwidgets)

# Function to Loop Through Outcomes and Generate GAMM Models
create_gamm_models <- function(outcomelist, df_all, save_dir) {
  for (outcome in outcomelist) {
    # Prepare variables dynamically
    outcome_log <- paste0(outcome, "log")
    df_all[[outcome_log]] <- log(df_all[[outcome]] + 1)

    # Plot histogram for transformed outcome
    hist(df_all[[outcome]], main = paste("Histogram of", outcome))
    hist(df_all[[outcome_log]], main = paste("Histogram of Log-transformed", outcome))

    # GAMM model with hormone z-scores and derivatives
    gamm_formula <- as.formula(paste0(
      outcome_log, " ~ s(id, bs = \"re\") + te(E2zd, P4zd)"
    ))

    gamm_model <- gam(
      gamm_formula,
      data = df_all,
      family = gaussian,
      method = "REML"
    )

    # Summary and plot of the model
    model_summary <- summary(gamm_model)
    print(model_summary)

    # Save the model summary as a text file
    summary_output <- capture.output(model_summary)
    summary_file_path <- file.path(save_dir, paste0(outcome, "_fixed__wzch_85_model_EP_summary.txt"))
    writeLines(summary_output, summary_file_path)
    cat("Model summary saved to:", summary_file_path, "\n")

    # Plot the GAMM effects
   # plot.gam(gamm_model, select = 2, main = paste("Estrogen Effect on", outcome))
   # plot.gam(gamm_model, select = 3, main = paste("Progesterone Effect on", outcome))
  # plot.gam(gamm_model, select = 5, main = paste("Estrogen Change Effect on", outcome))
  # plot.gam(gamm_model, select = 6, main = paste("Progesterone Change Effect on", outcome))

    # Simpler model without random effects for visualization
    gamm_simple_formula <- as.formula(paste0(
      outcome_log, " ~ te(E2zd, P4zd)"
    ))

    gamm_model_simple <- gam(
      gamm_simple_formula,
      data = df_all,
      family = gaussian,
      method = "REML"
    )

    # Create the grid for contour plot
    grid <- expand.grid(
      E2zd = seq(
        min(df_all$E2zd, na.rm = TRUE),
        max(df_all$E2zd, na.rm = TRUE),
        length.out = 100
      ),
      P4zd = seq(
        min(df_all$P4zd, na.rm = TRUE),
        max(df_all$P4zd, na.rm = TRUE),
        length.out = 100
      )
    )

    # Predict values
    grid$predicted <- predict(gamm_model_simple,
                              newdata = grid,
                              type = "response",
                              transform = function(x) exp(x) - 1)

    # Contour plot
    plot_title <- paste("Predicting", outcome, "from E2 and P4 (Person Standardized)")
    contour_plot <- ggplot() +
      geom_tile(data = grid, aes(x = E2zd, y = P4zd, fill = predicted)) +
      geom_contour(data = grid, aes(x = E2zd, y = P4zd, z = predicted), color = "black") +
      geom_point(data = df_all, aes(x = E2zd, y = P4zd), color = "black", alpha = 0.25) +
      scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = median(grid$predicted)) +
      labs(x = "E2zd", y = "P4zd", fill = "Outcome") +
      theme_minimal() +
      ggtitle(plot_title)

    print(contour_plot)

    # Save the contour plot
    plot_file_path <- file.path(save_dir, paste0(outcome, "_te_zd_fixed.png"))
    ggsave(filename = plot_file_path, plot = contour_plot, width = 8, height = 6)
    cat("Contour plot saved to:", plot_file_path, "\n")
    
     # Create the 3D surface plot using Plotly
    grid_matrix <- matrix(grid$predicted, nrow = 100, byrow = TRUE)
    surface_plot <- plot_ly(
      x = seq(min(df_all$E2zd, na.rm = TRUE), max(df_all$E2zd, na.rm = TRUE), length.out = 100),
      y = seq(min(df_all$P4zd, na.rm = TRUE), max(df_all$P4zd, na.rm = TRUE), length.out = 100),
      z = ~grid_matrix
    ) %>%
      add_surface() %>%
      layout(
        title = paste("3D Surface Plot of E2zd and P4zd Interaction for", outcome),
        scene = list(
          xaxis = list(title = "E2zd"),
          yaxis = list(title = "P4zd"),
          zaxis = list(title = "Predicted Outcome")
        )
      )

    # Save the interactive surface plot as an HTML file
    surface_plot_file_path <- file.path(save_dir, paste0(outcome, "_zd_3D_surface.html"))
    saveWidget(surface_plot, surface_plot_file_path)
    cat("3D surface plot saved to:", surface_plot_file_path, "\n")
  }
}


# CALL THE LOOP

outcomelist <- c("UPPS_NU_avg", "UPPS_Persev_avg", "UPPS_Premed_avg", "UPPS_Sens_avg", "UPPS_PU_avg", 
  "CSS_Fx_Total", "CSS_Inatt", "CSS_HypImp", "CSS_Inatt_Count", "CSS_HypImp_Count", 
  "score_pinball", "score_robot", "BDEFS_Total", "BDEFS_WM_avg", "BDEFS_RI_avg", 
  "UPPS_Total", "DEBQ_Total")

create_gamm_models(outcomelist, df, "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-10/")

```

"DRSP_1", "DRSP_2", "DRSP_3", "DRSP_4", "DRSP_5", "DRSP_6", "DRSP_7", "DRSP_8", "DRSP_9", 
  "DRSP_10", "DRSP_11", "DRSP_12", "DRSP_13", "DRSP_14", "DRSP_15", "DRSP_16", "DRSP_17", 
  "DRSP_18", "DRSP_19", "DRSP_20", "DRSP_21", "DRSP_22", "DRSP_23",

# Looping through hormone effects - FIXED - RAW - TENSOR PRODUCT SMOOTH

```{r}

# Function to Loop Through Outcomes and Generate GAMM Models
create_gamm_models <- function(outcomelist, df_all, save_dir) {
  for (outcome in outcomelist) {
    # Prepare variables dynamically
    outcome_log <- paste0(outcome, "log")
    df_all[[outcome_log]] <- log(df_all[[outcome]] + 1)

    # Plot histogram for transformed outcome
    hist(df_all[[outcome]], main = paste("Histogram of", outcome))
    hist(df_all[[outcome_log]], main = paste("Histogram of Log-transformed", outcome))

    # GAMM model with hormone z-scores and derivatives
    gamm_formula <- as.formula(paste0(
      outcome_log, " ~ s(id, bs = \"re\") + te(E2, P4)"
    ))

    gamm_model <- gam(
      gamm_formula,
      data = df_all,
      family = gaussian,
      method = "REML"
    )

    # Summary and plot of the model
    model_summary <- summary(gamm_model)
    print(model_summary)

    # Save the model summary as a text file
    summary_output <- capture.output(model_summary)
    summary_file_path <- file.path(save_dir, paste0(outcome, "_fixed_te_rawEP_model_EP_summary.txt"))
    writeLines(summary_output, summary_file_path)
    cat("Model summary saved to:", summary_file_path, "\n")

    # Plot the GAMM effects
   # plot.gam(gamm_model, select = 2, main = paste("Estrogen Effect on", outcome))
   # plot.gam(gamm_model, select = 3, main = paste("Progesterone Effect on", outcome))
  # plot.gam(gamm_model, select = 5, main = paste("Estrogen Change Effect on", outcome))
  # plot.gam(gamm_model, select = 6, main = paste("Progesterone Change Effect on", outcome))

    # Simpler model without random effects for visualization
    gamm_simple_formula <- as.formula(paste0(
      outcome_log, " ~ te(E2, P4)"
    ))

    gamm_model_simple <- gam(
      gamm_simple_formula,
      data = df_all,
      family = gaussian,
      method = "REML"
    )

    # Create the grid for contour plot
    grid <- expand.grid(
      E2 = seq(
        min(df_all$E2, na.rm = TRUE),
        max(df_all$E2, na.rm = TRUE),
        length.out = 100
      ),
      P4 = seq(
        min(df_all$P4, na.rm = TRUE),
        max(df_all$P4, na.rm = TRUE),
        length.out = 100
      )
    )

    # Predict values
    grid$predicted <- predict(gamm_model_simple,
                              newdata = grid,
                              type = "response",
                              transform = function(x) exp(x) - 1)

    # Contour plot
    plot_title <- paste("Predicting", outcome, "from E2 and P4 (Person Standardized)")
    contour_plot <- ggplot() +
      geom_tile(data = grid, aes(x = E2, y = P4, fill = predicted)) +
      geom_contour(data = grid, aes(x = E2, y = P4, z = predicted), color = "black") +
      geom_point(data = df_all, aes(x = E2, y = P4), color = "black", alpha = 0.25) +
      scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = median(grid$predicted)) +
      labs(x = "E2", y = "P4", fill = "Outcome") +
      theme_minimal() +
      ggtitle(plot_title)

    print(contour_plot)

    # Save the contour plot
    plot_file_path <- file.path(save_dir, paste0(outcome, "rawEP_te_fixed.png"))
    ggsave(filename = plot_file_path, plot = contour_plot, width = 8, height = 6)
    cat("Contour plot saved to:", plot_file_path, "\n")
    
    
  }
}


# CALL THE LOOP

#OutcomeList already exists above

create_gamm_models(outcomelist, df, "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-04/")

```


# Looping through hormone effects - FIXED - RAW LOG - TENSOR PRODUCT SMOOTH

```{r}

# Function to Loop Through Outcomes and Generate GAMM Models
create_gamm_models <- function(outcomelist, df_all, save_dir) {
  for (outcome in outcomelist) {
    # Prepare variables dynamically
    outcome_log <- paste0(outcome, "log")
    df_all[[outcome_log]] <- log(df_all[[outcome]] + 1)

    # Plot histogram for transformed outcome
    hist(df_all[[outcome]], main = paste("Histogram of", outcome))
    hist(df_all[[outcome_log]], main = paste("Histogram of Log-transformed", outcome))

    # GAMM model with hormone z-scores and derivatives
    gamm_formula <- as.formula(paste0(
      outcome_log, " ~ s(id, bs = \"re\") + te(logE2 , logP4)"
    ))

    gamm_model <- gam(
      gamm_formula,
      data = df_all,
      family = gaussian,
      method = "REML"
    )

    # Summary and plot of the model
    model_summary <- summary(gamm_model)
    print(model_summary)

    # Save the model summary as a text file
    summary_output <- capture.output(model_summary)
    summary_file_path <- file.path(save_dir, paste0(outcome, "_fixed_te_rawlogEP_model_EP_summary.txt"))
    writeLines(summary_output, summary_file_path)
    cat("Model summary saved to:", summary_file_path, "\n")

    # Plot the GAMM effects
   # plot.gam(gamm_model, select = 2, main = paste("Estrogen Effect on", outcome))
   # plot.gam(gamm_model, select = 3, main = paste("Progesterone Effect on", outcome))
  # plot.gam(gamm_model, select = 5, main = paste("Estrogen Change Effect on", outcome))
  # plot.gam(gamm_model, select = 6, main = paste("Progesterone Change Effect on", outcome))

    # Simpler model without random effects for visualization
    gamm_simple_formula <- as.formula(paste0(
      outcome_log, " ~ te(logE2 , logP4)"
    ))

    gamm_model_simple <- gam(
      gamm_simple_formula,
      data = df_all,
      family = gaussian,
      method = "REML"
    )

    # Create the grid for contour plot
    grid <- expand.grid(
      logE2  = seq(
        min(df_all$logE2 , na.rm = TRUE),
        max(df_all$logE2 , na.rm = TRUE),
        length.out = 100
      ),
      logP4 = seq(
        min(df_all$logP4, na.rm = TRUE),
        max(df_all$logP4, na.rm = TRUE),
        length.out = 100
      )
    )

    # Predict values
    grid$predicted <- predict(gamm_model_simple,
                              newdata = grid,
                              type = "response",
                              transform = function(x) exp(x) - 1)

    # Contour plot
    plot_title <- paste("Predicting", outcome, "from logE2  and logP4")
    contour_plot <- ggplot() +
      geom_tile(data = grid, aes(x = logE2 , y = logP4, fill = predicted)) +
      geom_contour(data = grid, aes(x = logE2 , y = logP4, z = predicted), color = "black") +
      geom_point(data = df_all, aes(x = logE2 , y = logP4), color = "black", alpha = 0.25) +
      scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = median(grid$predicted)) +
      labs(x = "logE2 ", y = "logP4", fill = "Outcome") +
      theme_minimal() +
      ggtitle(plot_title)

    print(contour_plot)

    # Save the contour plot
    plot_file_path <- file.path(save_dir, paste0(outcome, "rawlogEP_te_fixed.png"))
    ggsave(filename = plot_file_path, plot = contour_plot, width = 8, height = 6)
    cat("Contour plot saved to:", plot_file_path, "\n")
  }
}


# CALL THE LOOP

#OutcomeList already exists above

create_gamm_models(outcomelist, df, "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-04/")

# Simulate data for logE2, logP4, and outcome
logE2 <- seq(-2, 2, length.out = 100)
logP4 <- seq(-2, 2, length.out = 100)
z <- outer(logE2, logP4, function(e2, p4) sin(e2) * cos(p4)) # Replace with actual model estimates

plot_ly(x = ~logE2, y = ~logP4, z = ~z, type = "surface") %>%
  layout(title = "Interaction Effect of logE2 and logP4 on Outcome",
         scene = list(
           xaxis = list(title = "logE2"),
           yaxis = list(title = "logP4"),
           zaxis = list(title = "Predicted Outcome")
         ))


```






# Function for te(???? not sure) in 85 dataset with 3D html plot output
```{r}
# Load required libraries
library(dplyr)
library(tidyr)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(mgcv)
library(plotly)
library(htmlwidgets)

# Function to Loop Through Outcomes and Generate GAMM Models
create_gamm_models <- function(outcomelist, df_all, save_dir) {
  for (outcome in outcomelist) {
    # Prepare variables dynamically
    outcome_log <- paste0(outcome, "log")
    df_all <- df_all %>% mutate(!!outcome_log := log(!!sym(outcome) + 1))

    # Plot histogram for transformed outcome using ggplot2
    hist_orig <- ggplot(df_all, aes(x = !!sym(outcome))) +
      geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
      labs(title = paste("Histogram of", outcome), x = outcome, y = "Count") +
      theme_minimal()

    hist_log <- ggplot(df_all, aes(x = !!sym(outcome_log))) +
      geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
      labs(title = paste("Histogram of Log-transformed", outcome), x = paste0(outcome, " (log)"), y = "Count") +
      theme_minimal()

    # Arrange the two histograms side by side and save
    histograms <- grid.arrange(hist_orig, hist_log, ncol = 2)
    hist_file_path <- file.path(save_dir, paste0(outcome, "_histograms.png"))
    ggsave(filename = hist_file_path, plot = histograms, width = 12, height = 6)
    cat("Histograms saved to:", hist_file_path, "\n")

    # GAMM model with hormone z-scores and derivatives
    gamm_formula <- as.formula(paste0(
      outcome_log, " ~ s(id, bs = \"re\") + te(logE2 , logP4)"
    ))

    gamm_model <- gam(
      gamm_formula,
      data = df_all,
      family = gaussian(),
      method = "REML"
    )

    # Summary and plot of the model
    model_summary <- summary(gamm_model)
    print(model_summary)

    # Save the model summary as a text file
    summary_output <- capture.output(model_summary)
    summary_file_path <- file.path(save_dir, paste0(outcome, "_fixed_te_rawlogEP_model_EP_summary.txt"))
    writeLines(summary_output, summary_file_path)
    cat("Model summary saved to:", summary_file_path, "\n")

    # Simpler model without random effects for visualization
    gamm_simple_formula <- as.formula(paste0(
      outcome_log, " ~ te(logE2 , logP4)"
    ))

    gamm_model_simple <- gam(
      gamm_simple_formula,
      data = df_all,
      family = gaussian(),
      method = "REML"
    )

    # Create the grid for contour plot
    grid <- expand_grid(
      logE2 = seq(
        min(df_all$logE2, na.rm = TRUE),
        max(df_all$logE2, na.rm = TRUE),
        length.out = 100
      ),
      logP4 = seq(
        min(df_all$logP4, na.rm = TRUE),
        max(df_all$logP4, na.rm = TRUE),
        length.out = 100
      )
    )

    # Predict values for contour plot
    grid <- grid %>%
      mutate(predicted = predict(gamm_model_simple, newdata = ., type = "response"))

    # Contour plot using ggplot2
    plot_title <- paste("Predicting", outcome, "from logE2 and logP4")
    contour_plot <- ggplot(grid, aes(x = logE2, y = logP4, z = predicted)) +
      geom_tile(aes(fill = predicted)) +
      geom_contour(color = "black") +
      scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = median(grid$predicted, na.rm = TRUE)) +
      labs(x = "logE2", y = "logP4", fill = "Predicted Outcome") +
      theme_minimal() +
      ggtitle(plot_title)

    print(contour_plot)

    # Save the contour plot
    plot_file_path <- file.path(save_dir, paste0(outcome, "_rawlogEP_te_fixed.png"))
    ggsave(filename = plot_file_path, plot = contour_plot, width = 8, height = 6)
    cat("Contour plot saved to:", plot_file_path, "\n")

    # Create the 3D surface plot using Plotly
    grid_matrix <- matrix(grid$predicted, nrow = 100, byrow = TRUE)
    surface_plot <- plot_ly(
      x = seq(min(df_all$logE2, na.rm = TRUE), max(df_all$logE2, na.rm = TRUE), length.out = 100),
      y = seq(min(df_all$logP4, na.rm = TRUE), max(df_all$logP4, na.rm = TRUE), length.out = 100),
      z = ~grid_matrix
    ) %>%
      add_surface() %>%
      layout(
        title = paste("3D Surface Plot of logE2 and logP4 Interaction for", outcome),
        scene = list(
          xaxis = list(title = "logE2"),
          yaxis = list(title = "logP4"),
          zaxis = list(title = "Predicted Outcome")
        )
      )

    # Save the interactive surface plot as an HTML file
    surface_plot_file_path <- file.path(save_dir, paste0(outcome, "_3D_surface.html"))
    saveWidget(surface_plot, surface_plot_file_path)
    cat("3D surface plot saved to:", surface_plot_file_path, "\n")
  }
}

# CALL THE LOOP
create_gamm_models(outcomelist, df_all, "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-04/")

```

