---
title: "ADHDCYC_MLMs_July2024"
author: "Tory Eisenlohr-Moul"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 2024-07-16 - PLOTS and MLMs and IDIOGRAPHIC - BEFORE CYCLE SCALING

## Changes in careless responding across the menstrual cycle - BY ID

```{r}
# Define the output directory - DO THIS FOR EACH DAY YOU RUN IT

dailydir <- '~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-16'

# Loop through each participant ID and call the function - CSS_Fx_Total
for (participant_id in id_list) {
  plot_participant_data(df, participant_id, 'longstring', "Length of Identical Responding", dailydir)
}
```



## CALLING PERSON GRAPHING FUNCTIONS
### 1 - MAKE ALL CSS_Fx_Total + Hormone GRAPHS - Loop through each participant ID and call the function - CSS_Fx_Total
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}

# Define the output directory - DO THIS FOR EACH DAY YOU RUN IT

dailydir <- '~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024_07_09'

# Loop through each participant ID and call the function - CSS_Fx_Total
for (participant_id in id_list) {
  plot_participant_data(df, participant_id, 'CSS_Fx_Total', "CSS Functional Impact", dailydir)
}


```

### 2 - MAKE ALL HYPIMP + Hormone GRAPHS - Loop through each participant ID and call the function - CSS_HypImp
```{r warning=FALSE}
# Define the output directory - DO THIS FOR EACH DAY YOU RUN IT

dailydir <- '/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024_07_09'


# Loop through each participant ID and call the function - CSS_HypImp
for (participant_id in id_list) {
  plot_participant_data(df, participant_id, 'CSS_HypImp', "CSS Hyperactivity/Impulsivity", dailydir)
}

```

### 3 - MAKE ALL INATT + Hormone GRAPHS - Loop through each participant ID and call the function - CSS_Inatt
```{r warning=FALSE}

# Define the output directory - DO THIS FOR EACH DAY YOU RUN IT

dailydir <- '/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024_07_09'

# Loop through each participant ID and call the function - CSS_Inatt
for (participant_id in id_list) {
  plot_participant_data(df, participant_id, 'CSS_Inatt', "CSS Inattentiveness", dailydir)
}
```

### 4 - MAKE ALL INATT COUNT + Hormone GRAPHS - Loop through each participant ID and call the function
```{r warning=FALSE}

# Define the output directory - DO THIS FOR EACH DAY YOU RUN IT

dailydir <- '/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024_07_09'

# Loop through each participant ID and call the function - CSS_Inatt
for (participant_id in id_list) {
  plot_participant_data(df, participant_id, 'CSS_Inatt_Count', "CSS Inattentive Criteria (Count)", dailydir)
}
```

### 5 - MAKE ALL HYPIMP COUNT + Hormone GRAPHS - Loop through each participant ID and call the function
```{r warning=FALSE}

# Define the output directory - DO THIS FOR EACH DAY YOU RUN IT

dailydir <- '/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024_07_09'

# Loop through each participant ID and call the function
for (participant_id in id_list) {
  plot_participant_data(df, participant_id, 'CSS_HypImp_Count', "CSS Hyperactive/Impulsive Criteria (Count)", dailydir)
}
```




## Create plot_interaction function
```{r}

# Define the function
plot_interaction <- function(data, outcome, dailydir) {
  # Check for problematic values in the specified variables
  data_clean <- data %>%
    filter(!is.na(.data[[outcome]]) & !is.nan(.data[[outcome]]) & !is.infinite(.data[[outcome]]) &
           !is.na(E2) & !is.nan(E2) & !is.infinite(E2) &
           !is.na(P4) & !is.nan(P4) & !is.infinite(P4))
  
  # Scale predictors by random effect and calculate daily derivatives
  data_clean <- data_clean %>%
    group_by(id) %>%
    mutate(
      E2zd = scale(E2),
      P4zd = scale(P4),
      E2ch = c(NA, diff(E2)),  # Daily derivative for E2
      P4ch = c(NA, diff(P4))   # Daily derivative for P4
    ) %>%
    filter(!is.na(E2zd) & !is.nan(E2zd) & !is.infinite(E2zd) &
           !is.na(P4zd) & !is.nan(P4zd) & !is.infinite(P4zd) &
           !is.na(E2ch) & !is.nan(E2ch) & !is.infinite(E2ch) &
           !is.na(P4ch) & !is.nan(P4ch) & !is.infinite(P4ch)) %>%
    ungroup()
  
  # Create log-transformed outcome
  data_clean <- data_clean %>%
    mutate(log_outcome = log(.data[[outcome]])) %>%
    filter(!is.na(log_outcome) & !is.nan(log_outcome) & !is.infinite(log_outcome))
  
  # Fit the models with cleaned data
  rawmodel <- lmer(as.formula(paste(outcome, "~ E2zd * P4zd + (1 | id)")), data = data_clean)
  logmodel <- lmer(log_outcome ~ E2zd * P4zd + (1 | id), data = data_clean)
  model_prand <- lmer(log_outcome ~ E2zd * P4zd + (1 + P4zd | id), data = data_clean)
  model_erand <- lmer(log_outcome ~ E2zd * P4zd + (1 + E2zd | id), data = data_clean)
  model_derivs <- lmer(log_outcome ~ E2zd * P4zd + E2ch + P4ch + (1 | id), data = data_clean)
  
  # Print the model summaries
  # print(summary(rawmodel))
  # print(summary(logmodel))
  # print(summary(model_prand))
  # print(summary(model_erand))
  # print(summary(model_derivs))
  
  # Specify the file path for the HTML output - primary model of E2 * P4
  output_file <- file.path(dailydir, paste(outcome, "horm_tabmodel_output.html", sep = ""))
  cat("Output file path:", output_file, "\n")
  
  # Generate the HTML output from tab_model without saving to a file
  tab_model(rawmodel, logmodel, model_prand, model_erand, model_derivs)
  tab_model_output <- tab_model(rawmodel, logmodel, model_prand, model_erand, model_derivs)
  
  # Capture the HTML output as a string
  html_output <- as.character(tab_model_output)
  
  # Manually write the HTML output to the specified file
  writeLines(html_output, con = output_file)
  
  # Create a grid of values for predictors
  E2zd_seq <- seq(min(data_clean$E2zd), max(data_clean$E2zd), length.out = 100)
  P4zd_seq <- seq(min(data_clean$P4zd), max(data_clean$P4zd), length.out = 100)
  
  # Create a dataframe with all combinations of E2zd and P4zd
  grid_df <- expand.grid(E2zd = E2zd_seq, P4zd = P4zd_seq)
  
  # Add the mean values of the derivatives to the grid_df
  grid_df$E2ch <- mean(data_clean$E2ch, na.rm = TRUE)
  grid_df$P4ch <- mean(data_clean$P4ch, na.rm = TRUE)
  
  # Predict the outcome over the grid
  grid_df$predicted_log_outcome <- predict(logmodel, newdata = grid_df, re.form = NA)
  
  # Exponentiate the predicted values to return them to the original scale
  grid_df$predicted_outcome <- exp(grid_df$predicted_log_outcome)

# Create the interaction plot
intplot <- ggplot(grid_df, aes(x = E2zd, y = predicted_outcome, color = P4zd, group = P4zd)) +
    geom_line() +
    labs(
        title = paste("Interaction between E2zd and P4zd on", outcome),
        x = "E2zd",
        y = paste("Predicted", outcome),
        color = "P4zd"
    ) +
    theme_minimal(base_size = 15) +
    scale_color_gradient(low = "blue", high = "red") +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        panel.background = element_rect(fill = "white", color = NA), # White background
        plot.background = element_rect(fill = "white", color = NA),  # White background
        panel.grid.major = element_line(color = "gray90"), # Light gray major grid lines
        panel.grid.minor = element_line(color = "gray95")  # Lighter gray minor grid lines
    )

  # Save and return the plot
  ggsave(filename = file.path(dailydir, paste0("EPintplot_", outcome, ".tif")), plot = intplot, width = 12, height = 8, units = "in")
  
  return(intplot)
}


```

## Interaction Plots (running "plot_interaction")

```{r}

dailydir <- '~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-20'


plot_interaction(df_sens, "longstring", dailydir)


# Example usage
plot_interaction(df_sens, "CSS_Inatt", dailydir)
plot_interaction(df_sens, "CSS_Inatt_Count", dailydir) *SIG

plot_interaction(df_sens, "CSS_HypImp", dailydir)
plot_interaction(df_sens, "CSS_HypImp_Count", dailydir)

plot_interaction(df_sens, "CSS_Fx_Total", dailydir)

plot_interaction(df_sens, "CSS_B2_Total", dailydir)

plot_interaction(df_sens, "UPPS_Total", dailydir)

plot_interaction(df_sens, "BDEFS_Total", dailydir)

 plot_interaction(df_sens, "DEBQ_Total", dailydir)

```







##  Function to plot by cycle day ("plot_by_cycle_day")
```{r}
plot_by_cycle_day <- function(df, var) {
#Create individual overall means

df <- df %>%
group_by(id) %>%
mutate(var.roll.m = mean(.data[[var]], na.rm = TRUE))

dfcycdayplotdata <- df %>%
group_by(id, hybridcycleday) %>%
filter(!is.na(hybridcycleday)) %>%
summarize(
var.roll.m = mean(var.roll.m, na.rm = TRUE),
var.day.roll.m = mean(.data[[var]], na.rm = TRUE),
var.day.roll.d = var.day.roll.m - var.roll.m
) %>%
arrange(factor(hybridcycleday, levels = hybridcycleday_order))

#print(dfcycdayplotdata)

cycd_plot <- dfcycdayplotdata %>%
group_by(hybridcycleday) %>%
summarize(
personcentered.dev = mean(var.day.roll.d, na.rm = TRUE),
se = sd(var.day.roll.d, na.rm = TRUE) / sqrt(n())
) %>%
arrange(factor(hybridcycleday, levels = hybridcycleday_order)) %>%
ggplot(aes(x = factor(hybridcycleday, levels = hybridcycleday_order), y = personcentered.dev)) +
geom_line(group = 1) +
#geom_ribbon(aes(ymin = personcentered.dev - se, ymax = personcentered.dev + se), alpha = 0.2) +
geom_hline(yintercept = 0, linetype = "solid", col = "black") +

geom_errorbar(aes(ymin = personcentered.dev - se, ymax = personcentered.dev + se), size = 0.5, width = 0.5, position = position_dodge()) +
#scale_y_continuous(expand = c(0, 0.011)) +
theme_bw() +
xlab("Cycle Day\n(Relative to LH-surge and Menses Onset)") +
ylab(paste("Person-Centered", var)) +

  #Midfollicular Label
annotate("rect", ymin = .35, ymax = .45, xmin = "-7L", xmax = "-3L", alpha = .2) +
annotate("text", x = "-7L", y = .4, label = "Midfollicular", size = 3, hjust = -0.25) +
  
  #Periovulatory Label
annotate("rect", ymin = .35, ymax = .45, xmin = "-2L", xmax = "1L", alpha = .2) +
annotate("text", x = "-2L", y = .4, label = "Periovulatory", size = 3, hjust = 0) +
  
  #Early Luteal Label
annotate("rect", ymin = .35, ymax = .45, xmin = "2L", xmax = "5L", alpha = .2) +
annotate("text", x = "2L", y = .4, label = " Early Luteal", size = 3, hjust = 0) +
 
   #Midluteal Label
annotate("rect", ymin = .35, ymax = .45, xmin = "-9M", xmax = "-5M", alpha = .2) +
annotate("text", x = "-9M", y = .4, label = "Midluteal", size = 3, hjust = -0.5) +

  #Perimenstrual Label
annotate("rect", ymin = .35, ymax = .45, xmin = "-3M", xmax = "2M", alpha = .2) +
annotate("text", x = "-3M", y = .4, label = "Perimenstrual", size = 3, hjust = -0.15) +
 
   #Annotation of Ovulation
annotate("text", x = "0L", y = -.25, label = "*", size = 8, color = "forestgreen") +
 
   #Annotation of Menses
annotate("text", x = "1M", y = -.25, label = "*", size = 8, color = "red")
print(cycd_plot)
}

```

##  Function to plot by cycle phase ("plot_by_cycle_phase")
```{r warning=FALSE}

plot_by_cycle_phase <- function(df, var) {
  df$cyclephase_LH_graph <- as.factor(df$cyclephase_LH_graph)
  
  # Calculate overall mean for each individual
  df <- df %>%
    group_by(id) %>%
    mutate(overall_mean = mean(.data[[var]], na.rm = TRUE)) %>%
    ungroup()
  
  # Print intermediate data for debugging
  # print(head(df))
  
  # Calculate phase-specific deviation
  dfphaseplotdata <- df %>%
    filter(!is.na(cyclephase_LH_graph)) %>%
    group_by(id, cyclephase_LH_graph) %>%
    summarize(
      phase_mean = mean(.data[[var]], na.rm = TRUE),
      overall_mean = first(overall_mean)
    ) %>%
    mutate(phase_deviation = phase_mean - overall_mean) %>%
    ungroup()
  
  # Print intermediate data for debugging
  # print(head(dfphaseplotdata))
  
  # Calculate grand mean and standard error
  phase_plot_data <- dfphaseplotdata %>%
    group_by(cyclephase_LH_graph) %>%
    summarize(
      grand_mean_deviation = mean(phase_deviation, na.rm = TRUE),
      se = sd(phase_deviation, na.rm = TRUE) / sqrt(n())
    )
  
  # Print summarized data for debugging
 # print(head(phase_plot_data))
  
  # Create the plot
  phase_plot <- phase_plot_data %>%
    ggplot(aes(x = cyclephase_LH_graph, y = grand_mean_deviation, group = 1)) +
    geom_line() + # Add the line
    scale_x_discrete(labels = c(
      "1" = "Midfollicular",
      "2" = "Periovulatory", 
      "3" = "Early Luteal",
      "4" = "Midluteal",
      "5" = "Perimenstrual"
    )) +
    geom_ribbon(aes(ymin = grand_mean_deviation - se, ymax = grand_mean_deviation + se), alpha = 0.2) +
    geom_hline(yintercept = 0.0, color = "black", linetype = "solid") +
    theme_bw(base_size = 11) +
    xlab("Cycle Phase") +
    ylab(paste("Person-Centered", var))
  
  print(phase_plot)
}


```


##  FUNCTION TO RUN&SAVE PHASE & CYCLEDAY GRAPHS TO BOX
```{r}

# Define the function
run_and_save_plots <- function(variable_name, df, output_directory) {
  
  # Ensure the output directory exists
  dir.create(output_directory, recursive = TRUE, showWarnings = FALSE)
  
  # Create and save Cycle Day Plot
  day_plot <- plot_by_cycle_day(df, paste0(variable_name, ".roll.d"))
  ggsave(file.path(output_directory, paste0(variable_name, "_DAYPLOT.png")), 
         plot = day_plot, width = 10, height = 6, dpi = 300)
  
  # Create and save Cycle Phase Plot
  phase_plot <- plot_by_cycle_phase(df, paste0(variable_name, ".roll.d"))
  ggsave(file.path(output_directory, paste0(variable_name, "_PHSPLOT.png")), 
         plot = phase_plot, width = 10, height = 6, dpi = 300)
}


```


##  Function to run and save MLMs to html in dated Box output folder ("run_and_save_mlms")
```{r message=TRUE, warning=TRUE}

run_and_save_mlms <- function(variable_name, data, output_directory) {
  ## peri
  model_peri <- lmer(as.formula(paste(variable_name, "~ midfol_LH + periov_LH + earlylut_LH + midluteal_LH + (1 | id)")),
                     data = data)
  cat("Peri model fitted.\n")
  
  ## midfol
  model_midfol <- lmer(as.formula(paste(variable_name, "~ periov_LH + earlylut_LH + midluteal_LH + perimenstrual_LH + (1 | id)")),
                       data = data)
  cat("Midfol model fitted.\n")
  
  ## periov
  model_periov <- lmer(as.formula(paste(variable_name, "~ midfol_LH + earlylut_LH + midluteal_LH + perimenstrual_LH + (1 | id)")),
                       data = data)
  cat("Periov model fitted.\n")
  
  ## earlylut
  model_earlylut <- lmer(as.formula(paste(variable_name, "~ midfol_LH + periov_LH + midluteal_LH + perimenstrual_LH + (1 | id)")),
                         data = data)
  cat("Earlylut model fitted.\n")
  
  ## midlut
  model_midlut <- lmer(as.formula(paste(variable_name, "~ midfol_LH + periov_LH + earlylut_LH + perimenstrual_LH + (1 | id)")),
                       data = data)
  cat("Midlut model fitted.\n")
  
  # Create the directory if it doesn't exist
  if (!dir.exists(output_directory)) {
    dir.create(output_directory, recursive = TRUE, showWarnings = FALSE)
    cat("Directory created:", output_directory, "\n")
  } else {
    cat("Directory already exists:", output_directory, "\n")
  }
  
  # Specify the file path for the HTML output
  output_file <- file.path(output_directory, paste(variable_name, "_PHS_tab_model_output.html", sep = ""))
  cat("Output file path:", output_file, "\n")
  
  # Generate the HTML output from tab_model without saving to a file
  tab_model_output <- tab_model(model_midfol, model_periov, model_earlylut, model_midlut, model_peri)
  
  # Capture the HTML output as a string
  html_output <- as.character(tab_model_output)
  
  # Manually write the HTML output to the specified file
  writeLines(html_output, con = output_file)
  
  # Check if the file was created and has content
  if (file.exists(output_file)) {
    cat("HTML output generated and saved to:", output_file, "\n")
  } else {
    cat("Failed to save HTML output to:", output_file, "\n")
  }
}

# Run for all Outcomes

output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"

df <- df

```



##  CALL PLOTTING FUNCTIONS FOR CARELESS RESPONDING INDICES

```{r}

plot_by_cycle_day(df_sens, "longstring")
plot_by_cycle_phase(df_sens, "longstring")
```




##  CALL PLOTTING FUNCTIONS FOR HORMONES
```{r}

#E2
plot_by_cycle_day(df_sens, "E2")
plot_by_cycle_phase(df_sens, "E2")

#P4
plot_by_cycle_day(df_sens, "P4")
plot_by_cycle_phase(df_sens, "P4")

#LH
plot_by_cycle_day(df_sens, "LH")
plot_by_cycle_phase(df_sens, "LH")



```



##  CALL FUNCTIONS: E2

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("E2", df, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "E2"

#### NO NEED TO CHANGE ANYTHING BELOW

df <- df  # Replace with your actual dataframe
output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df, output_directory)

```
## # CALL FUNCTIONS: P4

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("P4", df, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "P4"

#### NO NEED TO CHANGE ANYTHING BELOW

df <- df  # Replace with your actual dataframe
output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df, output_directory)

```

## CALL FUNCTIONS: LH

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("LH", df, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "LH"

#### NO NEED TO CHANGE ANYTHING BELOW

df <- df  # Replace with your actual dataframe
output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-04"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df, output_directory)

```


## CALL FUNCTION FOR CARELESS RESPONDING VARS

```{r}

output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-20"


run_and_save_plots("longstring", df, output_directory)

```


##CALL FUNCTIONS: CSS_Inatt

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("CSS_Inatt", df_sens, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "CSS_Inatt"

#### NO NEED TO CHANGE ANYTHING BELOW

output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df_sens, output_directory)

```

## CALL FUNCTIONS: CSS_HypImp

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("CSS_HypImp", df_sens, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "CSS_HypImp"

#### NO NEED TO CHANGE ANYTHING BELOW

output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df_sens, output_directory)

```

## CALL FUNCTIONS: CSS_B2_Total

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("CSS_B2_Total", df_sens, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "CSS_B2_Total"

#### NO NEED TO CHANGE ANYTHING BELOW

output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df_sens, output_directory)

```

## CALL FUNCTIONS: CSS_Fx_Total

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("CSS_Fx_Total", df_sens, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "CSS_Fx_Total"

#### NO NEED TO CHANGE ANYTHING BELOW

output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df_sens, output_directory)

```




## CALL FUNCTIONS: CSS_Inatt_Count

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("CSS_Inatt_Count", df_sens, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "CSS_Inatt_Count"

#### NO NEED TO CHANGE ANYTHING BELOW

output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df_sens, output_directory)

```



## CALL FUNCTIONS: CSS_Imp_Count

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("CSS_HypImp_Count", df_sens, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "CSS_HypImp_Count"

#### NO NEED TO CHANGE ANYTHING BELOW

df <- df  # Replace with your actual dataframe
output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df_sens, output_directory)

```




## CALL FUNCTIONS: BDEFS_Total

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("BDEFS_Total", df_sens, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "BDEFS_Total"

#### NO NEED TO CHANGE ANYTHING BELOW

df <- df  # Replace with your actual dataframe
output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df_sens, output_directory)

```



## CALL FUNCTIONS: DEBQ_Total

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("DEBQ_Total", df_sens, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "DEBQ_Total"

#### NO NEED TO CHANGE ANYTHING BELOW

df <- df  # Replace with your actual dataframe
output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df_sens, output_directory)

```


## CALL FUNCTIONS: UPPS_Total

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("UPPS_Total", df_sens, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "UPPS_Total"

#### NO NEED TO CHANGE ANYTHING BELOW

df <- df  # Replace with your actual dataframe
output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df_sens, output_directory)

```

