---
title: "ADHDCYCLE-scoring-scales"
author: "Tory Eisenlohr-Moul"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
#Using IA_Count (as a known measure of more sx) to check direction of variable coding

```{r}

#---- Checking BDEFS WM - Yes, higher is more sx r=.66 ----
cor(df$CSS_Inatt_Count, df$BDEFS_WM_avg, use="pairwise.complete.obs")

#---- Checking BDEFS RI - Yes, higher is more sx r=.57----
cor(df$CSS_Inatt_Count, df$BDEFS_RI_avg, use="pairwise.complete.obs")

#---- Checking RAW BDEFS  - Yes, higher is more sx for all ----
cor(df$CSS_Inatt_Count, df$BDEFS_3, use="pairwise.complete.obs")

#---- Checking UPPS raw variables - Yes, higher is more sx for all UPPS vars (incl 7 and 8) ----
cor(df$CSS_Inatt_Count, df$UPPS_15, use="pairwise.complete.obs")

#---- Checking DEBQ_avg - Yes, higher is more sx r=.28----
cor(df$CSS_Inatt_Count, df$DEBQ_avg, use="pairwise.complete.obs")

#---- Checking DRSP - Yes, higher is more sx ----
cor(df$CSS_Inatt_Count, df$DRSP_1, use="pairwise.complete.obs")

#---- Checking score_robot - NO, higher is lower sx, so we will recode ----
cor(df$CSS_Inatt_Count, df$score_robot, use="pairwise.complete.obs")

#---- Checking score_pinball - NO, higher is lower sx, so we will recode ----
cor(df$CSS_Inatt_Count, df$score_pinball, use="pairwise.complete.obs")
```
#Notes re: how existing variables were coded --- but I'm not sure this is actually what was used because names don't match up 
<!-- **************************************. -->
<!-- Original Syntax by: Ashley Eng. -->
<!-- **************************************. -->
<!-- COMPUTE CSS_B_TotalIASymp=CSS_B_1 + CSS_B_3 + CSS_B_5 + CSS_B_7 + CSS_B_9 + CSS_B_11 + CSS_B_13 +  -->
<!--     CSS_B_15 + CSS_B_17. -->
<!-- VARIABLE LABELS CSS_B_TotalIASymp 'items 1, 3, 5, 7, 9. 11. 13. 15 .17'.  -->
<!-- EXECUTE. -->
<!-- COMPUTE CSS_B_TotalHISymp=CSS_B_2 + CSS_B_4 + CSS_B_6 + CSS_B_8 + CSS_B_10 + CSS_B_12 + CSS_B_14 +  -->
<!--     CSS_B_16 + CSS_B_18. -->
<!-- VARIABLE LABELS CSS_B_TotalHISymp 'items 2 4 6 8 10 12 14 16 18'.  -->
<!-- EXECUTE. -->
<!-- COMPUTE CSS_TotalFunction=CSS_Function_1 + CSS_Function_2 + CSS_Function_3 + CSS_Function_4 + CSS_Function_5 + CSS_Function_6 + CSS_Function_7 +  -->
<!--     CSS_Function_8 + CSS_Function_9 + CSS_Function_10. -->
<!-- EXECUTE. -->
<!-- COMPUTE CSS_B2_Total=CSS_B2_1 + CSS_B2_2 + CSS_B2_3 + CSS_B2_4 + CSS_B2_5 + CSS_B2_6 + CSS_B2_7 +  -->
<!--     CSS_B2_8. -->
<!-- EXECUTE. -->
<!-- COMPUTE DEBQ_Total=DEBQ_1 + DEBQ_2 + DEBQ_3 + DEBQ_4 + DEBQ_5 + DEBQ_6 + DEBQ_7 + DEBQ_8 + DEBQ_9 +  -->
<!--     DEBQ_10 + DEBQ_11 + DEBQ_12 + DEBQ_13. -->
<!-- EXECUTE. -->
<!-- COMPUTE BDEFS_Total=BDEFS_1 + BDEFS_2 + BDEFS_3 + BDEFS_4 + BDEFS_5 + BDEFS_6. -->
<!-- EXECUTE. -->
<!-- RECODE UPPS_7 UPPS_8 (1=4) (2=3) (3=2) (4=1). -->
<!-- EXECUTE. -->
<!-- COMPUTE UPPS_Total=UPPS_1 + UPPS_2 + UPPS_3 + UPPS_4 + UPPS_5 + UPPS_6 + UPPS_7 + UPPS_8 + UPPS_9 +  -->
<!--     UPPS_10 + UPPS_11 + UPPS_12 + UPPS_13 + UPPS_14 + UPPS_15. -->
<!-- VARIABLE LABELS UPPS_Total 'After recoding items 7 and 8'. -->
<!-- EXECUTE. -->


```{r}
# ---- Calculate ADHD, UPPS, BDEFS, and DEBQ Scores ----

df <- df %>%
  group_by(id) %>%
  mutate(
    # Convert columns to numeric to ensure proper summation
    CSS_Inatt = rowMeans(across(
      c(
        CSS_B_1,
        CSS_B_3,
        CSS_B_5,
        CSS_B_7,
        CSS_B_9,
        CSS_B_11,
        CSS_B_13,
        CSS_B_15,
        CSS_B_17
      ),
      ~ as.numeric(.)
    ), na.rm = TRUE),
    # Convert each column to numeric, handle missing values with na.rm = TRUE
    
    CSS_HypImp = rowMeans(across(
      c(
        CSS_B_2,
        CSS_B_4,
        CSS_B_6,
        CSS_B_8,
        CSS_B_10,
        CSS_B_12,
        CSS_B_14,
        CSS_B_16,
        CSS_B_18
      ),
      ~ as.numeric(.)
    ), na.rm = TRUE),
    # Same process for CSS_HypImp scores
    
    # Adding the counts without needing row-wise summation
    CSS_HypImp_Count = CSS_Imp_Count + CSS_Hyp_Count,
    
    # Direct summation of these variables, assuming they are already numeric
    
    # Functional Total Score: Summing specific function-related columns
    CSS_Fx_Total = rowMeans(across(
      c(
        CSS_Function_1,
        CSS_Function_2,
        CSS_Function_3,
        CSS_Function_4,
        CSS_Function_5,
        CSS_Function_6,
        CSS_Function_7,
        CSS_Function_8,
        CSS_Function_9,
        CSS_Function_10
      ),
      ~ as.numeric(.)
    ), na.rm = TRUE),
    
    # DEBQ Total Score summing all DEBQ columns
    DEBQ_Total = rowMeans(across(
      c(
        DEBQ_1,
        DEBQ_2,
        DEBQ_3,
        DEBQ_4,
        DEBQ_5,
        DEBQ_6,
        DEBQ_7,
        DEBQ_8,
        DEBQ_9,
        DEBQ_10,
        DEBQ_11,
        DEBQ_12,
        DEBQ_13
      ),
      ~ as.numeric(.)
    ), na.rm = TRUE),
    
    # BDEFS Total Score summing all BDEFS columns
    BDEFS_Total = rowMeans(across(
      c(BDEFS_1, BDEFS_2, BDEFS_3, BDEFS_4, BDEFS_5, BDEFS_6),
      ~ as.numeric(.)
    ), na.rm = TRUE)
  ) %>%
  ungroup()

# Reverse code pinball and robot

# Step 1: Find the maximum values for both variables
max_robot <- max(df$score_robot, na.rm = TRUE)
max_pinball <- max(df$score_pinball, na.rm = TRUE)

# Step 2: Reverse code the variables WITH SAME NAME
df$score_robot <- max_robot - df$score_robot
df$score_pinball <- max_pinball - df$score_pinball


# Check Histograms - they are all zero inflated

#hist(df$CSS_Inatt, breaks = 10, main = "Histogram of CSS_Inatt", xlab = "CSS_Inatt", col = "lightblue")
#hist(df$CSS_HypImp, breaks = 10, main = "Histogram of CSS_HypImp", xlab = "CSS_HypImp", col = "lightblue")
#hist(df$BDEFS_Total, breaks = 10, main = "Histogram of BDEFS_Total", xlab = "BDEFS_Total", col = "lightblue")
#hist(df$BDEFS_WM_avg, breaks = 10, main = "Histogram of BDEFS_WM_avg", xlab = "BDEFS_WM_avg", col = "lightblue")
#hist(df$BDEFS_RI_avg)


#hist(df$DEBQ_Total, breaks = 10, main = "Histogram of DEBQ_Total", xlab = "DEBQ_Total", col = "lightblue")

# Checking direction - They are all positively correlated

#cor(df$IA_Count, df$DEBQ_Total, use="pairwise.complete.obs")
#cor(df$BDEFS_Total, df$BDEFS_WM_avg, use="pairwise.complete.obs")
#cor(df$BDEFS_WM_avg, df$score_pinball, use="pairwise.complete.obs")
#cor(df$BDEFS_RI_avg, df$score_robot, use="pairwise.complete.obs")


```

