---
title: "prep_scaled_cycleday"
author: "Anisha Nagpal"
date: "`r Sys.Date()`"
output: html_document
---


```{r}
library(tidyverse)
library(readxl)
library(ggplot2)
library(dplyr)
library(tidyr)
library(haven)
```



```{r}
dat <- read_sav("C:/Users/Student/Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CLEAR2/02_datasets/CLEAR2_DAILY/03_cleaned_data/2024-07-15 CLEAR2DAILY.sav")
```

## Functions 

```{r}
calculate_mcyclength <- function(data) {
  # Calculate m2mcount: This counts from 1 menses onset to the next to give the length of a menses-to-menses cycle
  data <- data %>%
    group_by(id) %>%
    arrange(daterated, .by_group = TRUE) %>%
    mutate(
      m2mcount = NA,
      mcyclength = NA,
      cycle_incomplete = 0
    )
  
  for (i in seq_len(nrow(data))) {
    if (!is.na(data$A[i]) && data$A[i] == 1) {
      data$m2mcount[i] <- 1
      j <- i + 1
      
      while (j <= nrow(data) &&
             data$id[j] == data$id[i] && (is.na(data$A[j]) || data$A[j] != 1)) {
        data$m2mcount[j] <- data$m2mcount[j - 1] + 1
        j <- j + 1
      }
      
      if (j <= nrow(data) &&
          data$id[j] == data$id[i] && !is.na(data$A[j]) && data$A[j] == 1) {
        data$m2mcount[j] <- data$m2mcount[j - 1] + 1
      }
    }
  }
  
  # Identify incomplete cycles: This identifies instances where a complete menses-to-menses cycle was not captured
  data <- data %>%
    mutate(cycle_incomplete = ifelse(!is.na(m2mcount) &
                                       (is.na(lead(m2mcount)) & id != lead(id)), 1, 0))
  
  # New condition: Set cycle_incomplete = 1 if m2mcount restarts when id changes
  data <- data %>%
    group_by(id) %>%
    mutate(
      cycle_incomplete = ifelse(
        id != lag(id, default = first(id)) & m2mcount == 1, 1, cycle_incomplete
      )
    ) %>%
    ungroup()
  
  # Propagate cycle_incomplete within each group of m2mcount
  data <- data %>%
    group_by(id) %>%
    mutate(cycle_group = cumsum(!is.na(m2mcount) & m2mcount == 1)) %>%
    group_by(id, cycle_group) %>%
    mutate(
      cycle_incomplete = ifelse(any(cycle_incomplete == 1), 1, 0),
      mcyclength = max(m2mcount, na.rm = TRUE)
    ) %>%
    ungroup() %>%
    select(-cycle_group)
  
  data$cycle_incomplete = ifelse(is.na(data$cycle_incomplete), 1, 0)
  data$cycle_incomplete = ifelse(is.na(data$m2mcount), NA, data$cycle_incomplete)
  
  # Calculate cyclenum: calculates number of menses-to-menses cycles within a person
  data <- data %>%
    group_by(id) %>%
    mutate(cyclenum = cumsum(!is.na(m2mcount) & m2mcount == 1 & cycle_incomplete == 0)) %>%
    ungroup()
  
  data <- data %>%
    group_by(id) %>%
    mutate(cyclenum = ifelse(
      cycle_incomplete == 1,
      NA,
      cumsum(!is.na(m2mcount) & m2mcount == 1 & cycle_incomplete == 0)
    )) %>%
    ungroup()
  
  return(data)
}



#calculate_lutdaycount: counts length of luteal phase 

calculate_lutdaycount <- function(data, ovtoday) {
  last_id <- NULL
  lutdaycount1 <- rep(NA, nrow(data))
  lut_incomplete1 <- rep(NA, nrow(data))  # Initialize lut_incomplete
  count_started_row <- NA  # Track the starting row of a lutdaycount1 stretch
  active_count <- FALSE  # Track if lutdaycount1 is currently counting

  for (i in 1:nrow(data)) {
    # If id changes, reset lutdaycount1 and lut_incomplete
    if (is.null(last_id) || last_id != data$id[i]) {
      if (active_count) {
        # Mark the whole stretch as incomplete if stopped by id change
        lut_incomplete1[count_started_row:(i - 1)] <- 1
      }
      lutdaycount1[i] <- ifelse(ovtoday[i] == 1, 0, NA)
      active_count <- !is.na(lutdaycount1[i])  # Start counting if lutdaycount1 is initialized
      count_started_row <- ifelse(!is.na(lutdaycount1[i]), i, NA)  # Record start row
    } else if (!is.na(lutdaycount1[i - 1])) {
      lutdaycount1[i] <- lutdaycount1[i - 1] + 1
    }

    # If lutdaycount1 is active and stops because of A == 1, set lut_incomplete = 0 for the entire stretch
    if (!is.na(lutdaycount1[i]) && !is.na(data$A[i]) && data$A[i] == 1) {
      lutdaycount1[i] <- NA
      lut_incomplete1[count_started_row:(i - 1)] <- 0  # Mark stretch as complete due to A == 1
      active_count <- FALSE  # Stop counting due to A == 1
    } else if (ovtoday[i] == 1) {
      lutdaycount1[i] <- 0
      active_count <- TRUE  # Start counting at ovtoday == 1
      count_started_row <- i  # Record start row of the count
    }

    # Assign the current row to last_id for the next iteration
    last_id <- data$id[i]
  }

  # Update the dataset with lutdaycount1 and lut_incomplete
  data$lutdaycount1 <- lutdaycount1
  data$lut_incomplete1 <- lut_incomplete1
  
  return(data)
}

#calculate_foldaycount: counts length of follicular phase 

calculate_foldaycount <- function(data, ovtoday) {
  foldaycount <- NA
  fol_incomplete <- NA
  last_id <- NULL
  active_count <- FALSE  # Track if foldaycount is currently active
  count_started_row <- NULL  # Track the starting row of a foldaycount stretch
  
  for (i in 1:nrow(data)) {
    # If id changes, reset foldaycount and fol_incomplete
    if (is.null(last_id) || is.na(data$id[i]) || last_id != data$id[i]) {
      if (active_count) {
        # Mark the whole stretch as incomplete if stopped by id change
        data$fol_incomplete[count_started_row:i-1] <- 1
      }
      foldaycount <- ifelse(data$A[i] == 1, 0, NA)
      active_count <- !is.na(foldaycount)  # Start counting if foldaycount is initialized
      fol_incomplete <- NA  # Reset fol_incomplete when id changes
      count_started_row <- ifelse(!is.na(foldaycount), i, NA)  # Record start row
    } else if (!is.na(foldaycount)) {
      foldaycount <- foldaycount + 1
    }

    # If foldaycount is active and stopped because of ovtoday == 1, set fol_incomplete = 0 for the whole stretch
    if (!is.na(foldaycount) && i >= 2 && !is.na(ovtoday[i]) && ovtoday[i - 1] == 1) {
      foldaycount <- NA
      data$fol_incomplete[count_started_row:i-1] <- 0  # Mark stretch as complete due to ovulation
      active_count <- FALSE  # Stop counting due to ovulation
      fol_incomplete <- 0
    }

    # If A == 1, start or reset the count
    if (data$A[i] == 1) {
      foldaycount <- 0
      active_count <- TRUE  # Start counting
      count_started_row <- i  # Record start row of the count
    }

    # If foldaycount is actively counting and stops because of id change
    if (!is.na(last_id) && !is.na(data$id[i]) && last_id != data$id[i] && active_count) {
      data$fol_incomplete[count_started_row:i-1] <- 1  # Mark as incomplete because the count stopped due to id change
      active_count <- FALSE  # Stop counting as the id changed
    }

    # Assign foldaycount and fol_incomplete to the current row
    data$foldaycount[i] <- foldaycount
    if (is.na(fol_incomplete)) {
      data$fol_incomplete[i] <- NA
    }

    # Update last_id for the next iteration
    last_id <- data$id[i]
  }

  return(data)
}

```


```{r}
#subset to only include baseline, washout, and PBO 
dat = dat %>% filter(cleartrialphase == 0| cleartrialphase == 1| cleartrialphase == 4| cleartrialphase == 5| cleartrialphase == 8|cleartrialphase == 9|cleartrialphase == 12)

#create the variable 'A' which will represent menses onset 
dat$A <- ifelse(dat$cycleday == 1, 1, 0)
dat$A <- ifelse(is.na(dat$A), 0, dat$A)
#ovtoday = LH + 1
dat$ovtoday = lag(dat$posovtest)
#check data 
dat %>% select(daterated, ovtoday, posovtest)

```

# Calculate Menses-to-menses cycle lengths 

## m2mcount and cyclelength 

```{r}
#the mcyclength function, this will do forward count to give cycle lengths for menses-to-menses cycles, and count the number of complete cycles within a person 
dat <- calculate_mcyclength(dat)
#-Inf will occur if the cycle is incomplete (no bookend menses onset date) so we will change to NA
dat$mcyclength = ifelse(dat$mcyclength == -Inf , NA, dat$mcyclength)
#if cyclenum is NA, set cycle_incomplete to 1 
dat$cycle_incomplete = ifelse(is.na(dat$cyclenum), 1, dat$cycle_incomplete)
#check data
dat %>% select(id, daterated, A, m2mcount, ovtoday, cyclenum, mcyclength, cycle_incomplete) 
```



Here, I am doing scaling for baseline, washout, and PBO condition

# calculating ovtoday_impute
 
```{r}
# this is based on NC dataset 
dat <- dat %>% 
  mutate(lutlength_impute = case_when(mcyclength == 20 ~ (mcyclength* 0.46),
                                      mcyclength == 21 ~ (mcyclength*.476),
                                      mcyclength == 22 ~ (mcyclength*.491),
                                      mcyclength == 23 ~ (mcyclength*.491),
                                      mcyclength == 24 ~ (mcyclength*.492),
                                      mcyclength == 25 ~ (mcyclength*.484),
                                      mcyclength == 26 ~ (mcyclength*.481),
                                      mcyclength == 27 ~ (mcyclength*.470),
                                      mcyclength == 28 ~ (mcyclength*.461),
                                      mcyclength == 29 ~ (mcyclength*.448),
                                      mcyclength == 30 ~ (mcyclength*.437),
                                      mcyclength == 31 ~ (mcyclength*.426),
                                      mcyclength == 32 ~ (mcyclength*.416),
                                      mcyclength == 33 ~ (mcyclength*.40),
                                      mcyclength == 34 ~ (mcyclength*.391),
                                      mcyclength == 35 ~ (mcyclength*.377),
                                      mcyclength == 36 ~ (mcyclength*.369),
                                      mcyclength == 37 ~ (mcyclength*.359),
                                      TRUE ~ NA))

dat <- dat %>% 
  mutate(follength_impute = case_when(mcyclength == 20 ~ (mcyclength*.54),
                                      mcyclength == 21 ~ (mcyclength*.524),
                                      mcyclength == 22 ~ (mcyclength*.509),
                                      mcyclength == 23 ~ (mcyclength*.509),
                                      mcyclength == 24 ~ (mcyclength*.508),
                                      mcyclength == 25 ~ (mcyclength*.516),
                                      mcyclength == 26 ~ (mcyclength*.519),
                                      mcyclength == 27 ~ (mcyclength*.530),
                                      mcyclength == 28 ~ (mcyclength*.539),
                                      mcyclength == 29 ~ (mcyclength*.552),
                                      mcyclength == 30 ~ (mcyclength*.563),
                                      mcyclength == 31 ~ (mcyclength*.574),
                                      mcyclength == 32 ~ (mcyclength*.584),
                                      mcyclength == 33 ~ (mcyclength*.60),
                                      mcyclength == 34 ~ (mcyclength*.609),
                                      mcyclength == 35 ~ (mcyclength*.623),
                                      mcyclength == 36 ~ (mcyclength*.631),
                                      mcyclength == 37 ~ (mcyclength*.641),
                                      TRUE ~ NA))
#check data
dat %>% select(id, daterated, A, m2mcount, mcyclength, follength_impute)
#group by id, and make sure dates are in order 
dat <- dat %>%
  group_by(id) %>%
  arrange(daterated, .by_group = TRUE) 
#initialize new variable- follcount1
dat$follcount1 <- NA
for (i in 1:nrow(dat)) {
  if (dat$A[i] == 1 & !is.na(dat$follength_impute[i])) {
    follcount1 <- seq_len(round(dat$follength_impute[i]))
    dat$follcount1[i:(i + length(follcount1) - 1)] <- follcount1
  }
}
#follcount1 forward counts from menses onset based on imputed follength from NC data 

dat %>% select(id, A, follength_impute, follcount1)

#identify ovtoday_impute based on forward count from follcount1
dat <- dat %>% 
  mutate(ovtoday_impute = case_when(round(follength_impute) == follcount1 ~ 1,
                               TRUE ~ NA))

#if ovtoday is NA, change to 0 
dat$ovtoday <- ifelse(is.na(dat$ovtoday), 0, dat$ovtoday) 

#if ovtoday_impute is NA, change to 0 
dat$ovtoday_impute <- ifelse(is.na(dat$ovtoday_impute), 0, dat$ovtoday_impute) 

#check data 
dat %>% select(id, daterated, cleartrialphase, ovtoday, ovtoday_impute)
```


# percentage of luteal phase using ovtoday (LH + 1) 

 - variable is called perclut for menses-centered
 - variable is called perclut_ov for ovulation-centered 
 

```{r}
#group by id, make sure daterated is in order, initialize the variable lutmax 
dat <- dat  %>% 
  group_by(id) %>%
  arrange(daterated, .by_group = TRUE) %>%
  mutate(lutmax = NA)
#apply the calculate_lutdaycount function. which spits out the variable lutdaycount1. lutdaycount1 forward counts from ovtoday = 1 to the next menses onset, indexed at 0 
dat <- calculate_lutdaycount(dat, dat$ovtoday)
#check data 
dat %>% select(id, daterated, ovtoday, A, lutdaycount1)
#the below code creates lutdaycount which shifts lutdaycount1, and forward counts starting the day after ovtoday = 1, still indexed at 0 
dat <- dat %>%
  group_by(id) %>%
  mutate(
    lutdaycount = lag(lutdaycount1),
    lutdaycount = case_when(
      is.na(lutdaycount) | id != lag(id) ~ NA, 
      TRUE ~ lutdaycount
    )
  )

dat <- dat %>%
  group_by(id) %>%
  mutate(
    lut_incomplete = lag(lut_incomplete1),
    lut_incomplete = case_when(
      is.na(lut_incomplete) | id != lag(id) ~ NA, 
      TRUE ~ lut_incomplete
    )
  )


#check data 
dat %>% select(id, daterated, ovtoday, A, lutdaycount, lut_incomplete)
#lutmax corresponds to the length of the luteal phase in a complete menses-to-menses cycle (it is the max value of lutdaycount in a cycle)
for (i in 1:(nrow(dat)-1)) {
  if (is.na(dat$lutdaycount[i + 1] && !is.na(dat$lutdaycount[i]))) {
    dat$lutmax[(i-(dat$lutdaycount[i])):i] = as.numeric(dat$lutdaycount[i])
  }
}

dat$lut_incomplete = ifelse(dat$lutmax > 30, 1, dat$lut_incomplete)
#check data 
dat %>% select(id, daterated, ovtoday, A, lutdaycount, lutmax, lut_incomplete)

#calculates lutperc which is luteal phase scaled from 0 to 1, using ovtoday 
dat <- dat %>%
  mutate(lutperc = ifelse(lut_incomplete == 0, (lutdaycount / lutmax), NA))

#calculates perclut_base which is luteal phase scaled from -1 to 0, using ovtoday. This is the variable that will be used in the menses-centered scaled_cycleday 
dat$perclut = dat$lutperc -1 

# dat <- dat %>%
#   mutate(perclut = ifelse(A == 1, 0, perclut))

#check data
dat %>% select(id, daterated, ovtoday, A, lutmax, perclut, lut_incomplete)

#calculate lutdaycount_ov, which is lutdaycount but ovulation-centered 
dat <- dat %>%
  group_by(id) %>%
  mutate(
    lutdaycount_ov = lead(lutdaycount),
    lutdaycount_ov = case_when(is.na(lutdaycount_ov) |
                                 id != lead(id) ~ NA, TRUE ~ lutdaycount_ov)
  )

dat <- dat %>%
  mutate(perclut_ov = ifelse(lut_incomplete == 0, lutdaycount_ov / lutmax, NA))

# dat <- dat %>%
#   mutate(perclut_ov = ifelse(lutdaycount_ov == 0, 0, perclut_ov))


dat %>% select(id,
               daterated,
               ovtoday,
               A,
               lutdaycount,
               lutdaycount_ov,
               lutmax,
               perclut_ov, perclut)

#remove 'helper' variables
dat <- dat %>% select(-c(lutdaycount1, lutperc, lut_incomplete1))

```

# percentage of luteal phase using ovtoday_impute (based on NC norms)


```{r}
#make sure dates are in order for each id, initialize new variable called lutmax_impute
dat <- dat  %>% 
  group_by(id) %>%
  arrange(daterated, .by_group = TRUE) %>%
  mutate(lutmax_impute = NA)
#if A is NA, make it 0 (A is menses onset)
dat$A <- ifelse(is.na(dat$A), 0, dat$A)
#apply calculate_lutdaycount fxn which spits out the variable lutdaycount1. lutdaycount1 forward counts from ovtoday_impute = 1 to the next menses onset, indexed at 0 
dat <- calculate_lutdaycount(dat, dat$ovtoday_impute)
#check data
dat %>% select(id, daterated, ovtoday_impute, A, lutdaycount1)
#the below code creates lutdaycount_impute which shifts lutdaycount1, and forward counts starting the day after ovtoday_impute = 1, still indexed at 0 
dat <- dat %>%
  group_by(id) %>%
  mutate(
    lutdaycount_impute = lag(lutdaycount1),
    lutdaycount_impute = case_when(
      is.na(lutdaycount_impute) | id != lag(id) ~ NA, 
      TRUE ~ lutdaycount_impute
    )
  )
#check data 
dat %>% select(id, daterated, ovtoday_impute, A, lutdaycount_impute)

#lutmax_impute corresponds to the length of the luteal phase in a complete menses-to-menses cycle, using ovtoday_impute (it is the max value of lutdaycount_impute in a cycle)
for (i in 1:(nrow(dat)-1)) {
  if (is.na(dat$lutdaycount_impute[i + 1] && !is.na(dat$lutdaycount_impute[i]))) {
    dat$lutmax_impute[(i-(dat$lutdaycount_impute[i])):i] = as.numeric(dat$lutdaycount_impute[i])
  }
}
#lutlength1_impute is the imputed value of luteal phase based on NC norms, but lagged one day so it does align with any dates in the follicuar phase 
dat$lutlength1_impute = lag(dat$lutlength_impute)

#check data
dat %>% select(id, daterated, ovtoday_impute, A, lutdaycount_impute, lutmax_impute, lutlength_impute, lutlength1_impute)

#calculate lutperc_impute
dat <- dat %>%
  mutate(lutperc_impute = if_else(
    is.na(lutlength1_impute),
    NA,
    if_else(
      is.na(lutlength1_impute) &
        !is.na(lutmax_impute) & cycle_incomplete == 0,
      lutdaycount_impute / lutmax_impute,
      lutdaycount_impute / round(lutlength1_impute)
    )
  ))

#lutperc is scaled from 0 to 1, so substracting 1 so that it is scaled from -1 to 0 for menses-centered scaled_cycleday 
dat$perclut_impute = dat$lutperc_impute -1 

#check data
dat %>% select(id, daterated, ovtoday_impute, A, mcyclength, perclut_impute, perclut) 

#calculate lutdaycount_imp_ov, which is imputed lutdaycount but ovulation-centered 
dat <- dat %>%
  group_by(id) %>%
  mutate(
    lutdaycount_imp_ov = lead(lutdaycount_impute),
    lutdaycount_imp_ov = case_when(
      is.na(lutdaycount_imp_ov) | id != lead(id) ~ NA, 
      TRUE ~ lutdaycount_imp_ov
    )
  )

#calculate lutperc_imp_ov which is the variable that will go into ovulation-centered scaled cycleday 
dat <- dat %>%
  mutate(perclut_imp_ov = if_else(
    is.na(lutlength1_impute) ,
    NA,
    if_else(
      is.na(lutlength1_impute) & !is.na(lutmax) & cycle_incomplete == 0,
      lutdaycount1 / lutmax_impute,
      lutdaycount1 / round(lutlength1_impute)
    )
  ))


#if ovtoday_impute == 1, set lutperc_imp_ov to 0
# dat <- dat %>%
#   mutate(perclut_imp_ov = ifelse(ovtoday_impute == 1, 0, perclut_imp_ov))

#check data 
dat %>% select(id, daterated, ovtoday_impute, A, lutdaycount, lutlength1_impute, lutdaycount1, lutmax_impute, perclut_impute, perclut_imp_ov)

#remove helper variables
dat <- dat %>% select (-c(lutperc_impute, lutdaycount1))
```

# percentage of follicular phase using ovtoday (LH+1)

```{r}

#make sure daterated is in order by id, and initialize folmax variable 
dat <- dat %>%
  group_by(id) %>%
  arrange(daterated, .by_group = TRUE) %>%
  mutate(folmax = NA)
#apply calculate foldaycount function using ovtoday 
dat <- calculate_foldaycount(dat, dat$ovtoday)
#foldaycount forward counts starting at menses onset and ends at ovtoday 
dat %>% select(id, daterated, A, ovtoday, foldaycount)

#assigns folmax as the max value of foldaycount within a cycle 
for (i in 1:(nrow(dat)-1)) {
  if (is.na(dat$foldaycount[i + 1] && !is.na(dat$foldaycount[i]))) {
    dat$folmax[(i-(dat$foldaycount[i])):i] = as.numeric(dat$foldaycount[i])
  }
}

#follength is the length of the follicular phase that cycle (since folmax is indexed at 0)
dat$follength = dat$folmax + 1 

#check data
dat %>% select(id, daterated, ovtoday, A, foldaycount, folmax, follength, fol_incomplete)

dat <- dat %>%
  mutate(percfol = ifelse(fol_incomplete != 1, foldaycount / (folmax), NA))

dat$percfol_ov = dat$percfol - 1

dat %>% select(id, daterated, ovtoday, A, foldaycount, folmax, percfol, percfol_ov)

#renaming foldaycount based on ovtoday to be called foldaycount_ov. the variable name 'foldaycount' get overwritten in the code chunk below when the calculate_foldaycount() function is run 
dat <- dat %>% rename(foldaycount_ov = foldaycount)

```

# percentage of follicular phase using ovtoday_impute (based on NC norms)

```{r}
#yes, I'm constantly arranging by daterated because I'm paranoid. Initialize the variable folmax_impute 
dat <- dat %>%
  group_by(id) %>%
  arrange(daterated, .by_group = TRUE) %>%
  mutate(folmax_impute = NA)
#see comment on line 512
dat <- calculate_foldaycount(dat, dat$ovtoday_impute)
dat <- dat %>% rename(foldaycount_impute = foldaycount)
dat %>% select(id, daterated, ovtoday_impute, A, foldaycount_impute)

#create folmax_impute which is the max value of foldaycount_impute and matches the imputed follicular length in complete cycles 
for (i in 1:(nrow(dat)-1)) {
  if (is.na(dat$foldaycount_impute[i + 1] && !is.na(dat$foldaycount_impute[i]))) {
    dat$folmax_impute[(i-(dat$foldaycount_impute[i])):i] = as.numeric(dat$foldaycount_impute[i])
  }
}

dat <- dat %>%
  mutate(percfol_impute = ifelse(cycle_incomplete == 0, foldaycount_impute / (folmax_impute), NA))

# ovulation-centered percfol based on ovtoday_impute 
dat$percfol_imp_ov = dat$percfol_impute - 1


dat %>% select(id, daterated, ovtoday_impute, A, percfol_impute, percfol_imp_ov)
```




# prioritize LH test-based scaled values over imputed values 

```{r}
#functions to prioritize perc_base over perc

create_percent <- function(dat, percentvar, percvar, percvar_impute) {
  percentvar <- ensym(percentvar)
  percvar <- ensym(percvar)
  percvar_impute <- ensym(percvar_impute)
  
  dat %>%
    group_by(id, cyclenum) %>%
    mutate(
      !!percentvar := if (all(is.na(!!percvar) | !!percvar == 0)) {
        # if percvar contains only NA or 0, use percvar_impute
        !!percvar_impute
      } else {
        # otherwise, use percvar
        !!percvar
      }
    ) %>%
    ungroup()
}


dat = create_percent(dat, percentlut, perclut, perclut_impute)
dat = create_percent(dat, percentfol, percfol, percfol_impute)
dat = create_percent(dat, percentlut_ov, perclut_ov, perclut_imp_ov)
dat = create_percent(dat, percentfol_ov, percfol_ov, percfol_imp_ov)

dat %>% select(id, daterated, A, ovtoday, ovtoday_impute, cyclenum, mcyclength, percfol, percfol_impute, percentlut_ov)
```


# creating scaled_cycleday 

```{r}
dat <- dat %>%
  mutate(scaled_cycleday = ifelse(is.na(percentlut), percentfol, percentlut))
dat %>% select(id, daterated, percentlut, percentfol, scaled_cycleday)
dat <- dat %>%
  mutate(scaled_cycleday_ov = ifelse(is.na(percentlut_ov), percentfol_ov, percentlut_ov))
dat %>% select(id, daterated, percentlut_ov, percentfol_ov, scaled_cycleday_ov)

```

# how many non-missing: 

```{r}
dat %>%
  summarise(complete_cases = sum(!is.na(scaled_cycleday) & (!is.na("DRSP7_angirr"))))
```





# write file to Box

```{r}
dat %>% select(id, daterated, ovtoday_impute, scaled_cycleday, scaled_cycleday_ov, mcyclength) 


saveRDS(dat, "C:/Users/Student/Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CLEAR2/02_datasets/CLEAR2_DAILY/03_cleaned_data/clear2_baseline_washout_scaledcycleday.rds")
```



# plotting 

```{r}

dat <- dat %>%
  mutate(cycleday_perc = (scaled_cycleday + 1) / 2) %>%
  mutate(cycleday_perc_round = round(cycleday_perc, 1)) %>%
  mutate(cycleday_5perc = round(cycleday_perc / 0.05) * 0.05) %>%
  mutate(cycleday_2perc = round(cycleday_perc / 0.02) * 0.02)


df <- dat %>%
  group_by(id) %>%
  mutate(DRSP7_c =  DRSP7_angirr - mean(DRSP7_angirr, na.rm = T))



dvDRSP7 <- df %>% 
  group_by(cycleday_5perc) %>% 
  summarise(mean_dev = mean(DRSP7_c, na.rm = T))

 ggplot(dvDRSP7, aes(x = cycleday_5perc, y = mean_dev)) +
  geom_point(size = .2) +
  geom_line(size=.2) +
  geom_smooth(se=T, size = 0.3) 

```






