---
title: "prepping scaled cycleday"
author: "Anisha Nagpal"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(pdftools)
library(magick)

#importing data
dat <- readRDS("C:/Users/Student/Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/REVIEW_cycleday_modeling/02_analytic_workspace/07_16_24_bearsdailydaterated.rds")

```


# Manually adjusting LH confirm 

Here I am manually changing the date of LH_confirm to reflect peak LH surge. If LH pos test did not align with LH ZRTmax, we adjusted the date of LH_confirm to what would most appropriately reflect LH surge in the context of all of their hormone trajectories. This is documented in a spreadsheet as well. 

```{r}
dat$LH_confirm = dat$LHposdayfirst

#LH_confirms to manually change: 
dat %>% filter(ID == 107) %>% select(ID, daterated, LHposdayfirst, LH_confirm)

dat <- dat %>% 
  mutate(LH_confirm = case_when(
    ID == 107 & daterated == "2022-11-27" ~ 0,
    ID == 107 & daterated == "2022-11-28" ~ 1,
    TRUE ~ LH_confirm 
  ))
dat %>% filter(ID == 107) %>% select(ID, daterated, LHposdayfirst, LH_confirm)


#change 11/23/22 to 11/24/22 
dat <- dat %>% 
  mutate(LH_confirm = case_when(
    ID == 109 & daterated == "2022-11-23" ~ 0,
    ID == 109 & daterated == "2022-11-24" ~ 1,
    TRUE ~ LH_confirm 
  ))
dat %>% filter(ID == 109) %>% select(ID, daterated, LHposdayfirst, LH_confirm)

#move 2/20/23 to 2/18/23 
dat <- dat %>% 
  mutate(LH_confirm = case_when(
    ID == 115 & daterated == "2023-02-20" ~ 0,
    ID == 115 & daterated == "2023-02-18" ~ 1,
    TRUE ~ LH_confirm 
  ))
dat %>% filter(ID == 115) %>% select(ID, daterated, LHposdayfirst, LH_confirm)

#move 3/10/23 to 3/9/23
dat %>% filter(ID == 118) %>% select(ID, daterated, LHposdayfirst, LH_confirm)
dat <- dat %>% 
  mutate(LH_confirm = case_when(
    ID == 118 & daterated == "2023-03-10" ~ 0,
    ID == 118 & daterated == "2023-03-09" ~ 1,
    TRUE ~ LH_confirm 
  ))


#move 6/10/23 to 6/12/23
dat %>% filter(ID == 132) %>% select(ID, daterated, LHposdayfirst, LH_confirm)
dat <- dat %>% 
  mutate(LH_confirm = case_when(
    ID == 132 & daterated == "2023-06-10" ~ 0,
    ID == 132 & daterated == "2023-06-12" ~ 1,
    TRUE ~ LH_confirm 
  ))
dat %>% filter(ID == 132) %>% select(ID, daterated, LHposdayfirst, LH_confirm)


#move 6/3/23 to 6/4/23
dat %>% filter(ID == 137) %>% select(ID, daterated, LHposdayfirst, LH_confirm)
dat <- dat %>% 
  mutate(LH_confirm = case_when(
    ID == 137 & daterated == "2023-06-03" ~ 0,
    ID == 137 & daterated == "2023-06-04" ~ 1,
    TRUE ~ LH_confirm 
  ))
dat %>% filter(ID == 137) %>% select(ID, daterated, LHposdayfirst, LH_confirm)

#move 6/4/23 to 6/6/23
dat %>% filter(ID == 142) %>% select(ID, daterated, LHposdayfirst, LH_confirm)
dat <- dat %>% 
  mutate(LH_confirm = case_when(
    ID == 142 & daterated == "2023-06-04" ~ 0,
    ID == 142 & daterated == "2023-06-06" ~ 1,
    TRUE ~ LH_confirm 
  ))
dat %>% filter(ID == 142) %>% select(ID, daterated, LHposdayfirst, LH_confirm)


#move 8/7/23 to 8/8/23 
dat <- dat %>% 
  mutate(LH_confirm = case_when(
    ID == 143 & daterated == "2023-08-07" ~ 0,
    ID == 143 & daterated == "2023-08-08" ~ 1,
    TRUE ~ LH_confirm 
  ))
dat %>% filter(ID == 143) %>% select(ID, daterated, LHposdayfirst, LH_confirm)

#move 11/4/23 to 11/5/23
dat %>% filter(ID == 157) %>% select(ID, daterated, LHposdayfirst, LH_confirm)
dat <- dat %>% 
  mutate(LH_confirm = case_when(
    ID == 157 & daterated == "2023-11-04" ~ 0,
    ID == 157 & daterated == "2023-11-05" ~ 1,
    TRUE ~ LH_confirm 
  ))

#move 10/31/23 to 11/03/23
dat %>% filter(ID == 158) %>% select(ID, daterated, LHposdayfirst, LH_confirm)
dat <- dat %>% 
  mutate(LH_confirm = case_when(
    ID == 158 & daterated == "2023-10-31" ~ 0,
    ID == 158 & daterated == "2023-11-03" ~ 1,
    TRUE ~ LH_confirm 
  ))

#move 10/1/23 to 9/30/23
dat <- dat %>% 
  mutate(LH_confirm = case_when(
    ID == 161 & daterated == "2023-10-01" ~ 0,
    ID == 161 & daterated == "2023-09-30" ~ 1,
    TRUE ~ LH_confirm 
  ))
dat %>% filter(ID == 161) %>% select(ID, daterated, LHposdayfirst, LH_confirm)

#move 1/8/24 to 1/6/24
dat %>% filter(ID == 175) %>% select(ID, daterated, LHposdayfirst, LH_confirm)
dat <- dat %>% 
  mutate(LH_confirm = case_when(
    ID == 175 & daterated == "2024-01-08" ~ 0,
    ID == 175 & daterated == "2024-01-06" ~ 1,
    TRUE ~ LH_confirm 
  ))
dat %>% filter(ID == 175) %>% select(ID, daterated, LHposdayfirst, LH_confirm)

#move 2/17/24 to 2/18/24
dat %>% filter(ID == 178) %>% select(ID, daterated, LHposdayfirst, LH_confirm)
dat <- dat %>% 
  mutate(LH_confirm = case_when(
    ID == 178 & daterated == "2024-02-17" ~ 0,
    ID == 178 & daterated == "2024-02-18" ~ 1,
    TRUE ~ LH_confirm 
  ))

#move 2/19/24 to 2/18/24
dat %>% filter(ID == 184) %>% select(ID, daterated, LHposdayfirst, LH_confirm)
dat <- dat %>% 
  mutate(LH_confirm = case_when(
    ID == 184 & daterated == "2024-02-19" ~ 0,
    ID == 184 & daterated == "2024-02-18" ~ 1,
    TRUE ~ LH_confirm 
  ))

dat %>% filter(ID == 184) %>% select(ID, daterated, LHposdayfirst, LH_confirm)
dat <- dat %>% 
  mutate(LH_confirm = case_when(
    ID == 184 & daterated == "2024-03-16" ~ 0,
    ID == 184 & daterated == "2024-03-15" ~ 1,
    TRUE ~ LH_confirm 
  ))
dat %>% filter(ID == 184) %>% select(ID, daterated, LHposdayfirst, LH_confirm)


#fix 182

dat %>% filter(ID == 182) %>% select(ID, daterated, LHposdayfirst, LH_confirm)
dat <- dat %>% 
  mutate(LH_confirm = case_when(
    ID == 182 & daterated == "2024-01-30" ~ 0,
    ID == 182 & daterated == "2024-02-17" ~ 0,
    ID == 182 & daterated == "2024-03-01" ~ 1,
    TRUE ~ LH_confirm 
  ))
# creating ov_est (a bookend of the scaled_cycleday variable), to be LH + 1 
dat <- dat %>% arrange(ID, daterated) %>%
  group_by(ID) %>%
  mutate(ov_est = lag(LH_confirm))

dat %>% select(ID, daterated, LH_confirm, ov_est)
```

# Manually adding in menses onset dates 

On 8/22/24 Yunshu identified the following menses onset dates that were not available in the original dataset 

143
- 8/21/23

135
- 5/13/23


```{r}

dat <- dat %>% 
  mutate(mensdayonefirst = case_when(
    ID == 143 & daterated == "2023-08-21" ~ 1,
    ID == 135 & daterated == "2023-05-13" ~ 1,
    TRUE ~ mensdayonefirst 
  ))
```


# Scaled cycleday Time Coding

Follicular phase: the length from the 1st day of menses up to and including the estimated  ovulation, that is, from the 1st day of the cycle up to and including the 2nd peak day/nadir day (if using BBT). Ovulation is the two days of the peak and the following day or the nadir and the following day. 

Luteal phase: 1st day after the estimated day of ovulation up to and including the day before the next (menses) menstrual cycle.

## menses to menses cycle length 

```{r}
calculate_mcyclength <- function(data) {
  # Calculate m2mcount: This counts from 1 menses onset to the next to give you the length of a menses-to-menses cycle
  data <- data %>%
    group_by(ID) %>%
    arrange(daterated, .by_group = TRUE) %>%
    mutate(
      m2mcount = NA,
      mcyclength = NA,
      cycle_incomplete = 0
    )
  
  for (i in seq_len(nrow(data))) {
    if (!is.na(data$A[i]) && data$A[i] == 1) {
      data$m2mcount[i] <- 1
      j <- i + 1
      
      while (j <= nrow(data) &&
             data$ID[j] == data$ID[i] && (is.na(data$A[j]) || data$A[j] != 1)) {
        data$m2mcount[j] <- data$m2mcount[j - 1] + 1
        j <- j + 1
      }
      
      if (j <= nrow(data) &&
          data$ID[j] == data$ID[i] && !is.na(data$A[j]) && data$A[j] == 1) {
        data$m2mcount[j] <- data$m2mcount[j - 1] + 1
      }
    }
  }
  
  # Identify incomplete cycles: This identifies instances where a complete menses-to-menses cycle was not captured
  data <- data %>%
    mutate(cycle_incomplete = ifelse(!is.na(m2mcount) &
                                       (is.na(lead(
                                         m2mcount
                                       )) & ID != lead(ID)), 1, 0))
  
  ##  Propagate cycle_incomplete within each group of m2mcount
  data <- data %>%
    group_by(ID) %>%
    mutate(cycle_group = cumsum(!is.na(m2mcount) & m2mcount == 1)) %>%
    group_by(ID, cycle_group) %>%
    mutate(
      cycle_incomplete = ifelse(any(cycle_incomplete == 1), 1, 0),
      mcyclength = max(m2mcount, na.rm = TRUE)
    ) %>%
    ungroup() %>%
    select(-cycle_group)
  data$cycle_incomplete = ifelse(is.na(data$cycle_incomplete), 1, 0)
  data$cycle_incomplete = ifelse(is.na(data$m2mcount), NA, data$cycle_incomplete)
  # calculate cyclenum: calculates number of menses-to-menses cycles within a person 
  data <- data %>%
    group_by(ID) %>%
    mutate(cyclenum = cumsum(!is.na(m2mcount) &
                               m2mcount == 1 & cycle_incomplete == 0)) %>%
    ungroup()
  
  data <- data %>%
    group_by(ID) %>%
    mutate(cyclenum = ifelse(
      cycle_incomplete == 1,
      NA,
      cumsum(!is.na(m2mcount) &
               m2mcount == 1 & cycle_incomplete == 0)
    )) %>%
    ungroup()
  
  return(data)
  
}

#make sure daterated is date-type 
dat$daterated = as.Date(dat$daterated)
#fill in missing rows 
dat <- dat %>% group_by(ID) %>%
  complete(daterated = seq.Date(min(daterated), max(daterated), by = "day"))
#assign 'A' as the menses onset 
dat$A = dat$mensdayonefirst
#if A is NA, set A = to 0 
dat$A <- ifelse(is.na(dat$A), 0, dat$A)
#remove any duplicates (e.g. removing empty rows)
dat = unique(dat)
#taking a look at ID = 160
dat %>% filter(ID == 160 & A ==1) %>% select(ID, daterated, A)
#apply calculate_mcyclength to dat 
dat <- calculate_mcyclength(dat)
#dat <- dat[!(dat$mcyclength < 21 | dat$mcyclength > 35), ]
#if m2mcount = NA (did not start counting because menses onset did not occur), mcyclength = -Inf. This turns -Inf to NA
dat$mcyclength = ifelse(dat$mcyclength == -Inf , NA, dat$mcyclength)
#take a look at the outcomes of the function
dat %>% select(ID, daterated, A, m2mcount, mcyclength, cycle_incomplete, cyclenum)
dat %>% select(ID, daterated, A, m2mcount, mcyclength, cycle_incomplete, cyclenum) %>% filter(ID == 143)
```

## examine menses-to-menses cycle lengths 

```{r}
#create dataframe of cycle-lengths 
cyclength = dat %>% select(ID, mcyclength, cyclenum)
cyclength = unique(cyclength)
cyclength = cyclength %>% filter(!is.na(cyclenum))
cyclength = cyclength %>% filter(mcyclength < 42) #some cycles are very long, probably missing a menses onset 

ggplot(cyclength, aes(x = mcyclength)) +
  geom_histogram(binwidth = 1, fill = "green", color = "darkgreen") +
  labs(title = "Distribution of Cycle Length", x = "Menses-to-menses cycle length (days)", y = "Frequency") +
  theme_light() +
  theme(axis.title = element_text(size = 14), axis.text = element_text(size = 12))

summary(cyclength$mcyclength)
sd(cyclength$mcyclength)
```

## Luteal Phase

```{r}
dat$ov_est <- ifelse(is.na(dat$ov_est), 0, dat$ov_est) #if ov_est (estimated ovulation) = NA, set to 0

#create lutmax variable = length of the luteal phase
dat <- dat %>%
  group_by(ID) %>%
  arrange(daterated, .by_group = TRUE) %>%
  mutate(lutmax = NA)

#lutdaycount function (counts the length of the luteal phase)
#lutday count starts counting 1 day after ov_est and starts at 0 but ends the day of menses onset. Since it starts at 0, the max value of lutdaycount is = to the length of the luteal phase and the day of ov_est and the day of menses onset is not included in the luteal length counting

calculate_lutdaycount <- function(data, ovtoday) {
  last_id <- NULL
  lutdaycount1 <- rep(NA, nrow(data))
  
  for (i in 1:nrow(data)) {
    if (is.null(last_id) || last_id != data$ID[i]) {
      # Restart counting when id changes
      lutdaycount1[i] <- ifelse(ovtoday[i] == 1, 0, NA)
    } else if (!is.na(lutdaycount1[i - 1])) {
      lutdaycount1[i] <- lutdaycount1[i - 1] + 1
    }
    
    if (!is.na(lutdaycount1[i]) &&
        !is.na(data$A[i]) && data$A[i] == 1) {
      lutdaycount1[i] <- NA
    } else if (ovtoday[i] == 1) {
      lutdaycount1[i] <- 0
    }
    
    last_id <- data$ID[i]
  }
  
  data$lutdaycount1 <- lutdaycount1
  return(data)
}

dat <- calculate_lutdaycount(dat, dat$ov_est)
dat %>% select(ID, daterated, ov_est, A, lutdaycount1)

dat <- dat %>%
  group_by(ID) %>%
  mutate(
    lutdaycount = lag(lutdaycount1),
    lutdaycount = case_when(is.na(lutdaycount) |
                              ID != lag(ID) ~ NA, TRUE ~ lutdaycount)
  )

dat %>% select(ID, daterated, ov_est, A, lutdaycount)

#calculate luteal length from count -> lutmax
for (i in 1:(nrow(dat) - 1)) {
  if (is.na(dat$lutdaycount[i + 1] && !is.na(dat$lutdaycount[i]))) {
    dat$lutmax[(i - (dat$lutdaycount[i])):i] = as.numeric(dat$lutdaycount[i])
  }
}

dat %>% select(ID, daterated, LH_confirm, ov_est, A, lutdaycount, lutmax)

#check what luteal phase lengths we have
dat %>% select(ID, lutmax) %>% unique()
#check 115, 137 - no menses onset data for these IDs

dat %>% filter(ID == 135) %>% select(ID, daterated, ov_est, mensdayonefirst, A, lutdaycount, lutmax)

#only calculate lutperc when luteal length is <= 18 and >= 7 (Based on 2SD below and above means of luteal phase lengths in 21-24 day cycle and 31-35 day cycle)

dat <- dat %>%
  mutate(lutperc = ifelse(lutmax <= 18 &
                            lutmax >= 7, lutdaycount / lutmax, NA))
#check data
dat %>% select(ID, daterated, ov_est, A, lutdaycount, lutperc)
#lutperc1 is the variable that will be included in the scaled_cycleday. It equals -1 the day after ov_est and it equals 0 on the day of menses onset
dat$lutperc1 = dat$lutperc - 1
dat %>% select(ID, daterated, ov_est, A, lutdaycount, lutperc, lutperc1)

dat <- dat %>%
  group_by(ID) %>%
  mutate(
    lutdaycount_ov = lead(lutdaycount),
    lutdaycount_ov = case_when(is.na(lutdaycount_ov) |
                                 ID != lead(ID) ~ NA, TRUE ~ lutdaycount_ov)
  )


#calculating lutperc_ov. lutperc_ov will be included in the ovulation-centered scaled_cycleday variable
dat <- dat %>%
  mutate(lutperc_ov = ifelse(lutmax <= 18 &
                               lutmax >= 7, lutdaycount_ov / lutmax, NA))
dat <- dat %>%
  mutate(lutperc_ov = ifelse(lutdaycount_ov == 0, 0, lutperc_ov))
dat %>% select(ID,
               daterated,
               ov_est,
               A,
               lutdaycount,
               lutdaycount_ov,
               lutmax,
               lutperc_ov)
dat %>% select(ID, daterated, ov_est, A, lutperc1, lutperc_ov)
```
## examine luteal lengths 

```{r}

lutlength = dat %>% select(ID, mcyclength, cyclenum, lutmax)
lutlength = unique(lutlength)
lutlength = lutlength %>% filter(!is.na(cyclenum))
lutlength = lutlength %>% filter(mcyclength < 42)

ggplot(lutlength, aes(x = lutmax)) +
  geom_histogram(binwidth = 1, fill = "green", color = "darkgreen") +
  labs(title = "Distribution of Luteal Lengths", x = "Luteal Phase length (days)", y = "Frequency") +
  theme_light() +
  theme(axis.title = element_text(size = 14), axis.text = element_text(size = 12))

summary(lutlength$lutmax)
sd(lutlength$lutmax, na.rm = T)

```


## Follicular Phase 

folperc = % of follicular phase elapsed 

```{r}
#group by ID and sort by daterated. Also initialize folmax 
dat <- dat  %>%
  group_by(ID) %>%
  arrange(daterated, .by_group = TRUE) %>%
  mutate(folmax = NA)
#function to calculate foldaycount
## foldaycount strats counting when menses onset = 1, (starting at 0), and ends the day of ov_est 
calculate_foldaycount <- function(data, ovtoday) {
  foldaycount <- NA
  last_id <- NULL
  
  for (i in 1:nrow(data)) {
    if (is.null(last_id) || last_id != data$ID[i]) {
      # Restart counting when id changes
      foldaycount <- ifelse(data$A[i] == 1, 0, NA)
    } else if (!is.na(foldaycount)) {
      foldaycount <- foldaycount + 1
    }
    
    if (!is.na(foldaycount) &&
        i >= 3 && !is.na(ovtoday[i]) && ovtoday[i - 1] == 1) {
      # Stop counting one row after ovtoday == 1
      foldaycount <- NA
    } else if (data$A[i] == 1) {
      # Start counting when A == 1
      foldaycount <- 0
    }
    # Assign the folday_count value to the current row
    data$foldaycount[i] <- foldaycount
    
    # Update last_id
    last_id <- data$ID[i]
  }
  
  return(data)
}

dat <- calculate_foldaycount(dat, dat$ov_est)

dat %>% select(ID, daterated, ov_est, A, foldaycount)

#calculate folmax = follicular phase length from foldaycount 

#make folmax end when foldaycount ends and not the day after it ends 
for (i in 1:(nrow(dat)-1)) {
  if (is.na(dat$foldaycount[i + 1] && !is.na(dat$foldaycount[i]))) {
    dat$folmax[(i-(dat$foldaycount[i])):i] = as.numeric(dat$foldaycount[i])
  }
}

#folmax does not actually = the follicular length because folmax does not include menses onset. For this reason I made another variable called follength which is folmax + 1
#check data 
dat %>% select(ID, daterated, ov_est, A, foldaycount, folmax)
#create follength 
dat$follength = dat$folmax + 1

#only calculate folperc when follicular length is >= 8 and <= 25 (Based on 2SD below and above means of follicular phase lengths in 21-24 day cycle and 31-35 day cycle)
dat <- dat %>%
  mutate(folperc = ifelse(follength >= 8 & follength < 26, foldaycount / folmax, NA))
#check data
dat %>% select(ID, daterated, LH_confirm, ov_est, A, foldaycount, folmax, follength, folperc, lutperc1)
#momentarily saving foldaycount and folmax for menses centered as foldaycount_b and folmax_b 
dat$foldaycount_b = dat$foldaycount
dat$folmax_b = dat$folmax
```

# examine follicular lengths 

## imputing estimated ovulation

https://www.researchgate.net/publication/335422962_Real-world_menstrual_cycle_characteristics_of_more_than_600000_menstrual_cycles/fulltext/5d652535a6fdccc32cd48fa0/Real-world-menstrual-cycle-characteristics-of-more-than-600-000-menstrual-cycles.pdf?origin=publication_detail&_tp=eyJjb250ZXh0Ijp7ImZpcnN0UGFnZSI6InB1YmxpY2F0aW9uIiwicGFnZSI6InB1YmxpY2F0aW9uRG93bmxvYWQiLCJwcmV2aW91c1BhZ2UiOiJwdWJsaWNhdGlvbiJ9fQ

- if mcyclength is between 31-35 days, folphase is 60% of cycle
  - lutphase is 40%
- if mcyclength is between 25-30 days, folphase is 55% of cycle 
  - lutphase is 45%
- if mcyclength is between 21-24 days, folphase is 53% of cycle 
  - lutphase is 47%
  
  
UPDATE 8/22/24: 
We received more granular means!

- if mcyclength is:21 days 
  - folphase = 11/21, 52.4% of cycle 
  - lutphase = 10/21, 47.6% of cycle 
  
- 22 days 
  - folphase = 11.2/22, 50.9%
  - lutphase = 10.8/22, 49.1%
  
- 23 days 
  - folphase = 11.7/23, 50.9%
  - lutphase = 11.3/23, 49.1%

- 24 days 
  - folphase = 12.2/24, 50.8%
  - lutphase = 11.8/24, 49.2%
  
- 25 days 
  - folphase = 12.9/25, 51.6%
  - lutphase = 12.1/25, 48.4%
  
- 26 days 
  - folphase = 13.5/26, 51.9%
  - lutphase = 12.5/26, 48.1%
  
- 27 days 
  - folphase = 14.3/27, 53.0%
  - lutphase = 12.7/27, 47.0%
  
- 28 days 
  - folphase = 15.1/28, 53.9%
  - lutphase = 12.9/28, 46.1%
  
- 29 days 
  - folphase = 16/29, 55.2%
  - lutphase = 13/29, 44.8%
  
- 30 days 
  - folphase = 16.9/30, 56.3%
  - lutphase = 13.1/30, 43.7%
  
- 31 days 
  - folphase = 17.8/31, 57.4%
  - lutphase = 13.2/31, 42.6%
  
- 32 days 
  - folphase = 18.7/32, 58.4%
  - lutphase = 13.3/32, 41.6%
  
- 33 days 
  - folphase = 19.8/33, 60.0%
  - lutphase = 13.2/33, 40.0%
  
- 34 days 
  - folphase = 20.7/34, 60.9%
  - lutphase = 13.3/34, 39.1%
  
- 35 days 
  - folphase = 21.8/35, 62.3%
  - lutphase = 13.2/35, 37.7%


```{r}
#impute luteal and follicular phase lengths based on Bull 2019

dat <- dat %>% 
  mutate(lutlength_impute = case_when(mcyclength == 21 ~ (mcyclength*.476),
                                      mcyclength == 22 ~ (mcyclength*.491),
                                      mcyclength == 23 ~ (mcyclength*.491),
                                      mcyclength == 24 ~ (mcyclength*.492),
                                      mcyclength == 25 ~ (mcyclength*.484),
                                      mcyclength == 26 ~ (mcyclength*.481),
                                      mcyclength == 27 ~ (mcyclength*.470),
                                      mcyclength == 28 ~ (mcyclength*.461),
                                      mcyclength == 29 ~ (mcyclength*.448),
                                      mcyclength == 30 ~ (mcyclength*.437),
                                      mcyclength == 31 ~ (mcyclength*.426),
                                      mcyclength == 32 ~ (mcyclength*.416),
                                      mcyclength == 33 ~ (mcyclength*.40),
                                      mcyclength == 34 ~ (mcyclength*.391),
                                      mcyclength == 35 ~ (mcyclength*.377),
                                      TRUE ~ NA))

dat <- dat %>% 
  mutate(follength_impute = case_when(mcyclength == 21 ~ (mcyclength*.524),
                                      mcyclength == 22 ~ (mcyclength*.509),
                                      mcyclength == 23 ~ (mcyclength*.509),
                                      mcyclength == 24 ~ (mcyclength*.508),
                                      mcyclength == 25 ~ (mcyclength*.516),
                                      mcyclength == 26 ~ (mcyclength*.519),
                                      mcyclength == 27 ~ (mcyclength*.530),
                                      mcyclength == 28 ~ (mcyclength*.539),
                                      mcyclength == 29 ~ (mcyclength*.552),
                                      mcyclength == 30 ~ (mcyclength*.563),
                                      mcyclength == 31 ~ (mcyclength*.574),
                                      mcyclength == 32 ~ (mcyclength*.584),
                                      mcyclength == 33 ~ (mcyclength*.60),
                                      mcyclength == 34 ~ (mcyclength*.609),
                                      mcyclength == 35 ~ (mcyclength*.623),
                                      TRUE ~ NA))
#cehck data
dat %>% select(ID, daterated, A, m2mcount, mcyclength, follength_impute)

#extra check to make sure data is grouped by ID and ordered by date
dat <- dat %>%
  group_by(ID) %>%
  arrange(daterated, .by_group = TRUE) 

#follcount1_impute starts counting on menses onset and stops when it reaches the length of follength_impute 
dat$follcount1_impute <- NA
for (i in 1:nrow(dat)) {
  if (dat$A[i] == 1 & !is.na(dat$follength_impute[i])) {
    follcount1_impute <- seq_len(round(dat$follength_impute[i]))
    end_index <- i + length(follcount1_impute) - 1
    if (end_index <= nrow(dat)) {
      dat$follcount1_impute[i:end_index] <- follcount1_impute
    } else {
      dat$follcount1_impute[i:nrow(dat)] <- follcount1_impute[1:(nrow(dat) - i + 1)]
    }
  }
}
#check data 
dat %>% select(ID, A, follength_impute, follcount1_impute)
#ovtoday_impute = 1, when follength_impute and follcount1_impute are the same (or follcount1_impute rounds to follength_impute)
dat <- dat %>% 
  mutate(ovtoday_impute = case_when(round(follength_impute) == follcount1_impute ~ 1,
                               TRUE ~ NA))
#check data
dat %>% select(ID, daterated, A, follength_impute, follcount1_impute, ov_est, ovtoday_impute, cyclenum)

```

### perclut using ovtoday_impute 

- negative 1 starts the day after ov_est = 1 or ovtoday_impute = 1 (which is the 2nd day of ovulation)

```{r}
#group by ID and make sure sorted by daterated. Also initialize the variable lutmax_impute 
dat <- dat  %>% 
  group_by(ID) %>%
  arrange(daterated, .by_group = TRUE) %>%
  mutate(lutmax_impute = NA)
#if ovtoday_impute is NA, make it 0
dat$ovtoday_impute <- ifelse(is.na(dat$ovtoday_impute), 0, dat$ovtoday_impute) 
#applying calculate lutdaycount function 
dat <- calculate_lutdaycount(dat, dat$ovtoday_impute)
#check data 
dat %>% select(ID, daterated, ovtoday_impute, A, lutdaycount1)
#the function spits out the variable lutdaycount1, so here I am renaming it to lutdaycount1_impute
dat <- dat %>% rename(lutdaycount1_impute = lutdaycount1)
#this creates lutdaycount_impute, in which it starts couting the day after ov_est and ends on menses onset 
dat <- dat %>%
  group_by(ID) %>%
  mutate(
    lutdaycount_impute = lag(lutdaycount1_impute),
    lutdaycount_impute = case_when(
      is.na(lutdaycount_impute) | ID != lag(ID) ~ NA, 
      TRUE ~ lutdaycount_impute
    )
  )
#check data 
dat %>% select(ID, daterated, ovtoday_impute, A, lutdaycount_impute)
#lutmax_impute = max value of lutdaycount 
for (i in 1:(nrow(dat)-1)) {
  if (is.na(dat$lutdaycount_impute[i + 1] && !is.na(dat$lutdaycount_impute[i]))) {
    dat$lutmax_impute[(i-(dat$lutdaycount_impute[i])):i] = as.numeric(dat$lutdaycount_impute[i])
  }
}
#lutlength1_impute applies to the same rows as lutelength_impute 
dat$lutlength1_impute = lag(dat$lutlength_impute)

dat %>% select(ID, daterated, ovtoday_impute, A, lutdaycount_impute, lutmax_impute, lutlength_impute)
# calculate lutperc_impute only when lutdaycount_impute is >= 6 days
dat <- dat %>%
  mutate(lutperc_impute = if_else(is.na(lutlength1_impute) & lutmax_impute < 6,
                           NA,  
                           if_else(is.na(lutlength1_impute) & !is.na(lutmax_impute),
                                   lutdaycount_impute / lutmax_impute,
                                   lutdaycount_impute / round(lutlength1_impute))))
#calcualte perclut_impute, which is the variable that will be included in scaled_cycleday 
dat$perclut_impute = dat$lutperc_impute -1 
#check data 
dat %>% select(ID, daterated, ovtoday_impute, A, lutdaycount_impute, lutlength1_impute, lutmax_impute, perclut_impute) 

#calculate lutdaycount_imp_ov, which is imputed lutdaycount but ovulation-centered 
dat <- dat %>%
  group_by(ID) %>%
  mutate(
    lutdaycount_imp_ov = lead(lutdaycount_impute),
    lutdaycount_imp_ov = case_when(
      is.na(lutdaycount_imp_ov) | ID != lead(ID) ~ NA, 
      TRUE ~ lutdaycount_imp_ov
    )
  )
#check data
dat %>% select(ID, daterated, ovtoday_impute, A, lutdaycount_imp_ov, lutmax_impute) 
#calculate lutperc_imp_ov which is the variable that will go into ovulation-centered scaled cycleday 
dat <- dat %>%
  mutate(lutperc_imp_ov = ifelse(lutmax_impute <= 18 & lutmax_impute >= 6, lutdaycount_imp_ov / lutmax_impute, NA))
dat <- dat %>%
  mutate(lutperc_imp_ov = ifelse(ovtoday_impute == 1, 0, lutperc_imp_ov))

#check data 
dat %>% select(ID, daterated, ovtoday_impute, A, lutdaycount_impute, lutmax_impute, perclut_impute, lutperc_imp_ov)


```

### percfol using ovtoday_impute 

```{r}
#group by ID and sort daterated 
#initialize folmax_impute 
dat <- dat %>%
  group_by(ID) %>%
  arrange(daterated, .by_group = TRUE) %>%
  mutate(folmax_impute = NA)
#use function calculate_foldaycount to create foldaycount
dat <- calculate_foldaycount(dat, dat$ovtoday_impute)
#check data 
dat %>% select(ID, daterated, ovtoday_impute, A, foldaycount)
#rename foldaycount to foldaycount_impute
dat <- dat %>% rename(foldaycount_impute = foldaycount)
#create folmax_impute the highest number foldaycount_impute counts up to
for (i in 1:(nrow(dat)-1)) {
  if (is.na(dat$foldaycount_impute[i + 1] && !is.na(dat$foldaycount_impute[i]))) {
    dat$folmax_impute[(i-(dat$foldaycount_impute[i])):i] = as.numeric(dat$foldaycount_impute[i])
  }
}

#check data 
dat %>% select(ID, A, ovtoday_impute, foldaycount_impute, follength_impute, folmax_impute)
#create percfol_impute, no need to worry about length of follicular phase because imputed variables are based on Bull 2019 norms 
dat <- dat %>%
  mutate(percfol_impute = (foldaycount_impute / folmax_impute))

dat %>% select(ID, daterated, ovtoday_impute, A, foldaycount_impute, folmax_impute, percfol_impute, folperc)
```

## Scaled Cycleday 

```{r}
#create percfol_ov which will go into scaled_cycleday for ovulation-centered
dat <- dat %>% 
  mutate(
    percfol = case_when(!is.na(lutperc1) & !is.na(folperc) & folperc !=0 ~ NA, TRUE ~ folperc),
    percfol_ov = percfol - 1
  )
#checking 
dat %>% select(ID, daterated, folperc, lutperc1, percfol_ov)

#creating percentlut and percentfol 
dat <- dat %>%
  mutate(
    percentlut = case_when(
      !is.na(percfol) ~ 0,
      TRUE ~ lutperc1
    ),
    percentfol = case_when(
      !is.na(lutperc1) ~ 0,
      TRUE ~ percfol
    )
  )
#checking 
dat %>% select(ID, daterated, A, ov_est, percentlut, percentfol)
#create scaled_cycleday 
dat <- dat %>%
  mutate(scaled_cycleday = ifelse(percentlut==0, percentfol, percentlut))

### scaled cycleday ovulation centered 
dat <- dat %>%
  mutate(scaled_cycleday_ov = ifelse(is.na(percfol_ov), lutperc_ov, percfol_ov))
#checking 
dat %>% select(ID, daterated, A, ov_est, scaled_cycleday, scaled_cycleday_ov)

#creating cycleday_perc for plotting (rounded to the nearest 10th, 5th , and 2nd percent). In my experience, rounded to the nearest 5th has always worked the best 
dat <- dat %>%
  mutate(cycleday_perc = (scaled_cycleday + 1) / 2) %>%
  mutate(cycleday_perc_round = round(cycleday_perc, 1)) %>%
  mutate(cycleday_5perc = round(cycleday_perc / 0.05) * 0.05) %>%
  mutate(cycleday_2perc = round(cycleday_perc / 0.02) * 0.02)
#do the same for ovulation-centered 
dat <- dat %>%
  mutate(cycledayov_perc = (scaled_cycleday_ov + 1) / 2) %>%
  mutate(cycledayov_perc_round = round(cycledayov_perc, 1)) %>%
  mutate(cycledayov_5perc = round(cycledayov_perc / 0.05) * 0.05) %>%
  mutate(cycledayov_2perc = round(cycledayov_perc / 0.02) * 0.02)

```

## Scaled Cycleday using imputed ovulation 


- ovtoday_impute measures: percfol_impute and perclut_impute 
- ov_est measures: lutperc1 and percfol

```{r}
#these functions create_percentlut and create_percentfol prioritize measures created with ov_est over ovtoday_impute for menses centered 

create_percentlut <- function(dat) {
  dat %>%
    # Identify stretches of consecutive non-NA perclut values
    mutate(group = cumsum(is.na(perclut_impute) & lag(is.na(perclut_impute), default = TRUE) != is.na(perclut_impute))) %>%
    group_by(ID, group) %>%
    mutate(percentlut_impute = if (any(!is.na(lutperc1))) lutperc1 else lutperc1) %>%
    ungroup() %>%
    select(-group)
}

create_percentfol <- function(dat) {
  dat %>%
    # Identify stretches of consecutive non-NA perclut values
    mutate(group = cumsum(is.na(percfol_impute) & lag(is.na(percfol_impute), default = TRUE) != is.na(percfol_impute))) %>%
    group_by(ID, group) %>%
    mutate(percentfol_impute = if (any(!is.na(folperc))) folperc else percfol_impute) %>%
    ungroup() %>%
    select(-group)
}

#prioritizes luteal phase measures created with ov_est over ovtoday_impute for ovulation-centered
 dat <- dat %>%
    # Identify stretches of consecutive non-NA perclut values
    mutate(group = cumsum(is.na(lutperc_imp_ov) & lag(is.na(lutperc_imp_ov), default = TRUE) != is.na(lutperc_imp_ov))) %>%
    group_by(ID, group) %>%
    mutate(perclut_ov_imp = if (any(!is.na(lutperc_ov))) lutperc_ov else lutperc_ov) %>%
    ungroup() %>%
    select(-group)

#apply functions
dat <- create_percentlut(dat)
dat <- create_percentfol(dat)
#create ovulation-centered percfol_ov_imp from menses-centered percentfol_impute
dat <- dat %>%
  mutate(percfol_ov_imp = ifelse(!is.na(percentfol_impute), percentfol_impute-1, NA))
#created scaled_cycleday_impute from percentlut_impute and from percentfol_impute
dat <- dat %>%
  mutate(scaled_cycleday_impute = ifelse(is.na(percentlut_impute), percentfol_impute, percentlut_impute))
#turn NA to 0 in percentlut_impute and percentfol_impute
dat$percentlut_impute = ifelse(is.na(dat$percentlut_impute) & !is.na(dat$percentfol_impute), 0, dat$percentlut_impute)
dat$percentfol_impute = ifelse(is.na(dat$percentfol_impute) & !is.na(dat$percentlut_impute), 0, dat$percentfol_impute)
#create ovulation_centered scaled_cycleday with imputed measures too 
dat <- dat %>%
  mutate(scaled_cycleday_imp_ov = ifelse(is.na(percfol_ov_imp), perclut_ov_imp, percfol_ov_imp))
#check data 
dat %>% select(ID, ov_est, A, ovtoday_impute, scaled_cycleday, scaled_cycleday_impute, scaled_cycleday_imp_ov)


#percenatge variables for plotting 
dat <- dat %>%
  mutate(cycleday_perc_imp = (scaled_cycleday_impute + 1) / 2) %>%
  mutate(cycleday_perc_imp_round = round(cycleday_perc_imp, 1)) %>%
  mutate(cycleday_5perc_imp = round(cycleday_perc_imp / 0.05) * 0.05) %>%
  mutate(cycleday_2perc_imp = round(cycleday_perc_imp / 0.02) * 0.02)

dat <- dat %>%
  mutate(cycledayov_perc_imp = (scaled_cycleday_imp_ov + 1) / 2) %>%
  mutate(cycledayov_perc_imp_round = round(cycledayov_perc_imp, 1)) %>%
  mutate(cycledayov_5perc_imp = round(cycledayov_perc_imp / 0.05) * 0.05) %>%
  mutate(cycledayov_2perc_imp = round(cycledayov_perc_imp / 0.02) * 0.02)
```

# how many missing rows in scaled_cycleday vs scaled_cycleday_impute 

```{r}
sum(is.na(dat$scaled_cycleday)) #1801
sum(is.na(dat$scaled_cycleday_impute)) ##1395
sum(is.na(dat$cycleday)) #1191

#how many observations in which scaled_cycleday and hormones are non-NA?

#confirmed menses onsets and LH peaks
dat %>%
  summarise(complete_cases = sum(!is.na(scaled_cycleday) & (!is.na("E1G_Crtn_z")|!is.na("PDG_Crtn_z")))) #3024
#confirmed menses onsets
dat %>%
  summarise(complete_cases = sum(!is.na(cycleday) & (!is.na("E1G_Crtn_z")|!is.na("PDG_Crtn_z")))) #3634

#confirmed menses onsets, LH peaks, and imputed ovulation
dat %>%
  summarise(complete_cases = sum(!is.na(scaled_cycleday_impute) & (!is.na("E1G_Crtn_z")|!is.na("PDG_Crtn_z")))) #3430


dat %>%
  group_by(ID) %>%
  summarise(complete_cases = sum(!is.na(scaled_cycleday) & (!is.na("E1G_Crtn_z")|!is.na("PDG_Crtn_z"))))

dat %>%
  group_by(ID) %>%
  summarise(complete_cases = sum(!is.na(cycleday) & (!is.na("E1G_Crtn_z")|!is.na("PDG_Crtn_z"))))

dat %>%
  group_by(ID) %>%
  summarise(complete_cases = sum(!is.na(scaled_cycleday_impute) & (!is.na("E1G_Crtn_z")|!is.na("PDG_Crtn_z"))))

```

```{r}
saveRDS(dat, file = "C:/Users/Student/Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/BEARS/02_datasets/BEARS_DAILY/03_cleaned_data/BEARS_scaledcycleday.rds")
```





