---
title: "ADHDCYCLE_dailyprep"
output: html_document
date: "Last compiled on `r format(Sys.time(), '%B %d, %Y')`"
---

```{r setup}

knitr::opts_chunk$set(echo = TRUE)

library(dplyr) # Data manipulation
library(tidyr) # Tidying data
library(haven) # Reading files (SPSS, stata, SAS)
library(ggplot2) # Plotting
library(tidyverse) # Data management and manipulation
library(zoo) # Rolling averages
library(readxl) # Reading Excel files
library(lme4) # Mixed-effects models
library(lmerTest) # Mixed-effects models with p-values
library(emmeans) # Estimated marginal means
library(psych) # Descriptive statistics
library(lubridate) # Date and time manipulation
library(visdat) # Visualizing missing data
library(janitor) # Cleaning data
library(skimr) # Skimming data for summary statistics
library(ggdist) # Visualizing distributions and uncertainty
library(ggforce) # Extending ggplot2
library(sjPlot) # Plotting for statistical models
library(grid) # Base graphics
library(gridExtra) # Arranging grid-based graphics
library(usethis) # Setting up GitHub and other project tools
library(nlme) # Linear and nonlinear mixed-effects models
library(performance) # Model performance metrics
library(rmcorr) # Repeated measures correlations
library(gitcreds) # Managing Git credentials
library(see) # Visualizing model checking functions
library(corrplot) # Visualizing correlations
library(GPArotation) # Factor analysis rotations
library(broom.mixed) # Tidying mixed models
library(pbkrtest) # Parametric bootstrap and Kenward-Roger methods
library(ggrepel) # Adding text labels to ggplot2
library(zoo)

```

## Load raw daily data, merged to include surveys and hormones 

```{r load_data}
df <- read.csv("~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/02_datasets/CYCADHD_DAILY/01_raw_data/2024.04.24. Daily Master.csv")
```

# Change date_rated to be date format

```{r}
 # Convert date_rated to Date type
  df$date_rated <- as.Date(mdy(df$date_rated))

```

# Remove NA rows
```{r}
#remove rows where all data is NA
df <- df[rowSums(is.na(df)) != ncol(df), ]

#remove NA ID's from dataset
df <- df %>% filter(!is.na(id))

#remove NA daterated from dataset 
df <- df %>% filter(!is.na(date_rated))

# NUMBER OF ROWS

print(nrow(df))

```

# Sort by id and date_rated and save
```{r}
df <- df %>% arrange(id, date_rated, .na.last = FALSE)

#View(df)
```


#Hard-code changes to issues found later 
```{r}

#Add 208 Menses start date 
df$StartPeriod[df$id == 208 & df$TubeNumber == 27] <- 1
#df %>% filter(id==208) %>% View()

df$date_rated[df$id == 210 & df$date_rated == 11/20/2021] <- 11/20/2020
df$date_rated[df$id == 210 & df$date_rated == 11/21/2021] <- 11/21/2020
#df %>% filter(id==210) %>% View()

#View(df)

#Template Code to look at one pt's data

#df %>% filter(id==307) %>% View()

#ADDITIONAL DATA REVIEW NEEDED FOR HORMONES: 

# Filtering template: df <- df %>% filter(!is.na(id))

#Note that 307's ovulation might be a little off with their hormones
#Consider excluding participant 235
#Review participant 273's progesterone data
#Review participant 278's progesterone data
#Review participant 317's estrogen data
#Review participant 221's hormone data
#Review participant 243's hormone data
#Review participant 267's hormone data
#Review participant 272's hormone data
#Review participant 283's hormone data
#Review participant 305's hormone data
#Review participant 312's hormone data

badhorm_ids <- c(221, 235, 243, 267, 272, 273, 278, 283, 305, 307, 312, 317)

```


# Add row placeholders for missing survey/date_rated's
```{r}

df <- df %>% group_by(id) %>%
  complete(date_rated = seq.Date(min(date_rated), max(date_rated), by = "day"))

#View(df)

# NUMBER OF ROWS
print(nrow(df))

```

# Rename Hormones
```{r RENAME}
# Rename the columns
colnames(df)[colnames(df) == "Estradiol"] <- "E2"
colnames(df)[colnames(df) == "Progesterone"] <- "P4"
```

## Calculate BAARS ADHD Current Symptom Scale CSS Scores, UPPS, BDEFS, DEBQ
```{r calc CSS UPPS BDEFS DEBQ}

#CSS_Total - Overall total score on the current symptoms scale
#CSS_TotalSymp - Total ADHD symptom score (inattention + hyperactivity-impulsivity)
#CSS_B_TotalIASymp - Total inattention symptoms score 
#CSS_B_TotalHISymp - Total hyperactivity-impulsivity symptoms score 

#CSS_TotalCount - Total count of all ADHD symptoms (inattention + hyperactivity-impulsivity)
#CSS_Inatt_Count - Count of inattention symptoms
#CSS_Imp_Count - Count of impulsivity symptoms
#CSS_Hyp_Count - Count of hyperactivity symptoms

#CSS_Fx_Total - ???????
#CSS_B2_Total - ???????

# Reverse UPPS Items 7 and 8
df <- df %>%
  mutate(UPPS_7r=recode(UPPS_7, '1' = 4, '2' = 3, '3' = 2, '4' = 1), UPPS_8r=recode(UPPS_8, '1' = 4, '2' = 3, '3' = 2, '4' = 1))

# Drop original items UPPS_7 and UPPS_8

df <- df %>% select(-UPPS_7)
df <- df %>% select(-UPPS_8)


df <- df %>%
  group_by(id) %>%
  mutate(
    # CSS Symptom severity Scores
    
    # Calculate Inattention Severity: 
    CSS_Inatt = rowSums(across(c(CSS_B_1, CSS_B_3, CSS_B_5, CSS_B_7, CSS_B_9, CSS_B_11, CSS_B_13, CSS_B_15, CSS_B_17)), na.rm = TRUE),
    
    # Calculate Hyperactivity/Impulsivity Severity: 
    CSS_HypImp = rowSums(across(c(CSS_B_2, CSS_B_4, CSS_B_6, CSS_B_8, CSS_B_10, CSS_B_12, CSS_B_14, CSS_B_16, CSS_B_18)), na.rm = TRUE),
    
    # Function items 1-10
    CSS_Fx_Total = rowSums(across(starts_with("CSS_Function_")), na.rm = TRUE),
    
    # Not sure what this is (todo)
    CSS_B2_Total = rowSums(across(starts_with("CSS_B2_")), na.rm = TRUE),
    
    # CSS Count Variables
    CSS_Inatt_Count = rowSums(across(c(CSS_1_Count, CSS_3_Count, CSS_5_Count, CSS_7_Count, CSS_9_Count, CSS_11_Count, CSS_13_Count, CSS_15_Count, CSS_17_Count)), na.rm = TRUE),
    
    CSS_HypImp_Count = rowSums(across(c(CSS_2_Count, CSS_4_Count, CSS_6_Count, CSS_8_Count, CSS_10_Count, CSS_12_Count, CSS_14_Count, CSS_16_Count, CSS_18_Count)), na.rm = TRUE),
    
    # Other Daily Scored Scales
    DEBQ_Total = rowSums(across(starts_with("DEBQ_")), na.rm = TRUE),
    BDEFS_Total = rowSums(across(starts_with("BDEFS_")), na.rm = TRUE),
    UPPS_Total = rowSums(across(starts_with("UPPS_")), na.rm = TRUE)
  ) %>%
  ungroup()

```

### functions to create person means and person deviations
```{r echo=T}
#function to create person means
create.person.mean <- function(df, var, ...) {
  df %>%
    group_by(...) %>%
    mutate("{{var}}.m" := mean({{var}}, na.rm=T))
}
#function to create person deviations (note, must have person means already made)
create.deviation <- function(df, var, var.m) {
  df <- df %>%
    rowwise() %>%
    mutate("{{var}}.d" := {{var}} - {{var.m}})
}

create.3day.rolling.avg <- function(df, var, num) {
  df %>%
    group_by(id) %>%
    mutate("{{var}}.roll" := rollapply({{var}}, 3,  mean, align="center", fill=NA))
}


```


## create outcome lists to loop through functions
```{r echo=T}

#person-centered deviations
outcomelist <- c("E2",
           "P4",
           "LH",
           "CSS_Fx_Total",
           "CSS_B2_Total",
           "CSS_Inatt",
           "CSS_HypImp",
           "CSS_Inatt_Count",
           "CSS_HypImp_Count",
           "BDEFS_Total", 
           "UPPS_Total",
           "DEBQ_Total"
           ) %>% noquote()

#person-centered deviations
outcomelist.d <- c("E2.d",
           "P4.d",
           "LH.d",
           "CSS_Fx_Total.d",
           "CSS_B2_Total.d",
           "CSS_Inatt.d",
          "CSS_Inatt_Count.d",
           "CSS_HypImp.d",
           "CSS_HypImp_Count.d",
           "BDEFS_Total.d", 
           "UPPS_Total.d",
           "DEBQ_Total.d"
           ) %>% noquote()

#rolling averages
outcomelist.roll <- c("E2.roll",
"P4.roll",
"LH.roll",
"CSS_Fx_Total.roll",
"CSS_B2_Total.roll",
"CSS_Inatt.roll",
"CSS_HypImp.roll",
"CSS_Inatt_Count.roll",
"CSS_HypImp_Count.roll",
"BDEFS_Total.roll", 
           "UPPS_Total.roll",
           "DEBQ_Total.roll"
) %>% noquote()

View(df)


```

```{r}

#ROLLING AVGS

#create rolling averages on RAW variables
for (i in outcomelist) {
  df <- create.3day.rolling.avg(df, !!sym({{i}}), 3)
}

#MEANS

#execute for loop: run create.person.mean() on everything in "outcomelist.roll"
for (i in outcomelist.roll) {
  df <- create.person.mean(df, !!sym({{i}}), id)
}

#execute for loop: run create.person.mean() on everything in "outcomelist"
for (i in outcomelist) {
  df <- create.person.mean(df, !!sym({{i}}), id)
}

#PERSON-DEVIATIONS

#for raw deviations
#execute for loop: run create.deviation() on everything in list
for (i in outcomelist) {
  df <- create.deviation(df, !!sym({{i}}), !!sym(paste0({{i}}, ".m")))
}

#for rolling avgs deviations
#execute for loop: run create.deviation() on everything in list
for (i in outcomelist.roll) {
  df <- create.deviation(df, !!sym({{i}}), !!sym(paste0({{i}}, ".m")))
}


#Check 
df %>% select(id, E2, E2.m, E2.d, E2.roll, E2.roll.d) %>% View()

```

# Print Variable Names
```{r printvars}
variable_names <- names(df)
formatted_list <- paste(variable_names, collapse = ", ")
cat(formatted_list)
```

# Create dichotomous mensdayone/mensdayonefirst to indicate first day of period
```{r }


str(df$id) #check id format
str(df$date_rated) #check date_rated format

df <- df %>%
  filter(!is.na(id))

df <- df %>%
  arrange(id, date_rated)


df$StartPeriod <- as.numeric(df$StartPeriod)

df <- df %>%
  mutate(mensdayone = case_when(
    is.na(StartPeriod) ~ NA_real_,
    StartPeriod == 1 ~ 1,
    TRUE ~ 0
  ))


# Eliminate all but the first menses onset day in a new variable called "mensdayonefirst"

# Process the data
df <- df %>%
  arrange(id, date_rated) %>%  # Arrange the data by id and date_rated
  group_by(id) %>%  # Group the data by id
  mutate(
    mensdayonefirst = case_when(
      is.na(mensdayone) ~ NA,
      mensdayone == 1 & (lag(mensdayone, order_by = date_rated) != 1 | is.na(lag(mensdayone, order_by = date_rated))) ~ 1,
      row_number() == 1 & mensdayone == 1 ~ 1,
      TRUE ~ 0
    )
  ) %>% 
  ungroup() # Remove the grouping

#Check

#df %>% select(id, date_rated, mensdayone, mensdayonefirst, PosLHTest) %>% View()


```

# Create "LHposday" (0/1/NA) that represents the positive LH day.
```{r }

##View(df)
  
  df$PosLHTest <- as.numeric(df$PosLHTest)

df <- df %>%
  mutate(LHposday = case_when(
    is.na(PosLHTest) ~ NA_real_,
    PosLHTest == 1 ~ 1,
    TRUE ~ 0
  ))

##View(df)

# Eliminate all but the first pos Ov day in a new variable called "LHposdayfirst"
df <- df %>%
  arrange(id, date_rated) %>%  # Arrange the data by id and date_rated
  group_by(id) %>%  # Group the data by id
  mutate(
    LHposdayfirst = case_when(
      is.na(LHposday) ~ NA_real_,
      LHposday == 1 & lag(LHposday, order_by = date_rated) != 1 ~ 1,
      TRUE ~ 0
    )
  ) %>% 
  ungroup()  # Remove the grouping

#Check vars
#df %>% select(id, date_rated, mensdayonefirst, LHposdayfirst) %>% View()

```

# Remove people who never had a positive LH test
```{r}

# Create a person-level variable
idlhpos <- df %>%
  group_by(id) %>% 
  summarise(LHeverpos = ifelse(sum(LHposdayfirst, na.rm = TRUE) > 0, 1, 0)) %>% 
  ungroup()

# Merge it back into df
df <- df %>%
  left_join(idlhpos, by = "id")

# Filter out people with LHids = 0
df <- df %>%
  filter(LHeverpos == 1)

#Check vars
df %>% select(id, LHeverpos, date_rated, mensdayonefirst, LHposdayfirst) %>% head()

```

# Code Cycle Day Forward + Backward count from menses onset
```{r cycleday}

#make A as a variable copy of mensdayonefirst
df$A <- df$mensdayonefirst

cycleCount <- function(x) {
  #Get the index of 1
  inds <- which(x == 1)
  if(!length(inds)) return(0) 
  #Create a sequence with that index as 0
  num <- lapply(inds, function(i) {
    num <- seq_along(x) - i
    #Add 1 to values greater than equal to 0
    num[num >= 0] <- num[num >= 0] + 1
    num[num < -15 | num > 11] <- NA 
    num
  })
  #Select the first non-NA values from the sequence
  do.call(coalesce, num)
}

df <- df %>% group_by(id) %>% 
  mutate(cycleday = cycleCount(A))

df <- df %>% select(!A)

##View(df)

#Check vars
#df %>% select(id, LHeverpos, date_rated, mensdayonefirst, LHposdayfirst, cycleday) %>% View()

```

# Review Data, looking for cycleday=0 and removing those who never had any menses onsets
```{r}

#View in a separate window and manually scroll through
df %>% select(id, date_rated, mensdayonefirst, cycleday, LHposdayfirst, cycleday) %>% View()

#how many observations do you have for each cycle day?
nobs <- df %>% group_by(cycleday) %>% summarize(n=n())

#View(nobs)

#do you have any observations that got assigned a cycle day of 0? this is wrong, and you should check who that happened for
df %>% filter(cycleday==0) %>% pull(id)

# NOTE 2024-07-09: There are 2 people with lots of zeroes (see Jordan's note above in chunk) and it looks to me like they did not report a menses in the daily dataset: 331, 332. They have no menses onset dates

#View in a separate window and manually scroll through
#df %>% select(id, date_rated, mensdayonefirst, cycleday, LHposdayfirst, cycleday) %>% View()

# List of IDs to remove
ids_to_remove <- c(331, 332)

# Remove rows with specified IDs
df <- df %>% filter(!(id %in% ids_to_remove))


```

# Make daycountLH, based on positive LH test (LH test day = 0)
```{r daycountLH}
#make A a new temp column where pt received pos ov test
df$A <- df$LHposdayfirst

#FUNCTION for calculating the sequence
LHCount <- function(x) {
  #Get the index of 1
  inds <- which(x == 1)
  if(!length(inds)) return(0) 
  #Create a sequence with that index as 0
  num <- lapply(inds, function(i) {
    num <- seq_along(x) - i
    num[num < -7 | num > 15] <- NA 
    num
  })
  #Select the first non-NA values from the sequence
  do.call(coalesce, num)
}

#run the LHCount function and save it as as a new column called daycountLH
df<- df %>% group_by(id) %>% 
  mutate(daycountLH = LHCount(A))
#remove the temp column A
df <- df %>% select(!A)

#Check vars
#df %>% select(id, LHeverpos, date_rated, mensdayonefirst, LHposdayfirst, daycountLH) %>% View()
```

# View the number of observations for each cycleday
```{r daycountLH checking}

#how many daily ratings got assigned each cycle day?
nobs2 <- df %>% group_by(daycountLH) %>% summarize(n=n())

View(nobs2)

```

# Cycle Phase Coding - MENSES COUNT PART
reminder: dummy code comes from Schmalenberger 2021 appendix
```{r cyclephase_count}

df <- df %>% 
  mutate(midluteal_count = ifelse(cycleday >= -9 & cycleday <= -5, 1, 0),
         perimenstrual_count = ifelse(cycleday >=-3 & cycleday <=2, 1, 0),
         midfol_count = ifelse(cycleday >=4 & cycleday <=7, 1, 0),
         periov_count = ifelse(cycleday >= -15 & cycleday <= -12, 1, 0))

#add a non-dummy coded variable with all of cycle phase info above, which will be useful for future visualizations and summaries
df <- df %>% mutate(cyclephase_count = case_when(periov_count==1 ~ 1,
                                                     midluteal_count==1 ~ 2,
                                                     perimenstrual_count==1 ~ 3,
                                                     midfol_count==1 ~ 4,
                                                     TRUE ~ as.numeric(NA)))

#Check vars
#df %>% select(id, mensdayonefirst, LHposdayfirst, cycleday, daycountLH, ends_with("_count")) %>% View()

```

# Cycle Phase Coding - LH COUNT PART
```{r cyclephase_LH}

#note: the perimenstrual phase is still coded based on menses count variable, but for the purpose of keeping the dummy code labeling clear, we name it *perimenstrual_LH* to indicate that this variable gets included in the dummy code set of LH-based phasing; again, all of this information is from Schmalenberger 2021 appendix

df <- df %>% 
  mutate(midluteal_LH = ifelse(daycountLH >= 6 & daycountLH <= 11, 1, 0),
         perimenstrual_LH = ifelse(cycleday >=-3 & cycleday <=2, 1, 0), #note perimenstrual is based on cycleday
         midfol_LH = ifelse(daycountLH >=-7 & daycountLH <=-3, 1, 0),
         periov_LH = ifelse(daycountLH >= -2 & daycountLH <= 1, 1, 0),
         earlylut_LH = ifelse(daycountLH >= 2 & daycountLH <= 5, 1, 0))

#fill out the dummy code, so that if any phase is 1, all others should be 0 (not NA)
df <- df %>%
  rowwise() %>%
  mutate(sumdummy = sum(midfol_LH,
                        periov_LH,
                        earlylut_LH,
                        perimenstrual_LH,
                        midluteal_LH, na.rm = T)) %>% 
  #if all phases=0, make the midluteal_LH variable NA instead of 0,
  #if perimenstrual by menses count = 1, change midluteal_LH to 0 instead of NA to fill out structural set,
  #otherwise, keep as is
  mutate(midluteal_LH = case_when(sumdummy==0 ~ as.numeric(NA),
                                  sumdummy==1 & perimenstrual_LH==1 ~ 0,
                                  TRUE ~ midluteal_LH), 
         #same for midfol
         midfol_LH = case_when(sumdummy==0 ~ as.numeric(NA),
                               sumdummy==1 & perimenstrual_LH==1 ~ 0,
                               TRUE ~ midfol_LH), 
          #same for periov
         periov_LH = case_when(sumdummy==0 ~ as.numeric(NA),
                               sumdummy==1 & perimenstrual_LH==1 ~ 0,
                               TRUE ~ periov_LH),
         earlylut_LH = case_when(sumdummy==0 ~ as.numeric(NA),
                                  sumdummy==1 & perimenstrual_LH==1 ~ 0,
                                  TRUE ~ earlylut_LH), 
         #if any other phase is 1, fill out perimenstrual to be 0
         perimenstrual_LH = case_when(sumdummy==0 ~ as.numeric(NA), 
                                        sumdummy==1 & midfol_LH==1 ~ 0,
                                        sumdummy==1 & periov_LH==1 ~ 0,
                                      sumdummy==1 & earlylut_LH==1 ~ 0,
                                        sumdummy==1 & midluteal_LH==1 ~ 0,
                                        TRUE ~ perimenstrual_LH))

#add cyclephase_LH, a new categorical variable made from all of the new hybrid cycle phases above, which will be useful for future visualizations and summaries
df <- df %>% mutate(cyclephase_LH = case_when(periov_LH==1 ~ 1,
                                                                        earlylut_LH==1 ~ 2,
                                                     midluteal_LH==1 ~ 3,
                                                     perimenstrual_LH==1 ~ 4,
                                                     midfol_LH==1 ~ 5,
                                                     TRUE ~ as.numeric(NA)))

df <- df %>% mutate(cyclephase_LH_graph = case_when(midfol_LH==1 ~ 1,
                                                    periov_LH==1 ~ 2,
                                                    earlylut_LH==1 ~ 3,
                                                     midluteal_LH==1 ~ 4,
                                                     perimenstrual_LH==1 ~ 5,
                                                     TRUE ~ as.numeric(NA)))

#Check vars
#df %>% select(id, mensdayonefirst, LHposdayfirst, cycleday, daycountLH, ends_with("_LH")) %>% View()



#Coding additional binary variables for use in random effects that contrast each phase with all other phases for simplicity

#nonperi (1) vs. peri (0)
df <- df %>%
  mutate(NONPERI = case_when(cyclephase_LH=="1" ~ "1", #periov
                                    cyclephase_LH=="2" ~ "1", #earlylut
                                    cyclephase_LH=="3" ~ "1", #midlut
                                    cyclephase_LH=="4" ~ "0", #perimens
                                    cyclephase_LH=="5" ~ "1")) #midfol

#nonmid (1) vs. mid (0)
df <- df %>%
  mutate(NONMid = case_when(cyclephase_LH=="1" ~ "1", #periov
                                    cyclephase_LH=="2" ~ "1", #earlylut
                                    cyclephase_LH=="3" ~ "0", #midlut
                                    cyclephase_LH=="4" ~ "1", #perimens
                                    cyclephase_LH=="5" ~ "1")) #midfol

#nonel (1) vs. el (0)
df <- df %>%
  mutate(NONEL = case_when(cyclephase_LH=="1" ~ "1", #periov
                                    cyclephase_LH=="2" ~ "0", #earlylut
                                    cyclephase_LH=="3" ~ "1", #midlut
                                    cyclephase_LH=="4" ~ "1", #perimens
                                    cyclephase_LH=="5" ~ "1")) #midfol

#nonperiov (1) vs. periov (0)
df <- df %>%
  mutate(NONPERIOV = case_when(cyclephase_LH=="1" ~ "0", #periov
                                    cyclephase_LH=="2" ~ "1", #earlylut
                                    cyclephase_LH=="3" ~ "1", #midlut
                                    cyclephase_LH=="4" ~ "1", #perimens
                                    cyclephase_LH=="5" ~ "1")) #midfol

#nonmidfol (1) vs. midfol (0)
df <- df %>%
  mutate(NONMidFOL = case_when(cyclephase_LH=="1" ~ "1", #periov
                                    cyclephase_LH=="2" ~ "1", #earlylut
                                    cyclephase_LH=="3" ~ "1", #midlut
                                    cyclephase_LH=="4" ~ "1", #perimens
                                    cyclephase_LH=="5" ~ "0")) #midfol


# New vars = NONPERI, NONMid, NONEL, NONPERIOV, NONMidFOL

```

# REVIEW FOR OVERLAPPING CYCLE PHASE CODING
-when using LH-based phasing for all 4 phases, the most likely possibility for overlap is a day that got labeled midfollicular by LH-count and perimenstrual by menses count. *Perimenstrual_count trumps midfollicular by LH count* in this instance. any other overlaps within the dummy code are likely miscoded or an abnormal cycle.
```{r}

#check overlaps: goal is to only have 6 possible group combinations here (1 row per dummy coded phase, then a row where ALL are NA)

df %>%
  group_by(periov_LH, earlylut_LH, midluteal_LH, perimenstrual_LH, midfol_LH) %>%
  summarize(n=n())

#make note of errors - Tory 7/9/24
# peri and midfol = 9
# midlut and peri- 87
# earlylut and peri = 3
# periov and perimens = 64

#example code to review midfol_LH and perimenstrual_LH overlaps, if they exist

df %>%
  filter(midfol_LH==1) %>%
  filter(perimenstrual_LH==1) %>%
  select(id, date_rated, LHposdayfirst, daycountLH,
         mensdayonefirst, cycleday,
         midfol_LH, periov_LH, earlylut_LH,
         midluteal_LH, perimenstrual_LH) %>% View()

#in CLEAR studies, we are looking to see if the day in question is at the early end of the follicular phase as estimated by counting the days prior to ovulation
#example: if the day in question is -7 by LH count and +2 by menses count, this was a short follicular phase, and is probably OK to be included, but needs to be recoded as "perimenstrual" (see below)



```

# N=1 - MIFOL AND PERI --> REVIEWING AND FIXING OVERLAPS
```{r}
#View overlapping DAYS Dataset (N=9: )

df %>% filter(midfol_LH==1 & perimenstrual_LH==1) %>% 
  select(id, date_rated, daycountLH, cycleday, midfol_LH, perimenstrual_LH) %>% View()

# Output a list of IDs with overlap
midfol_peri_overlaps <- df %>% filter(midfol_LH==1 & perimenstrual_LH==1) %>% 
  distinct(id) %>% 
  pull(id)

print(midfol_peri_overlaps)

# 7/9/24: 216, 222, 241, 251, 324

# I will replace all of these such that peri wins

df  <- df  %>%
  mutate(midfol_LH = case_when(perimenstrual_LH==1 ~ 0,
                                        TRUE ~ midfol_LH)) 


# Filter the dataframe to check for leftover overlaps
filtered_df <- df %>% 
  filter(midfol_LH == 1 & perimenstrual_LH == 1)

# Output a list of ids that still overlap
id_list <- filtered_df %>% pull(id)
id_list

```

# NONE - MIDLUTEAL AND PERI --> REVIEWING AND FIXING OVERLAPS
```{r}
#review midlut and perimens overlaps - can override midluteal with perimens IF luteal phase is short but still seems normal (such as 9-11 days)

## make notes of any SUPER short luteal phases (like 4-5 days) that you might want to flag for anovulation

#df %>% filter(midluteal_LH==1 & perimenstrual_LH==1) %>% select(id, date_rated, daycountLH, cycleday, midluteal_LH, perimenstrual_LH) %>% View()

#OVERRIDE MIDLUTEALS WITH PERIs:

df  <- df  %>%
  #if perimenstrual by firstdayofperiod and midlut by count after LH test are both 1, perimen wins, make midlut=0
  mutate(midluteal_LH = case_when(perimenstrual_LH==1 ~ 0,
                                        TRUE ~ midluteal_LH)) 

# Filter the dataframe to check for leftover overlaps
filtered_df <- df %>% 
  filter(midfol_LH == 1 & perimenstrual_LH == 1)

# Output a list of ids that still overlap
id_list <- filtered_df %>% pull(id)
id_list

```

# NONE - EARLY-LUTEAL AND PERI --> REVIEWING AND FIXING OVERLAPS
```{r}

df %>% filter(earlylut_LH==1 & perimenstrual_LH==1) %>% 
  select(id, date_rated, mensdayonefirst, LHposdayfirst, daycountLH, cycleday, earlylut_LH, perimenstrual_LH) %>% View()

#df %>% filter(id==222) %>% View()

#good practice to review again and ensure that dummy code is fixed
df %>%
  group_by(periov_LH, earlylut_LH, midluteal_LH, perimenstrual_LH, midfol_LH) %>%
  summarize(n=n())
```

# SKIP: Identify participants who still have overlapping cycle phases
```{r eval=FALSE, include=FALSE}

participants_with_multiple_LH <- df %>%
  rowwise() %>%
  mutate(sum_LH = sum(c(midfol_LH, periov_LH, earlylut_LH, midluteal_LH, perimenstrual_LH) == 1)) %>%
  filter(sum_LH > 1) %>%
  ungroup() %>%
  distinct(id)

# Filter the original dataset to include all observations of the identified participants so you can view their cycle and count data to see where the errors are

review_these <- df %>%
  filter(id %in% participants_with_multiple_LH$id)

review_these %>% select(id, mensdayonefirst, LHposdayfirst, cycleday, daycountLH, midfol_LH, periov_LH, earlylut_LH, midluteal_LH, perimenstrual_LH) %>% View()

# NONE

# Extract the list of IDs with overlaps
id_list <- participants_with_multiple_LH$id

# Print the list of IDs as a comma-separated string
id_list_string <- paste(id_list, collapse = ", ")
cat(id_list_string)

```
 
# Count participants and rows
```{r CountIDs}

# Count how many remaining ids I have in the dataset
unique_id_count <- length(unique(df$id))
# Print the number of unique IDs
print(unique_id_count)

# NUMBER OF ROWS
print(nrow(df))
```

# Calculate "Hybrid" CycleDay 
```{r}
#create hybrid version of cycle day:
df <- df %>%
  mutate(hybridcycleday = case_when((daycountLH >= -7 & daycountLH <= 5) ~ 
                                      paste(as.character(daycountLH), "L", sep = ""),
                                 (cycleday >= -9 & cycleday <= 7) ~ 
                                   paste(as.character(cycleday), "M",  sep = ""),
                                 TRUE ~ NA_character_))

df <- df %>%
  filter(!is.na(hybridcycleday)) %>%
  arrange(id, date_rated)

hybridcycleday_order <- c("-7L",
                       "-6L",
                       "-5L",
                       "-4L",
                       "-3L",
                       "-2L",
                       "-1L",
                       "0L",
                       "1L",
                       "2L",
                       "3L",
                       "4L",
                       "5L",
                       "-9M",
                       "-8M",
                       "-7M",
                       "-6M",
                       "-5M",
                       "-4M",
                       "-3M",
                       "-2M",
                       "-1M",
                       "1M",
                       "2M",
                       "3M",
                       "4M",
                       "5M",
                       "6M",
                       "7M")
```



#FUNCTION TO PLOT EACH PERSON AS A GRAPH WITH SX AND HORMONES
```{r PERSON-PLOT, warning=FALSE}

plot_participant_data <- function(data, participant_id, outcome_var, outcome_label, output_dir) {
  data <- ungroup(data)
  if (!is.null(data$id)) {
    data$id <- as.numeric(as.character(data$id))
  } else {
    stop("Column 'id' not found in the dataset")
  }
  participant_id <- as.numeric(participant_id)
  participant_data <- data %>% filter(id == participant_id)
  if (nrow(participant_data) == 0) {
    stop(paste("No data available for participant ID:", participant_id))
  }
  participant_data$date_rated <- as.Date(participant_data$date_rated, format = "%m/%d/%Y")
  if (all(is.na(participant_data$date_rated))) {
    stop("Error: date_rated column could not be converted to Date type")
  }
  participant_data$date_numeric <- as.numeric(participant_data$date_rated - min(participant_data$date_rated, na.rm = TRUE))
  non_na_counts <- participant_data %>%
    summarise(across(c(E2, P4, LH, !!sym(outcome_var)), ~sum(!is.na(.))), total_days = n())
  safe_scale <- function(x) {
    if (all(is.na(x)) || length(unique(x)) == 1) return(rep(0, length(x)))
    as.vector(scale(x))
  }
  participant_data <- participant_data %>%
    mutate(across(c(E2, P4, LH, !!sym(outcome_var)), 
                  list(z = ~safe_scale(.), z_roll = ~rollapply(safe_scale(.), 3, mean, fill = NA, align = "center")),
                  .names = "{.col}_{.fn}")) %>%
    mutate(across(ends_with("_z_roll"), list(d = ~c(NA, diff(.))), .names = "{.col}_{.fn}"))
  safe_cor <- function(x, y) {
    if (all(is.na(x)) || all(is.na(y)) || length(unique(x)) == 1 || length(unique(y)) == 1) return(NA)
    cor(x, y, use = "pairwise.complete.obs")
  }
  corr_E2 <- safe_cor(participant_data$E2_z_roll, participant_data[[paste0(outcome_var, "_z_roll")]])
  corr_P4 <- safe_cor(participant_data$P4_z_roll, participant_data[[paste0(outcome_var, "_z_roll")]])
  corr_LH <- safe_cor(participant_data$LH_z_roll, participant_data[[paste0(outcome_var, "_z_roll")]])
  corr_dE2 <- safe_cor(participant_data$E2_z_roll_d, participant_data[[paste0(outcome_var, "_z_roll_d")]])
  corr_dP4 <- safe_cor(participant_data$P4_z_roll_d, participant_data[[paste0(outcome_var, "_z_roll_d")]])
  corr_dLH <- safe_cor(participant_data$LH_z_roll_d, participant_data[[paste0(outcome_var, "_z_roll_d")]])
  mondays <- participant_data$date_numeric[wday(participant_data$date_rated) == 2]
  all_y_values <- c(participant_data$P4_z_roll, participant_data$E2_z_roll, 
                    participant_data$LH_z_roll, participant_data[[paste0(outcome_var, "_z_roll")]])
  min_date <- min(participant_data$date_numeric[!is.na(participant_data$date_numeric)])
  max_date <- max(participant_data$date_numeric[!is.na(participant_data$date_numeric)])
  if (is.infinite(min_date) || is.infinite(max_date) || is.na(min_date) || is.na(max_date)) {
    stop("Error: 'min_date' or 'max_date' is invalid")
  }

  p <- ggplot(participant_data, aes(x = date_numeric)) +
    geom_line(aes(y = P4_z_roll, color = "P4"), linewidth = 1) +
    geom_point(aes(y = P4_z_roll), color = "#D62728", size = 3) +
    geom_line(aes(y = E2_z_roll, color = "E2"), linewidth = 1) +
    geom_point(aes(y = E2_z_roll), color = "#1F77B4", size = 3) +
    geom_line(aes(y = LH_z_roll, color = "LH"), linewidth = 1) +
    geom_point(aes(y = LH_z_roll), color = "#2CA02C", size = 3) +
    geom_line(aes(y = .data[[paste0(outcome_var, "_z_roll")]], color = "Clinical Outcome"), linewidth = 2) +
    geom_vline(xintercept = mondays, color = "gray", linetype = "dashed", alpha = 0.5) +
    geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +
    scale_color_manual(values = c("P4" = "#D62728", "E2" = "#1F77B4", "LH" = "#2CA02C", "Clinical Outcome" = "black")) +
    labs(title = paste("Menstrual Cycle Hormones and", outcome_label,"for Participant", participant_id),
         x = "Date", y = paste(outcome_label, "\nZ-scored, 3d Rolling Avg"), color = "Legend") +
    theme_minimal() +
    theme(plot.title = element_text(size = 14, face = "bold"), axis.title = element_text(size = 12),
          axis.text = element_text(size = 10), legend.title = element_blank(), legend.text = element_text(size = 10),
          legend.position = "right", legend.box = "horizontal", legend.margin = margin(0, 0, 10, 0),
          plot.margin = margin(10, 10, 30, 10)) +
    scale_x_continuous(breaks = seq(min_date, max_date, by = 3),
                       labels = format(min(participant_data$date_rated) + seq(min_date, max_date, by = 3), "%b %d"),
                       limits = c(min_date, max_date), expand = c(0.02, 0)) +
    scale_y_continuous(limits = c(min(all_y_values[is.finite(all_y_values)], na.rm = TRUE) - 1,
                                  max(all_y_values[is.finite(all_y_values)], na.rm = TRUE) + 1), expand = c(0.1, 0.1))

  mens_dates <- participant_data$date_numeric[participant_data$mensdayonefirst == 1]
  LH_dates <- participant_data$date_numeric[participant_data$LHposdayfirst == 1]

  for (date in mens_dates) {
    p <- p + annotate("rect", xmin = date, xmax = date + 4, ymin = -Inf, ymax = Inf, fill = "pink", alpha = 0.3)
  }
  for (date in LH_dates) {
    p <- p + annotate("rect", xmin = date, xmax = date + 1, ymin = -Inf, ymax = Inf, fill = "#2CA02C", alpha = 0.3)
  }

  min_y <- min(all_y_values[is.finite(all_y_values)], na.rm = TRUE)
  if (is.infinite(min_y) || is.na(min_y)) min_y <- -1
  min_y <- min_y - 0.5

  p <- p +
    geom_point(data = data.frame(x = mens_dates, y = rep(min_y, length(mens_dates))),
               aes(x = x, y = y), color = "#D62728", shape = 17, size = 4) +
    geom_point(data = data.frame(x = LH_dates, y = rep(min_y, length(LH_dates))),
               aes(x = x, y = y), color = "#2CA02C", shape = 17, size = 4)

  format_corr <- function(corr) {
    if (is.na(corr)) return("N/A")
    return(sprintf("%.2f", corr))
  }

  correlation_text <- sprintf(
    "Bivariate Correlations with %s:\nlevP4: r=%s, derP4: r=%s, levE2: r=%s, derE2: r=%s, levLH: r=%s, derLH: r=%s.",
    outcome_label, format_corr(corr_P4), format_corr(corr_dP4), format_corr(corr_E2),
    format_corr(corr_dE2), format_corr(corr_LH), format_corr(corr_dLH)
  )

  nonmissing_text <- sprintf(
    "Available Data for this Participant:\nP4=%d, E2=%d, LH=%d, %s=%d out of %d days.",
    non_na_counts$P4, non_na_counts$E2, non_na_counts$LH, outcome_label, 
    non_na_counts[[outcome_var]], non_na_counts$total_days
  )

  predictors <- c("P4_z_roll", "P4_z_roll_d", "E2_z_roll", "E2_z_roll_d")
  valid_predictors <- predictors[sapply(predictors, function(var) {
    sum(!is.na(participant_data[[var]])) > 1 && sd(participant_data[[var]], na.rm = TRUE) > 0
  })]

  predictor_labels <- c("P4_z_roll" = "levP4", "P4_z_roll_d" = "derP4",
                        "E2_z_roll" = "levE2", "E2_z_roll_d" = "derE2")

  reg_text <- "Regression results: Not enough valid data for regression."
  if (length(valid_predictors) > 0) {
    formula_str <- paste(outcome_var, "~", paste(valid_predictors, collapse = " + "))
    model <- lm(as.formula(formula_str), data = participant_data)
    model_summary <- summary(model)
    if (!is.null(model_summary$coefficients) && nrow(model_summary$coefficients) > 0) {
      coef_info <- model_summary$coefficients
      format_coef <- function(coef, p_val) {
        if (is.na(p_val)) return(sprintf("%.2f (NA)", coef))
        if (p_val < 0.001) return(sprintf("%.2f***", coef))
        if (p_val < 0.01) return(sprintf("%.2f**", coef))
        if (p_val < 0.05) return(sprintf("%.2f*", coef))
        if (p_val < 0.15) return(sprintf("%.2f+", coef))
        return(sprintf("%.2f", coef))
      }
      reg_text <- paste(
        "Regression results:",
        paste(sapply(valid_predictors, function(var) {
          label <- predictor_labels[[var]]
          if (var %in% rownames(coef_info)) {
            sprintf("%s: %s", label, format_coef(coef_info[var, "Estimate"], coef_info[var, "Pr(>|t|)"]))
          } else {
            sprintf("%s: N/A", label)
          }
        }), collapse = ", "),
        sep = "\n"
      )
    }
  }

  combined_text <- paste(nonmissing_text, correlation_text, reg_text, sep = "\n")
  print(combined_text)
  text_grob <- textGrob(combined_text, x = 0.05, y = 0.5, just = "left", gp = gpar(fontsize = 9))
  final_plot <- arrangeGrob(p, text_grob, ncol = 1, heights = c(4, 1))
  grid.newpage()
  grid.draw(final_plot)
  ggsave(filename = file.path(output_dir, paste0("participant_", participant_id, "_", outcome_var, "_hormplot.png")),
         plot = final_plot, width = 12, height = 8, units = "in")
  grid.draw(final_plot)
}

#TEST

#dailydir <- '/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024_07_09'

#plot_participant_data(df, 208, 'CSS_Fx_Total', "CSS Functional Impact", dailydir)


```

# Create "id_list": list of all remaining participant IDs to loop through
```{r}

# Load necessary library
library(dplyr)

# Assuming df is your dataframe
# Count how many remaining ids you have in the dataset
unique_id_count <- length(unique(df$id))

# Get a list of unique IDs in the dataset
id_list <- unique(df$id)

# Print the number of unique IDs and the list of unique IDs in a sentence
cat("There are", unique_id_count, "unique IDs in the dataset. The IDs are:", paste(id_list, collapse = ", "), ".\n")


#There are 66 unique IDs in the dataset. The IDs are: 203, 205, 206, 208, 209, 213, 214, 218, 221, 223, 224, 225, 227, 228, 230, 232, 233, 235, 237, 239, 241, 242, 243, 245, 246, 248, 249, 251, 252, 254, 256, 258, 259, 260, 261, 262, 263, 265, 266, 267, 268, 270, 271, 272, 273, 275, 277, 278, 279, 281, 283, 284, 286, 287, 289, 290, 293, 305, 307, 310, 312, 314, 317, 318, 321, 324.
```

```{r}

# There are 12 people whose hormones should probably be removed for sensitivity analysis
badhorm_ids <- c(221, 235, 243, 267, 272, 273, 278, 283, 305, 307, 312, 317)
#Filtering for sensitivity analyses: 
# Remove rows with specified IDs
df_sens <- df %>% filter(!(id %in% badhorm_ids))

```

# Count participants and rows - FULL DATASET

```{r}
# Count how many remaining ids I have in the dataset
unique_id_count <- length(unique(df$id))
# Print the number of unique IDs
print(unique_id_count)

# NUMBER OF ROWS
print(nrow(df))
```

# Count participants and rows - SENSITIVITY SMALLER DATASET

```{r}
# Count how many remaining ids I have in the dataset
unique_id_count <- length(unique(df_sens$id))
# Print the number of unique IDs
print(unique_id_count)

# NUMBER OF ROWS
print(nrow(df_sens))
```



# CALLING PERSON GRAPHING FUNCTIONS

## 1 - MAKE ALL CSS_Fx_Total + Hormone GRAPHS - Loop through each participant ID and call the function - CSS_Fx_Total
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}

# Define the output directory - DO THIS FOR EACH DAY YOU RUN IT

dailydir <- '~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024_07_09'

# Loop through each participant ID and call the function - CSS_Fx_Total
for (participant_id in id_list) {
  plot_participant_data(df, participant_id, 'CSS_Fx_Total', "CSS Functional Impact", dailydir)
}


```

## 2 - MAKE ALL HYPIMP + Hormone GRAPHS - Loop through each participant ID and call the function - CSS_HypImp
```{r warning=FALSE}
# Define the output directory - DO THIS FOR EACH DAY YOU RUN IT

dailydir <- '/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024_07_09'


# Loop through each participant ID and call the function - CSS_HypImp
for (participant_id in id_list) {
  plot_participant_data(df, participant_id, 'CSS_HypImp', "CSS Hyperactivity/Impulsivity", dailydir)
}

```

## 3 - MAKE ALL INATT + Hormone GRAPHS - Loop through each participant ID and call the function - CSS_Inatt
```{r warning=FALSE}

# Define the output directory - DO THIS FOR EACH DAY YOU RUN IT

dailydir <- '/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024_07_09'

# Loop through each participant ID and call the function - CSS_Inatt
for (participant_id in id_list) {
  plot_participant_data(df, participant_id, 'CSS_Inatt', "CSS Inattentiveness", dailydir)
}
```

## 4 - MAKE ALL INATT COUNT + Hormone GRAPHS - Loop through each participant ID and call the function
```{r warning=FALSE}

# Define the output directory - DO THIS FOR EACH DAY YOU RUN IT

dailydir <- '/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024_07_09'

# Loop through each participant ID and call the function - CSS_Inatt
for (participant_id in id_list) {
  plot_participant_data(df, participant_id, 'CSS_Inatt_Count', "CSS Inattentive Criteria (Count)", dailydir)
}
```

## 5 - MAKE ALL HYPIMP COUNT + Hormone GRAPHS - Loop through each participant ID and call the function
```{r warning=FALSE}

# Define the output directory - DO THIS FOR EACH DAY YOU RUN IT

dailydir <- '/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024_07_09'

# Loop through each participant ID and call the function
for (participant_id in id_list) {
  plot_participant_data(df, participant_id, 'CSS_HypImp_Count', "CSS Hyperactive/Impulsive Criteria (Count)", dailydir)
}
```


# Temporary Chunk to create the modeling code --INATT example
```{r}
# Check for problematic values in the original variables
df_clean <- df %>%
  filter(!is.na(CSS_Inatt) & !is.nan(CSS_Inatt) & !is.infinite(CSS_Inatt) &
         !is.na(E2) & !is.nan(E2) & !is.infinite(E2) &
         !is.na(P4) & !is.nan(P4) & !is.infinite(P4))

# Scale E2 and P4 by id
df_clean <- df_clean %>%
  group_by(id) %>%
  mutate(E2zd = scale(E2), P4zd = scale(P4)) %>%
  ungroup()

# Create log-transformed CSS_Inatt
df_clean <- df_clean %>%
  mutate(loginatt = log(CSS_Inatt))

# Check for problematic values after transformation
df_clean <- df_clean %>%
  filter(!is.na(loginatt) & !is.nan(loginatt) & !is.infinite(loginatt) &
         !is.na(E2zd) & !is.nan(E2zd) & !is.infinite(E2zd) &
         !is.na(P4zd) & !is.nan(P4zd) & !is.infinite(P4zd))


# Fit the model with cleaned data
inatt_int <- lmer(loginatt ~ E2zd * P4zd + (1 + E2zd + P4zd | id), data = df_clean)

summary(inatt_int)

# Create a grid of values for E2zd and P4zd
E2zd_seq <- seq(min(df_clean$E2zd), max(df_clean$E2zd), length.out = 100)
P4zd_seq <- seq(min(df_clean$P4zd), max(df_clean$P4zd), length.out = 100)

# Create a dataframe with all combinations of E2zd and P4zd
grid_df <- expand.grid(E2zd = E2zd_seq, P4zd = P4zd_seq)

# Predict loginatt over the grid
grid_df$predicted_loginatt <- predict(inatt_1, newdata = grid_df, re.form = NA)

# Plot the interaction
ggplot(grid_df, aes(x = E2zd, y = predicted_loginatt, color = P4zd, group = P4zd)) +
  geom_line() +
  labs(title = "Interaction between E2zd and P4zd on (log) Inattentiveness",
       x = "E2zd",
       y = "Predicted Inattentiveness (log)",
       color = "P4zd") +
  theme_minimal() +
  scale_color_gradient(low = "blue", high = "red")

```

# Create a function and run outcomes through
```{r}

# Define the function
plot_interaction <- function(data, outcome, dailydir) {
  # Check for problematic values in the specified variables
  data_clean <- data %>%
    filter(!is.na(.data[[outcome]]) & !is.nan(.data[[outcome]]) & !is.infinite(.data[[outcome]]) &
           !is.na(E2) & !is.nan(E2) & !is.infinite(E2) &
           !is.na(P4) & !is.nan(P4) & !is.infinite(P4))
  
  # Scale predictors by random effect and calculate daily derivatives
  data_clean <- data_clean %>%
    group_by(id) %>%
    mutate(
      E2zd = scale(E2),
      P4zd = scale(P4),
      E2ch = c(NA, diff(E2)),  # Daily derivative for E2
      P4ch = c(NA, diff(P4))   # Daily derivative for P4
    ) %>%
    filter(!is.na(E2zd) & !is.nan(E2zd) & !is.infinite(E2zd) &
           !is.na(P4zd) & !is.nan(P4zd) & !is.infinite(P4zd) &
           !is.na(E2ch) & !is.nan(E2ch) & !is.infinite(E2ch) &
           !is.na(P4ch) & !is.nan(P4ch) & !is.infinite(P4ch)) %>%
    ungroup()
  
  # Create log-transformed outcome
  data_clean <- data_clean %>%
    mutate(log_outcome = log(.data[[outcome]])) %>%
    filter(!is.na(log_outcome) & !is.nan(log_outcome) & !is.infinite(log_outcome))
  
  # Fit the models with cleaned data
  rawmodel <- lmer(as.formula(paste(outcome, "~ E2zd * P4zd + (1 | id)")), data = data_clean)
  logmodel <- lmer(log_outcome ~ E2zd * P4zd + (1 | id), data = data_clean)
  model_prand <- lmer(log_outcome ~ E2zd * P4zd + (1 + P4zd | id), data = data_clean)
  model_erand <- lmer(log_outcome ~ E2zd * P4zd + (1 + E2zd | id), data = data_clean)
  model_derivs <- lmer(log_outcome ~ E2zd * P4zd + E2ch + P4ch + (1 | id), data = data_clean)
  
  # Print the model summaries
  # print(summary(rawmodel))
  # print(summary(logmodel))
  # print(summary(model_prand))
  # print(summary(model_erand))
  # print(summary(model_derivs))
  
  # Specify the file path for the HTML output - primary model of E2 * P4
  output_file <- file.path(dailydir, paste(outcome, "horm_tabmodel_output.html", sep = ""))
  cat("Output file path:", output_file, "\n")
  
  # Generate the HTML output from tab_model without saving to a file
  tab_model(rawmodel, logmodel, model_prand, model_erand, model_derivs)
  tab_model_output <- tab_model(rawmodel, logmodel, model_prand, model_erand, model_derivs)
  
  # Capture the HTML output as a string
  html_output <- as.character(tab_model_output)
  
  # Manually write the HTML output to the specified file
  writeLines(html_output, con = output_file)
  
  # Create a grid of values for predictors
  E2zd_seq <- seq(min(data_clean$E2zd), max(data_clean$E2zd), length.out = 100)
  P4zd_seq <- seq(min(data_clean$P4zd), max(data_clean$P4zd), length.out = 100)
  
  # Create a dataframe with all combinations of E2zd and P4zd
  grid_df <- expand.grid(E2zd = E2zd_seq, P4zd = P4zd_seq)
  
  # Add the mean values of the derivatives to the grid_df
  grid_df$E2ch <- mean(data_clean$E2ch, na.rm = TRUE)
  grid_df$P4ch <- mean(data_clean$P4ch, na.rm = TRUE)
  
  # Predict the outcome over the grid
  grid_df$predicted_log_outcome <- predict(logmodel, newdata = grid_df, re.form = NA)
  
  # Exponentiate the predicted values to return them to the original scale
  grid_df$predicted_outcome <- exp(grid_df$predicted_log_outcome)
  
  # Plot the interaction
 library(ggplot2)

# Create the interaction plot
intplot <- ggplot(grid_df, aes(x = E2zd, y = predicted_outcome, color = P4zd, group = P4zd)) +
    geom_line() +
    labs(
        title = paste("Interaction between E2zd and P4zd on", outcome),
        x = "E2zd",
        y = paste("Predicted", outcome),
        color = "P4zd"
    ) +
    theme_minimal(base_size = 15) +
    scale_color_gradient(low = "blue", high = "red") +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        panel.background = element_rect(fill = "white", color = NA), # White background
        plot.background = element_rect(fill = "white", color = NA),  # White background
        panel.grid.major = element_line(color = "gray90"), # Light gray major grid lines
        panel.grid.minor = element_line(color = "gray95")  # Lighter gray minor grid lines
    )

  # Save and return the plot
  ggsave(filename = file.path(dailydir, paste0("EPintplot_", outcome, ".tif")), plot = intplot, width = 12, height = 8, units = "in")
  
  return(intplot)
}


```


```{r}

dailydir <- '~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024_07_09'

# Example usage
plot_interaction(df_sens, "CSS_Inatt", dailydir)
plot_interaction(df_sens, "CSS_Inatt_Count", dailydir) *SIG

plot_interaction(df_sens, "CSS_HypImp", dailydir)
plot_interaction(df_sens, "CSS_HypImp_Count", dailydir)

plot_interaction(df_sens, "CSS_Fx_Total", dailydir)

plot_interaction(df_sens, "CSS_B2_Total", dailydir)

plot_interaction(df_sens, "UPPS_Total", dailydir)

plot_interaction(df_sens, "BDEFS_Total", dailydir)

 plot_interaction(df_sens, "DEBQ_Total", dailydir)

```







# Function to plot by cycle day ("plot_by_cycle_day")
```{r}
plot_by_cycle_day <- function(df, var) {
#Create individual overall means

df <- df %>%
group_by(id) %>%
mutate(var.roll.m = mean(.data[[var]], na.rm = TRUE))

dfcycdayplotdata <- df %>%
group_by(id, hybridcycleday) %>%
filter(!is.na(hybridcycleday)) %>%
summarize(
var.roll.m = mean(var.roll.m, na.rm = TRUE),
var.day.roll.m = mean(.data[[var]], na.rm = TRUE),
var.day.roll.d = var.day.roll.m - var.roll.m
) %>%
arrange(factor(hybridcycleday, levels = hybridcycleday_order))

#print(dfcycdayplotdata)

cycd_plot <- dfcycdayplotdata %>%
group_by(hybridcycleday) %>%
summarize(
personcentered.dev = mean(var.day.roll.d, na.rm = TRUE),
se = sd(var.day.roll.d, na.rm = TRUE) / sqrt(n())
) %>%
arrange(factor(hybridcycleday, levels = hybridcycleday_order)) %>%
ggplot(aes(x = factor(hybridcycleday, levels = hybridcycleday_order), y = personcentered.dev)) +
geom_line(group = 1) +
#geom_ribbon(aes(ymin = personcentered.dev - se, ymax = personcentered.dev + se), alpha = 0.2) +
geom_hline(yintercept = 0, linetype = "solid", col = "grey") +

geom_errorbar(aes(ymin = personcentered.dev - se, ymax = personcentered.dev + se), size = 0.5, width = 0.5, position = position_dodge()) +
#scale_y_continuous(expand = c(0, 0.011)) +
theme_bw() +
xlab("Cycle Day\n(Relative to LH-surge and Menses Onset)") +
ylab(paste("Person-Centered", var)) +

  #Midfollicular Label
annotate("rect", ymin = .35, ymax = .45, xmin = "-7L", xmax = "-3L", alpha = .2) +
annotate("text", x = "-7L", y = .4, label = "Midfollicular", size = 3, hjust = -0.25) +
  
  #Periovulatory Label
annotate("rect", ymin = .35, ymax = .45, xmin = "-2L", xmax = "1L", alpha = .2) +
annotate("text", x = "-2L", y = .4, label = "Periovulatory", size = 3, hjust = 0) +
  
  #Early Luteal Label
annotate("rect", ymin = .35, ymax = .45, xmin = "2L", xmax = "5L", alpha = .2) +
annotate("text", x = "2L", y = .4, label = " Early Luteal", size = 3, hjust = 0) +
 
   #Midluteal Label
annotate("rect", ymin = .35, ymax = .45, xmin = "-9M", xmax = "-5M", alpha = .2) +
annotate("text", x = "-9M", y = .4, label = "Midluteal", size = 3, hjust = -0.5) +

  #Perimenstrual Label
annotate("rect", ymin = .35, ymax = .45, xmin = "-3M", xmax = "2M", alpha = .2) +
annotate("text", x = "-3M", y = .4, label = "Perimenstrual", size = 3, hjust = -0.15) +
 
   #Annotation of Ovulation
annotate("text", x = "0L", y = -.25, label = "*", size = 8, color = "forestgreen") +
 
   #Annotation of Menses
annotate("text", x = "1M", y = -.25, label = "*", size = 8, color = "red")
print(cycd_plot)
}

```

# Function to plot by cycle phase ("plot_by_cycle_phase")
```{r warning=FALSE}

plot_by_cycle_phase <- function(df, var) {
  df$cyclephase_LH_graph <- as.factor(df$cyclephase_LH_graph)
  
  # Calculate overall mean for each individual
  df <- df %>%
    group_by(id) %>%
    mutate(overall_mean = mean(.data[[var]], na.rm = TRUE)) %>%
    ungroup()
  
  # Print intermediate data for debugging
  # print(head(df))
  
  # Calculate phase-specific deviation
  dfphaseplotdata <- df %>%
    filter(!is.na(cyclephase_LH_graph)) %>%
    group_by(id, cyclephase_LH_graph) %>%
    summarize(
      phase_mean = mean(.data[[var]], na.rm = TRUE),
      overall_mean = first(overall_mean)
    ) %>%
    mutate(phase_deviation = phase_mean - overall_mean) %>%
    ungroup()
  
  # Print intermediate data for debugging
  # print(head(dfphaseplotdata))
  
  # Calculate grand mean and standard error
  phase_plot_data <- dfphaseplotdata %>%
    group_by(cyclephase_LH_graph) %>%
    summarize(
      grand_mean_deviation = mean(phase_deviation, na.rm = TRUE),
      se = sd(phase_deviation, na.rm = TRUE) / sqrt(n())
    )
  
  # Print summarized data for debugging
 # print(head(phase_plot_data))
  
  # Create the plot
  phase_plot <- phase_plot_data %>%
    ggplot(aes(x = cyclephase_LH_graph, y = grand_mean_deviation, group = 1)) +
    geom_line() + # Add the line
    scale_x_discrete(labels = c(
      "1" = "Midfollicular",
      "2" = "Periovulatory", 
      "3" = "Early Luteal",
      "4" = "Midluteal",
      "5" = "Perimenstrual"
    )) +
    geom_ribbon(aes(ymin = grand_mean_deviation - se, ymax = grand_mean_deviation + se), alpha = 0.2) +
    geom_hline(yintercept = 0.0, color = "grey", linetype = "solid") +
    theme_bw(base_size = 11) +
    xlab("Cycle Phase") +
    ylab(paste("Person-Centered", var))
  
  print(phase_plot)
}


```


# FUNCTION TO RUN&SAVE PHASE & CYCLEDAY GRAPHS TO BOX
```{r}

# Define the function
run_and_save_plots <- function(variable_name, df, output_directory) {
  
  # Ensure the output directory exists
  dir.create(output_directory, recursive = TRUE, showWarnings = FALSE)
  
  # Create and save Cycle Day Plot
  day_plot <- plot_by_cycle_day(df, paste0(variable_name, ".roll.d"))
  ggsave(file.path(output_directory, paste0(variable_name, "_DAYPLOT.png")), 
         plot = day_plot, width = 10, height = 6, dpi = 300)
  
  # Create and save Cycle Phase Plot
  phase_plot <- plot_by_cycle_phase(df, paste0(variable_name, ".roll.d"))
  ggsave(file.path(output_directory, paste0(variable_name, "_PHSPLOT.png")), 
         plot = phase_plot, width = 10, height = 6, dpi = 300)
}


```


# Function to run and save MLMs to html in dated Box output folder ("run_and_save_mlms")
```{r message=TRUE, warning=TRUE}

run_and_save_mlms <- function(variable_name, data, output_directory) {
  ## peri
  model_peri <- lmer(as.formula(paste(variable_name, "~ midfol_LH + periov_LH + earlylut_LH + midluteal_LH + (1 | id)")),
                     data = data)
  cat("Peri model fitted.\n")
  
  ## midfol
  model_midfol <- lmer(as.formula(paste(variable_name, "~ periov_LH + earlylut_LH + midluteal_LH + perimenstrual_LH + (1 | id)")),
                       data = data)
  cat("Midfol model fitted.\n")
  
  ## periov
  model_periov <- lmer(as.formula(paste(variable_name, "~ midfol_LH + earlylut_LH + midluteal_LH + perimenstrual_LH + (1 | id)")),
                       data = data)
  cat("Periov model fitted.\n")
  
  ## earlylut
  model_earlylut <- lmer(as.formula(paste(variable_name, "~ midfol_LH + periov_LH + midluteal_LH + perimenstrual_LH + (1 | id)")),
                         data = data)
  cat("Earlylut model fitted.\n")
  
  ## midlut
  model_midlut <- lmer(as.formula(paste(variable_name, "~ midfol_LH + periov_LH + earlylut_LH + perimenstrual_LH + (1 | id)")),
                       data = data)
  cat("Midlut model fitted.\n")
  
  # Create the directory if it doesn't exist
  if (!dir.exists(output_directory)) {
    dir.create(output_directory, recursive = TRUE, showWarnings = FALSE)
    cat("Directory created:", output_directory, "\n")
  } else {
    cat("Directory already exists:", output_directory, "\n")
  }
  
  # Specify the file path for the HTML output
  output_file <- file.path(output_directory, paste(variable_name, "_PHS_tab_model_output.html", sep = ""))
  cat("Output file path:", output_file, "\n")
  
  # Generate the HTML output from tab_model without saving to a file
  tab_model_output <- tab_model(model_midfol, model_periov, model_earlylut, model_midlut, model_peri)
  
  # Capture the HTML output as a string
  html_output <- as.character(tab_model_output)
  
  # Manually write the HTML output to the specified file
  writeLines(html_output, con = output_file)
  
  # Check if the file was created and has content
  if (file.exists(output_file)) {
    cat("HTML output generated and saved to:", output_file, "\n")
  } else {
    cat("Failed to save HTML output to:", output_file, "\n")
  }
}

# Run for all Outcomes

output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"

df <- df

```

# CALL PLOTTING FUNCTIONS FOR HORMONES
```{r}

#E2
plot_by_cycle_day(df_sens, "E2")
plot_by_cycle_phase(df_sens, "E2")

#P4
plot_by_cycle_day(df_sens, "P4")
plot_by_cycle_phase(df_sens, "P4")

#LH
plot_by_cycle_day(df_sens, "LH")
plot_by_cycle_phase(df_sens, "LH")


```

# CALL FUNCTIONS: E2

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("E2", df, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "E2"

#### NO NEED TO CHANGE ANYTHING BELOW

df <- df  # Replace with your actual dataframe
output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df, output_directory)

```
# CALL FUNCTIONS: P4

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("P4", df, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "P4"

#### NO NEED TO CHANGE ANYTHING BELOW

df <- df  # Replace with your actual dataframe
output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df, output_directory)

```

# CALL FUNCTIONS: LH

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("LH", df, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "LH"

#### NO NEED TO CHANGE ANYTHING BELOW

df <- df  # Replace with your actual dataframe
output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-04"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df, output_directory)

```


# CALL FUNCTIONS: CSS_Inatt

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("CSS_Inatt", df_sens, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "CSS_Inatt"

#### NO NEED TO CHANGE ANYTHING BELOW

output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df_sens, output_directory)

```

# CALL FUNCTIONS: CSS_HypImp

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("CSS_HypImp", df_sens, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "CSS_HypImp"

#### NO NEED TO CHANGE ANYTHING BELOW

output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df_sens, output_directory)

```

# CALL FUNCTIONS: CSS_B2_Total

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("CSS_B2_Total", df_sens, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "CSS_B2_Total"

#### NO NEED TO CHANGE ANYTHING BELOW

output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df_sens, output_directory)

```

# CALL FUNCTIONS: CSS_Fx_Total

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("CSS_Fx_Total", df_sens, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "CSS_Fx_Total"

#### NO NEED TO CHANGE ANYTHING BELOW

output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df_sens, output_directory)

```


CSS_Inatt_Count, CSS_Hyp_Count, CSS_Imp_Count, DEBQ_Total, BDEFS_Total, UPPS_Total


# CALL FUNCTIONS: CSS_Inatt_Count

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("CSS_Inatt_Count", df_sens, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "CSS_Inatt_Count"

#### NO NEED TO CHANGE ANYTHING BELOW

output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df_sens, output_directory)

```



# CALL FUNCTIONS: CSS_Imp_Count

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("CSS_HypImp_Count", df_sens, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "CSS_HypImp_Count"

#### NO NEED TO CHANGE ANYTHING BELOW

df <- df  # Replace with your actual dataframe
output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df_sens, output_directory)

```




# CALL FUNCTIONS: BDEFS_Total

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("BDEFS_Total", df_sens, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "BDEFS_Total"

#### NO NEED TO CHANGE ANYTHING BELOW

df <- df  # Replace with your actual dataframe
output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df_sens, output_directory)

```



# CALL FUNCTIONS: DEBQ_Total

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("DEBQ_Total", df_sens, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "DEBQ_Total"

#### NO NEED TO CHANGE ANYTHING BELOW

df <- df  # Replace with your actual dataframe
output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df_sens, output_directory)

```


# CALL FUNCTIONS: UPPS_Total

```{r}

#SPECIFY VARIABLE NAME FOR MODELS FUNCTION 

run_and_save_mlms("UPPS_Total", df_sens, output_directory)

# SPECIFY VARIABLE NAME FOR GRAPHING FUNCTION

variable_name <- "UPPS_Total"

#### NO NEED TO CHANGE ANYTHING BELOW

df <- df  # Replace with your actual dataframe
output_directory <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-07-09"
# Run the plotting function (Phase, Day)
run_and_save_plots(variable_name, df_sens, output_directory)

```




