---
title: "ADHDCYCLE_5PhaseCoding"
author: "Tory Eisenlohr-Moul"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Cycle Phase Coding - MENSES COUNT PART
reminder: dummy code comes from Schmalenberger 2021 appendix
```{r cyclephase_count}

df <- df %>% 
  mutate(midluteal_count = ifelse(cycleday >= -9 & cycleday <= -5, 1, 0),
         perimenstrual_count = ifelse(cycleday >=-3 & cycleday <=2, 1, 0),
         midfol_count = ifelse(cycleday >=4 & cycleday <=7, 1, 0),
         periov_count = ifelse(cycleday >= -15 & cycleday <= -12, 1, 0))

#add a non-dummy coded variable with all of cycle phase info above, which will be useful for future visualizations and summaries
df <- df %>% mutate(cyclephase_count = case_when(periov_count==1 ~ 1,
                                                     midluteal_count==1 ~ 2,
                                                     perimenstrual_count==1 ~ 3,
                                                     midfol_count==1 ~ 4,
                                                     TRUE ~ as.numeric(NA)))

#Check vars
#df %>% dplyr::select(id, mensdayonefirst, LHposdayfirst, cycleday, daycountLH, ends_with("_count")) %>% View()

```

# Cycle Phase Coding - LH COUNT PART
```{r cyclephase_LH}

#note: the perimenstrual phase is still coded based on menses count variable, but for the purpose of keeping the dummy code labeling clear, we name it *perimenstrual_LH* to indicate that this variable gets included in the dummy code set of LH-based phasing; again, all of this information is from Schmalenberger 2021 appendix

df <- df %>% 
  mutate(midluteal_LH = ifelse(daycountLH >= 6 & daycountLH <= 11, 1, 0),
         perimenstrual_LH = ifelse(cycleday >=-3 & cycleday <=2, 1, 0), #note perimenstrual is based on cycleday
         midfol_LH = ifelse(daycountLH >=-7 & daycountLH <=-3, 1, 0),
         periov_LH = ifelse(daycountLH >= -2 & daycountLH <= 1, 1, 0),
         earlylut_LH = ifelse(daycountLH >= 2 & daycountLH <= 5, 1, 0))

#fill out the dummy code, so that if any phase is 1, all others should be 0 (not NA)
df <- df %>%
  rowwise() %>%
  mutate(sumdummy = sum(midfol_LH,
                        periov_LH,
                        earlylut_LH,
                        perimenstrual_LH,
                        midluteal_LH, na.rm = T)) %>% 
  #if all phases=0, make the midluteal_LH variable NA instead of 0,
  #if perimenstrual by menses count = 1, change midluteal_LH to 0 instead of NA to fill out structural set,
  #otherwise, keep as is
  mutate(midluteal_LH = case_when(sumdummy==0 ~ as.numeric(NA),
                                  sumdummy==1 & perimenstrual_LH==1 ~ 0,
                                  TRUE ~ midluteal_LH), 
         #same for midfol
         midfol_LH = case_when(sumdummy==0 ~ as.numeric(NA),
                               sumdummy==1 & perimenstrual_LH==1 ~ 0,
                               TRUE ~ midfol_LH), 
          #same for periov
         periov_LH = case_when(sumdummy==0 ~ as.numeric(NA),
                               sumdummy==1 & perimenstrual_LH==1 ~ 0,
                               TRUE ~ periov_LH),
         earlylut_LH = case_when(sumdummy==0 ~ as.numeric(NA),
                                  sumdummy==1 & perimenstrual_LH==1 ~ 0,
                                  TRUE ~ earlylut_LH), 
         #if any other phase is 1, fill out perimenstrual to be 0
         perimenstrual_LH = case_when(sumdummy==0 ~ as.numeric(NA), 
                                        sumdummy==1 & midfol_LH==1 ~ 0,
                                        sumdummy==1 & periov_LH==1 ~ 0,
                                      sumdummy==1 & earlylut_LH==1 ~ 0,
                                        sumdummy==1 & midluteal_LH==1 ~ 0,
                                        TRUE ~ perimenstrual_LH))

#add cyclephase_LH, a new categorical variable made from all of the new hybrid cycle phases above, which will be useful for future visualizations and summaries
df <- df %>% mutate(cyclephase_LH = case_when(periov_LH==1 ~ 1,
                                                                        earlylut_LH==1 ~ 2,
                                                     midluteal_LH==1 ~ 3,
                                                     perimenstrual_LH==1 ~ 4,
                                                     midfol_LH==1 ~ 5,
                                                     TRUE ~ as.numeric(NA)))

df <- df %>% mutate(cyclephase_LH_graph = case_when(midfol_LH==1 ~ 1,
                                                    periov_LH==1 ~ 2,
                                                    earlylut_LH==1 ~ 3,
                                                     midluteal_LH==1 ~ 4,
                                                     perimenstrual_LH==1 ~ 5,
                                                     TRUE ~ as.numeric(NA)))

#Check vars
#df %>% dplyr::select(id, mensdayonefirst, LHposdayfirst, cycleday, daycountLH, ends_with("_LH")) %>% View()



#Coding additional binary variables for use in random effects that contrast each phase with all other phases for simplicity

#nonperi (1) vs. peri (0)
df <- df %>%
  mutate(NONPERI = case_when(cyclephase_LH=="1" ~ "1", #periov
                                    cyclephase_LH=="2" ~ "1", #earlylut
                                    cyclephase_LH=="3" ~ "1", #midlut
                                    cyclephase_LH=="4" ~ "0", #perimens
                                    cyclephase_LH=="5" ~ "1")) #midfol

#nonmid (1) vs. mid (0)
df <- df %>%
  mutate(NONMid = case_when(cyclephase_LH=="1" ~ "1", #periov
                                    cyclephase_LH=="2" ~ "1", #earlylut
                                    cyclephase_LH=="3" ~ "0", #midlut
                                    cyclephase_LH=="4" ~ "1", #perimens
                                    cyclephase_LH=="5" ~ "1")) #midfol

#nonel (1) vs. el (0)
df <- df %>%
  mutate(NONEL = case_when(cyclephase_LH=="1" ~ "1", #periov
                                    cyclephase_LH=="2" ~ "0", #earlylut
                                    cyclephase_LH=="3" ~ "1", #midlut
                                    cyclephase_LH=="4" ~ "1", #perimens
                                    cyclephase_LH=="5" ~ "1")) #midfol

#nonperiov (1) vs. periov (0)
df <- df %>%
  mutate(NONPERIOV = case_when(cyclephase_LH=="1" ~ "0", #periov
                                    cyclephase_LH=="2" ~ "1", #earlylut
                                    cyclephase_LH=="3" ~ "1", #midlut
                                    cyclephase_LH=="4" ~ "1", #perimens
                                    cyclephase_LH=="5" ~ "1")) #midfol

#nonmidfol (1) vs. midfol (0)
df <- df %>%
  mutate(NONMidFOL = case_when(cyclephase_LH=="1" ~ "1", #periov
                                    cyclephase_LH=="2" ~ "1", #earlylut
                                    cyclephase_LH=="3" ~ "1", #midlut
                                    cyclephase_LH=="4" ~ "1", #perimens
                                    cyclephase_LH=="5" ~ "0")) #midfol


# New vars = NONPERI, NONMid, NONEL, NONPERIOV, NONMidFOL

```

# REVIEW FOR OVERLAPPING CYCLE PHASE CODING
-when using LH-based phasing for all 4 phases, the most likely possibility for overlap is a day that got labeled midfollicular by LH-count and perimenstrual by menses count. *Perimenstrual_count trumps midfollicular by LH count* in this instance. any other overlaps within the dummy code are likely miscoded or an abnormal cycle.
```{r}

#check overlaps: goal is to only have 6 possible group combinations here (1 row per dummy coded phase, then a row where ALL are NA)

df %>%
  group_by(periov_LH, earlylut_LH, midluteal_LH, perimenstrual_LH, midfol_LH) %>%
  summarize(n=n())

#make note of errors - Tory 7/9/24
# peri and midfol = 9
# midlut and peri- 87
# earlylut and peri = 3
# periov and perimens = 64

#example code to review midfol_LH and perimenstrual_LH overlaps, if they exist

#BROKEN

df %>%
  filter(midfol_LH==1) %>%
  filter(perimenstrual_LH==1) %>%
  dplyr::select(id, date_rated, LHposdayfirst, daycountLH,
         mensdayonefirst, cycleday,
         midfol_LH, periov_LH, earlylut_LH,
         midluteal_LH, perimenstrual_LH) %>% View()

#in CLEAR studies, we are looking to see if the day in question is at the early end of the follicular phase as estimated by counting the days prior to ovulation
#example: if the day in question is -7 by LH count and +2 by menses count, this was a short follicular phase, and is probably OK to be included, but needs to be recoded as "perimenstrual" (see below)



```

# N=1 - MIFOL AND PERI --> REVIEWING AND FIXING OVERLAPS
```{r}
#View overlapping DAYS Dataset (N=9: )

df %>% filter(midfol_LH==1 & perimenstrual_LH==1) %>% 
  dplyr::select(id, date_rated, daycountLH, cycleday, midfol_LH, perimenstrual_LH) %>% View()

# Output a list of IDs with overlap
midfol_peri_overlaps <- df %>% filter(midfol_LH==1 & perimenstrual_LH==1) %>% 
  distinct(id) %>% 
  pull(id)

print(midfol_peri_overlaps)

# 7/9/24: 216, 222, 241, 251, 324

# I will replace all of these such that peri wins

df  <- df  %>%
  mutate(midfol_LH = case_when(perimenstrual_LH==1 ~ 0,
                                        TRUE ~ midfol_LH)) 


# Filter the dataframe to check for leftover overlaps
filtered_df <- df %>% 
  filter(midfol_LH == 1 & perimenstrual_LH == 1)

# Output a list of ids that still overlap
id_list <- filtered_df %>% pull(id)
id_list

```

# NONE - MIDLUTEAL AND PERI --> REVIEWING AND FIXING OVERLAPS
```{r}

#review midlut and perimens overlaps - can override midluteal with perimens IF luteal phase is short but still seems normal (such as 9-11 days)

## make notes of any SUPER short luteal phases (like 4-5 days) that you might want to flag for anovulation

#df %>% filter(midluteal_LH==1 & perimenstrual_LH==1) %>% dplyr::select(id, date_rated, daycountLH, cycleday, midluteal_LH, perimenstrual_LH) %>% View()

#OVERRIDE MIDLUTEALS WITH PERIs:

df  <- df  %>%
  #if perimenstrual by firstdayofperiod and midlut by count after LH test are both 1, perimen wins, make midlut=0
  mutate(midluteal_LH = case_when(perimenstrual_LH==1 ~ 0,
                                        TRUE ~ midluteal_LH)) 

# Filter the dataframe to check for leftover overlaps
filtered_df <- df %>% 
  filter(midfol_LH == 1 & perimenstrual_LH == 1)

# Output a list of ids that still overlap
id_list <- filtered_df %>% pull(id)
id_list

```

# NONE - EARLY-LUTEAL AND PERI --> REVIEWING AND FIXING OVERLAPS
```{r}

df %>% filter(earlylut_LH==1 & perimenstrual_LH==1) %>% 
  dplyr::select(id, date_rated, mensdayonefirst, LHposdayfirst, daycountLH, cycleday, earlylut_LH, perimenstrual_LH) %>% View()

#df %>% filter(id==222) %>% View()

#good practice to review again and ensure that dummy code is fixed

df %>%
  group_by(periov_LH, earlylut_LH, midluteal_LH, perimenstrual_LH, midfol_LH) %>%
  summarize(n=n())
```

# Identify participants who still have overlapping cycle phases
```{r}


participants_with_multiple_LH <- df %>%
  rowwise() %>%
  mutate(sum_LH = sum(c(midfol_LH, periov_LH, earlylut_LH, midluteal_LH, perimenstrual_LH) == 1)) %>%
  filter(sum_LH > 1) %>%
  ungroup() %>%
  distinct(id)

# Filter the original dataset to include all observations of the identified participants so you can view their cycle and count data to see where the errors are

review_these <- df %>%
  filter(id %in% participants_with_multiple_LH$id)

review_these %>% dplyr::select(id, mensdayonefirst, LHposdayfirst, cycleday, daycountLH, midfol_LH, periov_LH, earlylut_LH, midluteal_LH, perimenstrual_LH) %>% View()

# NONE

# Extract the list of IDs with overlaps
id_list <- participants_with_multiple_LH$id

# Print the list of IDs as a comma-separated string
id_list_string <- paste(id_list, collapse = ", ")
cat(id_list_string)

```
 

# Calculate "Hybrid" CycleDay 
```{r}

#create hybrid version of cycle day:
df <- df %>%
  mutate(hybridcycleday = case_when((daycountLH >= -7 & daycountLH <= 5) ~ 
                                      paste(as.character(daycountLH), "L", sep = ""),
                                 (cycleday >= -9 & cycleday <= 7) ~ 
                                   paste(as.character(cycleday), "M",  sep = ""),
                                 TRUE ~ NA_character_))

df <- df %>%
  filter(!is.na(hybridcycleday)) %>%
  arrange(id, date_rated)

hybridcycleday_order <- c("-7L",
                       "-6L",
                       "-5L",
                       "-4L",
                       "-3L",
                       "-2L",
                       "-1L",
                       "0L",
                       "1L",
                       "2L",
                       "3L",
                       "4L",
                       "5L",
                       "-9M",
                       "-8M",
                       "-7M",
                       "-6M",
                       "-5M",
                       "-4M",
                       "-3M",
                       "-2M",
                       "-1M",
                       "1M",
                       "2M",
                       "3M",
                       "4M",
                       "5M",
                       "6M",
                       "7M")
```
