---
title: "ADHDCYCLE_dailyprep"
output: 
  html_document: 
    toc: true
    fig_caption: true
    number_sections: true
    df_print: default
date: "Last compiled on `r format(Sys.time(), '%B %d, %Y')`"
---

# Setup

```{r Packages}

knitr::opts_chunk$set(echo = TRUE)

library(dplyr) # Data manipulation
library(tidyr) # Tidying data
library(haven) # Reading files (SPSS, stata, SAS)
library(ggplot2) # Plotting
library(tidyverse) # Data management and manipulation
library(zoo) # Rolling averages
library(readxl) # Reading Excel files
library(lme4) # Mixed-effects models
library(lmerTest) # Mixed-effects models with p-values
library(emmeans) # Estimated marginal means
library(psych) # Descriptive statistics
library(lubridate) # Date and time manipulation
library(visdat) # Visualizing missing data
library(janitor) # Cleaning data
library(skimr) # Skimming data for summary statistics
library(ggdist) # Visualizing distributions and uncertainty
library(ggforce) # Extending ggplot2
library(sjPlot) # Plotting for statistical models
library(grid) # Base graphics
library(gridExtra) # Arranging grid-based graphics
library(usethis) # Setting up GitHub and other project tools
library(nlme) # Linear and nonlinear mixed-effects models
library(performance) # Model performance metrics
library(rmcorr) # Repeated measures correlations
library(gitcreds) # Managing Git credentials
library(see) # Visualizing model checking functions
library(corrplot) # Visualizing correlations
library(GPArotation) # Factor analysis rotations
library(broom.mixed) # Tidying mixed models
library(pbkrtest) # Parametric bootstrap and Kenward-Roger methods
library(ggrepel) # Adding text labels to ggplot2
library(zoo)
library(careless)
library(responsePatterns)
library(dplyr)
library(tidyverse)
library(readxl)
library(tidyverse)
library(readxl)
library(ggplot2)
library(dplyr)
library(tidyr)
library(haven)

```

# Data
```{r Load Data}

## Raw daily data, merged to include surveys and hormones 

df <- read.csv("~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/02_datasets/CYCADHD_DAILY/01_raw_data/2024.04.24. Daily Master.csv")

```

```{r}
 # Convert date_rated to Date type
  df$date_rated <- as.Date(mdy(df$date_rated))

```

## Remove NA rows
```{r}
#remove rows where all data is NA
df <- df[rowSums(is.na(df)) != ncol(df), ]

#remove NA ID's from dataset
df <- df %>% filter(!is.na(id))

#remove NA daterated from dataset 
df <- df %>% filter(!is.na(date_rated))

# NUMBER OF ROWS

print(nrow(df))

```

## Sort by id and date_rated and save
```{r}
df <- df %>% arrange(id, date_rated, .na.last = FALSE)

##View(df)
```


#Hard-code changes to issues found later 
```{r}

#Add 208 Menses start date 
df$StartPeriod[df$id == 208 & df$TubeNumber == 27] <- 1
#df %>% filter(id==208) %>% View()

df$date_rated[df$id == 210 & df$date_rated == 11/20/2021] <- 11/20/2020
df$date_rated[df$id == 210 & df$date_rated == 11/21/2021] <- 11/21/2020
#df %>% filter(id==210) %>% View()

##View(df)

#Template Code to look at one pt's data

#df %>% filter(id==307) %>% View()

#ADDITIONAL DATA REVIEW NEEDED FOR HORMONES: 

# Filtering template: df <- df %>% filter(!is.na(id))

#Note that 307's ovulation might be a little off with their hormones
#Consider excluding participant 235
#Review participant 273's progesterone data
#Review participant 278's progesterone data
#Review participant 317's estrogen data
#Review participant 221's hormone data
#Review participant 243's hormone data
#Review participant 267's hormone data
#Review participant 272's hormone data
#Review participant 283's hormone data
#Review participant 305's hormone data
#Review participant 312's hormone data

badhorm_ids <- c(221, 235, 243, 267, 272, 273, 278, 283, 305, 307, 312, 317)

```


# Add row placeholders for missing survey/date_rated's
```{r}

df <- df %>% group_by(id) %>%
  complete(date_rated = seq.Date(min(date_rated), max(date_rated), by = "day"))

##View(df)

# NUMBER OF ROWS
print(nrow(df))

```

# Rename Hormones
```{r RENAME}
# Rename the columns
colnames(df)[colnames(df) == "Estradiol"] <- "E2"
colnames(df)[colnames(df) == "Progesterone"] <- "P4"
```

## Calculate BAARS ADHD Current Symptom Scale CSS Scores, UPPS, BDEFS, DEBQ
```{r calc CSS UPPS BDEFS DEBQ}

#CSS_Total - Overall total score on the current symptoms scale
#CSS_TotalSymp - Total ADHD symptom score (inattention + hyperactivity-impulsivity)
#CSS_B_TotalIASymp - Total inattention symptoms score 
#CSS_B_TotalHISymp - Total hyperactivity-impulsivity symptoms score 

#CSS_TotalCount - Total count of all ADHD symptoms (inattention + hyperactivity-impulsivity)
#CSS_Inatt_Count - Count of inattention symptoms
#CSS_Imp_Count - Count of impulsivity symptoms
#CSS_Hyp_Count - Count of hyperactivity symptoms

#CSS_Fx_Total - ???????
#CSS_B2_Total - ???????

# Reverse UPPS Items 7 and 8
df <- df %>%
  mutate(UPPS_7r=recode(UPPS_7, '1' = 4, '2' = 3, '3' = 2, '4' = 1), UPPS_8r=recode(UPPS_8, '1' = 4, '2' = 3, '3' = 2, '4' = 1))

# Drop original items UPPS_7 and UPPS_8

df <- df %>% select(-UPPS_7)
df <- df %>% select(-UPPS_8)


df <- df %>%
  group_by(id) %>%
  mutate(
    # CSS Symptom severity Scores
    
    # Calculate Inattention Severity: 
    CSS_Inatt = rowSums(across(c(CSS_B_1, CSS_B_3, CSS_B_5, CSS_B_7, CSS_B_9, CSS_B_11, CSS_B_13, CSS_B_15, CSS_B_17)), na.rm = TRUE),
    
    # Calculate Hyperactivity/Impulsivity Severity: 
    CSS_HypImp = rowSums(across(c(CSS_B_2, CSS_B_4, CSS_B_6, CSS_B_8, CSS_B_10, CSS_B_12, CSS_B_14, CSS_B_16, CSS_B_18)), na.rm = TRUE),
    
    # Function items 1-10
    CSS_Fx_Total = rowSums(across(starts_with("CSS_Function_")), na.rm = TRUE),
    
    # Not sure what this is (todo)
    CSS_B2_Total = rowSums(across(starts_with("CSS_B2_")), na.rm = TRUE),
    
    # CSS Count Variables
    CSS_Inatt_Count = rowSums(across(c(CSS_1_Count, CSS_3_Count, CSS_5_Count, CSS_7_Count, CSS_9_Count, CSS_11_Count, CSS_13_Count, CSS_15_Count, CSS_17_Count)), na.rm = TRUE),
    
    CSS_HypImp_Count = rowSums(across(c(CSS_2_Count, CSS_4_Count, CSS_6_Count, CSS_8_Count, CSS_10_Count, CSS_12_Count, CSS_14_Count, CSS_16_Count, CSS_18_Count)), na.rm = TRUE),
    
    # Other Daily Scored Scales
    DEBQ_Total = rowSums(across(starts_with("DEBQ_")), na.rm = TRUE),
    BDEFS_Total = rowSums(across(starts_with("BDEFS_")), na.rm = TRUE),
    UPPS_Total = rowSums(across(starts_with("UPPS_")), na.rm = TRUE)
  ) %>%
  ungroup()

```

# SCREENING FOR CARELESS RESPONDING OR RESPONSE SETS
```{r}

# The Longstring Index:

df <- df %>% mutate(longstring = longstring(df))
df <- df %>% mutate(avgstring = (longstring(df, avg = T)))


#df %>% arrange(desc(longstring)) %>% select(id, date_rated, longstring, avgstring) %>% View()


careless_long <- longstring(df)
boxplot(careless_long, main = "Boxplot of Longstring index")

#careless_long <- longstring(df, avg = T)
#boxplot(careless_long$avg, main = "Boxplot of Averagestring index")

# The IRV: The intra-individual response variability (IRV) is similiar in spirit to the Longstring index. It is defined as the “standard deviation of responses across a set of consecutive item responses for an individual” (Dunn et al. 2018).

#careless_irv <- irv(df, split = TRUE, num.split = 4)
#head(careless_irv)

```

### Create functions to create person means and person deviations
```{r echo=T}

#function to create person means
create.person.mean <- function(df, var, ...) {
  df %>%
    group_by(...) %>%
    mutate("{{var}}.m" := mean({{var}}, na.rm=T))
}
#function to create person deviations (note, must have person means already made)
create.deviation <- function(df, var, var.m) {
  df <- df %>%
    rowwise() %>%
    mutate("{{var}}.d" := {{var}} - {{var.m}})
}

create.3day.rolling.avg <- function(df, var, num) {
  df %>%
    group_by(id) %>%
    mutate("{{var}}.roll" := rollapply({{var}}, 3,  mean, align="center", fill=NA))
}

```


## create outcome lists to loop through functions
```{r echo=T}

#person-centered deviations
outcomelist <- c("E2",
           "P4",
           "LH",
           "CSS_Fx_Total",
           "CSS_B2_Total",
           "CSS_Inatt",
           "CSS_HypImp",
           "CSS_Inatt_Count",
           "CSS_HypImp_Count",
           "BDEFS_Total", 
           "UPPS_Total",
          "longstring",
           "DEBQ_Total") %>% noquote()

#person-centered deviations
outcomelist.d <- c("E2.d",
           "P4.d",
           "LH.d",
           "CSS_Fx_Total.d",
           "CSS_B2_Total.d",
           "CSS_Inatt.d",
          "CSS_Inatt_Count.d",
           "CSS_HypImp.d",
           "CSS_HypImp_Count.d",
           "BDEFS_Total.d", 
           "UPPS_Total.d",
          "longstring.d",
           "DEBQ_Total.d"
           ) %>% noquote()

#rolling averages
outcomelist.roll <- c("E2.roll",
"P4.roll",
"LH.roll",
"CSS_Fx_Total.roll",
"CSS_B2_Total.roll",
"CSS_Inatt.roll",
"CSS_HypImp.roll",
"CSS_Inatt_Count.roll",
"CSS_HypImp_Count.roll",
"BDEFS_Total.roll", 
           "UPPS_Total.roll",
"longstring.roll",
           "DEBQ_Total.roll"
) %>% noquote()

#View(df)


```

```{r}

#ROLLING AVGS

#create rolling averages on RAW variables
for (i in outcomelist) {
  df <- create.3day.rolling.avg(df, !!sym({{i}}), 3)
}

#MEANS

#execute for loop: run create.person.mean() on everything in "outcomelist.roll"
for (i in outcomelist.roll) {
  df <- create.person.mean(df, !!sym({{i}}), id)
}

#execute for loop: run create.person.mean() on everything in "outcomelist"
for (i in outcomelist) {
  df <- create.person.mean(df, !!sym({{i}}), id)
}

#PERSON-DEVIATIONS

#for raw deviations
#execute for loop: run create.deviation() on everything in list
for (i in outcomelist) {
  df <- create.deviation(df, !!sym({{i}}), !!sym(paste0({{i}}, ".m")))
}

#for rolling avgs deviations
#execute for loop: run create.deviation() on everything in list
for (i in outcomelist.roll) {
  df <- create.deviation(df, !!sym({{i}}), !!sym(paste0({{i}}, ".m")))
}


#Check 
df %>% select(id, E2, E2.m, E2.d, E2.roll, E2.roll.d) %>% View()

```

# Print Variable Names
```{r printvars}
variable_names <- names(df)
formatted_list <- paste(variable_names, collapse = ", ")
cat(formatted_list)
```

# Create dichotomous mensdayone/mensdayonefirst to indicate first day of period
```{r }


str(df$id) #check id format
str(df$date_rated) #check date_rated format

df <- df %>%
  filter(!is.na(id))

df <- df %>%
  arrange(id, date_rated)


df$StartPeriod <- as.numeric(df$StartPeriod)

df <- df %>%
  mutate(mensdayone = case_when(
    is.na(StartPeriod) ~ NA_real_,
    StartPeriod == 1 ~ 1,
    TRUE ~ 0
  ))


# Eliminate all but the first menses onset day in a new variable called "mensdayonefirst"

# Process the data
df <- df %>%
  arrange(id, date_rated) %>%  # Arrange the data by id and date_rated
  group_by(id) %>%  # Group the data by id
  mutate(
    mensdayonefirst = case_when(
      is.na(mensdayone) ~ NA,
      mensdayone == 1 & (lag(mensdayone, order_by = date_rated) != 1 | is.na(lag(mensdayone, order_by = date_rated))) ~ 1,
      row_number() == 1 & mensdayone == 1 ~ 1,
      TRUE ~ 0
    )
  ) %>% 
  ungroup() # Remove the grouping

#Check

#df %>% select(id, date_rated, mensdayone, mensdayonefirst, PosLHTest) %>% View()


```

# Create "LHposday" (0/1/NA) that represents the positive LH day.
```{r }

###View(df)
  
  df$PosLHTest <- as.numeric(df$PosLHTest)

df <- df %>%
  mutate(LHposday = case_when(
    is.na(PosLHTest) ~ NA_real_,
    PosLHTest == 1 ~ 1,
    TRUE ~ 0
  ))

###View(df)

# Eliminate all but the first pos Ov day in a new variable called "LHposdayfirst"
df <- df %>%
  arrange(id, date_rated) %>%  # Arrange the data by id and date_rated
  group_by(id) %>%  # Group the data by id
  mutate(
    LHposdayfirst = case_when(
      is.na(LHposday) ~ NA_real_,
      LHposday == 1 & lag(LHposday, order_by = date_rated) != 1 ~ 1,
      TRUE ~ 0
    )
  ) %>% 
  ungroup()  # Remove the grouping

#Check vars
#df %>% select(id, date_rated, mensdayonefirst, LHposdayfirst) %>% View()

```

# Remove people who never had a positive LH test
```{r}

# Create a person-level variable
idlhpos <- df %>%
  group_by(id) %>% 
  summarise(LHeverpos = ifelse(sum(LHposdayfirst, na.rm = TRUE) > 0, 1, 0)) %>% 
  ungroup()

# Merge it back into df
df <- df %>%
  left_join(idlhpos, by = "id")

# Filter out people with LHids = 0
df <- df %>%
  filter(LHeverpos == 1)

#Check vars
df %>% select(id, LHeverpos, date_rated, mensdayonefirst, LHposdayfirst) %>% head()

```

# Code Cycle Day Forward + Backward count from menses onset
```{r cycleday}

#make A as a variable copy of mensdayonefirst
df$A <- df$mensdayonefirst

cycleCount <- function(x) {
  #Get the index of 1
  inds <- which(x == 1)
  if(!length(inds)) return(0) 
  #Create a sequence with that index as 0
  num <- lapply(inds, function(i) {
    num <- seq_along(x) - i
    #Add 1 to values greater than equal to 0
    num[num >= 0] <- num[num >= 0] + 1
    num[num < -15 | num > 11] <- NA 
    num
  })
  #Select the first non-NA values from the sequence
  do.call(coalesce, num)
}

df <- df %>% group_by(id) %>% 
  mutate(cycleday = cycleCount(A))

#df <- df %>% select(!A)

###View(df)

#Check vars
#df %>% select(id, LHeverpos, date_rated, mensdayonefirst, LHposdayfirst, cycleday) %>% View()

```

# Review Data, looking for cycleday=0 and removing those who never had any menses onsets
```{r}

#View in a separate window and manually scroll through
df %>% select(id, date_rated, mensdayonefirst, cycleday, LHposdayfirst, cycleday) %>% View()

#how many observations do you have for each cycle day?
nobs <- df %>% group_by(cycleday) %>% summarize(n=n())

##View(nobs)

#do you have any observations that got assigned a cycle day of 0? this is wrong, and you should check who that happened for
df %>% filter(cycleday==0) %>% pull(id)

# NOTE 2024-07-09: There are 2 people with lots of zeroes (see Jordan's note above in chunk) and it looks to me like they did not report a menses in the daily dataset: 331, 332. They have no menses onset dates

#View in a separate window and manually scroll through
#df %>% select(id, date_rated, mensdayonefirst, cycleday, LHposdayfirst, cycleday) %>% View()

# List of IDs to remove
ids_to_remove <- c(331, 332)

# Remove rows with specified IDs
df <- df %>% filter(!(id %in% ids_to_remove))


```

# Make daycountLH, based on positive LH test (LH test day = 0)
```{r daycountLH}
#make A a new temp column where pt received pos ov test
df$L <- df$LHposdayfirst

#FUNCTION for calculating the sequence
LHCount <- function(x) {
  #Get the index of 1
  inds <- which(x == 1)
  if(!length(inds)) return(0) 
  #Create a sequence with that index as 0
  num <- lapply(inds, function(i) {
    num <- seq_along(x) - i
    num[num < -7 | num > 15] <- NA 
    num
  })
  #Select the first non-NA values from the sequence
  do.call(coalesce, num)
}

#run the LHCount function and save it as as a new column called daycountLH
df<- df %>% group_by(id) %>% 
  mutate(daycountLH = LHCount(L))
#remove the temp column L
df <- df %>% select(!L)

#Check vars
#df %>% select(id, LHeverpos, date_rated, mensdayonefirst, LHposdayfirst, daycountLH) %>% View()
```

# View the number of observations for each cycleday
```{r daycountLH checking}

#how many daily ratings got assigned each cycle day?
nobs2 <- df %>% group_by(daycountLH) %>% summarize(n=n())

#View(nobs2)

```

# Cycle Phase Coding - MENSES COUNT PART
reminder: dummy code comes from Schmalenberger 2021 appendix
```{r cyclephase_count}

df <- df %>% 
  mutate(midluteal_count = ifelse(cycleday >= -9 & cycleday <= -5, 1, 0),
         perimenstrual_count = ifelse(cycleday >=-3 & cycleday <=2, 1, 0),
         midfol_count = ifelse(cycleday >=4 & cycleday <=7, 1, 0),
         periov_count = ifelse(cycleday >= -15 & cycleday <= -12, 1, 0))

#add a non-dummy coded variable with all of cycle phase info above, which will be useful for future visualizations and summaries
df <- df %>% mutate(cyclephase_count = case_when(periov_count==1 ~ 1,
                                                     midluteal_count==1 ~ 2,
                                                     perimenstrual_count==1 ~ 3,
                                                     midfol_count==1 ~ 4,
                                                     TRUE ~ as.numeric(NA)))

#Check vars
#df %>% select(id, mensdayonefirst, LHposdayfirst, cycleday, daycountLH, ends_with("_count")) %>% View()

```

# Cycle Phase Coding - LH COUNT PART
```{r cyclephase_LH}

#note: the perimenstrual phase is still coded based on menses count variable, but for the purpose of keeping the dummy code labeling clear, we name it *perimenstrual_LH* to indicate that this variable gets included in the dummy code set of LH-based phasing; again, all of this information is from Schmalenberger 2021 appendix

df <- df %>% 
  mutate(midluteal_LH = ifelse(daycountLH >= 6 & daycountLH <= 11, 1, 0),
         perimenstrual_LH = ifelse(cycleday >=-3 & cycleday <=2, 1, 0), #note perimenstrual is based on cycleday
         midfol_LH = ifelse(daycountLH >=-7 & daycountLH <=-3, 1, 0),
         periov_LH = ifelse(daycountLH >= -2 & daycountLH <= 1, 1, 0),
         earlylut_LH = ifelse(daycountLH >= 2 & daycountLH <= 5, 1, 0))

#fill out the dummy code, so that if any phase is 1, all others should be 0 (not NA)
df <- df %>%
  rowwise() %>%
  mutate(sumdummy = sum(midfol_LH,
                        periov_LH,
                        earlylut_LH,
                        perimenstrual_LH,
                        midluteal_LH, na.rm = T)) %>% 
  #if all phases=0, make the midluteal_LH variable NA instead of 0,
  #if perimenstrual by menses count = 1, change midluteal_LH to 0 instead of NA to fill out structural set,
  #otherwise, keep as is
  mutate(midluteal_LH = case_when(sumdummy==0 ~ as.numeric(NA),
                                  sumdummy==1 & perimenstrual_LH==1 ~ 0,
                                  TRUE ~ midluteal_LH), 
         #same for midfol
         midfol_LH = case_when(sumdummy==0 ~ as.numeric(NA),
                               sumdummy==1 & perimenstrual_LH==1 ~ 0,
                               TRUE ~ midfol_LH), 
          #same for periov
         periov_LH = case_when(sumdummy==0 ~ as.numeric(NA),
                               sumdummy==1 & perimenstrual_LH==1 ~ 0,
                               TRUE ~ periov_LH),
         earlylut_LH = case_when(sumdummy==0 ~ as.numeric(NA),
                                  sumdummy==1 & perimenstrual_LH==1 ~ 0,
                                  TRUE ~ earlylut_LH), 
         #if any other phase is 1, fill out perimenstrual to be 0
         perimenstrual_LH = case_when(sumdummy==0 ~ as.numeric(NA), 
                                        sumdummy==1 & midfol_LH==1 ~ 0,
                                        sumdummy==1 & periov_LH==1 ~ 0,
                                      sumdummy==1 & earlylut_LH==1 ~ 0,
                                        sumdummy==1 & midluteal_LH==1 ~ 0,
                                        TRUE ~ perimenstrual_LH))

#add cyclephase_LH, a new categorical variable made from all of the new hybrid cycle phases above, which will be useful for future visualizations and summaries
df <- df %>% mutate(cyclephase_LH = case_when(periov_LH==1 ~ 1,
                                                                        earlylut_LH==1 ~ 2,
                                                     midluteal_LH==1 ~ 3,
                                                     perimenstrual_LH==1 ~ 4,
                                                     midfol_LH==1 ~ 5,
                                                     TRUE ~ as.numeric(NA)))

df <- df %>% mutate(cyclephase_LH_graph = case_when(midfol_LH==1 ~ 1,
                                                    periov_LH==1 ~ 2,
                                                    earlylut_LH==1 ~ 3,
                                                     midluteal_LH==1 ~ 4,
                                                     perimenstrual_LH==1 ~ 5,
                                                     TRUE ~ as.numeric(NA)))

#Check vars
#df %>% select(id, mensdayonefirst, LHposdayfirst, cycleday, daycountLH, ends_with("_LH")) %>% View()



#Coding additional binary variables for use in random effects that contrast each phase with all other phases for simplicity

#nonperi (1) vs. peri (0)
df <- df %>%
  mutate(NONPERI = case_when(cyclephase_LH=="1" ~ "1", #periov
                                    cyclephase_LH=="2" ~ "1", #earlylut
                                    cyclephase_LH=="3" ~ "1", #midlut
                                    cyclephase_LH=="4" ~ "0", #perimens
                                    cyclephase_LH=="5" ~ "1")) #midfol

#nonmid (1) vs. mid (0)
df <- df %>%
  mutate(NONMid = case_when(cyclephase_LH=="1" ~ "1", #periov
                                    cyclephase_LH=="2" ~ "1", #earlylut
                                    cyclephase_LH=="3" ~ "0", #midlut
                                    cyclephase_LH=="4" ~ "1", #perimens
                                    cyclephase_LH=="5" ~ "1")) #midfol

#nonel (1) vs. el (0)
df <- df %>%
  mutate(NONEL = case_when(cyclephase_LH=="1" ~ "1", #periov
                                    cyclephase_LH=="2" ~ "0", #earlylut
                                    cyclephase_LH=="3" ~ "1", #midlut
                                    cyclephase_LH=="4" ~ "1", #perimens
                                    cyclephase_LH=="5" ~ "1")) #midfol

#nonperiov (1) vs. periov (0)
df <- df %>%
  mutate(NONPERIOV = case_when(cyclephase_LH=="1" ~ "0", #periov
                                    cyclephase_LH=="2" ~ "1", #earlylut
                                    cyclephase_LH=="3" ~ "1", #midlut
                                    cyclephase_LH=="4" ~ "1", #perimens
                                    cyclephase_LH=="5" ~ "1")) #midfol

#nonmidfol (1) vs. midfol (0)
df <- df %>%
  mutate(NONMidFOL = case_when(cyclephase_LH=="1" ~ "1", #periov
                                    cyclephase_LH=="2" ~ "1", #earlylut
                                    cyclephase_LH=="3" ~ "1", #midlut
                                    cyclephase_LH=="4" ~ "1", #perimens
                                    cyclephase_LH=="5" ~ "0")) #midfol


# New vars = NONPERI, NONMid, NONEL, NONPERIOV, NONMidFOL

```

# REVIEW FOR OVERLAPPING CYCLE PHASE CODING
-when using LH-based phasing for all 4 phases, the most likely possibility for overlap is a day that got labeled midfollicular by LH-count and perimenstrual by menses count. *Perimenstrual_count trumps midfollicular by LH count* in this instance. any other overlaps within the dummy code are likely miscoded or an abnormal cycle.
```{r}

#check overlaps: goal is to only have 6 possible group combinations here (1 row per dummy coded phase, then a row where ALL are NA)

df %>%
  group_by(periov_LH, earlylut_LH, midluteal_LH, perimenstrual_LH, midfol_LH) %>%
  summarize(n=n())

#make note of errors - Tory 7/9/24
# peri and midfol = 9
# midlut and peri- 87
# earlylut and peri = 3
# periov and perimens = 64

#example code to review midfol_LH and perimenstrual_LH overlaps, if they exist

df %>%
  filter(midfol_LH==1) %>%
  filter(perimenstrual_LH==1) %>%
  select(id, date_rated, LHposdayfirst, daycountLH,
         mensdayonefirst, cycleday,
         midfol_LH, periov_LH, earlylut_LH,
         midluteal_LH, perimenstrual_LH) %>% View()

#in CLEAR studies, we are looking to see if the day in question is at the early end of the follicular phase as estimated by counting the days prior to ovulation
#example: if the day in question is -7 by LH count and +2 by menses count, this was a short follicular phase, and is probably OK to be included, but needs to be recoded as "perimenstrual" (see below)



```

# N=1 - MIFOL AND PERI --> REVIEWING AND FIXING OVERLAPS
```{r}
#View overlapping DAYS Dataset (N=9: )

df %>% filter(midfol_LH==1 & perimenstrual_LH==1) %>% 
  select(id, date_rated, daycountLH, cycleday, midfol_LH, perimenstrual_LH) %>% View()

# Output a list of IDs with overlap
midfol_peri_overlaps <- df %>% filter(midfol_LH==1 & perimenstrual_LH==1) %>% 
  distinct(id) %>% 
  pull(id)

print(midfol_peri_overlaps)

# 7/9/24: 216, 222, 241, 251, 324

# I will replace all of these such that peri wins

df  <- df  %>%
  mutate(midfol_LH = case_when(perimenstrual_LH==1 ~ 0,
                                        TRUE ~ midfol_LH)) 


# Filter the dataframe to check for leftover overlaps
filtered_df <- df %>% 
  filter(midfol_LH == 1 & perimenstrual_LH == 1)

# Output a list of ids that still overlap
id_list <- filtered_df %>% pull(id)
id_list

```

# NONE - MIDLUTEAL AND PERI --> REVIEWING AND FIXING OVERLAPS
```{r}
#review midlut and perimens overlaps - can override midluteal with perimens IF luteal phase is short but still seems normal (such as 9-11 days)

## make notes of any SUPER short luteal phases (like 4-5 days) that you might want to flag for anovulation

#df %>% filter(midluteal_LH==1 & perimenstrual_LH==1) %>% select(id, date_rated, daycountLH, cycleday, midluteal_LH, perimenstrual_LH) %>% View()

#OVERRIDE MIDLUTEALS WITH PERIs:

df  <- df  %>%
  #if perimenstrual by firstdayofperiod and midlut by count after LH test are both 1, perimen wins, make midlut=0
  mutate(midluteal_LH = case_when(perimenstrual_LH==1 ~ 0,
                                        TRUE ~ midluteal_LH)) 

# Filter the dataframe to check for leftover overlaps
filtered_df <- df %>% 
  filter(midfol_LH == 1 & perimenstrual_LH == 1)

# Output a list of ids that still overlap
id_list <- filtered_df %>% pull(id)
id_list

```

# NONE - EARLY-LUTEAL AND PERI --> REVIEWING AND FIXING OVERLAPS
```{r}

df %>% filter(earlylut_LH==1 & perimenstrual_LH==1) %>% 
  select(id, date_rated, mensdayonefirst, LHposdayfirst, daycountLH, cycleday, earlylut_LH, perimenstrual_LH) %>% View()

#df %>% filter(id==222) %>% View()

#good practice to review again and ensure that dummy code is fixed

df %>%
  group_by(periov_LH, earlylut_LH, midluteal_LH, perimenstrual_LH, midfol_LH) %>%
  summarize(n=n())
```

# Identify participants who still have overlapping cycle phases
```{r eval=FALSE, include=FALSE}

participants_with_multiple_LH <- df %>%
  rowwise() %>%
  mutate(sum_LH = sum(c(midfol_LH, periov_LH, earlylut_LH, midluteal_LH, perimenstrual_LH) == 1)) %>%
  filter(sum_LH > 1) %>%
  ungroup() %>%
  distinct(id)

# Filter the original dataset to include all observations of the identified participants so you can view their cycle and count data to see where the errors are

review_these <- df %>%
  filter(id %in% participants_with_multiple_LH$id)

review_these %>% select(id, mensdayonefirst, LHposdayfirst, cycleday, daycountLH, midfol_LH, periov_LH, earlylut_LH, midluteal_LH, perimenstrual_LH) %>% View()

# NONE

# Extract the list of IDs with overlaps
id_list <- participants_with_multiple_LH$id

# Print the list of IDs as a comma-separated string
id_list_string <- paste(id_list, collapse = ", ")
cat(id_list_string)

```
 
# Count participants and rows
```{r CountIDs}

# Count how many remaining ids I have in the dataset
unique_id_count <- length(unique(df$id))
# Print the number of unique IDs
print(unique_id_count)

# NUMBER OF ROWS
print(nrow(df))
```

# Calculate "Hybrid" CycleDay 
```{r}
#create hybrid version of cycle day:
df <- df %>%
  mutate(hybridcycleday = case_when((daycountLH >= -7 & daycountLH <= 5) ~ 
                                      paste(as.character(daycountLH), "L", sep = ""),
                                 (cycleday >= -9 & cycleday <= 7) ~ 
                                   paste(as.character(cycleday), "M",  sep = ""),
                                 TRUE ~ NA_character_))

df <- df %>%
  filter(!is.na(hybridcycleday)) %>%
  arrange(id, date_rated)

hybridcycleday_order <- c("-7L",
                       "-6L",
                       "-5L",
                       "-4L",
                       "-3L",
                       "-2L",
                       "-1L",
                       "0L",
                       "1L",
                       "2L",
                       "3L",
                       "4L",
                       "5L",
                       "-9M",
                       "-8M",
                       "-7M",
                       "-6M",
                       "-5M",
                       "-4M",
                       "-3M",
                       "-2M",
                       "-1M",
                       "1M",
                       "2M",
                       "3M",
                       "4M",
                       "5M",
                       "6M",
                       "7M")
```






#FUNCTION TO PLOT EACH PERSON AS A GRAPH WITH SX AND HORMONES
```{r PERSON-PLOT, warning=FALSE}

plot_participant_data <- function(data, participant_id, outcome_var, outcome_label, output_dir) {
  data <- ungroup(data)
  if (!is.null(data$id)) {
    data$id <- as.numeric(as.character(data$id))
  } else {
    stop("Column 'id' not found in the dataset")
  }
  participant_id <- as.numeric(participant_id)
  participant_data <- data %>% filter(id == participant_id)
  if (nrow(participant_data) == 0) {
    stop(paste("No data available for participant ID:", participant_id))
  }
  participant_data$date_rated <- as.Date(participant_data$date_rated, format = "%m/%d/%Y")
  if (all(is.na(participant_data$date_rated))) {
    stop("Error: date_rated column could not be converted to Date type")
  }
  participant_data$date_numeric <- as.numeric(participant_data$date_rated - min(participant_data$date_rated, na.rm = TRUE))
  non_na_counts <- participant_data %>%
    summarise(across(c(E2, P4, LH, !!sym(outcome_var)), ~sum(!is.na(.))), total_days = n())
  safe_scale <- function(x) {
    if (all(is.na(x)) || length(unique(x)) == 1) return(rep(0, length(x)))
    as.vector(scale(x))
  }
  participant_data <- participant_data %>%
    mutate(across(c(E2, P4, LH, !!sym(outcome_var)), 
                  list(z = ~safe_scale(.), z_roll = ~rollapply(safe_scale(.), 3, mean, fill = NA, align = "center")),
                  .names = "{.col}_{.fn}")) %>%
    mutate(across(ends_with("_z_roll"), list(d = ~c(NA, diff(.))), .names = "{.col}_{.fn}"))
  safe_cor <- function(x, y) {
    if (all(is.na(x)) || all(is.na(y)) || length(unique(x)) == 1 || length(unique(y)) == 1) return(NA)
    cor(x, y, use = "pairwise.complete.obs")
  }
  corr_E2 <- safe_cor(participant_data$E2_z_roll, participant_data[[paste0(outcome_var, "_z_roll")]])
  corr_P4 <- safe_cor(participant_data$P4_z_roll, participant_data[[paste0(outcome_var, "_z_roll")]])
  corr_LH <- safe_cor(participant_data$LH_z_roll, participant_data[[paste0(outcome_var, "_z_roll")]])
  corr_dE2 <- safe_cor(participant_data$E2_z_roll_d, participant_data[[paste0(outcome_var, "_z_roll_d")]])
  corr_dP4 <- safe_cor(participant_data$P4_z_roll_d, participant_data[[paste0(outcome_var, "_z_roll_d")]])
  corr_dLH <- safe_cor(participant_data$LH_z_roll_d, participant_data[[paste0(outcome_var, "_z_roll_d")]])
  mondays <- participant_data$date_numeric[wday(participant_data$date_rated) == 2]
  all_y_values <- c(participant_data$P4_z_roll, participant_data$E2_z_roll, 
                    participant_data$LH_z_roll, participant_data[[paste0(outcome_var, "_z_roll")]])
  min_date <- min(participant_data$date_numeric[!is.na(participant_data$date_numeric)])
  max_date <- max(participant_data$date_numeric[!is.na(participant_data$date_numeric)])
  if (is.infinite(min_date) || is.infinite(max_date) || is.na(min_date) || is.na(max_date)) {
    stop("Error: 'min_date' or 'max_date' is invalid")
  }

  p <- ggplot(participant_data, aes(x = date_numeric)) +
    geom_line(aes(y = P4_z_roll, color = "P4"), linewidth = 1) +
    geom_point(aes(y = P4_z_roll), color = "#D62728", size = 3) +
    geom_line(aes(y = E2_z_roll, color = "E2"), linewidth = 1) +
    geom_point(aes(y = E2_z_roll), color = "#1F77B4", size = 3) +
    geom_line(aes(y = LH_z_roll, color = "LH"), linewidth = 1) +
    geom_point(aes(y = LH_z_roll), color = "#2CA02C", size = 3) +
    geom_line(aes(y = .data[[paste0(outcome_var, "_z_roll")]], color = "Clinical Outcome"), linewidth = 2) +
    geom_vline(xintercept = mondays, color = "gray", linetype = "dashed", alpha = 0.5) +
    geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +
    scale_color_manual(values = c("P4" = "#D62728", "E2" = "#1F77B4", "LH" = "#2CA02C", "Clinical Outcome" = "black")) +
    labs(title = paste("Menstrual Cycle Hormones and", outcome_label,"for Participant", participant_id),
         x = "Date", y = paste(outcome_label, "\nZ-scored, 3d Rolling Avg"), color = "Legend") +
    theme_minimal() +
    theme(plot.title = element_text(size = 14, face = "bold"), axis.title = element_text(size = 12),
          axis.text = element_text(size = 10), legend.title = element_blank(), legend.text = element_text(size = 10),
          legend.position = "right", legend.box = "horizontal", legend.margin = margin(0, 0, 10, 0),
          plot.margin = margin(10, 10, 30, 10)) +
    scale_x_continuous(breaks = seq(min_date, max_date, by = 3),
                       labels = format(min(participant_data$date_rated) + seq(min_date, max_date, by = 3), "%b %d"),
                       limits = c(min_date, max_date), expand = c(0.02, 0)) +
    scale_y_continuous(limits = c(min(all_y_values[is.finite(all_y_values)], na.rm = TRUE) - 1,
                                  max(all_y_values[is.finite(all_y_values)], na.rm = TRUE) + 1), expand = c(0.1, 0.1))

  mens_dates <- participant_data$date_numeric[participant_data$mensdayonefirst == 1]
  LH_dates <- participant_data$date_numeric[participant_data$LHposdayfirst == 1]

  for (date in mens_dates) {
    p <- p + annotate("rect", xmin = date, xmax = date + 4, ymin = -Inf, ymax = Inf, fill = "pink", alpha = 0.3)
  }
  for (date in LH_dates) {
    p <- p + annotate("rect", xmin = date, xmax = date + 1, ymin = -Inf, ymax = Inf, fill = "#2CA02C", alpha = 0.3)
  }

  min_y <- min(all_y_values[is.finite(all_y_values)], na.rm = TRUE)
  if (is.infinite(min_y) || is.na(min_y)) min_y <- -1
  min_y <- min_y - 0.5

  p <- p +
    geom_point(data = data.frame(x = mens_dates, y = rep(min_y, length(mens_dates))),
               aes(x = x, y = y), color = "#D62728", shape = 17, size = 4) +
    geom_point(data = data.frame(x = LH_dates, y = rep(min_y, length(LH_dates))),
               aes(x = x, y = y), color = "#2CA02C", shape = 17, size = 4)

  format_corr <- function(corr) {
    if (is.na(corr)) return("N/A")
    return(sprintf("%.2f", corr))
  }

  correlation_text <- sprintf(
    "Bivariate Correlations with %s:\nlevP4: r=%s, derP4: r=%s, levE2: r=%s, derE2: r=%s, levLH: r=%s, derLH: r=%s.",
    outcome_label, format_corr(corr_P4), format_corr(corr_dP4), format_corr(corr_E2),
    format_corr(corr_dE2), format_corr(corr_LH), format_corr(corr_dLH)
  )

  nonmissing_text <- sprintf(
    "Available Data for this Participant:\nP4=%d, E2=%d, LH=%d, %s=%d out of %d days.",
    non_na_counts$P4, non_na_counts$E2, non_na_counts$LH, outcome_label, 
    non_na_counts[[outcome_var]], non_na_counts$total_days
  )

  predictors <- c("P4_z_roll", "P4_z_roll_d", "E2_z_roll", "E2_z_roll_d")
  valid_predictors <- predictors[sapply(predictors, function(var) {
    sum(!is.na(participant_data[[var]])) > 1 && sd(participant_data[[var]], na.rm = TRUE) > 0
  })]

  predictor_labels <- c("P4_z_roll" = "levP4", "P4_z_roll_d" = "derP4",
                        "E2_z_roll" = "levE2", "E2_z_roll_d" = "derE2")

  reg_text <- "Regression results: Not enough valid data for regression."
  if (length(valid_predictors) > 0) {
    formula_str <- paste(outcome_var, "~", paste(valid_predictors, collapse = " + "))
    model <- lm(as.formula(formula_str), data = participant_data)
    model_summary <- summary(model)
    if (!is.null(model_summary$coefficients) && nrow(model_summary$coefficients) > 0) {
      coef_info <- model_summary$coefficients
      format_coef <- function(coef, p_val) {
        if (is.na(p_val)) return(sprintf("%.2f (NA)", coef))
        if (p_val < 0.001) return(sprintf("%.2f***", coef))
        if (p_val < 0.01) return(sprintf("%.2f**", coef))
        if (p_val < 0.05) return(sprintf("%.2f*", coef))
        if (p_val < 0.15) return(sprintf("%.2f+", coef))
        return(sprintf("%.2f", coef))
      }
      reg_text <- paste(
        "Regression results:",
        paste(sapply(valid_predictors, function(var) {
          label <- predictor_labels[[var]]
          if (var %in% rownames(coef_info)) {
            sprintf("%s: %s", label, format_coef(coef_info[var, "Estimate"], coef_info[var, "Pr(>|t|)"]))
          } else {
            sprintf("%s: N/A", label)
          }
        }), collapse = ", "),
        sep = "\n"
      )
    }
  }

  combined_text <- paste(nonmissing_text, correlation_text, reg_text, sep = "\n")
  print(combined_text)
  text_grob <- textGrob(combined_text, x = 0.05, y = 0.5, just = "left", gp = gpar(fontsize = 9))
  final_plot <- arrangeGrob(p, text_grob, ncol = 1, heights = c(4, 1))
  grid.newpage()
  grid.draw(final_plot)
  ggsave(filename = file.path(output_dir, paste0("participant_", participant_id, "_", outcome_var, "_hormplot.png")),
         plot = final_plot, width = 12, height = 8, units = "in")
  grid.draw(final_plot)
}

#TEST

#dailydir <- '/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024_07_09'

#plot_participant_data(df, 208, 'CSS_Fx_Total', "CSS Functional Impact", dailydir)


```






# Create "id_list": list of all remaining participant IDs to loop through
```{r}

# Assuming df is your dataframe
# Count how many remaining ids you have in the dataset
unique_id_count <- length(unique(df$id))

# Get a list of unique IDs in the dataset
id_list <- unique(df$id)

# Print the number of unique IDs and the list of unique IDs in a sentence
cat("There are", unique_id_count, "unique IDs in the dataset. The IDs are:", paste(id_list, collapse = ", "), ".\n")


#There are 66 unique IDs in the dataset. The IDs are: 203, 205, 206, 208, 209, 213, 214, 218, 221, 223, 224, 225, 227, 228, 230, 232, 233, 235, 237, 239, 241, 242, 243, 245, 246, 248, 249, 251, 252, 254, 256, 258, 259, 260, 261, 262, 263, 265, 266, 267, 268, 270, 271, 272, 273, 275, 277, 278, 279, 281, 283, 284, 286, 287, 289, 290, 293, 305, 307, 310, 312, 314, 317, 318, 321, 324.
```

```{r}

# There are 12 people whose hormones should probably be removed for sensitivity analysis
badhorm_ids <- c(221, 235, 243, 267, 272, 273, 278, 283, 305, 307, 312, 317)
#Filtering for sensitivity analyses: 
# Remove rows with specified IDs
df_sens <- df %>% filter(!(id %in% badhorm_ids))

```

# Count participants and rows - FULL DATASET

```{r}
# Count how many remaining ids I have in the dataset
unique_id_count <- length(unique(df$id))
# Print the number of unique IDs
print(unique_id_count)

# NUMBER OF ROWS
print(nrow(df))
```

# Count participants and rows - SENSITIVITY SMALLER DATASET

```{r}
# Count how many remaining ids I have in the dataset
unique_id_count <- length(unique(df_sens$id))
# Print the number of unique IDs
print(unique_id_count)

# NUMBER OF ROWS
print(nrow(df_sens))
```

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

##### END OF DATA PREP 


# Save out Prepped Dataset for Easier Use Later On

## DATASET (L2N=66, L1N=2122)


```{r}

write.csv(df, "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-09-25/adhd_daily_df.csv", row.names = FALSE)

variable_names <- names(df)
formatted_list <- paste(variable_names, collapse = ", ")
cat(formatted_list)

```
## CLEANER SENSITIVITY DATASET (L2N=54, L1N=1730)


```{r}
write.csv(df_sens, "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-09-25/adhd_daily_df_sens.csv", row.names = FALSE)

variable_names <- names(df_sens)
formatted_list <- paste(variable_names, collapse = ", ")
cat(formatted_list)
```






############################################################################# CYCLE SCALING


## Using CYCLEDAY SCALING TEMPLATE - from Anisha updated September 14, 2024



```{r}

# Renaming Datasets
dat <- df

# Renaming Variables
dat$ID <-dat$id
dat$id <-dat$id
dat$daterated <-dat$date_rated


#df$LHposdayfirst
#df$mensdayonefirst
```

## Functions 

```{r}
calculate_mcyclength <- function(data) {
  # Calculate m2mcount: This counts from 1 menses onset to the next to give the length of a menses-to-menses cycle
  data <- data %>%
    group_by(id) %>%
    arrange(daterated, .by_group = TRUE) %>%
    mutate(
      m2mcount = NA,
      mcyclength = NA,
      cycle_incomplete = 0
    )
  
  for (i in seq_len(nrow(data))) {
    if (!is.na(data$A[i]) && data$A[i] == 1) {
      data$m2mcount[i] <- 1
      j <- i + 1
      
      while (j <= nrow(data) &&
             data$id[j] == data$id[i] && (is.na(data$A[j]) || data$A[j] != 1)) {
        data$m2mcount[j] <- data$m2mcount[j - 1] + 1
        j <- j + 1
      }
      
      if (j <= nrow(data) &&
          data$id[j] == data$id[i] && !is.na(data$A[j]) && data$A[j] == 1) {
        data$m2mcount[j] <- data$m2mcount[j - 1] + 1
      }
    }
  }
  
  # Identify incomplete cycles: This identifies instances where a complete menses-to-menses cycle was not captured
  data <- data %>%
    mutate(cycle_incomplete = ifelse(!is.na(m2mcount) &
                                       (is.na(lead(m2mcount)) & id != lead(id)), 1, 0))
  
  # New condition: Set cycle_incomplete = 1 if m2mcount restarts when id changes
  data <- data %>%
    group_by(id) %>%
    mutate(
      cycle_incomplete = ifelse(
        id != lag(id, default = first(id)) & m2mcount == 1, 1, cycle_incomplete
      )
    ) %>%
    ungroup()
  
  # Propagate cycle_incomplete within each group of m2mcount
  data <- data %>%
    group_by(id) %>%
    mutate(cycle_group = cumsum(!is.na(m2mcount) & m2mcount == 1)) %>%
    group_by(id, cycle_group) %>%
    mutate(
      cycle_incomplete = ifelse(any(cycle_incomplete == 1), 1, 0),
      # Fix: handle cases where all values in m2mcount are NA
      mcyclength = ifelse(all(is.na(m2mcount)), NA, max(m2mcount, na.rm = TRUE))
    ) %>%
    ungroup() %>%
    select(-cycle_group)
  
  data$cycle_incomplete = ifelse(is.na(data$cycle_incomplete), 1, 0)
  data$cycle_incomplete = ifelse(is.na(data$m2mcount), NA, data$cycle_incomplete)
  
  # Calculate cyclenum: calculates number of menses-to-menses cycles within a person
  data <- data %>%
    group_by(id) %>%
    mutate(cyclenum = cumsum(!is.na(m2mcount) & m2mcount == 1 & cycle_incomplete == 0)) %>%
    ungroup()
  
  data <- data %>%
    group_by(id) %>%
    mutate(cyclenum = ifelse(
      cycle_incomplete == 1,
      NA,
      cumsum(!is.na(m2mcount) & m2mcount == 1 & cycle_incomplete == 0)
    )) %>%
    ungroup()
  
  return(data)
}



#calculate_lutdaycount: counts length of luteal phase 

calculate_lutdaycount <- function(data, ovtoday) {
  last_id <- NULL
  lutdaycount1 <- rep(NA, nrow(data))
  lut_incomplete1 <- rep(NA, nrow(data))  # Initialize lut_incomplete
  count_started_row <- NA  # Track the starting row of a lutdaycount1 stretch
  active_count <- FALSE  # Track if lutdaycount1 is currently counting

  for (i in 1:nrow(data)) {
    # If id changes, reset lutdaycount1 and lut_incomplete
    if (is.null(last_id) || last_id != data$id[i]) {
      if (active_count) {
        # Mark the whole stretch as incomplete if stopped by id change
        lut_incomplete1[count_started_row:(i - 1)] <- 1
      }
      lutdaycount1[i] <- ifelse(ovtoday[i] == 1, 0, NA)
      active_count <- !is.na(lutdaycount1[i])  # Start counting if lutdaycount1 is initialized
      count_started_row <- ifelse(!is.na(lutdaycount1[i]), i, NA)  # Record start row
    } else if (!is.na(lutdaycount1[i - 1])) {
      lutdaycount1[i] <- lutdaycount1[i - 1] + 1
    }

    # If lutdaycount1 is active and stops because of A == 1, set lut_incomplete = 0 for the entire stretch
    if (!is.na(lutdaycount1[i]) && !is.na(data$A[i]) && data$A[i] == 1) {
      lutdaycount1[i] <- NA
      lut_incomplete1[count_started_row:(i - 1)] <- 0  # Mark stretch as complete due to A == 1
      active_count <- FALSE  # Stop counting due to A == 1
    } else if (ovtoday[i] == 1) {
      lutdaycount1[i] <- 0
      active_count <- TRUE  # Start counting at ovtoday == 1
      count_started_row <- i  # Record start row of the count
    }

    # Assign the current row to last_id for the next iteration
    last_id <- data$id[i]
  }

  # Update the dataset with lutdaycount1 and lut_incomplete
  data$lutdaycount1 <- lutdaycount1
  data$lut_incomplete1 <- lut_incomplete1
  
  return(data)
}

#calculate_foldaycount: counts length of follicular phase 

calculate_foldaycount <- function(data, ovtoday) {
  foldaycount <- NA
  fol_incomplete <- NA
  last_id <- NULL
  active_count <- FALSE  # Track if foldaycount is currently active
  count_started_row <- NULL  # Track the starting row of a foldaycount stretch
  
  for (i in 1:nrow(data)) {
    # If id changes, reset foldaycount and fol_incomplete
    if (is.null(last_id) || is.na(data$id[i]) || last_id != data$id[i]) {
      if (active_count) {
        # Mark the whole stretch as incomplete if stopped by id change
        data$fol_incomplete[count_started_row:i-1] <- 1
      }
      foldaycount <- ifelse(data$A[i] == 1, 0, NA)
      active_count <- !is.na(foldaycount)  # Start counting if foldaycount is initialized
      fol_incomplete <- NA  # Reset fol_incomplete when id changes
      count_started_row <- ifelse(!is.na(foldaycount), i, NA)  # Record start row
    } else if (!is.na(foldaycount)) {
      foldaycount <- foldaycount + 1
    }

    # If foldaycount is active and stopped because of ovtoday == 1, set fol_incomplete = 0 for the whole stretch
    if (!is.na(foldaycount) && i >= 2 && !is.na(ovtoday[i]) && ovtoday[i - 1] == 1) {
      foldaycount <- NA
      data$fol_incomplete[count_started_row:i-1] <- 0  # Mark stretch as complete due to ovulation
      active_count <- FALSE  # Stop counting due to ovulation
      fol_incomplete <- 0
    }

    # If A == 1, start or reset the count
    if (data$A[i] == 1) {
      foldaycount <- 0
      active_count <- TRUE  # Start counting
      count_started_row <- i  # Record start row of the count
    }

    # If foldaycount is actively counting and stops because of id change
    if (!is.na(last_id) && !is.na(data$id[i]) && last_id != data$id[i] && active_count) {
      data$fol_incomplete[count_started_row:i-1] <- 1  # Mark as incomplete because the count stopped due to id change
      active_count <- FALSE  # Stop counting as the id changed
    }

    # Assign foldaycount and fol_incomplete to the current row
    data$foldaycount[i] <- foldaycount
    if (is.na(fol_incomplete)) {
      data$fol_incomplete[i] <- NA
    }

    # Update last_id for the next iteration
    last_id <- data$id[i]
  }

  return(data)
}

```


<!-- ```{r} -->
<!-- #subset to only include baseline, washout, and PBO  -->
<!-- dat = dat %>% filter(cleartrialphase == 0| cleartrialphase == 1| cleartrialphase == 4| cleartrialphase == 5| cleartrialphase == 8|cleartrialphase == 9|cleartrialphase == 12) -->


```{r}

dat <- dat %>%
  group_by(id) %>%
  arrange(daterated, .by_group = TRUE) %>%
  mutate(folmax = NA)

#create the variable 'A' which will represent menses onset
dat$A <- ifelse(dat$cycleday == 1, 1, 0)
dat$A <- ifelse(is.na(dat$A), 0, dat$A)
#ovtoday = LH + 1
dat$ovtoday = lag(dat$LHposdayfirst) #changed this from "posovtest"
#check data
dat %>% select(daterated, ovtoday, LHposdayfirst)
```






## Functions 

```{r}



calculate_mcyclength <- function(data) {
  # Calculate m2mcount: This counts from 1 menses onset to the next to give the length of a menses-to-menses cycle
  data <- data %>%
    group_by(id) %>%
    arrange(daterated, .by_group = TRUE) %>%
    mutate(
      m2mcount = NA,
      mcyclength = NA,
      cycle_incomplete = 0
    )
  
  for (i in seq_len(nrow(data))) {
    if (!is.na(data$A[i]) && data$A[i] == 1) {
      data$m2mcount[i] <- 1
      j <- i + 1
      
      while (j <= nrow(data) &&
             data$id[j] == data$id[i] && (is.na(data$A[j]) || data$A[j] != 1)) {
        data$m2mcount[j] <- data$m2mcount[j - 1] + 1
        j <- j + 1
      }
      
      if (j <= nrow(data) &&
          data$id[j] == data$id[i] && !is.na(data$A[j]) && data$A[j] == 1) {
        data$m2mcount[j] <- data$m2mcount[j - 1] + 1
      }
    }
  }
  
  # Identify incomplete cycles: This identifies instances where a complete menses-to-menses cycle was not captured
  data <- data %>%
    mutate(cycle_incomplete = ifelse(!is.na(m2mcount) &
                                       (is.na(lead(m2mcount)) & id != lead(id)), 1, 0))
  
  # New condition: Set cycle_incomplete = 1 if m2mcount restarts when id changes
  data <- data %>%
    group_by(id) %>%
    mutate(
      cycle_incomplete = ifelse(
        id != lag(id, default = first(id)) & m2mcount == 1, 1, cycle_incomplete
      )
    ) %>%
    ungroup()
  
  # Propagate cycle_incomplete within each group of m2mcount
  data <- data %>%
    group_by(id) %>%
    mutate(cycle_group = cumsum(!is.na(m2mcount) & m2mcount == 1)) %>%
    group_by(id, cycle_group) %>%
    mutate(
      cycle_incomplete = ifelse(any(cycle_incomplete == 1), 1, 0),
      # Fix: handle cases where all values in m2mcount are NA or max is missing
      mcyclength = ifelse(sum(!is.na(m2mcount)) > 0, max(m2mcount, na.rm = TRUE), NA)
    ) %>%
    ungroup() %>%
    select(-cycle_group)
  
  data$cycle_incomplete = ifelse(is.na(data$cycle_incomplete), 1, 0)
  data$cycle_incomplete = ifelse(is.na(data$m2mcount), NA, data$cycle_incomplete)
  
  # Calculate cyclenum: calculates number of menses-to-menses cycles within a person
  data <- data %>%
    group_by(id) %>%
    mutate(cyclenum = cumsum(!is.na(m2mcount) & m2mcount == 1 & cycle_incomplete == 0)) %>%
    ungroup()
  
  data <- data %>%
    group_by(id) %>%
    mutate(cyclenum = ifelse(
      cycle_incomplete == 1,
      NA,
      cumsum(!is.na(m2mcount) & m2mcount == 1 & cycle_incomplete == 0)
    )) %>%
    ungroup()
  
  return(data)
}


#calculate_lutdaycount: counts length of luteal phase 

calculate_lutdaycount <- function(data, ovtoday) {
  last_id <- NULL
  lutdaycount1 <- rep(NA, nrow(data))
  lut_incomplete1 <- rep(NA, nrow(data))  # Initialize lut_incomplete
  count_started_row <- NA  # Track the starting row of a lutdaycount1 stretch
  active_count <- FALSE  # Track if lutdaycount1 is currently counting

  for (i in 1:nrow(data)) {
    # If id changes, reset lutdaycount1 and lut_incomplete
    if (is.null(last_id) || last_id != data$id[i]) {
      if (active_count) {
        # Mark the whole stretch as incomplete if stopped by id change
        lut_incomplete1[count_started_row:(i - 1)] <- 1
      }
      lutdaycount1[i] <- ifelse(ovtoday[i] == 1, 0, NA)
      active_count <- !is.na(lutdaycount1[i])  # Start counting if lutdaycount1 is initialized
      count_started_row <- ifelse(!is.na(lutdaycount1[i]), i, NA)  # Record start row
    } else if (!is.na(lutdaycount1[i - 1])) {
      lutdaycount1[i] <- lutdaycount1[i - 1] + 1
    }

    # If lutdaycount1 is active and stops because of A == 1, set lut_incomplete = 0 for the entire stretch
    if (!is.na(lutdaycount1[i]) && !is.na(data$A[i]) && data$A[i] == 1) {
      lutdaycount1[i] <- NA
      lut_incomplete1[count_started_row:(i - 1)] <- 0  # Mark stretch as complete due to A == 1
      active_count <- FALSE  # Stop counting due to A == 1
    } else if (ovtoday[i] == 1) {
      lutdaycount1[i] <- 0
      active_count <- TRUE  # Start counting at ovtoday == 1
      count_started_row <- i  # Record start row of the count
    }

    # Assign the current row to last_id for the next iteration
    last_id <- data$id[i]
  }

  # Update the dataset with lutdaycount1 and lut_incomplete
  data$lutdaycount1 <- lutdaycount1
  data$lut_incomplete1 <- lut_incomplete1
  
  return(data)
}

calculate_foldaycount <- function(data, ovtoday) {
  # Initialize the foldaycount and fol_incomplete columns
  data$foldaycount <- NA
  data$fol_incomplete <- NA
  
  foldaycount <- NA
  fol_incomplete <- NA
  last_id <- NULL
  active_count <- FALSE  # Track if foldaycount is currently active
  count_started_row <- NA  # Track the starting row of a foldaycount stretch
  
  for (i in 1:nrow(data)) {
    
    # If id changes or this is the first row, reset foldaycount and fol_incomplete
    if (is.null(last_id) || is.na(data$id[i]) || (!is.na(last_id) && last_id != data$id[i])) {
      if (active_count) {
        # Mark the whole stretch as incomplete if stopped by id change
        data$fol_incomplete[count_started_row:(i-1)] <- 1
      }
      foldaycount <- ifelse(data$A[i] == 1, 0, NA)
      active_count <- !is.na(foldaycount)  # Start counting if foldaycount is initialized
      fol_incomplete <- NA  # Reset fol_incomplete when id changes
      count_started_row <- ifelse(!is.na(foldaycount), i, NA)  # Record start row
    } else if (!is.na(foldaycount)) {
      foldaycount <- foldaycount + 1
    }

    # If foldaycount is active and stopped because ovtoday == 1 (indicating ovulation)
    if (active_count && i > 1 && !is.na(ovtoday[i]) && ovtoday[i] == 1) {
      foldaycount <- NA
      data$fol_incomplete[count_started_row:(i-1)] <- 0  # Mark stretch as complete due to ovulation
      active_count <- FALSE  # Stop counting due to ovulation
      fol_incomplete <- 0
    }

    # If A == 1, start or reset the count
    if (!is.na(data$A[i]) && data$A[i] == 1) {
      foldaycount <- 0
      active_count <- TRUE  # Start counting
      count_started_row <- i  # Record start row of the count
    }

    # Assign foldaycount and fol_incomplete to the current row
    data$foldaycount[i] <- foldaycount
    data$fol_incomplete[i] <- ifelse(is.na(fol_incomplete), NA, fol_incomplete)

    # Update last_id for the next iteration, but only if it's not missing
    if (!is.na(data$id[i])) {
      last_id <- data$id[i]
    }
  }

  return(data)
}


```


# Removed the cleartrialphase piece here

# Calculate Menses-to-menses cycle lengths 

## m2mcount and cyclelength 

```{r}

#the mcyclength function, this will do forward count to give cycle lengths for menses-to-menses cycles, and count the number of complete cycles within a person 
dat <- calculate_mcyclength(dat)

#-Inf will occur if the cycle is incomplete (no bookend menses onset date) so we will change to NA
dat$mcyclength = ifelse(dat$mcyclength == -Inf , NA, dat$mcyclength)

#if cyclenum is NA, set cycle_incomplete to 1 
dat$cycle_incomplete = ifelse(is.na(dat$cyclenum), 1, dat$cycle_incomplete)

#check data
dat %>% select(id, daterated, A, m2mcount, ovtoday, cyclenum, mcyclength, cycle_incomplete) 

```



Here, I am doing scaling for baseline, washout, and PBO condition

# calculating ovtoday_impute
 
```{r}
# this is based on NC dataset 
dat <- dat %>% 
  mutate(lutlength_impute = case_when(mcyclength == 20 ~ (mcyclength* 0.46),
                                      mcyclength == 21 ~ (mcyclength*.476),
                                      mcyclength == 22 ~ (mcyclength*.491),
                                      mcyclength == 23 ~ (mcyclength*.491),
                                      mcyclength == 24 ~ (mcyclength*.492),
                                      mcyclength == 25 ~ (mcyclength*.484),
                                      mcyclength == 26 ~ (mcyclength*.481),
                                      mcyclength == 27 ~ (mcyclength*.470),
                                      mcyclength == 28 ~ (mcyclength*.461),
                                      mcyclength == 29 ~ (mcyclength*.448),
                                      mcyclength == 30 ~ (mcyclength*.437),
                                      mcyclength == 31 ~ (mcyclength*.426),
                                      mcyclength == 32 ~ (mcyclength*.416),
                                      mcyclength == 33 ~ (mcyclength*.40),
                                      mcyclength == 34 ~ (mcyclength*.391),
                                      mcyclength == 35 ~ (mcyclength*.377),
                                      mcyclength == 36 ~ (mcyclength*.369),
                                      mcyclength == 37 ~ (mcyclength*.359),
                                      TRUE ~ NA))

dat <- dat %>% 
  mutate(follength_impute = case_when(mcyclength == 20 ~ (mcyclength*.54),
                                      mcyclength == 21 ~ (mcyclength*.524),
                                      mcyclength == 22 ~ (mcyclength*.509),
                                      mcyclength == 23 ~ (mcyclength*.509),
                                      mcyclength == 24 ~ (mcyclength*.508),
                                      mcyclength == 25 ~ (mcyclength*.516),
                                      mcyclength == 26 ~ (mcyclength*.519),
                                      mcyclength == 27 ~ (mcyclength*.530),
                                      mcyclength == 28 ~ (mcyclength*.539),
                                      mcyclength == 29 ~ (mcyclength*.552),
                                      mcyclength == 30 ~ (mcyclength*.563),
                                      mcyclength == 31 ~ (mcyclength*.574),
                                      mcyclength == 32 ~ (mcyclength*.584),
                                      mcyclength == 33 ~ (mcyclength*.60),
                                      mcyclength == 34 ~ (mcyclength*.609),
                                      mcyclength == 35 ~ (mcyclength*.623),
                                      mcyclength == 36 ~ (mcyclength*.631),
                                      mcyclength == 37 ~ (mcyclength*.641),
                                      TRUE ~ NA))
#check data
dat %>% select(id, daterated, A, m2mcount, mcyclength, follength_impute)
#group by id, and make sure dates are in order 
dat <- dat %>%
  group_by(id) %>%
  arrange(daterated, .by_group = TRUE) 
#initialize new variable- follcount1
dat$follcount1 <- NA
for (i in 1:nrow(dat)) {
  if (dat$A[i] == 1 & !is.na(dat$follength_impute[i])) {
    follcount1 <- seq_len(round(dat$follength_impute[i]))
    dat$follcount1[i:(i + length(follcount1) - 1)] <- follcount1
  }
}
#follcount1 forward counts from menses onset based on imputed follength from NC data 

dat %>% select(id, A, follength_impute, follcount1)

#identify ovtoday_impute based on forward count from follcount1
dat <- dat %>% 
  mutate(ovtoday_impute = case_when(round(follength_impute) == follcount1 ~ 1,
                               TRUE ~ NA))

#if ovtoday is NA, change to 0 
dat$ovtoday <- ifelse(is.na(dat$ovtoday), 0, dat$ovtoday) 

#if ovtoday_impute is NA, change to 0 
dat$ovtoday_impute <- ifelse(is.na(dat$ovtoday_impute), 0, dat$ovtoday_impute) 

#check data 
dat %>% select(id, daterated, ovtoday, ovtoday_impute) # Removed cleartrialphase
```


# percentage of luteal phase using ovtoday (LH + 1) 

 - variable is called perclut for menses-centered
 - variable is called perclut_ov for ovulation-centered 
 

```{r}
#group by id, make sure daterated is in order, initialize the variable lutmax 
dat <- dat  %>% 
  group_by(id) %>%
  arrange(daterated, .by_group = TRUE) %>%
  mutate(lutmax = NA)
#apply the calculate_lutdaycount function. which spits out the variable lutdaycount1. lutdaycount1 forward counts from ovtoday = 1 to the next menses onset, indexed at 0 
dat <- calculate_lutdaycount(dat, dat$ovtoday)
#check data 
dat %>% select(id, daterated, ovtoday, A, lutdaycount1)
#the below code creates lutdaycount which shifts lutdaycount1, and forward counts starting the day after ovtoday = 1, still indexed at 0 
dat <- dat %>%
  group_by(id) %>%
  mutate(
    lutdaycount = lag(lutdaycount1),
    lutdaycount = case_when(
      is.na(lutdaycount) | id != lag(id) ~ NA, 
      TRUE ~ lutdaycount
    )
  )

dat <- dat %>%
  group_by(id) %>%
  mutate(
    lut_incomplete = lag(lut_incomplete1),
    lut_incomplete = case_when(
      is.na(lut_incomplete) | id != lag(id) ~ NA, 
      TRUE ~ lut_incomplete
    )
  )


#check data 
dat %>% select(id, daterated, ovtoday, A, lutdaycount, lut_incomplete)
#lutmax corresponds to the length of the luteal phase in a complete menses-to-menses cycle (it is the max value of lutdaycount in a cycle)
for (i in 1:(nrow(dat)-1)) {
  if (is.na(dat$lutdaycount[i + 1] && !is.na(dat$lutdaycount[i]))) {
    dat$lutmax[(i-(dat$lutdaycount[i])):i] = as.numeric(dat$lutdaycount[i])
  }
}

dat$lut_incomplete = ifelse(dat$lutmax > 30, 1, dat$lut_incomplete)
#check data 
dat %>% select(id, daterated, ovtoday, A, lutdaycount, lutmax, lut_incomplete)

#calculates lutperc which is luteal phase scaled from 0 to 1, using ovtoday 
dat <- dat %>%
  mutate(lutperc = ifelse(lut_incomplete == 0, (lutdaycount / lutmax), NA))

#calculates perclut_base which is luteal phase scaled from -1 to 0, using ovtoday. This is the variable that will be used in the menses-centered scaled_cycleday 
dat$perclut = dat$lutperc -1 

# dat <- dat %>%
#   mutate(perclut = ifelse(A == 1, 0, perclut))

#check data
dat %>% select(id, daterated, ovtoday, A, lutmax, perclut, lut_incomplete)

#calculate lutdaycount_ov, which is lutdaycount but ovulation-centered 
dat <- dat %>%
  group_by(id) %>%
  mutate(
    lutdaycount_ov = lead(lutdaycount),
    lutdaycount_ov = case_when(is.na(lutdaycount_ov) |
                                 id != lead(id) ~ NA, TRUE ~ lutdaycount_ov)
  )

dat <- dat %>%
  mutate(perclut_ov = ifelse(lut_incomplete == 0, lutdaycount_ov / lutmax, NA))

# dat <- dat %>%
#   mutate(perclut_ov = ifelse(lutdaycount_ov == 0, 0, perclut_ov))


dat %>% select(id,
               daterated,
               ovtoday,
               A,
               lutdaycount,
               lutdaycount_ov,
               lutmax,
               perclut_ov, perclut)

#remove 'helper' variables
dat <- dat %>% select(-c(lutdaycount1, lutperc, lut_incomplete1))

```

# percentage of luteal phase using ovtoday_impute (based on NC norms)


```{r}
#make sure dates are in order for each id, initialize new variable called lutmax_impute
dat <- dat  %>% 
  group_by(id) %>%
  arrange(daterated, .by_group = TRUE) %>%
  mutate(lutmax_impute = NA)
#if A is NA, make it 0 (A is menses onset)
dat$A <- ifelse(is.na(dat$A), 0, dat$A)
#apply calculate_lutdaycount fxn which spits out the variable lutdaycount1. lutdaycount1 forward counts from ovtoday_impute = 1 to the next menses onset, indexed at 0 
dat <- calculate_lutdaycount(dat, dat$ovtoday_impute)
#check data
dat %>% select(id, daterated, ovtoday_impute, A, lutdaycount1)
#the below code creates lutdaycount_impute which shifts lutdaycount1, and forward counts starting the day after ovtoday_impute = 1, still indexed at 0 
dat <- dat %>%
  group_by(id) %>%
  mutate(
    lutdaycount_impute = lag(lutdaycount1),
    lutdaycount_impute = case_when(
      is.na(lutdaycount_impute) | id != lag(id) ~ NA, 
      TRUE ~ lutdaycount_impute
    )
  )
#check data 
dat %>% select(id, daterated, ovtoday_impute, A, lutdaycount_impute)

#lutmax_impute corresponds to the length of the luteal phase in a complete menses-to-menses cycle, using ovtoday_impute (it is the max value of lutdaycount_impute in a cycle)
for (i in 1:(nrow(dat)-1)) {
  if (is.na(dat$lutdaycount_impute[i + 1] && !is.na(dat$lutdaycount_impute[i]))) {
    dat$lutmax_impute[(i-(dat$lutdaycount_impute[i])):i] = as.numeric(dat$lutdaycount_impute[i])
  }
}
#lutlength1_impute is the imputed value of luteal phase based on NC norms, but lagged one day so it does align with any dates in the follicuar phase 
dat$lutlength1_impute = lag(dat$lutlength_impute)

#check data
dat %>% select(id, daterated, ovtoday_impute, A, lutdaycount_impute, lutmax_impute, lutlength_impute, lutlength1_impute)

#calculate lutperc_impute
dat <- dat %>%
  mutate(lutperc_impute = if_else(
    is.na(lutlength1_impute),
    NA,
    if_else(
      is.na(lutlength1_impute) &
        !is.na(lutmax_impute) & cycle_incomplete == 0,
      lutdaycount_impute / lutmax_impute,
      lutdaycount_impute / round(lutlength1_impute)
    )
  ))

#lutperc is scaled from 0 to 1, so substracting 1 so that it is scaled from -1 to 0 for menses-centered scaled_cycleday 
dat$perclut_impute = dat$lutperc_impute -1 

#check data
dat %>% select(id, daterated, ovtoday_impute, A, mcyclength, perclut_impute, perclut) 

#calculate lutdaycount_imp_ov, which is imputed lutdaycount but ovulation-centered 
dat <- dat %>%
  group_by(id) %>%
  mutate(
    lutdaycount_imp_ov = lead(lutdaycount_impute),
    lutdaycount_imp_ov = case_when(
      is.na(lutdaycount_imp_ov) | id != lead(id) ~ NA, 
      TRUE ~ lutdaycount_imp_ov
    )
  )

#calculate lutperc_imp_ov which is the variable that will go into ovulation-centered scaled cycleday 
dat <- dat %>%
  mutate(perclut_imp_ov = if_else(
    is.na(lutlength1_impute) ,
    NA,
    if_else(
      is.na(lutlength1_impute) & !is.na(lutmax) & cycle_incomplete == 0,
      lutdaycount1 / lutmax_impute,
      lutdaycount1 / round(lutlength1_impute)
    )
  ))


#if ovtoday_impute == 1, set lutperc_imp_ov to 0
# dat <- dat %>%
#   mutate(perclut_imp_ov = ifelse(ovtoday_impute == 1, 0, perclut_imp_ov))

#check data 
dat %>% select(id, daterated, ovtoday_impute, A, lutdaycount, lutlength1_impute, lutdaycount1, lutmax_impute, perclut_impute, perclut_imp_ov)

#remove helper variables
dat <- dat %>% select (-c(lutperc_impute, lutdaycount1))
```

# percentage of follicular phase using ovtoday (LH+1)

```{r}
# Check for missing values
table(dat$id)
table(dat$ovtoday)

#remove obs with missing id or ovtoday
dat <- dat %>%
  filter(!is.na(id) & !is.na(ovtoday))

# Ensure daterated is in order by id and initialize folmax variable
dat <- dat %>%
  group_by(id) %>%
  arrange(daterated, .by_group = TRUE) %>%
  mutate(folmax = NA)

#make sure daterated is in order by id, and initialize folmax variable 
dat <- dat %>%
  group_by(id) %>%
  arrange(daterated, .by_group = TRUE) %>%
  mutate(folmax = NA)

#apply calculate foldaycount function using ovtoday 
dat <- calculate_foldaycount(dat, dat$ovtoday)

#foldaycount forward counts starting at menses onset and ends at ovtoday 
dat %>% select(id, daterated, A, ovtoday, foldaycount)

#assigns folmax as the max value of foldaycount within a cycle 
for (i in 1:(nrow(dat)-1)) {
  if (is.na(dat$foldaycount[i + 1] && !is.na(dat$foldaycount[i]))) {
    dat$folmax[(i-(dat$foldaycount[i])):i] = as.numeric(dat$foldaycount[i])
  }
}

#follength is the length of the follicular phase that cycle (since folmax is indexed at 0)
dat$follength = dat$folmax + 1 

#check data
dat %>% select(id, daterated, ovtoday, A, foldaycount, folmax, follength, fol_incomplete)

dat <- dat %>%
  mutate(percfol = ifelse(fol_incomplete != 1, foldaycount / (folmax), NA))

dat$percfol_ov = dat$percfol - 1

dat %>% select(id, daterated, ovtoday, A, foldaycount, folmax, percfol, percfol_ov)

#renaming foldaycount based on ovtoday to be called foldaycount_ov. the variable name 'foldaycount' get overwritten in the code chunk below when the calculate_foldaycount() function is run 
dat <- dat %>% rename(foldaycount_ov = foldaycount)

```

# percentage of follicular phase using ovtoday_impute (based on NC norms)

```{r}

#yes, I'm constantly arranging by daterated because I'm paranoid. Initialize the variable folmax_impute 
dat <- dat %>%
  group_by(id) %>%
  arrange(daterated, .by_group = TRUE) %>%
  mutate(folmax_impute = NA)
#see comment on line 512
dat <- calculate_foldaycount(dat, dat$ovtoday_impute)
dat <- dat %>% rename(foldaycount_impute = foldaycount)
dat %>% select(id, daterated, ovtoday_impute, A, foldaycount_impute)

#create folmax_impute which is the max value of foldaycount_impute and matches the imputed follicular length in complete cycles 
for (i in 1:(nrow(dat)-1)) {
  if (is.na(dat$foldaycount_impute[i + 1] && !is.na(dat$foldaycount_impute[i]))) {
    dat$folmax_impute[(i-(dat$foldaycount_impute[i])):i] = as.numeric(dat$foldaycount_impute[i])
  }
}

dat <- dat %>%
  mutate(percfol_impute = ifelse(cycle_incomplete == 0, foldaycount_impute / (folmax_impute), NA))

# ovulation-centered percfol based on ovtoday_impute 
dat$percfol_imp_ov = dat$percfol_impute - 1


dat %>% select(id, daterated, ovtoday_impute, A, percfol_impute, percfol_imp_ov)
```




# prioritize LH test-based scaled values over imputed values 

```{r}

#functions to prioritize perc_base over perc

create_percent <- function(dat, percentvar, percvar, percvar_impute) {
  percentvar <- ensym(percentvar)
  percvar <- ensym(percvar)
  percvar_impute <- ensym(percvar_impute)
  
  dat %>%
    group_by(id, cyclenum) %>%
    mutate(
      !!percentvar := if (all(is.na(!!percvar) | !!percvar == 0)) {
        # if percvar contains only NA or 0, use percvar_impute
        !!percvar_impute
      } else {
        # otherwise, use percvar
        !!percvar
      }
    ) %>%
    ungroup()
}


dat = create_percent(dat, percentlut, perclut, perclut_impute)
dat = create_percent(dat, percentfol, percfol, percfol_impute)
dat = create_percent(dat, percentlut_ov, perclut_ov, perclut_imp_ov)
dat = create_percent(dat, percentfol_ov, percfol_ov, percfol_imp_ov)

dat %>% select(id, daterated, A, ovtoday, ovtoday_impute, cyclenum, mcyclength, percfol, percfol_impute, percentlut_ov)
```


# creating scaled_cycleday 

```{r}

dat <- dat %>%
  mutate(scaled_cycleday = ifelse(is.na(percentlut), percentfol, percentlut))
dat %>% select(id, daterated, percentlut, percentfol, scaled_cycleday)

```

# creating scaled_cycleday_ov

```{r}
dat <- dat %>%
  mutate(scaled_cycleday_ov = ifelse(is.na(percentlut_ov), percentfol_ov, percentlut_ov))
dat %>% select(id, daterated, percentlut_ov, percentfol_ov, scaled_cycleday_ov)

```


## Rounding Percents for LUTEAL-FOLLICULAR ORDER

```{r}

dat <- dat %>%
  mutate(cycleday_perc = (scaled_cycleday + 1) / 2) %>%
  mutate(cycleday_perc_round = round(cycleday_perc, 1)) %>%
  mutate(cycleday_10perc = round(cycleday_perc / 0.10) * 0.10) %>%
  mutate(cycleday_2perc = round(cycleday_perc / 0.02) * 0.02)


```


## Rounding Percents for FOLLICULAR-LUTEAL ORDER

```{r}

dat <- dat %>%
  mutate(cycledayov_perc = (scaled_cycleday_ov + 1) / 2) %>%
  mutate(cycledayov_perc_round = round(cycledayov_perc, 1)) %>%
  mutate(cycledayov_10perc = round(cycledayov_perc / 0.10) * 0.10) %>%
  mutate(cycledayov_2perc = round(cycledayov_perc / 0.02) * 0.02)


```


#RE-CALCULATE DEVATIONS AND STD AND ROLLING AVGS


```{r}

#create rolling averages on RAW variables
for (i in outcomelist) {
  dat <- create.3day.rolling.avg(dat, !!sym({{i}}), 3)
}

#MEANS

#execute for loop: run create.person.mean() on everything in "outcomelist.roll"
for (i in outcomelist.roll) {
  dat <- create.person.mean(dat, !!sym({{i}}), id)
}

#execute for loop: run create.person.mean() on everything in "outcomelist"
for (i in outcomelist) {
  dat <- create.person.mean(dat, !!sym({{i}}), id)
}

#PERSON-DEVIATIONS

#for raw deviations
#execute for loop: run create.deviation() on everything in list
for (i in outcomelist) {
  dat <- create.deviation(dat, !!sym({{i}}), !!sym(paste0({{i}}, ".m")))
}

#for rolling avgs deviations
#execute for loop: run create.deviation() on everything in list
for (i in outcomelist.roll) {
  dat <- create.deviation(dat, !!sym({{i}}), !!sym(paste0({{i}}, ".m")))
}


#Check 
#dat %>% select(id, E2, E2.m, E2.d, E2.roll, E2.roll.d) %>% View()

```

# 2024-09-25 - PLOTS - LUTEAL then FOLLICULAR - (OVULATION to OVULATION)
```{r}

# Define the folder location for saving the plots
output_folder <- "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-09-25"

# Define the function to generate plots for a list of DVs
plot_DVs <- function(DV_list, dat) {
  
  # Create person-standardized (z-scored within each participant) versions of each DV
  for (DV in DV_list) {
    zd_var <- paste0(DV, ".zd")  # Name for the person-standardized version
    
    # Add the person-standardized columns for each DV
    dat <- dat %>%
      group_by(id) %>%
      mutate(!!zd_var := scale(!!sym(DV), center = TRUE, scale = TRUE)) %>%
      ungroup()
  }
  
  # Calculate person-rescaled E2 and P4 (scaled within each person, sd = 0.5)
  dat <- dat %>%
    group_by(id) %>%
    mutate(E2.rescale = scale(E2, center = TRUE, scale = sd(E2, na.rm=TRUE)/0.5), #changed here
           P4.rescale = scale(P4, center = TRUE, scale = sd(P4, na.rm=TRUE)/0.5)) %>% #changed here
    ungroup()
  
  # Indicators for menstrual bleeding
  bleeding_indicator <- data.frame(xmin = .5, xmax = 0.7, ymin = -Inf, ymax = Inf)
  
  for (DV in DV_list) {
    
    if (DV %in% c("E2", "P4", "LH")) {
      # Raw plot for E2, P4, LH
      dv <- dat %>%
        group_by(cycleday_10perc) %>%
        summarise(mean_value = mean(!!sym(DV), na.rm = TRUE),
                  se_value = sd(!!sym(DV), na.rm = TRUE) / sqrt(n()))  # SE calculation
      
      p1 <- ggplot(dv, aes(x = cycleday_10perc, y = mean_value)) +
        geom_ribbon(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                    fill = "black", alpha = 0.1) +
        geom_point(size = 1) +
        geom_line(size = 1) +
        ggtitle(paste("Raw", DV, "Plot")) +
        xlab("Cycle Day (10% bins)") + ylab(paste(DV, "Mean")) +
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5, size = 14),
              plot.margin = margin(15, 5, 50, 5)) +
        coord_cartesian(clip = "off") +
        geom_rect(data = bleeding_indicator, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf), 
                  fill = "red", alpha = 0.2, inherit.aes = FALSE)
      
      # Plot the data
      print(p1)
      
      ggsave(filename = paste0(output_folder, "/", DV, "_", "raw", ".png"), plot = last_plot(), width = 8, height = 6)

      
    } else {
      # For DVs that are NOT E2, P4, LH
      
      # Raw data points and line for the dependent variable
      dv <- dat %>%
        group_by(cycleday_10perc) %>%
        summarise(mean_value = mean(!!sym(DV), na.rm = TRUE),
                  se_value = sd(!!sym(DV), na.rm = TRUE) / sqrt(n()))  # SE calculation
      
    # Person-rescaled E2 and P4 lines (thin dotted)  <- Changed variable names here
      e2_std <- dat %>%
        group_by(cycleday_10perc) %>%
        summarise(mean_value = mean(E2.rescale, na.rm = TRUE)) #changed here

      p4_std <- dat %>%
        group_by(cycleday_10perc) %>%
        summarise(mean_value = mean(P4.rescale, na.rm = TRUE))  #changed here
      
      p1 <- ggplot(dv, aes(x = cycleday_10perc, y = mean_value)) +
        geom_ribbon(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                    fill = "black", alpha = 0.3) +
        geom_point(size = 1) +
        geom_line(size = 2) +
        # Thin dotted lines for E2 and P4
        # geom_line(data = e2_std, aes(x = cycleday_10perc, y = mean_value), 
        #           linetype = "dotted", color = "blue", size = 1, alpha = 0.6) +  # E2
        # geom_line(data = p4_std, aes(x = cycleday_10perc, y = mean_value), 
        #           linetype = "dotted", color = "forestgreen", size = 1, alpha = 0.6) +  # P4
        ggtitle(paste("Raw", DV, "Plot")) +
        xlab("Cycle Day (10% bins)") + ylab(paste(DV, "Mean")) +
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5, size = 14),
              plot.margin = margin(15, 5, 50, 5)) +
        coord_cartesian(clip = "off") +
        geom_rect(data = bleeding_indicator, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf), 
                  fill = "red", alpha = 0.2, inherit.aes = FALSE)
      
      # Plot the data
      print(p1)
      
      # Deviations plot
      dev_var <- paste0(DV, ".d")
      dv <- dat %>%
        group_by(cycleday_10perc) %>%
        summarise(mean_value = mean(!!sym(dev_var), na.rm = TRUE),
                  se_value = sd(!!sym(dev_var), na.rm = TRUE) / sqrt(n()))  # SE calculation
      
      p2 <- ggplot(dv, aes(x = cycleday_10perc, y = mean_value)) +
        geom_ribbon(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                    fill = "black", alpha = 0.3) +
        geom_point(size = 1) +
        geom_line(size = 2) +
        # geom_line(data = e2_std, aes(x = cycleday_10perc, y = mean_value), 
        #           linetype = "dotted", color = "blue", size = 1, alpha = 0.6) +  # E2
        # geom_line(data = p4_std, aes(x = cycleday_10perc, y = mean_value), 
        #           linetype = "dotted", color = "forestgreen", size = 1, alpha = 0.6) +  # P4
        ggtitle(paste("Deviations", DV, "Plot")) +
        xlab("Cycle Day (10% bins)") + ylab(paste(DV, "Deviation Mean")) +
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5, size = 14),
              plot.margin = margin(15, 5, 50, 5)) +
        coord_cartesian(clip = "off") +
        geom_rect(data = bleeding_indicator, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf), 
                  fill = "red", alpha = 0.2, inherit.aes = FALSE)
      
      print(p2)
      
            ggsave(filename = paste0(output_folder, "/", DV, "_", "dev", ".png"), plot = last_plot(), width = 8, height = 6)

      
      # Person-standardized outcome (zd)
      zd_var <- paste0(DV, ".zd")
      dv <- dat %>%
        group_by(cycleday_10perc) %>%
        summarise(mean_value = mean(!!sym(zd_var), na.rm = TRUE),
                  se_value = sd(!!sym(zd_var), na.rm = TRUE) / sqrt(n()))  # SE calculation
      
      p3 <- ggplot(dv, aes(x = cycleday_10perc, y = mean_value)) +
        geom_ribbon(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                    fill = "black", alpha = 0.3) +
        geom_point(size = 1) +
        geom_line(size = 2) +
        geom_line(data = e2_std, aes(x = cycleday_10perc, y = mean_value), 
                  linetype = "solid", color = "forestgreen", size = 1, alpha = 0.9) +  # E2
        geom_line(data = p4_std, aes(x = cycleday_10perc, y = mean_value), 
                  linetype = "solid", color = "red", size = 1, alpha = 0.8) +  # P4
        ggtitle(paste("Person-standardized", DV, "Plot")) +
        xlab("Cycle Day (10% bins)") + ylab(paste(DV, "Person-standardized Mean")) +
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5, size = 14),
              plot.margin = margin(15, 5, 50, 5)) +
        coord_cartesian(clip = "off") +
        geom_rect(data = bleeding_indicator, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf), 
                  fill = "red", alpha = 0.2, inherit.aes = FALSE)
      
      print(p3)
                  ggsave(filename = paste0(output_folder, "/", DV, "_", "zd", ".png"), plot = last_plot(), width = 8, height = 6)

      
      # Rolling deviations plot
      roll_dev_var <- paste0(DV, ".roll.d")
      dv <- dat %>%
        group_by(cycleday_10perc) %>%
        summarise(mean_value = mean(!!sym(roll_dev_var), na.rm = TRUE),
                  se_value = sd(!!sym(roll_dev_var), na.rm = TRUE) / sqrt(n()))  # SE calculation
      
      p4 <- ggplot(dv, aes(x = cycleday_10perc, y = mean_value)) +
        geom_ribbon(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                    fill = "black", alpha = 0.3) +
        geom_point(size = 1) +
        geom_line(size = 2) +
        # geom_line(data = e2_std, aes(x = cycleday_10perc, y = mean_value), 
        #           linetype = "dotted", color = "blue", size = 1, alpha = 0.6) +  # E2
        # geom_line(data = p4_std, aes(x = cycleday_10perc, y = mean_value), 
        #           linetype = "dotted", color = "forestgreen", size = 1, alpha = 0.6) +  # P4
        ggtitle(paste("Rolling Deviations", DV, "Plot")) +
        xlab("Cycle Day (10% bins)") + ylab(paste(DV, "Rolling Deviation Mean")) +
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5, size = 14),
              plot.margin = margin(15, 5, 50, 5)) +
        coord_cartesian(clip = "off") +
        geom_rect(data = bleeding_indicator, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf), 
                  fill = "red", alpha = 0.2, inherit.aes = FALSE)
      
      # Plot the data
      print(p4)
                  ggsave(filename = paste0(output_folder, "/", DV, "_", "rolld", ".png"), plot = last_plot(), width = 8, height = 6)

    }
  }
}

# Call the function with the list of dependent variables (DVs)
DVs <- c("E2", "P4", "LH", "CSS_Inatt", "CSS_Inatt_Count", "CSS_HypImp", "CSS_HypImp_Count",
         "CSS_Fx_Total", "CSS_B2_Total", "UPPS_Total", "BDEFS_Total", "DEBQ_Total")

plot_DVs(DVs, dat)





```

# 2024-09-25 - PLOTS - FOLLICULAR then LUTEAL - (MENSES to MENSES)
## NOTE: These plots aren't looking right yet. Need to figure out if I'm losing observations somewhere. - taem, 20240925
```{r}

# Define the folder location for saving the plots
output_folder <- "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-09-25"

# Define the function to generate plots for a list of DVs
plot_DVs_M2M <- function(DV_list, dat) {
  
  # Create person-standardized (z-scored within each participant) versions of each DV
  for (DV in DV_list) {
    zd_var <- paste0(DV, ".zd")  # Name for the person-standardized version
    
    # Add the person-standardized columns for each DV
    dat <- dat %>%
      group_by(id) %>%
      mutate(!!zd_var := scale(!!sym(DV), center = TRUE, scale = TRUE)) %>%
      ungroup()
  }
  
  # Calculate person-rescaled E2 and P4 (scaled within each person, sd = 0.5)
  dat <- dat %>%
    group_by(id) %>%
    mutate(E2.rescale = scale(E2, center = TRUE, scale = sd(E2, na.rm=TRUE)/0.5), #changed here
           P4.rescale = scale(P4, center = TRUE, scale = sd(P4, na.rm=TRUE)/0.5)) %>% #changed here
    ungroup()
  
  # Indicators for menstrual bleeding
  bleeding_indicator <- data.frame(xmin = 0, xmax = 0.2, ymin = -Inf, ymax = Inf)
  
  for (DV in DV_list) {
    
    if (DV %in% c("E2", "P4", "LH")) {
      # Raw plot for E2, P4, LH
      dv <- dat %>%
        group_by(cycledayov_10perc) %>%
        summarise(mean_value = mean(!!sym(DV), na.rm = TRUE),
                  se_value = sd(!!sym(DV), na.rm = TRUE) / sqrt(n()))  # SE calculation
      
      p1 <- ggplot(dv, aes(x = cycledayov_10perc, y = mean_value)) +
        geom_ribbon(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                    fill = "black", alpha = 0.1) +
        geom_point(size = 1) +
        geom_line(size = 1) +
        ggtitle(paste("Raw", DV, "Plot")) +
        xlab("Cycle Day (10% bins)") + ylab(paste(DV, "Mean")) +
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5, size = 14),
              plot.margin = margin(15, 5, 50, 5)) +
        coord_cartesian(clip = "off") +
        geom_rect(data = bleeding_indicator, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf), 
                  fill = "red", alpha = 0.2, inherit.aes = FALSE)
      
      # Plot the data
      print(p1)
      
      ggsave(filename = paste0(output_folder, "/", DV, "_", "ov_", ".png"), plot = last_plot(), width = 8, height = 6)

      
    } else {
      # For DVs that are NOT E2, P4, LH
      
      # Raw data points and line for the dependent variable
      dv <- dat %>%
        group_by(cycledayov_10perc) %>%
        summarise(mean_value = mean(!!sym(DV), na.rm = TRUE),
                  se_value = sd(!!sym(DV), na.rm = TRUE) / sqrt(n()))  # SE calculation
      
    # Person-rescaled E2 and P4 lines (thin dotted)  <- Changed variable names here
      e2_std <- dat %>%
        group_by(cycledayov_10perc) %>%
        summarise(mean_value = mean(E2.rescale, na.rm = TRUE)) #changed here

      p4_std <- dat %>%
        group_by(cycledayov_10perc) %>%
        summarise(mean_value = mean(P4.rescale, na.rm = TRUE))  #changed here
      
      p1 <- ggplot(dv, aes(x = cycledayov_10perc, y = mean_value)) +
        geom_ribbon(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                    fill = "black", alpha = 0.3) +
        geom_point(size = 1) +
        geom_line(size = 2) +
        # Thin dotted lines for E2 and P4
        # geom_line(data = e2_std, aes(x = cycledayov_10perc, y = mean_value), 
        #           linetype = "dotted", color = "blue", size = 1, alpha = 0.6) +  # E2
        # geom_line(data = p4_std, aes(x = cycledayov_10perc, y = mean_value), 
        #           linetype = "dotted", color = "forestgreen", size = 1, alpha = 0.6) +  # P4
        ggtitle(paste("Raw", DV, "Plot")) +
        xlab("Cycle Day (10% bins)") + ylab(paste(DV, "Mean")) +
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5, size = 14),
              plot.margin = margin(15, 5, 50, 5)) +
        coord_cartesian(clip = "off") +
        geom_rect(data = bleeding_indicator, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf), 
                  fill = "red", alpha = 0.2, inherit.aes = FALSE)
      
      # Plot the data
      print(p1)
      
      # Deviations plot
      dev_var <- paste0(DV, ".d")
      dv <- dat %>%
        group_by(cycledayov_10perc) %>%
        summarise(mean_value = mean(!!sym(dev_var), na.rm = TRUE),
                  se_value = sd(!!sym(dev_var), na.rm = TRUE) / sqrt(n()))  # SE calculation
      
      p2 <- ggplot(dv, aes(x = cycledayov_10perc, y = mean_value)) +
        geom_ribbon(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                    fill = "black", alpha = 0.3) +
        geom_point(size = 1) +
        geom_line(size = 2) +
        # geom_line(data = e2_std, aes(x = cycledayov_10perc, y = mean_value), 
        #           linetype = "dotted", color = "blue", size = 1, alpha = 0.6) +  # E2
        # geom_line(data = p4_std, aes(x = cycledayov_10perc, y = mean_value), 
        #           linetype = "dotted", color = "forestgreen", size = 1, alpha = 0.6) +  # P4
        ggtitle(paste("Deviations", DV, "Plot")) +
        xlab("Cycle Day (10% bins)") + ylab(paste(DV, "Deviation Mean")) +
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5, size = 14),
              plot.margin = margin(15, 5, 50, 5)) +
        coord_cartesian(clip = "off") +
        geom_rect(data = bleeding_indicator, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf), 
                  fill = "red", alpha = 0.2, inherit.aes = FALSE)
      
      print(p2)
      
            ggsave(filename = paste0(output_folder, "/", DV, "_", "ov_dev", ".png"), plot = last_plot(), width = 8, height = 6)

      
      # Person-standardized outcome (zd)
      zd_var <- paste0(DV, ".zd")
      dv <- dat %>%
        group_by(cycledayov_10perc) %>%
        summarise(mean_value = mean(!!sym(zd_var), na.rm = TRUE),
                  se_value = sd(!!sym(zd_var), na.rm = TRUE) / sqrt(n()))  # SE calculation
      
      p3 <- ggplot(dv, aes(x = cycledayov_10perc, y = mean_value)) +
        geom_ribbon(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                    fill = "black", alpha = 0.3) +
        geom_point(size = 1) +
        geom_line(size = 2) +
        geom_line(data = e2_std, aes(x = cycledayov_10perc, y = mean_value), 
                  linetype = "solid", color = "forestgreen", size = 1, alpha = 0.9) +  # E2
        geom_line(data = p4_std, aes(x = cycledayov_10perc, y = mean_value), 
                  linetype = "solid", color = "red", size = 1, alpha = 0.8) +  # P4
        ggtitle(paste("Person-standardized", DV, "Plot")) +
        xlab("Cycle Day (10% bins)") + ylab(paste(DV, "Person-standardized Mean")) +
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5, size = 14),
              plot.margin = margin(15, 5, 50, 5)) +
        coord_cartesian(clip = "off") +
        geom_rect(data = bleeding_indicator, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf), 
                  fill = "red", alpha = 0.2, inherit.aes = FALSE)
      
      print(p3)
                  ggsave(filename = paste0(output_folder, "/", DV, "_", "ov_zd", ".png"), plot = last_plot(), width = 8, height = 6)

      
      # Rolling deviations plot
      roll_dev_var <- paste0(DV, ".roll.d")
      dv <- dat %>%
        group_by(cycledayov_10perc) %>%
        summarise(mean_value = mean(!!sym(roll_dev_var), na.rm = TRUE),
                  se_value = sd(!!sym(roll_dev_var), na.rm = TRUE) / sqrt(n()))  # SE calculation
      
      p4 <- ggplot(dv, aes(x = cycledayov_10perc, y = mean_value)) +
        geom_ribbon(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                    fill = "black", alpha = 0.3) +
        geom_point(size = 1) +
        geom_line(size = 2) +
        # geom_line(data = e2_std, aes(x = cycledayov_10perc, y = mean_value), 
        #           linetype = "dotted", color = "blue", size = 1, alpha = 0.6) +  # E2
        # geom_line(data = p4_std, aes(x = cycledayov_10perc, y = mean_value), 
        #           linetype = "dotted", color = "forestgreen", size = 1, alpha = 0.6) +  # P4
        ggtitle(paste("Rolling Deviations", DV, "Plot")) +
        xlab("Cycle Day (10% bins)") + ylab(paste(DV, "Rolling Deviation Mean")) +
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5, size = 14),
              plot.margin = margin(15, 5, 50, 5)) +
        coord_cartesian(clip = "off") +
        geom_rect(data = bleeding_indicator, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf), 
                  fill = "red", alpha = 0.2, inherit.aes = FALSE)
      
      # Plot the data
      print(p4)
                  ggsave(filename = paste0(output_folder, "/", DV, "_", "ov_rolld", ".png"), plot = last_plot(), width = 8, height = 6)

    }
  }
}

# Call the function with the list of dependent variables (DVs)
DVs <- c("E2", "P4", "LH", "CSS_Inatt", "CSS_Inatt_Count", "CSS_HypImp", "CSS_HypImp_Count",
         "CSS_Fx_Total", "CSS_B2_Total", "UPPS_Total", "BDEFS_Total", "DEBQ_Total")

plot_DVs_M2M(DVs, dat)



```





# FITTING GAMMs 

```{r}
library(mgcv)
library(marginaleffects)

gam3ahsd2 <- gam(x3a_hsd2.w.log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re") + s(scaled_cycleday, k = 20), data = dat3ahsd2, method = "REML")

summary(gam3ahsd2)

plot.gam(gam3ahsd2, select = 3)

dat3ahsd2plot <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gam3ahsd2, newdata = dat3ahsd2plot, type = "response", transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 
dat3ahsd2plot$estimate = pred$estimate
dat3ahsd2plot$conf.low = pred$conf.low
dat3ahsd2plot$conf.high = pred$conf.high


# Plotting
x3ahsd2plot <- ggplot(dat3ahsd2plot, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "3a hsd2") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

x3ahsd2plot
```

# Fitting Scaled CycleDay GAMMs
```{r}
library(mgcv)
library(marginaleffects)

#dat$daysinstudy <- as.numeric(dat$TubeNumber)

```

# GAMM and initial plot - E2 - Ov to Ov
```{r}

hist(dat$E2)
dat$E2log <- log(dat$E2+1)
hist(dat$E2log)

gamm_e2 <- gam(E2log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re") + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_e2)

plot.gam(gamm_e2, select = 3)
```

# Model-Implied Plot - E2 - Ov to Ov
```{r}

e2dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_e2, newdata = e2dat, type = "response", transform = function(x) exp(x) - 1)

#don't need to include anything for 'transform' if your outcome is not log-transformed. This undoes the logtransform so model-implied values are not on the log scale 

e2dat$estimate = pred$estimate
e2dat$conf.low = pred$conf.low
e2dat$conf.high = pred$conf.high


# Plotting
e2plot <- ggplot(e2dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Salivary Estradiol") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

e2plot
```

# GAMM and initial plot - E2 - M2M
```{r}

hist(dat$E2)
dat$E2log <- log(dat$E2+1)
hist(dat$E2log)

gamm_e2_m2m <- gam(E2log ~ s(id, bs = "re") + s(scaled_cycleday_ov, id, bs = "re") + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_e2_m2m)

plot.gam(gamm_e2_m2m, select = 3)
```




# GAMM and initial plot - LH
```{r}
hist(dat$LH)
dat$LHlog <- log(dat$LH+1)
hist(dat$LHlog)
gamm_LH <- gam(LHlog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_LH)

plot.gam(gamm_LH, select = 3)

hist(dat$LH)
```


# Model-Implied Plot - LH
```{r}

LHdat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_LH, newdata = LHdat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

LHdat$estimate = pred$estimate
LHdat$conf.low = pred$conf.low
LHdat$conf.high = pred$conf.high

# Plotting
LHplot <- ggplot(LHdat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "LH") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

LHplot
```


# GAMM and initial plot - P4
```{r}
hist(dat$P4)
dat$P4log <- log(dat$P4+1)
hist(dat$P4log)

gamm_p4 <- gam(P4log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 10), data = dat, method = "REML")

summary(gamm_p4)

plot.gam(gamm_p4, select = 3)

```

# Model-Implied Plot - P4
```{r}

P4dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_p4, newdata = P4dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

P4dat$estimate = pred$estimate
P4dat$conf.low = pred$conf.low
P4dat$conf.high = pred$conf.high


# Plotting
P4plot <- ggplot(P4dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Salivary P4") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

P4plot
```



# GAMM and initial plot - Inatt
```{r}

hist(dat$CSS_Inatt)
dat$CSS_Inattlog <- log(dat$CSS_Inatt+1)
hist(dat$CSS_Inattlog)

gamm_CSS_Inatt <- gam(CSS_Inattlog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re") + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_CSS_Inatt)

plot.gam(gamm_CSS_Inatt, select = 3)

```


# Model-Implied Plot - Inatt
```{r}

CSS_Inattdat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_CSS_Inatt, newdata = CSS_Inattdat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

CSS_Inattdat$estimate = pred$estimate
CSS_Inattdat$conf.low = pred$conf.low
CSS_Inattdat$conf.high = pred$conf.high


# Plotting
CSS_Inattplot <- ggplot(CSS_Inattdat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "CSS Inattention Symptoms") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

CSS_Inattplot
```






# GAMM and initial plot - HypImp
```{r}


hist(dat$CSS_HypImp)
dat$CSS_HypImplog <- log(dat$CSS_HypImp+1)
hist(dat$CSS_HypImplog)

gamm_CSS_HypImp <- gam(CSS_HypImplog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re") + s(scaled_cycleday, k = 30), data = dat, method = "REML")

summary(gamm_CSS_HypImp)

plot.gam(gamm_CSS_HypImp, select = 3)

hist(dat$CSS_HypImp)
```

# Model-Implied Plot - HypImp
```{r}

CSS_HypImpdat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_CSS_HypImp, newdata = CSS_HypImpdat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

CSS_HypImpdat$estimate = pred$estimate
CSS_HypImpdat$conf.low = pred$conf.low
CSS_HypImpdat$conf.high = pred$conf.high


# Plotting
CSS_HypImpplot <- ggplot(CSS_HypImpdat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "CSS Hyperactivity/Impulsivity") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

CSS_HypImpplot
```
  # Calculate Inattention Severity: 
    CSS_Inatt = rowSums(across(c(CSS_B_1, CSS_B_3, CSS_B_5, CSS_B_7, CSS_B_9, CSS_B_11, CSS_B_13, CSS_B_15, CSS_B_17)), na.rm = TRUE),
    
    # Calculate Hyperactivity/Impulsivity Severity: 
    CSS_HypImp = rowSums(across(c(CSS_B_2, CSS_B_4, CSS_B_6, CSS_B_8, CSS_B_10, CSS_B_12, CSS_B_14, CSS_B_16, CSS_B_18)), na.rm = TRUE),
    
    # Function items 1-10
    CSS_Fx_Total = rowSums(across(starts_with("CSS_Function_")), na.rm = TRUE),
    
    # Not sure what this is (todo)
    CSS_B2_Total = rowSums(across(starts_with("CSS_B2_")), na.rm = TRUE),
    
    # CSS Count Variables
    CSS_Inatt_Count = rowSums(across(c(CSS_1_Count, CSS_3_Count, CSS_5_Count, CSS_7_Count, CSS_9_Count, CSS_11_Count, CSS_13_Count, CSS_15_Count, CSS_17_Count)), na.rm = TRUE),
    
    CSS_HypImp_Count = rowSums(across(c(CSS_2_Count, CSS_4_Count, CSS_6_Count, CSS_8_Count, CSS_10_Count, CSS_12_Count, CSS_14_Count, CSS_16_Count, CSS_18_Count)), na.rm = TRUE),
    
    # Other Daily Scored Scales
    DEBQ_Total = rowSums(across(starts_with("DEBQ_")), na.rm = TRUE),
    BDEFS_Total = rowSums(across(starts_with("BDEFS_")), na.rm = TRUE),
    UPPS_Total = rowSums(across(starts_with("UPPS_")), na.rm = TRUE)


# GAMM and initial plot - Inatt Sx Count
```{r}
hist(dat$CSS_Inatt_Count)
dat$CSS_Inatt_Countlog <- log(dat$CSS_Inatt_Count+1)
hist(dat$CSS_Inatt_Countlog)

gamm_CSS_Inatt_Count <- gam(CSS_Inatt_Countlog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_CSS_Inatt_Count)

plot.gam(gamm_CSS_Inatt_Count, select = 3)

hist(dat$CSS_Inatt_Count)
```


# Model-Implied Plot - Inatt Sx Count
```{r}

CSS_Inatt_Countdat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_CSS_Inatt_Count, newdata = CSS_Inatt_Countdat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

CSS_Inatt_Countdat$estimate = pred$estimate
CSS_Inatt_Countdat$conf.low = pred$conf.low
CSS_Inatt_Countdat$conf.high = pred$conf.high

# Plotting
CSS_Inatt_Countplot <- ggplot(CSS_Inatt_Countdat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "CSS Inattention Sx Count") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

CSS_Inatt_Countplot
```




# GAMM and initial plot - HypImp Count
```{r}
hist(dat$CSS_HypImp_Count)
dat$CSS_HypImp_Countlog <- log(dat$CSS_HypImp_Count+1)
hist(dat$CSS_HypImp_Countlog)
gamm_CSS_HypImp_Count <- gam(CSS_HypImp_Countlog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_CSS_HypImp_Count)

plot.gam(gamm_CSS_HypImp_Count, select = 3)

hist(dat$CSS_HypImp_Count)
```


# Model-Implied Plot - TEMPLATE
```{r}

CSS_HypImp_Countdat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_CSS_HypImp_Count, newdata = CSS_HypImp_Countdat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

CSS_HypImp_Countdat$estimate = pred$estimate
CSS_HypImp_Countdat$conf.low = pred$conf.low
CSS_HypImp_Countdat$conf.high = pred$conf.high

# Plotting
CSS_HypImp_Countplot <- ggplot(CSS_HypImp_Countdat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "CSS_HypImp_Count") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

CSS_HypImp_Countplot
```



# GAMM and initial plot - CSS_Fx_Total
```{r}
hist(dat$CSS_Fx_Total)
dat$CSS_Fx_Totallog <- log(dat$CSS_Fx_Total+1)
hist(dat$CSS_Fx_Totallog)
gamm_CSS_Fx_Total <- gam(CSS_Fx_Totallog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_CSS_Fx_Total)

plot.gam(gamm_CSS_Fx_Total, select = 3)

hist(dat$CSS_Fx_Total)
```


# Model-Implied Plot - CSS_Fx_Total
```{r}

CSS_Fx_Totaldat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_CSS_Fx_Total, newdata = CSS_Fx_Totaldat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

CSS_Fx_Totaldat$estimate = pred$estimate
CSS_Fx_Totaldat$conf.low = pred$conf.low
CSS_Fx_Totaldat$conf.high = pred$conf.high

# Plotting
CSS_Fx_Totalplot <- ggplot(CSS_Fx_Totaldat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "CSS_Fx_Total") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

CSS_Fx_Totalplot
```
# GAMM and initial plot - CSS_B2_Total
```{r}
hist(dat$CSS_B2_Total)
dat$CSS_B2_Totallog <- log(dat$CSS_B2_Total+1)
hist(dat$CSS_B2_Totallog)
gamm_CSS_B2_Total <- gam(CSS_B2_Totallog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_CSS_B2_Total)

plot.gam(gamm_CSS_B2_Total, select = 3)

hist(dat$CSS_B2_Total)
```


# Model-Implied Plot - CSS_B2_Total
```{r}

CSS_B2_Totaldat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_CSS_B2_Total, newdata = CSS_B2_Totaldat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

CSS_B2_Totaldat$estimate = pred$estimate
CSS_B2_Totaldat$conf.low = pred$conf.low
CSS_B2_Totaldat$conf.high = pred$conf.high

# Plotting
CSS_B2_Totalplot <- ggplot(CSS_B2_Totaldat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "CSS_B2_Total") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

CSS_B2_Totalplot
```

# GAMM and initial plot - BDEFS_Total
```{r}
hist(dat$BDEFS_Total)
dat$BDEFS_Totallog <- log(dat$BDEFS_Total+1)
hist(dat$BDEFS_Totallog)
gamm_BDEFS_Total <- gam(BDEFS_Totallog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_BDEFS_Total)

plot.gam(gamm_BDEFS_Total, select = 3)

hist(dat$BDEFS_Total)
```


# Model-Implied Plot - BDEFS_Total
```{r}

BDEFS_Totaldat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_BDEFS_Total, newdata = BDEFS_Totaldat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

BDEFS_Totaldat$estimate = pred$estimate
BDEFS_Totaldat$conf.low = pred$conf.low
BDEFS_Totaldat$conf.high = pred$conf.high

# Plotting
BDEFS_Totalplot <- ggplot(BDEFS_Totaldat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "BDEFS_Total") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

BDEFS_Totalplot
```


# GAMM and initial plot - UPPS_Total
```{r}
hist(dat$UPPS_Total)
dat$UPPS_Totallog <- log(dat$UPPS_Total+1)
hist(dat$UPPS_Totallog)
gamm_UPPS_Total <- gam(UPPS_Totallog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_UPPS_Total)

plot.gam(gamm_UPPS_Total, select = 3)

hist(dat$UPPS_Total)
```


# Model-Implied Plot - UPPS_Total
```{r}

UPPS_Totaldat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_UPPS_Total, newdata = UPPS_Totaldat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

UPPS_Totaldat$estimate = pred$estimate
UPPS_Totaldat$conf.low = pred$conf.low
UPPS_Totaldat$conf.high = pred$conf.high

# Plotting
UPPS_Totalplot <- ggplot(UPPS_Totaldat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "UPPS_Total") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

UPPS_Totalplot
```


# GAMM and initial plot - DEBQ_Total
```{r}
hist(dat$DEBQ_Total)
dat$DEBQ_Totallog <- log(dat$DEBQ_Total+1)
hist(dat$DEBQ_Totallog)
gamm_DEBQ_Total <- gam(DEBQ_Totallog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DEBQ_Total)

plot.gam(gamm_DEBQ_Total, select = 3)

hist(dat$DEBQ_Total)
```


# Model-Implied Plot - DEBQ_Total
```{r}

DEBQ_Totaldat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DEBQ_Total, newdata = DEBQ_Totaldat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DEBQ_Totaldat$estimate = pred$estimate
DEBQ_Totaldat$conf.low = pred$conf.low
DEBQ_Totaldat$conf.high = pred$conf.high

# Plotting
DEBQ_Totalplot <- ggplot(DEBQ_Totaldat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "DEBQ_Total") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DEBQ_Totalplot
```

# GAMM and initial plot - longstring
```{r}
hist(dat$longstring)
dat$longstringlog <- log(dat$longstring+1)
hist(dat$longstringlog)
gamm_longstring <- gam(longstringlog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_longstring)

plot.gam(gamm_longstring, select = 3)

hist(dat$longstring)
```


# Model-Implied Plot - longstring
```{r}

longstringdat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_longstring, newdata = longstringdat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

longstringdat$estimate = pred$estimate
longstringdat$conf.low = pred$conf.low
longstringdat$conf.high = pred$conf.high

# Plotting
longstringplot <- ggplot(longstringdat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "longstring") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

longstringplot
```

# GAMM and initial plot - BDEFS_WM_avg
```{r}
hist(dat$BDEFS_WM_avg)
dat$BDEFS_WM_avglog <- log(dat$BDEFS_WM_avg+1)
hist(dat$BDEFS_WM_avglog)
gamm_BDEFS_WM_avg <- gam(BDEFS_WM_avglog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_BDEFS_WM_avg)

plot.gam(gamm_BDEFS_WM_avg, select = 3)

hist(dat$BDEFS_WM_avg)
```


# Model-Implied Plot - BDEFS_WM_avg
```{r}

BDEFS_WM_avgdat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_BDEFS_WM_avg, newdata = BDEFS_WM_avgdat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

BDEFS_WM_avgdat$estimate = pred$estimate
BDEFS_WM_avgdat$conf.low = pred$conf.low
BDEFS_WM_avgdat$conf.high = pred$conf.high

# Plotting
BDEFS_WM_avgplot <- ggplot(BDEFS_WM_avgdat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "BDEFS_WM_avg") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

BDEFS_WM_avgplot
```


# GAMM and initial plot - BDEFS_RI_avg
```{r}
hist(dat$BDEFS_RI_avg)
dat$BDEFS_RI_avglog <- log(dat$BDEFS_RI_avg+1)
hist(dat$BDEFS_RI_avglog)
gamm_BDEFS_RI_avg <- gam(BDEFS_RI_avglog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_BDEFS_RI_avg)

plot.gam(gamm_BDEFS_RI_avg, select = 3)

hist(dat$BDEFS_RI_avg)
```


# Model-Implied Plot - BDEFS_RI_avg
```{r}

BDEFS_RI_avgdat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_BDEFS_RI_avg, newdata = BDEFS_RI_avgdat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

BDEFS_RI_avgdat$estimate = pred$estimate
BDEFS_RI_avgdat$conf.low = pred$conf.low
BDEFS_RI_avgdat$conf.high = pred$conf.high

# Plotting
BDEFS_RI_avgplot <- ggplot(BDEFS_RI_avgdat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "BDEFS_RI_avg") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

BDEFS_RI_avgplot
```


# GAMM and initial plot - UPPS_NU_avg
```{r}
hist(dat$UPPS_NU_avg)
dat$UPPS_NU_avglog <- log(dat$UPPS_NU_avg+1)
hist(dat$UPPS_NU_avglog)
gamm_UPPS_NU_avg <- gam(UPPS_NU_avglog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_UPPS_NU_avg)

plot.gam(gamm_UPPS_NU_avg, select = 3)

hist(dat$UPPS_NU_avg)
```


# Model-Implied Plot - UPPS_NU_avg
```{r}

UPPS_NU_avgdat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_UPPS_NU_avg, newdata = UPPS_NU_avgdat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

UPPS_NU_avgdat$estimate = pred$estimate
UPPS_NU_avgdat$conf.low = pred$conf.low
UPPS_NU_avgdat$conf.high = pred$conf.high

# Plotting
UPPS_NU_avgplot <- ggplot(UPPS_NU_avgdat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "UPPS_NU_avg") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

UPPS_NU_avgplot
```


# GAMM and initial plot - UPPS_Persev_avg
```{r}
hist(dat$UPPS_Persev_avg)
dat$UPPS_Persev_avglog <- log(dat$UPPS_Persev_avg+1)
hist(dat$UPPS_Persev_avglog)
gamm_UPPS_Persev_avg <- gam(UPPS_Persev_avglog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_UPPS_Persev_avg)

plot.gam(gamm_UPPS_Persev_avg, select = 3)

hist(dat$UPPS_Persev_avg)
```


# Model-Implied Plot - TEMPLATE
```{r}

UPPS_Persev_avgdat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_UPPS_Persev_avg, newdata = UPPS_Persev_avgdat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

UPPS_Persev_avgdat$estimate = pred$estimate
UPPS_Persev_avgdat$conf.low = pred$conf.low
UPPS_Persev_avgdat$conf.high = pred$conf.high

# Plotting
UPPS_Persev_avgplot <- ggplot(UPPS_Persev_avgdat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "UPPS_Persev_avg") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

UPPS_Persev_avgplot
```

# GAMM and initial plot - UPPS_Premed_avg
```{r}
hist(dat$UPPS_Premed_avg)
dat$UPPS_Premed_avglog <- log(dat$UPPS_Premed_avg+1)
hist(dat$UPPS_Premed_avglog)
gamm_UPPS_Premed_avg <- gam(UPPS_Premed_avglog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_UPPS_Premed_avg)

plot.gam(gamm_UPPS_Premed_avg, select = 3)

hist(dat$UPPS_Premed_avg)
```


# Model-Implied Plot - UPPS_Premed_avg
```{r}

UPPS_Premed_avgdat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_UPPS_Premed_avg, newdata = UPPS_Premed_avgdat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

UPPS_Premed_avgdat$estimate = pred$estimate
UPPS_Premed_avgdat$conf.low = pred$conf.low
UPPS_Premed_avgdat$conf.high = pred$conf.high

# Plotting
UPPS_Premed_avgplot <- ggplot(UPPS_Premed_avgdat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "UPPS_Premed_avg") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

UPPS_Premed_avgplot
```

# GAMM and initial plot - UPPS_Sens_avg
```{r}
hist(dat$UPPS_Sens_avg)
dat$UPPS_Sens_avglog <- log(dat$UPPS_Sens_avg+1)
hist(dat$UPPS_Sens_avglog)
gamm_UPPS_Sens_avg <- gam(UPPS_Sens_avglog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_UPPS_Sens_avg)

plot.gam(gamm_UPPS_Sens_avg, select = 3)

hist(dat$UPPS_Sens_avg)
```


# Model-Implied Plot - UPPS_Sens_avg
```{r}

UPPS_Sens_avgdat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_UPPS_Sens_avg, newdata = UPPS_Sens_avgdat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

UPPS_Sens_avgdat$estimate = pred$estimate
UPPS_Sens_avgdat$conf.low = pred$conf.low
UPPS_Sens_avgdat$conf.high = pred$conf.high

# Plotting
UPPS_Sens_avgplot <- ggplot(UPPS_Sens_avgdat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "UPPS_Sens_avg") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

UPPS_Sens_avgplot
```


# GAMM and initial plot - UPPS_PU_avg
```{r}
hist(dat$UPPS_PU_avg)
dat$UPPS_PU_avglog <- log(dat$UPPS_PU_avg+1)
hist(dat$UPPS_PU_avglog)
gamm_UPPS_PU_avg <- gam(UPPS_PU_avglog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_UPPS_PU_avg)

plot.gam(gamm_UPPS_PU_avg, select = 3)

hist(dat$UPPS_PU_avg)
```


# Model-Implied Plot - UPPS_PU_avg
```{r}

UPPS_PU_avgdat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_UPPS_PU_avg, newdata = UPPS_PU_avgdat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

UPPS_PU_avgdat$estimate = pred$estimate
UPPS_PU_avgdat$conf.low = pred$conf.low
UPPS_PU_avgdat$conf.high = pred$conf.high

# Plotting
UPPS_PU_avgplot <- ggplot(UPPS_PU_avgdat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "UPPS_PU_avg") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

UPPS_PU_avgplot
```

# GAMM and initial plot - score_pinball
```{r}
hist(dat$score_pinball)
dat$score_pinballlog <- log(dat$score_pinball+1)
hist(dat$score_pinballlog)
gamm_score_pinball <- gam(score_pinballlog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_score_pinball)

plot.gam(gamm_score_pinball, select = 3)

hist(dat$score_pinball)
```


# Model-Implied Plot - score_pinball
```{r}

score_pinballdat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_score_pinball, newdata = score_pinballdat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

score_pinballdat$estimate = pred$estimate
score_pinballdat$conf.low = pred$conf.low
score_pinballdat$conf.high = pred$conf.high

# Plotting
score_pinballplot <- ggplot(score_pinballdat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "score_pinball") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

score_pinballplot
```



# GAMM and initial plot - score_robot
```{r}
hist(dat$score_robot)
dat$score_robotlog <- log(dat$score_robot+1)
hist(dat$score_robotlog)
gamm_score_robot <- gam(score_robotlog ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_score_robot)

plot.gam(gamm_score_robot, select = 3)

hist(dat$score_robot)
```


# Model-Implied Plot - score_robot
```{r}

score_robotdat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_score_robot, newdata = score_robotdat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

score_robotdat$estimate = pred$estimate
score_robotdat$conf.low = pred$conf.low
score_robotdat$conf.high = pred$conf.high

# Plotting
score_robotplot <- ggplot(score_robotdat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "score_robot") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

score_robotplot
```



# GAMM and initial plot - DRSP_1
```{r}
hist(dat$DRSP_1)
dat$DRSP_1log <- log(dat$DRSP_1+1)
hist(dat$DRSP_1log)
gamm_DRSP_1 <- gam(DRSP_1log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_1)

plot.gam(gamm_DRSP_1, select = 3)

hist(dat$DRSP_1)
```


# Model-Implied Plot - DRSP_1
```{r}

DRSP_1dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_1, newdata = DRSP_1dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_1dat$estimate = pred$estimate
DRSP_1dat$conf.low = pred$conf.low
DRSP_1dat$conf.high = pred$conf.high

# Plotting
DRSP_1plot <- ggplot(DRSP_1dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Depressed Mood") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_1plot
```


# GAMM and initial plot - DRSP_2
```{r}
hist(dat$DRSP_2)
dat$DRSP_2log <- log(dat$DRSP_2+1)
hist(dat$DRSP_2log)
gamm_DRSP_2 <- gam(DRSP_2log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_2)

plot.gam(gamm_DRSP_2, select = 3)

hist(dat$DRSP_2)
```


# Model-Implied Plot - TEMPLATE
```{r}

DRSP_2dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_2, newdata = DRSP_2dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_2dat$estimate = pred$estimate
DRSP_2dat$conf.low = pred$conf.low
DRSP_2dat$conf.high = pred$conf.high

# Plotting
DRSP_2plot <- ggplot(DRSP_2dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Hopelessness") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_2plot
```

# GAMM and initial plot - DRSP_3
```{r}
hist(dat$DRSP_3)
dat$DRSP_3log <- log(dat$DRSP_3+1)
hist(dat$DRSP_3log)
gamm_DRSP_3 <- gam(DRSP_3log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_3)

plot.gam(gamm_DRSP_3, select = 3)

hist(dat$DRSP_3)
```


# Model-Implied Plot - DRSP_3
```{r}

DRSP_3dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_3, newdata = DRSP_3dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_3dat$estimate = pred$estimate
DRSP_3dat$conf.low = pred$conf.low
DRSP_3dat$conf.high = pred$conf.high

# Plotting
DRSP_3plot <- ggplot(DRSP_3dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Worthlessness & Guilt") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_3plot
```


# GAMM and initial plot - DRSP_4
```{r}
hist(dat$DRSP_4)
dat$DRSP_4log <- log(dat$DRSP_4+1)
hist(dat$DRSP_4log)
gamm_DRSP_4 <- gam(DRSP_4log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_4)

plot.gam(gamm_DRSP_4, select = 3)

hist(dat$DRSP_4)
```


# Model-Implied Plot - DRSP_4
```{r}

DRSP_4dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_4, newdata = DRSP_4dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_4dat$estimate = pred$estimate
DRSP_4dat$conf.low = pred$conf.low
DRSP_4dat$conf.high = pred$conf.high

# Plotting
DRSP_4plot <- ggplot(DRSP_4dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Anxiety & Tension") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_4plot
```


# GAMM and initial plot - DRSP_5
```{r}
hist(dat$DRSP_5)
dat$DRSP_5log <- log(dat$DRSP_5+1)
hist(dat$DRSP_5log)
gamm_DRSP_5 <- gam(DRSP_5log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_5)

plot.gam(gamm_DRSP_5, select = 3)

hist(dat$DRSP_5)
```


# Model-Implied Plot - DRSP_5
```{r}

DRSP_5dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_5, newdata = DRSP_5dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_5dat$estimate = pred$estimate
DRSP_5dat$conf.low = pred$conf.low
DRSP_5dat$conf.high = pred$conf.high

# Plotting
DRSP_5plot <- ggplot(DRSP_5dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Mood Swings") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_5plot
```


# GAMM and initial plot - DRSP_6
```{r}
hist(dat$DRSP_6)
dat$DRSP_6log <- log(dat$DRSP_6+1)
hist(dat$DRSP_6log)
gamm_DRSP_6 <- gam(DRSP_6log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_6)

plot.gam(gamm_DRSP_6, select = 3)

hist(dat$DRSP_6)
```


# Model-Implied Plot - DRSP_6
```{r}

DRSP_6dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_6, newdata = DRSP_6dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_6dat$estimate = pred$estimate
DRSP_6dat$conf.low = pred$conf.low
DRSP_6dat$conf.high = pred$conf.high

# Plotting
DRSP_6plot <- ggplot(DRSP_6dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Rejection Sensitivity") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_6plot
```



# GAMM and initial plot - DRSP_7
```{r}
hist(dat$DRSP_7)
dat$DRSP_7log <- log(dat$DRSP_7+1)
hist(dat$DRSP_7log)
gamm_DRSP_7 <- gam(DRSP_7log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_7)

plot.gam(gamm_DRSP_7, select = 3)

hist(dat$DRSP_7)
```


# Model-Implied Plot - DRSP_7
```{r}

DRSP_7dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_7, newdata = DRSP_7dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_7dat$estimate = pred$estimate
DRSP_7dat$conf.low = pred$conf.low
DRSP_7dat$conf.high = pred$conf.high

# Plotting
DRSP_7plot <- ggplot(DRSP_7dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Anger or Irritability") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_7plot
```


# GAMM and initial plot - DRSP_8
```{r}
hist(dat$DRSP_8)
dat$DRSP_8log <- log(dat$DRSP_8+1)
hist(dat$DRSP_8log)
gamm_DRSP_8 <- gam(DRSP_8log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_8)

plot.gam(gamm_DRSP_8, select = 3)

hist(dat$DRSP_8)
```


# Model-Implied Plot - DRSP_8
```{r}

DRSP_8dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_8, newdata = DRSP_8dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_8dat$estimate = pred$estimate
DRSP_8dat$conf.low = pred$conf.low
DRSP_8dat$conf.high = pred$conf.high

# Plotting
DRSP_8plot <- ggplot(DRSP_8dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Interpersonal Conflict") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_8plot
```

# GAMM and initial plot - DRSP_9
```{r}
hist(dat$DRSP_9)
dat$DRSP_9log <- log(dat$DRSP_9+1)
hist(dat$DRSP_9log)
gamm_DRSP_9 <- gam(DRSP_9log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_9)

plot.gam(gamm_DRSP_9, select = 3)

hist(dat$DRSP_9)
```


# Model-Implied Plot - DRSP_9
```{r}

DRSP_9dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_9, newdata = DRSP_9dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_9dat$estimate = pred$estimate
DRSP_9dat$conf.low = pred$conf.low
DRSP_9dat$conf.high = pred$conf.high

# Plotting
DRSP_9plot <- ggplot(DRSP_9dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Lack of Interest or Motivation") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_9plot
```

# GAMM and initial plot - DRSP_10
```{r}
hist(dat$DRSP_10)
dat$DRSP_10log <- log(dat$DRSP_10+1)
hist(dat$DRSP_10log)
gamm_DRSP_10 <- gam(DRSP_10log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_10)

plot.gam(gamm_DRSP_10, select = 3)

hist(dat$DRSP_10)
```


# Model-Implied Plot - DRSP_10
```{r}

DRSP_10dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_10, newdata = DRSP_10dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_10dat$estimate = pred$estimate
DRSP_10dat$conf.low = pred$conf.low
DRSP_10dat$conf.high = pred$conf.high

# Plotting
DRSP_10plot <- ggplot(DRSP_10dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Difficulty Concentrating") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_10plot
```

# GAMM and initial plot - DRSP_11
```{r}
hist(dat$DRSP_11)
dat$DRSP_11log <- log(dat$DRSP_11+1)
hist(dat$DRSP_11log)
gamm_DRSP_11 <- gam(DRSP_11log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_11)

plot.gam(gamm_DRSP_11, select = 3)

hist(dat$DRSP_11)
```


# Model-Implied Plot - DRSP_11
```{r}

DRSP_11dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_11, newdata = DRSP_11dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_11dat$estimate = pred$estimate
DRSP_11dat$conf.low = pred$conf.low
DRSP_11dat$conf.high = pred$conf.high

# Plotting
DRSP_11plot <- ggplot(DRSP_11dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Lethargy, Fatigue") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_11plot
```


# GAMM and initial plot - DRSP_12
```{r}
hist(dat$DRSP_12)
dat$DRSP_12log <- log(dat$DRSP_12+1)
hist(dat$DRSP_12log)
gamm_DRSP_12 <- gam(DRSP_12log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_12)

plot.gam(gamm_DRSP_12, select = 3)

hist(dat$DRSP_12)
```


# Model-Implied Plot - DRSP_12
```{r}

DRSP_12dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_12, newdata = DRSP_12dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_12dat$estimate = pred$estimate
DRSP_12dat$conf.low = pred$conf.low
DRSP_12dat$conf.high = pred$conf.high

# Plotting
DRSP_12plot <- ggplot(DRSP_12dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Increased Appetite, Overeating") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_12plot
```

# GAMM and initial plot - DRSP_13
```{r}
hist(dat$DRSP_13)
dat$DRSP_13log <- log(dat$DRSP_13+1)
hist(dat$DRSP_13log)
gamm_DRSP_13 <- gam(DRSP_13log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_13)

plot.gam(gamm_DRSP_13, select = 3)

hist(dat$DRSP_13)
```


# Model-Implied Plot - DRSP_13
```{r}

DRSP_13dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_13, newdata = DRSP_13dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_13dat$estimate = pred$estimate
DRSP_13dat$conf.low = pred$conf.low
DRSP_13dat$conf.high = pred$conf.high

# Plotting
DRSP_13plot <- ggplot(DRSP_13dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "DRSP_13") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_13plot
```

# GAMM and initial plot - DRSP_14
```{r}
hist(dat$DRSP_14)
dat$DRSP_14log <- log(dat$DRSP_14+1)
hist(dat$DRSP_14log)
gamm_DRSP_14 <- gam(DRSP_14log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_14)

plot.gam(gamm_DRSP_14, select = 3)

hist(dat$DRSP_14)
```


# Model-Implied Plot - DRSP_14
```{r}

DRSP_14dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_14, newdata = DRSP_14dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_14dat$estimate = pred$estimate
DRSP_14dat$conf.low = pred$conf.low
DRSP_14dat$conf.high = pred$conf.high

# Plotting
DRSP_14plot <- ggplot(DRSP_14dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Hypersomnia") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_14plot
```


# GAMM and initial plot - DRSP_15
```{r}
hist(dat$DRSP_15)
dat$DRSP_15log <- log(dat$DRSP_15+1)
hist(dat$DRSP_15log)
gamm_DRSP_15 <- gam(DRSP_15log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_15)

plot.gam(gamm_DRSP_15, select = 3)

hist(dat$DRSP_15)
```


# Model-Implied Plot - TEMPLATE
```{r}

DRSP_15dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_15, newdata = DRSP_15dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_15dat$estimate = pred$estimate
DRSP_15dat$conf.low = pred$conf.low
DRSP_15dat$conf.high = pred$conf.high

# Plotting
DRSP_15plot <- ggplot(DRSP_15dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Insomnia") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_15plot
```


# GAMM and initial plot - DRSP_16
```{r}
hist(dat$DRSP_16)
dat$DRSP_16log <- log(dat$DRSP_16+1)
hist(dat$DRSP_16log)
gamm_DRSP_16 <- gam(DRSP_16log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_16)

plot.gam(gamm_DRSP_16, select = 3)

hist(dat$DRSP_16)
```


# Model-Implied Plot - DRSP_16
```{r}

DRSP_16dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_16, newdata = DRSP_16dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_16dat$estimate = pred$estimate
DRSP_16dat$conf.low = pred$conf.low
DRSP_16dat$conf.high = pred$conf.high

# Plotting
DRSP_16plot <- ggplot(DRSP_16dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Overwhelmed") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_16plot
```


# GAMM and initial plot - DRSP_17
```{r}
hist(dat$DRSP_17)
dat$DRSP_17log <- log(dat$DRSP_17+1)
hist(dat$DRSP_17log)
gamm_DRSP_17 <- gam(DRSP_17log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_17)

plot.gam(gamm_DRSP_17, select = 3)

hist(dat$DRSP_17)
```


# Model-Implied Plot - TEMPLATE
```{r}

DRSP_17dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_17, newdata = DRSP_17dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_17dat$estimate = pred$estimate
DRSP_17dat$conf.low = pred$conf.low
DRSP_17dat$conf.high = pred$conf.high

# Plotting
DRSP_17plot <- ggplot(DRSP_17dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Out of Control") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_17plot
```



# GAMM and initial plot - DRSP_18
```{r}
hist(dat$DRSP_18)
dat$DRSP_18log <- log(dat$DRSP_18+1)
hist(dat$DRSP_18log)
gamm_DRSP_18 <- gam(DRSP_18log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_18)

plot.gam(gamm_DRSP_18, select = 3)

hist(dat$DRSP_18)
```


# Model-Implied Plot - DRSP_18
```{r}

DRSP_18dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_18, newdata = DRSP_18dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_18dat$estimate = pred$estimate
DRSP_18dat$conf.low = pred$conf.low
DRSP_18dat$conf.high = pred$conf.high

# Plotting
DRSP_18plot <- ggplot(DRSP_18dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Breast Tenderness") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_18plot
```


# GAMM and initial plot - DRSP_19
```{r}
hist(dat$DRSP_19)
dat$DRSP_19log <- log(dat$DRSP_19+1)
hist(dat$DRSP_19log)
gamm_DRSP_19 <- gam(DRSP_19log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_19)

plot.gam(gamm_DRSP_19, select = 3)

hist(dat$DRSP_19)
```


# Model-Implied Plot - DRSP_19
```{r}

DRSP_19dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_19, newdata = DRSP_19dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_19dat$estimate = pred$estimate
DRSP_19dat$conf.low = pred$conf.low
DRSP_19dat$conf.high = pred$conf.high

# Plotting
DRSP_19plot <- ggplot(DRSP_19dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Swelling, Bloating, or Weight Gain") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_19plot
```


# GAMM and initial plot - DSRP_20
```{r}
hist(dat$DRSP_20)
dat$DRSP_20log <- log(dat$DRSP_20+1)
hist(dat$DRSP_20log)
gamm_DRSP_20 <- gam(DRSP_20log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_20)

plot.gam(gamm_DRSP_20, select = 3)

hist(dat$DRSP_20)
```


# Model-Implied Plot - DSRP_20
```{r}

DRSP_20dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_20, newdata = DRSP_20dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_20dat$estimate = pred$estimate
DRSP_20dat$conf.low = pred$conf.low
DRSP_20dat$conf.high = pred$conf.high

# Plotting
DRSP_20plot <- ggplot(DRSP_20dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Headache") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_20plot
```




# GAMM and initial plot - DRSP_21
```{r}
hist(dat$DRSP_21)
dat$DRSP_21log <- log(dat$DRSP_21+1)
hist(dat$DRSP_21log)
gamm_DRSP_21 <- gam(DRSP_21log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_21)

plot.gam(gamm_DRSP_21, select = 3)

hist(dat$DRSP_21)
```


# Model-Implied Plot - DRSP_21
```{r}

DRSP_21dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_21, newdata = DRSP_21dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_21dat$estimate = pred$estimate
DRSP_21dat$conf.low = pred$conf.low
DRSP_21dat$conf.high = pred$conf.high

# Plotting
DRSP_21plot <- ggplot(DRSP_21dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Joint or Muscle Pain") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_21plot
```

# GAMM and initial plot - DRSP_22
```{r}
hist(dat$DRSP_22)
dat$DRSP_22log <- log(dat$DRSP_22+1)
hist(dat$DRSP_22log)
gamm_DRSP_22 <- gam(DRSP_22log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_22)

plot.gam(gamm_DRSP_22, select = 3)

hist(dat$DRSP_22)
```


# Model-Implied Plot - DRSP_22
```{r}

DRSP_22dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_22, newdata = DRSP_22dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_22dat$estimate = pred$estimate
DRSP_22dat$conf.low = pred$conf.low
DRSP_22dat$conf.high = pred$conf.high

# Plotting
DRSP_22plot <- ggplot(DRSP_22dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "School or Work Impairment") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_22plot
```



# GAMM and initial plot - DRSP_23
```{r}
hist(dat$DRSP_23)
dat$DRSP_23log <- log(dat$DRSP_23+1)
hist(dat$DRSP_23log)
gamm_DRSP_23 <- gam(DRSP_23log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_DRSP_23)

plot.gam(gamm_DRSP_23, select = 3)

hist(dat$DRSP_23)
```


# Model-Implied Plot - DRSP_23
```{r}

DRSP_23dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_DRSP_23, newdata = DRSP_23dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

DRSP_23dat$estimate = pred$estimate
DRSP_23dat$conf.low = pred$conf.low
DRSP_23dat$conf.high = pred$conf.high

# Plotting
DRSP_23plot <- ggplot(DRSP_23dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "Relationship Interference") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

DRSP_23plot
```






##### Reference and template info below


# OUTCOMES

BDEFS_1, BDEFS_2, BDEFS_3, BDEFS_4, BDEFS_5, BDEFS_6, DRSP_1, DRSP_2, DRSP_3, DRSP_4, DRSP_5, DRSP_6, DRSP_7, DRSP_8, DRSP_9, DRSP_10, DRSP_11, DRSP_12, DRSP_13, DRSP_14, DRSP_15, DRSP_16, DRSP_17, DRSP_18, DRSP_19, DRSP_20, DRSP_21, DRSP_22, DRSP_23, CSS_B_1, CSS_B_2, CSS_B_3, CSS_B_4, CSS_B_5, CSS_B_6, CSS_B_7, CSS_B_8, CSS_B_9, CSS_B_10, CSS_B_11, CSS_B_12, CSS_B_13, CSS_B_14, CSS_B_15, CSS_B_16, CSS_B_17, CSS_B_18, CSS_Function_1, CSS_Function_2, CSS_Function_3, CSS_Function_4, CSS_Function_5, CSS_Function_6, CSS_Function_7, CSS_Function_8, CSS_Function_9, CSS_Function_10, CSS_B2_1, CSS_B2_2, CSS_B2_3, CSS_B2_4, CSS_B2_5, CSS_B2_6, CSS_B2_7, CSS_B2_8, UPPS_1, UPPS_2, UPPS_3, UPPS_4, UPPS_5, UPPS_6, UPPS_9, UPPS_10, UPPS_11, UPPS_12, UPPS_13, UPPS_14, UPPS_15, DEBQ_1, DEBQ_2, DEBQ_3, DEBQ_4, DEBQ_5, DEBQ_6, DEBQ_7, DEBQ_8, DEBQ_9, DEBQ_10, DEBQ_11, DEBQ_12, DEBQ_13

BDEFS_WM_avg, BDEFS_RI_avg, UPPS_NU_avg, UPPS_Persev_avg, UPPS_Premed_avg, UPPS_Sens_avg, UPPS_PU_avg, game_pinball, game_name_pinball, score_pinball, user_level_pinball, session_level_pinball, game_nth_pinball, game_lpi_pinball, RecordedDate_pinball, recorded_date_pinball, hour_recorded_pinball, midnight_to_seven_pinball, game_type_pinball, date_rated_dontuse_pinball, game_robot, game_name_robot, score_robot, user_level_robot, session_level_robot, game_nth_robot, game_lpi_robot, RecordedDate_robot, recorded_date_robot, hour_recorded_robot, midnight_to_seven_robot, game_type_robot, date_rated_dontuse_robot, nth_pinball_squared, nth_robot_squared, RES_robot, RES_pinball,




# GAMM and initial plot - TEMPLATE
```{r}
hist(dat$P4)
dat$P4log <- log(dat$P4+1)
hist(dat$P4log)
gamm_p4 <- gam(P4log ~ s(id, bs = "re") + s(scaled_cycleday, id, bs = "re")  + s(scaled_cycleday, k = 20), data = dat, method = "REML")

summary(gamm_p4)

plot.gam(gamm_p4, select = 3)

hist(dat$P4)
```


# Model-Implied Plot - TEMPLATE
```{r}

P4dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1),
                      id = 0) 

# Predict using the model for each dataset and add predictions
pred <- predictions(gamm_p4, newdata = P4dat, type = "response", transform = function(x) exp(x) - 1)

#, transform = function(x) exp(x) - 1) #don't need to include anything for 'transform' if your outcome is not log-transformed. This undo's the logtransform so model-implied values are not on the log scale 

P4dat$estimate = pred$estimate
P4dat$conf.low = pred$conf.low
P4dat$conf.high = pred$conf.high

# Plotting
P4plot <- ggplot(P4dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
  labs(x = "", y = "P4") +
  geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf,
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf,
            fill = "grey87", alpha = 0.2, color = "white") +
  #geom_point(size = 3) +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal()

P4plot
```



#GAMM Deviation effects of Hormones - Inatt 

```{r}

gamm_hormd_CSS_Inatt <- gam(CSS_Inattlog ~ s(id, bs = "re") + s(E2.d, id, bs = "re") + s(P4.d, id, bs = "re") + s(E2.d, k = 10) + s(P4.d, k = 10) + ti(E2.d, P4.d, k=10), data = dat, method = "REML")

summary(gamm_hormd_CSS_Inatt)

plot.gam(gamm_hormd_CSS_Inatt, select = 6)

```


#GAMM Deviation effects of Hormones - Inatt 

```{r}

gamm_horm_CSS_Inatt <- gam(CSS_Inattlog ~ s(id, bs = "re") + s(E2, id, bs = "re") + s(P4, id, bs = "re") + s(E2, k = 10) + s(P4, k = 10) + ti(E2, P4, k=10), data = dat, method = "REML")

summary(gamm_horm_CSS_Inatt)

plot.gam(gamm_horm_CSS_Inatt, select = 6)

```

#Create person-standardized values and derivatives 

```{r}
dat <- dat %>%
    group_by(id) %>%
    mutate(
      E2zd = scale(E2),
      P4zd = scale(P4),
      E2ch = c(NA, diff(E2)),  # Daily derivative for E2
      P4ch = c(NA, diff(P4))   # Daily derivative for P4
    ) %>%
    filter(!is.na(E2zd) & !is.nan(E2zd) & !is.infinite(E2zd) &
           !is.na(P4zd) & !is.nan(P4zd) & !is.infinite(P4zd) &
           !is.na(E2ch) & !is.nan(E2ch) & !is.infinite(E2ch) &
           !is.na(P4ch) & !is.nan(P4ch) & !is.infinite(P4ch)) %>%
    ungroup()
```

 
#GAMM Deviations and Derivative effects of Hormones - Inatt 

```{r}

gamm_horm_CSS_Inatt <- gam(CSS_Inattlog ~ s(id, bs = "re") + s(E2zd) + s(E2zd, id, bs = "re") + s(P4zd) + s(P4zd, id, bs = "re") + ti(E2zd, P4zd) + s(E2ch) + s(E2ch, id, bs = "re") + s(P4ch) + s(P4ch, id, bs = "re"), data = dat, method = "REML")

summary(gamm_horm_CSS_Inatt)

plot.gam(gamm_horm_CSS_Inatt) #Plot E2zd
plot.gam(gamm_horm_CSS_Inatt, select = 4) #Plot Pz4d
plot.gam(gamm_horm_CSS_Inatt, select = 6) #Plot E2zd*P4zd
plot.gam(gamm_horm_CSS_Inatt, select = 7) #Plot E2ch
plot.gam(gamm_horm_CSS_Inatt, select = 9) #Plot P4ch

# Install and load necessary packages
install.packages("plot3D")
library(plot3D)
library(mgcv)

# Refit the model without random effects for visualization
gamm_horm_CSS_Inatt_simple <- gam(CSS_Inattlog ~ s(E2zd) + s(P4zd) + ti(E2zd, P4zd), 
                                  data = dat, method = "REML")

# Create grid of E2.d and P4.d values
e2_grid <- seq(min(dat$E2zd, na.rm = TRUE), max(dat$E2zd, na.rm = TRUE), length.out = 100)
p4_grid <- seq(min(dat$P4zd, na.rm = TRUE), max(dat$P4zd, na.rm = TRUE), length.out = 100)
grid <- expand.grid(E2zd = e2_grid, P4zd = p4_grid)

# Predict the outcome for each grid point using the simpler model
grid$predicted <- predict(gamm_horm_CSS_Inatt_simple, newdata = grid)

# Create 3D perspective plot
persp3D(x = e2_grid, y = p4_grid, z = matrix(grid$predicted, 100, 100),
        xlab = "E2zd", ylab = "P4zd", zlab = "Predicted Outcome",
        theta = 30, phi = 20, col = "lightblue")

# CONTOUR PLOT

library(ggplot2)

# Create a data frame for plotting
grid$predicted <- predict(gamm_horm_CSS_Inatt_simple, newdata = grid)

# Create contour plot with color gradients
ggplot(grid, aes(x = E2.d, y = P4.d, z = predicted)) +
  geom_tile(aes(fill = predicted)) +
  geom_contour(color = "black") +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(x = "E2.d", y = "P4.d", fill = "Predicted Outcome") +
  theme_minimal()

```

#EXAMPLE: FACET PLOT

```{r}

# Create groups for P4.d to break into facets
grid$P4.d_group <- cut(grid$P4.d, breaks = 4, labels = c("Low", "Medium-Low", "Medium-High", "High"))

# Plot the effect of E2.d at different levels of P4.d
ggplot(grid, aes(x = E2.d, y = predicted)) +
  geom_line() +
  facet_wrap(~P4.d_group) +
  labs(x = "E2.d", y = "Predicted Outcome", title = "Effect of E2.d at different levels of P4.d") +
  theme_minimal()
```


