---
title: "ADHDCYC_FailedFunctions"
author: "Tory Eisenlohr-Moul"
date: "`r Sys.Date()`"
output: html_document
---

#prep_cycle_LH_mens(): Process Data Function for Menstrual and LH Variable Preparation

# Example usage
# df <- prep_cycle_LH_mens(df, remove_participants_no_menses = TRUE, remove_participants_no_LH = TRUE, test_mode = FALSE)


# This function processes a data frame containing participant menstrual and LH test data. It performs the following operations:
# 1. Creates variables to identify first day of menses and first positive LH test.
# 2. Removes participants based on optional criteria (e.g., no positive LH test, no menses onset).
# 3. Calculates cycle days relative to menses onset and LH test.
# 4. Logs participant removals, potential issues, and outputs descriptive information.

# Parameters:
# - df: The data frame containing participant data. Must have columns: 'id', 'daterated', 'StartPeriod', 'LHposday'.
# - remove_participants_no_menses (default TRUE): If TRUE, removes participants without any menses onset.
# - remove_participants_no_LH (default TRUE): If TRUE, removes participants without any positive LH test.
# - test_mode (default FALSE): If TRUE, only processes a small subset of the data for testing purposes.

# Required Columns:
# - id: Unique identifier for each participant.
# - daterated: Date of the rating, must be in 'YYYY-MM-DD' format.
# - StartPeriod: Numeric flag indicating the start of the menstrual period (1 for start, otherwise 0).
# - LHposday: Numeric flag indicating a positive LH test (1 for positive, otherwise 0).


```{r prep_cycle_LH_mens}

prep_cycle_LH_mens <- function(df, remove_participants_no_menses = TRUE, remove_participants_no_LH = TRUE, test_mode = FALSE) {
  # Input Validation
  if (!all(c("id", "daterated", "StartPeriod", "PosLHTest") %in% names(df))) {
    stop("Error: Missing one or more required columns in the input data.")
  }

  df <- df %>%
    mutate(
      id = as.character(id),
      daterated = as.Date(daterated, format = "%Y-%m-%d"),
      StartPeriod = as.numeric(StartPeriod),
      LHposday = as.numeric(PosLHTest)
    )

  if (test_mode) {
    df <- df %>% sample_n(100)  # Process only a subset for testing
  }

  # Step 1: Create first_day_of_menses variable (1 if first day of period, otherwise 0 or NA)
  log_message("Step 1: Creating first_day_of_menses variable")
  df <- df %>%
    filter(!is.na(id)) %>%
    arrange(id, daterated) %>%
    mutate(
      first_day_of_menses = ifelse(StartPeriod == 1, 1, 0),
      mensdayone = first_day_of_menses  # Original variable name for compatibility
    )
  log_message(paste("Step 1 completed: Number of unique participant IDs remaining:", length(unique(df$id))))

  # Step 2: Create first_menses_day_flag variable (only the first menses onset day per participant)
  log_message("Step 2: Creating first_menses_day_flag variable")
  df <- df %>%
    group_by(id) %>%
    mutate(
      first_menses_day_flag = ifelse(
        first_day_of_menses == 1 & (lag(first_day_of_menses, default = 0) != 1),
        1, 0
      ),
      mensdayonefirst = first_menses_day_flag  # Original variable name for compatibility
    ) %>%
    ungroup()

  # Step 3: Create positive_LH_test_day variable (1 if positive LH test, otherwise 0 or NA)
  log_message("Step 3: Creating positive_LH_test_day variable")
  df <- df %>%
    mutate(
      positive_LH_test_day = ifelse(LHposday == 1, 1, 0),
      LHposday = positive_LH_test_day  # Original variable name for compatibility
    )

  # Step 4: Create first_positive_LH_test_day_flag variable (only the first positive LH test day per participant)
  log_message("Step 4: Creating first_positive_LH_test_day_flag variable")
  df <- df %>%
    group_by(id) %>%
    mutate(
      first_positive_LH_test_day_flag = ifelse(
        positive_LH_test_day == 1 & (lag(positive_LH_test_day, default = 0) != 1),
        1, 0
      ),
      LHposdayfirst = first_positive_LH_test_day_flag  # Original variable name for compatibility
    ) %>%
    ungroup()

  # Step 5: Create participant_level_positive_LH_test variable
  log_message("Step 5: Creating participant_level_positive_LH_test variable")
  participant_LH_test_summary <- df %>%
    group_by(id) %>%
    summarise(participant_level_positive_LH_test = ifelse(sum(first_positive_LH_test_day_flag, na.rm = TRUE) > 0, 1, 0)) %>%
    ungroup()

  df <- df %>%
    left_join(participant_LH_test_summary, by = "id")

  # Step 6: Optionally remove participants with no positive LH test
  if (remove_participants_no_LH) {
    log_message("Step 6: Removing participants with no positive LH test")
    participants_with_LH <- participant_LH_test_summary %>%
      filter(participant_level_positive_LH_test == 1) %>%
      pull(id)
    df <- df %>% filter(id %in% participants_with_LH)
  }

  # Step 7: Create cycle_day_from_menses variable (count forward and backward from first menses onset)
  log_message("Step 7: Creating cycle_day_from_menses variable")
  backward_limit <- -15
  forward_limit <- 11
  df <- df %>%
    group_by(id) %>%
    mutate(
      cycle_day_from_menses = ifelse(first_menses_day_flag == 1, 0, NA),
      cycle_day_from_menses = cumsum(!is.na(cycle_day_from_menses)) - 1,
      cycle_day_from_menses = ifelse(cycle_day_from_menses < backward_limit | cycle_day_from_menses > forward_limit, NA, cycle_day_from_menses),
      cycleday = cycle_day_from_menses  # Original variable name for compatibility
    ) %>%
    ungroup()

  # Step 8: Optionally remove participants with no menses onset
  if (remove_participants_no_menses) {
    log_message("Step 8: Removing participants with no menses onset")
    participants_with_menses <- df %>%
      group_by(id) %>%
      summarise(has_menses = any(first_menses_day_flag == 1)) %>%
      filter(has_menses) %>%
      pull(id)
    df <- df %>% filter(id %in% participants_with_menses)
  }

  # Step 9: Create cycle_day_from_positive_LH variable (count forward and backward from first positive LH test)
  log_message("Step 9: Creating cycle_day_from_positive_LH variable")
  lh_backward_limit <- -7
  lh_forward_limit <- 15
  df <- df %>%
    group_by(id) %>%
    mutate(
      cycle_day_from_positive_LH = ifelse(first_positive_LH_test_day_flag == 1, 0, NA),
      cycle_day_from_positive_LH = cumsum(!is.na(cycle_day_from_positive_LH)) - 1,
      cycle_day_from_positive_LH = ifelse(cycle_day_from_positive_LH < lh_backward_limit | cycle_day_from_positive_LH > lh_forward_limit, NA, cycle_day_from_positive_LH),
      daycountLH = cycle_day_from_positive_LH  # Original variable name for compatibility
    ) %>%
    ungroup()

  return(df)
}

# Helper function for logging messages to console and optionally to a file
log_message <- function(message) {
  cat(message, "\n")
  # Uncomment the line below to log messages to a file
  # write(message, file = "process_data_log.txt", append = TRUE)
}

# Example usage
df <- prep_cycle_LH_mens(df, remove_participants_no_menses = TRUE, remove_participants_no_LH = TRUE, test_mode = FALSE)

# Check the resulting variables
df %>% dplyr::select(id, daterated, first_menses_day_flag, first_positive_LH_test_day_flag, cycle_day_from_menses, cycle_day_from_positive_LH) %>% View()




# Looping through hormone effects - FIXED: E2.zd, P4.zd, E2.zd*P4.zd, E2zch, P4zch

```{r}

# Function to Loop Through Outcomes and Generate GAMM Models
create_gamm_models <- function(dv_list, df, save_dir) {
  for (outcome in dv_list) {
    # Prepare variables dynamically
    outcome_log <- paste0(outcome, "log")
    df[[outcome_log]] <- log(df[[outcome]] + 1)

    # Plot histogram for transformed outcome
    hist(df[[outcome]], main = paste("Histogram of", outcome))
    hist(df[[outcome_log]], main = paste("Histogram of Log-transformed", outcome))

    # GAMM model with hormone z-scores and derivatives
    gamm_formula <- as.formula(paste0(
      outcome_log, " ~ s(id, bs = \"re\") + s(E2.zd) +",
      " s(P4.zd)  + ti(E2.zd, P4.zd) +",
      " s(E2zch) + s(P4zch)"
    ))

    gamm_model <- gam(
      gamm_formula,
      data = df,
      family = gaussian,
      method = "REML"
    )

    # Summary and plot of the model
    model_summary <- summary(gamm_model)
    print(model_summary)

    # Save the model summary as a text file
    summary_output <- capture.output(model_summary)
    summary_file_path <- file.path(save_dir, paste0(outcome, "_fixed__wzch_85_model_EP_summary.txt"))
    writeLines(summary_output, summary_file_path)
    cat("Model summary saved to:", summary_file_path, "\n")

    # Plot the GAMM effects
    plot.gam(gamm_model, select = 2, main = paste("Estrogen Effect on", outcome))
    plot.gam(gamm_model, select = 3, main = paste("Progesterone Effect on", outcome))
   plot.gam(gamm_model, select = 5, main = paste("Estrogen Change Effect on", outcome))
   plot.gam(gamm_model, select = 6, main = paste("Progesterone Change Effect on", outcome))

    # Simpler model without random effects for visualization
    gamm_simple_formula <- as.formula(paste0(
      outcome_log, " ~ s(E2.zd) + s(P4.zd) + ti(E2.zd, P4.zd)"
    ))

    gamm_model_simple <- gam(
      gamm_simple_formula,
      data = df,
      family = gaussian,
      method = "REML"
    )

    # Create the grid for contour plot
    grid <- expand.grid(
      E2.zd = seq(
        min(df$E2.zd, na.rm = TRUE),
        max(df$E2.zd, na.rm = TRUE),
        length.out = 100
      ),
      P4.zd = seq(
        min(df$P4.zd, na.rm = TRUE),
        max(df$P4.zd, na.rm = TRUE),
        length.out = 100
      )
    )

    # Predict values
    grid$predicted <- predict(gamm_model_simple,
                              newdata = grid,
                              type = "response",
                              transform = function(x) exp(x) - 1)

    # Contour plot
    plot_title <- paste("Predicting", outcome, "from E2 and P4 (Person Standardized)")
    contour_plot <- ggplot() +
      geom_tile(data = grid, aes(x = E2.zd, y = P4.zd, fill = predicted)) +
      geom_contour(data = grid, aes(x = E2.zd, y = P4.zd, z = predicted), color = "black") +
      geom_point(data = df, aes(x = E2.zd, y = P4.zd), color = "black", alpha = 0.25) +
      scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = median(grid$predicted)) +
      labs(x = "E2.zd", y = "P4.zd", fill = "Outcome") +
      theme_minimal() +
      ggtitle(plot_title)

    print(contour_plot)

    # Save the contour plot
    plot_file_path <- file.path(save_dir, paste0(outcome, "_85_wzch_fixed.png"))
    ggsave(filename = plot_file_path, plot = contour_plot, width = 8, height = 6)
    cat("Contour plot saved to:", plot_file_path, "\n")
  }
}

# CALL THE LOOP


create_gamm_models(dv_list, df, "/Users/taem-macbook-air/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/2024-10-03/EP_Models_Plots")

```
