---
title: "ADHDCYCLE_daily_analysis"
author: "Tory Eisenlohr-Moul"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    number_sections: true
---

# ---- Setup ----
- Turns of Scientific Notation
- Saves 'current_date' as a string in YYYYMMDD format
- Sets the number of digits (3) to display in output
- Saves Current Date as a string in YYYYMMDD format
- Sets the 'output_folder' as path to the output folder for saving files
```{r setup}

source("scripts/options_packages.R")
source("scripts/set-output-folder.R")

```


```{r set-output-folder, quiet=TRUE, message=FALSE, warning=FALSE}
source("scripts/set-output-folder.R")

# Create the output folder if needed
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
}
```



# ---- Store Global Lists of final variable names and labels ----
- Create Lists of Variables for Analysis
-   @dv_list: List of dependent variables for analysis
-   @hormlist: List of hormone variables for analysis
-   @alldailyvars: dv_list and hormlist combined

```{r make lists, cache=TRUE}

dv_list <- c(
  "CSS_Inatt",
  "CSS_HypImp",
  "score_pinball",
  "score_robot",
  "BDEFS_Total",
  "BDEFS_WM_avg",
  "BDEFS_RI_avg",
  "UPPS_NU_avg",
  "UPPS_PU_avg",
  "UPPS_Premed_avg",
  "UPPS_Persev_avg",
  "UPPS_Sens_avg",
  "DEBQ_Total",
  "CSS_Inatt_Count",
  "CSS_Hyp_Count",
  "CSS_Imp_Count",
  "CSS_HypImp_Count",
  "DRSP_1",
  "DRSP_2",
  "DRSP_3",
  "DRSP_4",
  "DRSP_5",
  "DRSP_6",
  "DRSP_7",
  "DRSP_8",
  "DRSP_9",
  "DRSP_10",
  "DRSP_11",
  "DRSP_12",
  "DRSP_13",
  "DRSP_14",
  "DRSP_15",
  "DRSP_16",
  "DRSP_17",
  "DRSP_18",
  "DRSP_19",
  "DRSP_20",
  "DRSP_21",
  "DRSP_22",
  "DRSP_23"
) %>% noquote()



# Rename the variables based on DRSP items
# This creates a named vector to assign meaningful labels to each variable for better interpretation in the plots
names(dv_list) <- c(
  "Inattention Sx", 
  "Hyperactivity/Imp Sx", 
  "Pinball Score",
  "Robot Score",
  "BDEFS Total",
  "BDEFS Working Memory",
  "BDEFS Resp Inhibition",
  "Negative Urgency",
  "Positive Urgency",
  "Lack of Premeditation",
  "Lack of Perseverance",
  "Sensation Seeking",
  "DEBQ Total",
  "Inattention Sx Count",
  "Hyperactivity Sx Count",
  "Impulsivity Sx Count",
  "Hyp/Imp Sx Count",
  "Depressed Mood", # DRSP_1
  "Hopelessness", # DRSP_2
  "Worthlessness/Guilt", # DRSP_3
  "Anxiety/Tension", # DRSP_4
  "Mood Swings", # DRSP_5
  "Rejection Sensitivity", # DRSP_6
  "Anger/Irritability", # DRSP_7
  "Interpersonal Conflict", # DRSP_8
  "Less Interest", # DRSP_9
  "Difficulty Concentrating", # DRSP_10
  "Lethargy/Fatigue", # DRSP_11
  "Incr Appetite/Overate", # DRSP_12
  "Food Cravings", # DRSP_13
  "Hypersomnia", # DRSP_14
  "Insomnia", # DRSP_15
  "Overwhelm/Can't Cope", # DRSP_16
  "Out of Control", # DRSP_17
  "Breast Tenderness", # DRSP_18
  "Swelling/Bloat/Wt Gain", # DRSP_19
  "Joint/Muscle Pain", # DRSP_20
  "Headache", # DRSP_21
  "Work/School Imp", 
  "Relational Imp"
)

# Rename the variables based on DRSP items
names(dv_list) <- c(
  "Depressed Mood", # DRSP_1
  "hopelessness", # DRSP_2
  "Worthlessness/Guilt", # DRSP_3
  "Anxiety/Tension", # DRSP_4
  "Mood Swings", # DRSP_5
  "Rejection Sensitivity", # DRSP_6
  "Anger/Irritability", # DRSP_7
  "Interpersonal Conflict", # DRSP_8
  "Less Interest", # DRSP_9
  "Difficulty Concentrating", # DRSP_10
  "Lethargy/Fatigue", # DRSP_11
  "Increased Appetite/Overate", # DRSP_12
  "Food Cravings", # DRSP_13
  "Hypersomnia", # DRSP_14
  "Insomnia", # DRSP_15
  "Overwhelm/Can't Cope", # DRSP_16
  "Out of Control", # DRSP_17
  "Breast Tenderness", # DRSP_18
  "Swelling/Bloating/Weight Gain", # DRSP_19
  "Joint/Muscle Pain", # DRSP_20
  "Headache", # DRSP_21
  "Work/School Impairment", 
  "Relational Impairment",
  "Inattention Symptoms", 
  "Hyperactivity/Impulsivity Symptoms", 
  "Pinball Score",
  "Robot Score",
  "BDEFS Total",
  "BDEFS Working Memory",
  "BDEFS Response Inhibition",
  "Negative Urgency",
  "Positive Urgency",
  "Lack of Premeditation",
  "Lack of Perseverance",
  "Sensation Seeking",
  "DEBQ Total",
  "Inattention Sx Count",
  "Hyperactivity Sx Count",
  "Impulsivity Sx Count",
  "Hyperactivity/Impulsivity Sx Count"
)

hormlist <- c(
  "E2",
  "P4",
  "LH"
) %>% noquote()

#Create another list that combines both dv_list and hormlist above
alldailyvars <- c(dv_list, hormlist) %>% noquote()
```

# Input dataset - CLEAN AS OF 2024-10-20

```{r}
# Load your CSV file (adjust the path to your actual file location)
df <- read_csv("~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CYCLEADHD/03_analytic_projects/CYCADHD_PRIMARY/03_code_dataedits_output/adhd_daily_df_20241020.csv")
source("scripts/print.variable.names.R")
print.variable.names(df)

str(df$id)
 #set id as factor
df$id <- as.factor(df$id)
```

# ---- Setting Variable Formats  ----
- Convert daterated to Date format using lubridate if necessary (ensure date format matches)
```{r updating formats, include=TRUE}

convert_date_if_needed <- function(df, date_column) {
  # Check if the date column is in Date format
  if (!inherits(df[[date_column]], "Date")) {
    df <- df %>%
      mutate({{ date_column }} := lubridate::mdy(!!sym(date_column))) # Convert to mm/dd/yyyy format if not already in Date format
  }
  return(df)
}

# Call
df <- convert_date_if_needed(df, "daterated")

```

# ---- Sort df by id and daterated ----
```{r sort df}
df <- df %>%
  arrange(id, daterated)

source("scripts/count_rows_ids.R")
count_rows_ids(df)
```


# NOTE: Cristina Pinheiro EF Analyses Happen about here; see RMD file. 



# ---- END SETUP AND PREP  ----

# ---- START ANALYSES   ----

# Check Association between cycleday variable and days since starting

```{r, cache=TRUE}

df$TubeNumber <- as.numeric(df$TubeNumber) # Convert TubeNumber to numeric

df_first <- df %>%
  group_by(id) %>%
  filter(TubeNumber == 1) %>%
  ungroup()

hist(df_first$scaled_cycleday, breaks = 10)
hist(df_first$scaled_cycleday_ov, breaks = 10)



#cor(x=as.numeric(df$TubeNumber), y=df$scaled_cycleday, use="complete.obs")

#plot(x=as.numeric(df$TubeNumber), y=df$scaled_cycleday, use="complete.obs")

```

# 2024-10-20 - PLOTS - LUTEAL then FOLLICULAR - (OVULATION to OVULATION)
# Saved in 20241020 folder
```{r plot-ov-ov-means, eval=FALSE, cache=TRUE}

# Define the function to generate plots for a list of DVs
plot_DVs <- function(dv_list, df) {
  
  # Create person-standardized (z-scored within each participant) versions of each DV
  for (DV in dv_list) {
    zd_var <- paste0(DV, ".zd")  # Name for the person-standardized version
    
    # Add the person-standardized columns for each DV
    df <- df %>%
      group_by(id) %>%
      mutate(!!zd_var := scale(!!sym(DV), center = TRUE, scale = TRUE)) %>%
      ungroup()
  }
  
  # Calculate person-rescaled E2 and P4 (scaled within each person, sd = 0.5)
  df <- df %>%
    group_by(id) %>%
    mutate(E2.rescale = scale(E2, center = TRUE, scale = sd(E2, na.rm=TRUE)/0.5), #changed here
           P4.rescale = scale(P4, center = TRUE, scale = sd(P4, na.rm=TRUE)/0.5)) %>% #changed here
    ungroup()
  
  # Indicators for menstrual bleeding
  bleeding_indicator <- data.frame(xmin = .5, xmax = 0.7, ymin = -Inf, ymax = Inf)
  
  for (DV in dv_list) {
    
    if (DV %in% c("E2", "P4", "LH")) {
      # Raw plot for E2, P4, LH
      dv <- df %>%
        group_by(cycleday_10perc) %>%
        summarise(mean_value = mean(!!sym(DV), na.rm = TRUE),
                  se_value = sd(!!sym(DV), na.rm = TRUE) / sqrt(n()))  # SE calculation
      
      p1 <- ggplot(dv, aes(x = cycleday_10perc, y = mean_value)) +
        geom_ribbon(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                    fill = "black", alpha = 0.1) +
        geom_point(size = 1) +
        geom_line(size = 1) +
        ggtitle(paste("Raw", DV, "Plot")) +
        xlab("Cycle Day (10% bins)") + ylab(paste(DV, "Mean")) +
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5, size = 14),
              plot.margin = margin(15, 5, 50, 5)) +
        coord_cartesian(clip = "off") +
        geom_rect(data = bleeding_indicator, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf), 
                  fill = "red", alpha = 0.2, inherit.aes = FALSE)
      
      # Plot the data
      print(p1)
      
      ggsave(filename = paste0(output_folder, "/", DV, "_", "raw", ".png"), plot = last_plot(), width = 8, height = 6)

      
    } else {
      # For DVs that are NOT E2, P4, LH
      
      # Raw data points and line for the dependent variable
      dv <- df %>%
        group_by(cycleday_10perc) %>%
        summarise(mean_value = mean(!!sym(DV), na.rm = TRUE),
                  se_value = sd(!!sym(DV), na.rm = TRUE) / sqrt(n()))  # SE calculation
      
    # Person-rescaled E2 and P4 lines (thin dotted)  <- Changed variable names here
      E2_std <- df %>%
        group_by(cycleday_10perc) %>%
        summarise(mean_value = mean(E2.rescale, na.rm = TRUE)) #changed here

      p4_std <- df %>%
        group_by(cycleday_10perc) %>%
        summarise(mean_value = mean(P4.rescale, na.rm = TRUE))  #changed here
      
      p1 <- ggplot(dv, aes(x = cycleday_10perc, y = mean_value)) +
        geom_ribbon(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                    fill = "black", alpha = 0.3) +
        geom_point(size = 1) +
        geom_line(size = 2) +
        # Thin dotted lines for E2 and P4
        # geom_line(data = E2_std, aes(x = cycleday_10perc, y = mean_value), 
        #           linetype = "dotted", color = "blue", size = 1, alpha = 0.6) +  # E2
        # geom_line(data = p4_std, aes(x = cycleday_10perc, y = mean_value), 
        #           linetype = "dotted", color = "forestgreen", size = 1, alpha = 0.6) +  # P4
        ggtitle(paste("Raw", DV, "Plot")) +
        xlab("Cycle Day (10% bins)") + ylab(paste(DV, "Mean")) +
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5, size = 14),
              plot.margin = margin(15, 5, 50, 5)) +
        coord_cartesian(clip = "off") +
        geom_rect(data = bleeding_indicator, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf), 
                  fill = "red", alpha = 0.2, inherit.aes = FALSE)
      
      # Plot the data
      print(p1)
      
      # Deviations plot
      dev_var <- paste0(DV, ".d")
      dv <- df %>%
        group_by(cycleday_10perc) %>%
        summarise(mean_value = mean(!!sym(dev_var), na.rm = TRUE),
                  se_value = sd(!!sym(dev_var), na.rm = TRUE) / sqrt(n()))  # SE calculation
      
      p2 <- ggplot(dv, aes(x = cycleday_10perc, y = mean_value)) +
        geom_ribbon(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                    fill = "black", alpha = 0.3) +
        geom_point(size = 1) +
        geom_line(size = 2) +
        # geom_line(data = E2_std, aes(x = cycleday_10perc, y = mean_value), 
        #           linetype = "dotted", color = "blue", size = 1, alpha = 0.6) +  # E2
        # geom_line(data = p4_std, aes(x = cycleday_10perc, y = mean_value), 
        #           linetype = "dotted", color = "forestgreen", size = 1, alpha = 0.6) +  # P4
        ggtitle(paste("Deviations", DV, "Plot")) +
        xlab("Cycle Day (10% bins)") + ylab(paste(DV, "Deviation Mean")) +
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5, size = 14),
              plot.margin = margin(15, 5, 50, 5)) +
        coord_cartesian(clip = "off") +
        geom_rect(data = bleeding_indicator, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf), 
                  fill = "red", alpha = 0.2, inherit.aes = FALSE)
      
      print(p2)
      
            ggsave(filename = paste0(output_folder, "/", DV, "_", "dev", ".png"), plot = last_plot(), width = 8, height = 6)

      
      # Person-standardized outcome (zd)
      zd_var <- paste0(DV, ".zd")
      dv <- df %>%
        group_by(cycleday_10perc) %>%
        summarise(mean_value = mean(!!sym(zd_var), na.rm = TRUE),
                  se_value = sd(!!sym(zd_var), na.rm = TRUE) / sqrt(n()))  # SE calculation
      
      p3 <- ggplot(dv, aes(x = cycleday_10perc, y = mean_value)) +
        geom_ribbon(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                    fill = "black", alpha = 0.3) +
        geom_point(size = 1) +
        geom_line(size = 2) +
        geom_line(data = E2_std, aes(x = cycleday_10perc, y = mean_value), 
                  linetype = "solid", color = "forestgreen", size = 1, alpha = 0.9) +  # E2
        geom_line(data = p4_std, aes(x = cycleday_10perc, y = mean_value), 
                  linetype = "solid", color = "red", size = 1, alpha = 0.8) +  # P4
        ggtitle(paste("Person-standardized", DV, "Plot")) +
        xlab("Cycle Day (10% bins)") + ylab(paste(DV, "Person-standardized Mean")) +
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5, size = 14),
              plot.margin = margin(15, 5, 50, 5)) +
        coord_cartesian(clip = "off") +
        geom_rect(data = bleeding_indicator, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf), 
                  fill = "red", alpha = 0.2, inherit.aes = FALSE)
      
      print(p3)
                  ggsave(filename = paste0(output_folder, "/", DV, "_", "zd", ".png"), plot = last_plot(), width = 8, height = 6)

      
      # Rolling deviations plot
      roll_dev_var <- paste0(DV, ".d.5roll")
      dv <- df %>%
        group_by(cycleday_10perc) %>%
        summarise(mean_value = mean(!!sym(roll_dev_var), na.rm = TRUE),
                  se_value = sd(!!sym(roll_dev_var), na.rm = TRUE) / sqrt(n()))  # SE calculation
      
      p4 <- ggplot(dv, aes(x = cycleday_10perc, y = mean_value)) +
        geom_ribbon(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                    fill = "black", alpha = 0.3) +
        geom_point(size = 1) +
        geom_line(size = 2) +
        # geom_line(data = E2_std, aes(x = cycleday_10perc, y = mean_value), 
        #           linetype = "dotted", color = "blue", size = 1, alpha = 0.6) +  # E2
        # geom_line(data = p4_std, aes(x = cycleday_10perc, y = mean_value), 
        #           linetype = "dotted", color = "forestgreen", size = 1, alpha = 0.6) +  # P4
        ggtitle(paste("Rolling Deviations", DV, "Plot")) +
        xlab("Cycle Day (10% bins)") + ylab(paste(DV, "Rolling Deviation Mean")) +
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5, size = 14),
              plot.margin = margin(15, 5, 50, 5)) +
        coord_cartesian(clip = "off") +
        geom_rect(data = bleeding_indicator, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf), 
                  fill = "red", alpha = 0.2, inherit.aes = FALSE)
      
      # Plot the data
      print(p4)
                  ggsave(filename = paste0(output_folder, "/", DV, "_", "rolld", ".png"), plot = last_plot(), width = 8, height = 6)

    }
  }
}

plot_DVs(dv_list, df)

```

# 2024-10-20 - PLOTS - FOLLICULAR then LUTEAL - (MENSES to MENSES)
# Saved in 20241020 folder
```{r plot-mens-mens-means, eval=FALSE, cache=TRUE}


# Define the function to generate plots for a list of DVs
# Args:
#   dv_list: A named vector of dependent variables (DVs) to be plotted.
#   df: A data frame containing the data for analysis and plotting.
#   output_folder: The folder path where the generated plots should be saved.
plot_DVs_M2M <- function(dv_list, df, output_folder) {
  
  # Create person-standardized (z-scored within each participant) versions of each DV
  for (DV in dv_list) {
    zd_var <- paste0(DV, ".zd")  # Name for the person-standardized version
    
    # Add the person-standardized columns for each DV
    df <- df %>%
      group_by(id) %>%
      mutate(!!zd_var := scale(!!sym(DV), center = TRUE, scale = TRUE)) %>%
      ungroup()
  }
  
  # Calculate person-rescaled E2 and P4 (scaled within each person, sd = 0.5)
  # This transformation helps in visual comparison by keeping the SD consistent across participants
  df <- df %>%
    group_by(id) %>%
    mutate(E2.rescale = scale(E2, center = TRUE, scale = (sd(E2, na.rm = TRUE)*1.75)),
           P4.rescale = scale(P4, center = TRUE, scale = (sd(P4, na.rm = TRUE)*1.75))) %>%
    ungroup()
  
  # Indicators for menstrual bleeding (static values for highlighting cycle day range)
  bleeding_indicator <- data.frame(xmin = 0, xmax = 0.2, ymin = -Inf, ymax = Inf)
  
  # Loop through each DV to generate plots
  for (DV in dv_list) {
    # Get the label for the DV from dv_list
    DV_label <- names(dv_list)[which(dv_list == DV)]
    
    # Calculate mean and standard error for the raw dependent variable
    dv <- df %>%
      group_by(cycledayov_10perc) %>%
      summarise(mean_value = mean(!!sym(DV), na.rm = TRUE),
                se_value = sd(!!sym(DV), na.rm = TRUE) / sqrt(n()))  # SE calculation
    
    # Calculate mean and standard error for the person-z-scored outcome (zd)
    zd_var <- paste0(DV, ".zd")
    dv_zd <- df %>%
      group_by(cycledayov_10perc) %>%
      summarise(mean_value = mean(!!sym(zd_var), na.rm = TRUE),
                se_value = sd(!!sym(zd_var), na.rm = TRUE) / sqrt(n()))  # SE calculation
    
    # Generate plot for the person-standardized outcome (zd)
p3 <- ggplot(dv_zd, aes(x = cycledayov_10perc, y = mean_value)) +
  # Plotting the mean values for the person-standardized outcome with LOESS smoothing
  geom_smooth(method = "loess", span = .7, size = 1.5, se = FALSE, color = "black") +  # Use geom_smooth for the main DV with CI
  # Plotting the rescaled E2 and P4 values with LOESS smoothing
  geom_smooth(data = df, aes(x = cycledayov_10perc, y = E2.rescale, color = "E2"), method = "loess", span = 0.5, 
              linetype = "dashed", size = 1, alpha = 0.9, se = FALSE) +
  geom_smooth(data = df, aes(x = cycledayov_10perc, y = P4.rescale, color = "P4"), method = "loess", span = 0.5, 
              linetype = "dashed", size = 1, alpha = 0.8, se = FALSE) +
  # Manual color settings for different lines in the plot
  scale_color_manual(values = c("E2" = "forestgreen", "P4" = "red")) +
  # Set axis labels and plot title
  xlab("Cycle Day (10% bins)") + ylab(paste("Smoothed Person-Z-Scored", DV_label)) +
  # Apply minimal theme for cleaner look
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        plot.margin = margin(15, 5, 50, 5),
        legend.title = element_blank(),
        legend.position = "right") +
  coord_cartesian(clip = "off") +
  ggtitle(label=DV_label)+
  # Highlight menstrual bleeding days with a red rectangle
  geom_rect(data = bleeding_indicator, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), 
            fill = "red", alpha = 0.2, inherit.aes = FALSE)

# Print the plot to the graphics device
print(p3)
    
    # Save the plot to the specified output folder
    ggsave(filename = file.path(output_folder, paste0(DV, "_", "ov_zd_noSE", ".png")), plot = p3, width = 8, height = 6)
  }
}

# Call the function with the dataframe and output folder
plot_DVs_M2M(dv_list, df, output_folder)

```


# 2024-02-14 - FITTING GAMMs for E2, P4, LH - Have to do this first because otherwise the function below breaks
# Saved in 20250214 Folder with same dataset as in October but with factor set to ID. 

```{r}

#############################
### 1. E2 Analysis Section ###
#############################

# Plot and transform the E2 data
hist(df$E2, main = "Histogram of E2", xlab = "E2")
df$E2log <- log(df$E2 + 1)   # Log-transform E2
hist(df$E2log, main = "Histogram of log(E2+1)", xlab = "log(E2+1)")

# Ensure 'id' is a factor
df$id <- as.factor(df$id)

# Fit the GAMM model
gamm_E2 <- gam(E2log ~ s(id, bs = "re") + 
                 s(id, scaled_cycleday, bs = "re") + 
                 s(scaled_cycleday), 
               data = df, method = "REML")

# Save model summary
summary_E2 <- capture.output(summary(gamm_E2))
writeLines(summary_E2, file.path(output_folder, "E2_gamm_summary.txt"))

# Hierarchical partitioning
var_part_E2 <- gam.hp::gam.hp(gamm_E2)
var_part_E2_output <- capture.output(var_part_E2$hierarchical.partitioning)
writeLines(var_part_E2_output, file.path(output_folder, "E2_var_partitioning.txt"))

# Create a dataset for predictions
E2dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1), id = 0)
pred_E2 <- predictions(gamm_E2, newdata = E2dat, type = "response",
                       transform = function(x) exp(x) - 1)
E2dat$estimate <- pred_E2$estimate
E2dat$conf.low <- pred_E2$conf.low
E2dat$conf.high <- pred_E2$conf.high

# Create the E2 plot
E2plot <- ggplot(E2dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(
    limits = c(-1, 1),
    breaks = seq(-1, 1, by = 0.50),
    labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")
  ) +
  labs(x = "", y = "Salivary Estradiol") +
  geom_rect(aes(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf),
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(aes(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf),
            fill = "grey87", alpha = 0.2, color = "white") +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal(base_size = 16)

print(E2plot)
ggsave(
  filename = file.path(output_folder, "E2_gamm_plot.png"),
  plot = E2plot,
  width = 10,
  height = 8,
  dpi = 300,
  device = ragg::agg_png,
  bg = "white"
)

#############################
### 2. P4 Analysis Section ###
#############################

hist(df$P4, main = "Histogram of P4", xlab = "P4")
df$P4log <- log(df$P4 + 1)
hist(df$P4log, main = "Histogram of log(P4+1)", xlab = "log(P4+1)")

gamm_P4 <- gam(P4log ~ s(id, bs = "re") + 
                 s(scaled_cycleday, id, bs = "re") + 
                 s(scaled_cycleday, k = 20),
               data = df, method = "REML")

summary_P4 <- capture.output(summary(gamm_P4))
writeLines(summary_P4, file.path(output_folder, "P4_gamm_summary.txt"))

var_part_P4 <- gam.hp::gam.hp(gamm_P4)
var_part_P4_output <- capture.output(var_part_P4$hierarchical.partitioning)
writeLines(var_part_P4_output, file.path(output_folder, "P4_var_partitioning.txt"))

P4dat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1), id = 0)
pred_P4 <- predictions(gamm_P4, newdata = P4dat, type = "response",
                       transform = function(x) exp(x) - 1)
P4dat$estimate <- pred_P4$estimate
P4dat$conf.low <- pred_P4$conf.low
P4dat$conf.high <- pred_P4$conf.high

P4plot <- ggplot(P4dat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(
    limits = c(-1, 1),
    breaks = seq(-1, 1, by = 0.50),
    labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")
  ) +
  labs(x = "", y = "Salivary Progesterone") +
  geom_rect(aes(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf),
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(aes(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf),
            fill = "grey87", alpha = 0.2, color = "white") +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal(base_size = 16)

print(P4plot)
ggsave(
  filename = file.path(output_folder, "P4_gamm_plot.png"),
  plot = P4plot,
  width = 10,
  height = 8,
  dpi = 300,
  device = ragg::agg_png,
  bg = "white"
)

#############################
### 3. LH Analysis Section ###
#############################

hist(df$LH, main = "Histogram of LH", xlab = "LH")
df$LHlog <- log(df$LH + 1)
hist(df$LHlog, main = "Histogram of log(LH+1)", xlab = "log(LH+1)")

gamm_LH <- gam(LHlog ~ s(id, bs = "re") + 
                 s(scaled_cycleday, id, bs = "re") + 
                 s(scaled_cycleday, k = 20),
               data = df, method = "REML")

summary_LH <- capture.output(summary(gamm_LH))
writeLines(summary_LH, file.path(output_folder, "LH_gamm_summary.txt"))

var_part_LH <- gam.hp::gam.hp(gamm_LH)
var_part_LH_output <- capture.output(var_part_LH$hierarchical.partitioning)
writeLines(var_part_LH_output, file.path(output_folder, "LH_var_partitioning.txt"))

LHdat <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1), id = 0)
pred_LH <- predictions(gamm_LH, newdata = LHdat, type = "response",
                       transform = function(x) exp(x) - 1)
LHdat$estimate <- pred_LH$estimate
LHdat$conf.low <- pred_LH$conf.low
LHdat$conf.high <- pred_LH$conf.high

LHplot <- ggplot(LHdat, aes(x = scaled_cycleday, y = estimate)) +
  scale_x_continuous(
    limits = c(-1, 1),
    breaks = seq(-1, 1, by = 0.50),
    labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")
  ) +
  labs(x = "", y = "Salivary Progesterone") +
  geom_rect(aes(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf),
            fill = "grey70", alpha = 0.2, color = "white") +
  geom_rect(aes(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf),
            fill = "grey87", alpha = 0.2, color = "white") +
  geom_line(size = 1, show.legend = TRUE) +
  theme_minimal(base_size = 16)

print(LHplot)
ggsave(
  filename = file.path(output_folder, "LH_gamm_plot.png"),
  plot = LHplot,
  width = 10,
  height = 8,
  dpi = 300,
  device = ragg::agg_png,
  bg = "white"
)



```





# ---- Scaled Cycle Day GAMMs Predicting all outcomes on dv_list
- For each variable in @dv_list, 
- Runs GAMM and Hierarchical Variance partitioning hp.gamm()
- Saves it to a .txt file in @current_date's @output_folder in the project's Box Folder for 03_code_dataedits_output. 
- Creates two plots for each variable in @dv_list:
  - Plot 1: Predicted values and confidence intervals for the variable
  - Plot 2: Z-Scored predicted values and confidence intervals for the variable with E2 and P4 curves overlaid
```{r gamm_models_plots}
install.packages("cowplot")  # Install cowplot if needed
library(cowplot)             # Load cowplot for annotations

gamm_models_plots <- function(dv, df, output_folder, plot_path) {
  
  # Step 1: Log-transform variable
  log_dv <- paste0(dv, "_log")
  df[[log_dv]] <- log(df[[dv]] + 1)  

  # Step 2: Fit GAMM model
  gamm_model <- gam(as.formula(paste0(log_dv, " ~ s(id, bs = 're') + s(scaled_cycleday, id, bs = 're') + s(scaled_cycleday, k = 30)")), 
                     data = df, method = "REML")

  # Step 3: Extract model summary & hierarchical partitioning
  model_summary <- summary(gamm_model)
  var_part <- gam.hp::gam.hp(gamm_model)
  
  # Extract p-value and EDF
  scaled_cycleday_row <- which(rownames(model_summary$s.table) == "s(scaled_cycleday)")
  p_value <- model_summary$s.table[scaled_cycleday_row, "p-value"]
  edf <- model_summary$s.table[scaled_cycleday_row, "edf"]

  # Determine significance symbol
  significance <- ifelse(p_value < 0.001, "***", 
                  ifelse(p_value < 0.01, "**", 
                  ifelse(p_value < 0.05, "*", 
                  ifelse(p_value < 0.15, "†", ""))))

  # Save model summary and partitioning results to a file
  summary_output <- capture.output(model_summary)
  partitioning_output <- capture.output(var_part$hierarchical.partitioning)
  combined_output <- c(summary_output, "", "Hierarchical Partitioning:", partitioning_output)
  writeLines(combined_output, file.path(output_folder, paste0("gamm_hp_", dv, ".txt")))

  # Step 4: Create dataset for predictions
  pred_data <- expand.grid(scaled_cycleday = seq(-1, 1, by = 0.1), id = 0)
  pred <- predictions(gamm_model, newdata = pred_data, type = "response", transform = function(x) exp(x) - 1)

  # Add predictions and confidence intervals
  pred_data$estimate <- scale(pred$estimate)
  pred_data$conf.low <- scale(pred$conf.low)
  pred_data$conf.high <- scale(pred$conf.high)

  # Step 5: Create ggplot for predictions
  dv_plot <- ggplot(pred_data, aes(x = scaled_cycleday)) +
    scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), 
                       labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
    labs(x = "Cycle Day (Scaled)", y = paste0(dv, " (Z, Pred)")) +
    
    # Background shading for key regions
    geom_rect(aes(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf), fill = "grey70", alpha = 0.2) +
    geom_rect(aes(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf), fill = "grey87", alpha = 0.2) +
    
    # Confidence intervals
    geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "black", alpha = 0.1) +
    
    # Main predicted line
    geom_line(aes(y = estimate, color = "dv"), size = 1) +
    
    # Add E2 and P4 curves
    geom_line(data = E2dat, aes(x = scaled_cycleday, y = scale(estimate), color = "E2"), size = 1, linetype = "dashed") +
    geom_line(data = P4dat, aes(x = scaled_cycleday, y = scale(estimate), color = "P4"), size = 1, linetype = "dashed") +
    
    # Customize legend
    scale_color_manual(name = "Key", 
                       values = c("dv" = "black", "E2" = "forestgreen", "P4" = "red"),
                       labels = c("dv" = dv, "E2" = "E2", "P4" = "P4"),
                       breaks = c("dv", "E2", "P4")) +
    
    # Styling with balanced text sizes
    theme_minimal() +
    theme(
      axis.text = element_text(size = 14),    # Slightly reduced from 18
      axis.title = element_text(size = 16),   # Slightly reduced from 20
      legend.text = element_text(size = 14),  # Reduced for better fit
      legend.title = element_text(size = 16)  # Keep title slightly larger
    )

  # Step 6: Annotate with EDF & p-value
  annotation_text <- paste("EDF:", round(edf, 2), significance)
  annotated_plot <- ggdraw(dv_plot) +
    draw_label(annotation_text, x = 0.85, y = 0.2, hjust = 0, size = 10, fontface = "bold")

  # Save the plot as PDF with slightly increased size
  plot_file_path <- file.path(plot_path, paste0(dv, "_scaled_with_horm.pdf"))
  ggsave(filename = plot_file_path, plot = annotated_plot, width = 10, height = 7.5, device = cairo_pdf)
  cat("Plot with Hormone overlays saved to:", plot_file_path, "\n")

  # Step 7: Create another plot without scaling
  pred_data$estimate <- pred$estimate
  pred_data$conf.low <- pred$conf.low
  pred_data$conf.high <- pred$conf.high

  dv_plot_orig <- ggplot(pred_data, aes(x = scaled_cycleday)) +
    scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.50), 
                       labels = c("0%L", "50%L", "Menses Onset", "50%F", "Ovulation")) +
    labs(x = "Cycle Day (Scaled)", y = paste0(dv, " (Original Scale)")) +
    
    # Background shading
    geom_rect(aes(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf), fill = "grey70", alpha = 0.2) +
    geom_rect(aes(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf), fill = "grey87", alpha = 0.2) +
    
    # Confidence intervals
    geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "black", alpha = 0.1) +
    
    # Main predicted line
    geom_line(aes(y = estimate), color = "black", size = 1) +
    
    theme_minimal() +
    theme(
      axis.text = element_text(size = 14),
      axis.title = element_text(size = 16)
    )

  # Save as PDF
  plot_file_path_orig <- file.path(plot_path, paste0(dv, "_origscale.pdf"))
  ggsave(filename = plot_file_path_orig, plot = dv_plot_orig, width = 10, height = 7.5, device = cairo_pdf)
  cat("Original Scale Plot saved to:", plot_file_path_orig, "\n")
}

# Example Usage
plot_path <- file.path(output_folder, "GAMM_Plots")
for (dv in dv_list) {
  gamm_models_plots(dv, df, output_folder, plot_path)
}



```







## Prepare Hormones

```{r}

#df$logE2 = log(df$E2 + 1)
#df$logP4 = log(df$P4 + 1)

df <- df %>%
  group_by(id) %>%
  mutate(
    E2zd = scale(E2),
    P4zd = scale(P4),
    
    # Calculate 5th and 95th percentiles for E2 and P4 within each id
    E2_lower = quantile(E2zd, 0.05, na.rm = TRUE),
    E2_upper = quantile(E2zd, 0.95, na.rm = TRUE),
    P4_lower = quantile(P4zd, 0.05, na.rm = TRUE),
    P4_upper = quantile(P4zd, 0.95, na.rm = TRUE),
    
    # Manually Winsorize by clamping values to the 5th and 95th percentiles
    E2zd = pmin(pmax(E2zd, E2_lower), E2_upper),
    P4zd = pmin(pmax(P4zd, P4_lower), P4_upper),
    
    # Daily derivatives for E2 and P4
    E2zch = c(NA, diff(E2zd)),  # Winsorized E2 derivative
    P4zch = c(NA, diff(P4zd))   # Winsorized P4 derivative
  ) %>%
  filter(
    !is.na(E2zd) & !is.nan(E2zd) & !is.infinite(E2zd) &
      !is.na(P4zd) & !is.nan(P4zd) & !is.infinite(P4zd) &
      !is.na(E2zch) & !is.nan(E2zch) & !is.infinite(E2zch) &
      !is.na(P4zch) & !is.nan(P4zch) & !is.infinite(P4zch)
  ) %>%
  ungroup()


```

# ---- E2 and P4 GAMMs Predicting all outcomes on dv_list
- Several Variables were created in the dailyprep file: 
- E2.m, E2.sd, E2.d, E2.zd, E2.zd.3roll, E2.zd.5roll (for E2, P4, and LH, as well as everything in @alldailyvars)
-TODO: We need to add lags and consider derivatives and using rolling averages 
-TODO:  HOWEVER, I need the winsorization below. We will need to add the winsorization to the dailyprep file and re-run it.

# Looping through hormone effects - FIXED: E2zd, P4zd, E2zd*P4zd, E2zch, P4zch

```{r}

# Function to Loop Through Outcomes and Generate GAMM Models
create_gamm_models <- function(dv_list, df, save_dir) {
  for (outcome in dv_list) {
    # Prepare variables dynamically
    outcome_log <- paste0(outcome, "log")
    df[[outcome_log]] <- log(df[[outcome]] + 1)

    # Plot histogram for transformed outcome
    hist(df[[outcome]], main = paste("Histogram of", outcome))
    hist(df[[outcome_log]], main = paste("Histogram of Log-transformed", outcome))

    # GAMM model with hormone z-scores and derivatives
    gamm_formula <- as.formula(paste0(
      outcome_log, " ~ s(id, bs = \"re\") + s(E2zd) +",
      " s(P4zd)  + ti(E2zd, P4zd) +",
      " s(E2zch) + s(P4zch)"
    ))

    gamm_model <- gam(
      gamm_formula,
      data = df,
      family = gaussian,
      method = "REML"
    )

    # Summary and plot of the model
    model_summary <- summary(gamm_model)
    print(model_summary)

    # Save the model summary as a text file
    summary_output <- capture.output(model_summary)
    summary_file_path <- file.path(save_dir, paste0(outcome, "_fixed__wzch_85_model_EP_summary.txt"))
    writeLines(summary_output, summary_file_path)
    cat("Model summary saved to:", summary_file_path, "\n")

    # Plot the GAMM effects
    plot.gam(gamm_model, select = 2, main = paste("Estrogen Effect on", outcome))
    plot.gam(gamm_model, select = 3, main = paste("Progesterone Effect on", outcome))
   plot.gam(gamm_model, select = 5, main = paste("Estrogen Change Effect on", outcome))
   plot.gam(gamm_model, select = 6, main = paste("Progesterone Change Effect on", outcome))

    # Simpler model without random effects for visualization
    gamm_simple_formula <- as.formula(paste0(
      outcome_log, " ~ s(E2zd) + s(P4zd) + ti(E2zd, P4zd)"
    ))

    gamm_model_simple <- gam(
      gamm_simple_formula,
      data = df,
      family = gaussian,
      method = "REML"
    )

    # Create the grid for contour plot
    grid <- expand.grid(
      E2zd = seq(
        min(df$E2zd, na.rm = TRUE),
        max(df$E2zd, na.rm = TRUE),
        length.out = 100
      ),
      P4zd = seq(
        min(df$P4zd, na.rm = TRUE),
        max(df$P4zd, na.rm = TRUE),
        length.out = 100
      )
    )

    # Predict values
    grid$predicted <- predict(gamm_model_simple,
                              newdata = grid,
                              type = "response",
                              transform = function(x) exp(x) - 1)

    # Contour plot
    plot_title <- paste("Predicting", outcome, "from E2 and P4 (Person Standardized)")
    contour_plot <- ggplot() +
      geom_tile(data = grid, aes(x = E2zd, y = P4zd, fill = predicted)) +
      geom_contour(data = grid, aes(x = E2zd, y = P4zd, z = predicted), color = "black") +
      geom_point(data = df, aes(x = E2zd, y = P4zd), color = "black", alpha = 0.25) +
      scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = median(grid$predicted)) +
      labs(x = "E2zd", y = "P4zd", fill = "Outcome") +
      theme_minimal() +
      ggtitle(plot_title)

    print(contour_plot)

    # Save the contour plot
    plot_file_path <- file.path(save_dir, paste0(outcome, "_85_wzch_fixed.png"))
    ggsave(filename = plot_file_path, plot = contour_plot, width = 8, height = 6)
    cat("Contour plot saved to:", plot_file_path, "\n")
  }
}

# CALL THE LOOP


create_gamm_models(dv_list, df, output_folder)

```

# Looping through hormone effects - FIXED E2zd*P4zd TENSOR PRODUCT SMOOTH
#TODO - need to figure out how to add the intercept back to the 3dplot because the estimates are wrong due to that

```{r}
library(plotly)
library(htmlwidgets)

# Function to Loop Through Outcomes and Generate GAMM Models
create_gamm_models <- function(dv_list, df, save_dir) {
  for (outcome in dv_list) {
    # Prepare variables dynamically
    outcome_log <- paste0(outcome, "log")
    df[[outcome_log]] <- log(df[[outcome]] + 1)

    # Plot histogram for transformed outcome
    hist(df[[outcome]], main = paste("Histogram of", outcome))
    hist(df[[outcome_log]], main = paste("Histogram of Log-transformed", outcome))

    # GAMM model with hormone z-scores and derivatives
    gamm_formula <- as.formula(paste0(
      outcome_log, " ~ s(id, bs = \"re\") + te(E2zd, P4zd)"
    ))

    gamm_model <- gam(
      gamm_formula,
      data = df,
      family = gaussian,
      method = "REML"
    )

    # Summary and plot of the model
    model_summary <- summary(gamm_model)
    print(model_summary)
    
    # Step 5: Perform hierarchical partitioning
  var_part <- gam.hp::gam.hp(gamm_model)
  print(model_summary)
  print(var_part$hierarchical.partitioning)

  # Step 6: Capture the output for both model summary and hierarchical partitioning
  summary_output <- capture.output(model_summary)
  partitioning_output <- capture.output(var_part$hierarchical.partitioning)

  # Step 7: Combine the model summary and hierarchical partitioning results
  combined_output <- c(summary_output, "", "Hierarchical Partitioning:", partitioning_output)

  # Step 8: Write to file
  summary_file_path <- file.path(output_folder, paste0("gamm_HORMS_hp_", dv, ".txt"))
  writeLines(combined_output, summary_file_path)
    cat("Model summary saved to:", summary_file_path, "\n")

    # Plot the GAMM effects
   # plot.gam(gamm_model, select = 2, main = paste("Estrogen Effect on", outcome))
   # plot.gam(gamm_model, select = 3, main = paste("Progesterone Effect on", outcome))
  # plot.gam(gamm_model, select = 5, main = paste("Estrogen Change Effect on", outcome))
  # plot.gam(gamm_model, select = 6, main = paste("Progesterone Change Effect on", outcome))

    # Simpler model without random effects for visualization
    gamm_simple_formula <- as.formula(paste0(
      outcome_log, " ~ te(E2zd, P4zd)"
    ))

    gamm_model_simple <- gam(
      gamm_simple_formula,
      data = df,
      family = gaussian,
      method = "REML"
    )

    # Create the grid for contour plot
    grid <- expand.grid(
      E2zd = seq(
        min(df$E2zd, na.rm = TRUE),
        max(df$E2zd, na.rm = TRUE),
        length.out = 100
      ),
      P4zd = seq(
        min(df$P4zd, na.rm = TRUE),
        max(df$P4zd, na.rm = TRUE),
        length.out = 100
      )
    )

    # Predict values
    grid$predicted <- predict(gamm_model_simple,
                              newdata = grid,
                              type = "response",
                              transform = function(x) exp(x) - 1)

    # Contour plot
    plot_title <- paste("Predicting", outcome, "from E2 and P4 (Person Standardized)")
    contour_plot <- ggplot() +
      geom_tile(data = grid, aes(x = E2zd, y = P4zd, fill = predicted)) +
      geom_contour(data = grid, aes(x = E2zd, y = P4zd, z = predicted), color = "black") +
      geom_point(data = df, aes(x = E2zd, y = P4zd), color = "black", alpha = 0.25) +
      scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = median(grid$predicted)) +
      labs(x = "E2zd", y = "P4zd", fill = "Outcome") +
      theme_minimal() +
      ggtitle(plot_title)

    print(contour_plot)

    # Save the contour plot
    plot_file_path <- file.path(save_dir, paste0(outcome, "_te_zd_fixed.png"))
    ggsave(filename = plot_file_path, plot = contour_plot, width = 8, height = 6)
    cat("Contour plot saved to:", plot_file_path, "\n")
    
     # Create the 3D surface plot using Plotly
    grid_matrix <- matrix(grid$predicted, nrow = 100, byrow = TRUE)
    surface_plot <- plot_ly(
      x = seq(min(df$E2zd, na.rm = TRUE), max(df$E2zd, na.rm = TRUE), length.out = 100),
      y = seq(min(df$P4zd, na.rm = TRUE), max(df$P4zd, na.rm = TRUE), length.out = 100),
      z = ~grid_matrix
    ) %>%
      add_surface() %>%
      layout(
        title = paste("3D Surface Plot of E2zd and P4zd Interaction for", outcome),
        scene = list(
          xaxis = list(title = "E2zd"),
          yaxis = list(title = "P4zd"),
          zaxis = list(title = "Predicted Outcome")
        )
      )

    # Save the interactive surface plot as an HTML file
    surface_plot_file_path <- file.path(save_dir, paste0(outcome, "_zd_3D_surface.html"))
    saveWidget(surface_plot, surface_plot_file_path)
    cat("3D surface plot saved to:", surface_plot_file_path, "\n")
  }
}


# CALL THE LOOP

create_gamm_models(dv_list, df=df, save_dir=output_folder)

```

